### 2014-03-17 project-mutual-modeling.tex
Creation of the proposal Mutual Modeling for Human Robot Interaction
Currently, mostly a Latex port of the 'Is partner modeling mutual?' paper from Pierre

\documentclass{report}

\usepackage{tikz}
\usetikzlibrary{shapes}

\usepackage{subcaption}

\newcommand{\ie}{{\textit{i.e.\ }}}
\newcommand{\cf}{{\textit{cf\ }}}
\newcommand{\eg}{{\textit{e.g.\ }}}
\newcommand{\al}{{\textit{et al.\ }}}

\begin{document}

\title{Mutual Modeling in Human-Robot Interaction - Scientific part}

\chapter{Scientific Part}
\section{Summary}

\paragraph{Mutual Modelling}

In order to build a shared understanding, do partners have to build a
representation of each other's knowledge? We refer to \emph{mutual modeling} (MM) as
the process of inferring one's partner mental states. Any claim that students
carry out a detailed monitoring of their peers would be as incorrect as any
claim that they do not maintain any representation at all. If mutual modeling
had to be permanently detailed and accurate, subjects would obviously face a
huge cognitive load. Conversely, peers could not collaborate without some
minimal amount of mutual modeling. For instance, A cannot disagree with B
without knowing that B has a different opinion. The mutual model can be implicit
(A is not aware of what he knows about B), by-default (I believe that B beliefs
what I believe unless contrary evidence), opportunistic (A does not model B
unless the conversation requires it), global (A infers B's beliefs based on
categories such as age, culture or profession) and, of course, it can be
incorrect… but it can not remain empty. Dialogues include many instances of
utterances such as "I thought he would do that" (first level of MM) or even "He
thought I would do that but I intended something else." (second level of MM).

The content of mutual models ranges from 'dispositional' versus 'situational'
aspects. The 'dispositional' aspects refer to A's representation of B's long
term knowledge, skills or traits. It is thus closely related to the notion of
transactive memory (Wegner, 1987; Moreland, 2000). 'Situational' aspects refer
to A's representation of B's knowledge, behavior or intentions specifically
activated in the situation in which A and B are collaborating, some of them
being valid for 2 seconds, other ones for 2 hours. Here are examples of
fragments that constitute A's model of B regarding to aspects X, noted
$Model(A,B,X)$:

\begin{itemize}

    \item $Model(A,B, knowledge)$: What does A know about B's knowledge with
        respect to the task at hand or, inversely, about B's knowledge gaps?
        When can A consider B's statements as reliable?

    \item $Model(A,B, skills)$: What does A know about B's skills with respect to
        the task at hand? May A expect B to perform well in a specific subtask?
        The effectiveness of division of labor depends on the quality of this
        mutual model.

    \item $Model(A,B, goals)$: What does A know about B's intentions with respect
        to the project, including B' motivation and commitment? Can A trust B
        when B promises to deliver?

    \item $Model(A,B, task)$: What does A know about B's representation of the
        situation and the task: does A knows whether B has the same
        understanding of the problem at stake?

    \item $Model(A,B, plans)$: What does A know about B's strategy. Does A
        understand why B did what he did? Is A able to anticipate what B will do
        next?

    \item $Model(A,B, "urgent")$: What does A about know B's understanding of A's
        last utterance: does ‘urgent’ means now, ASAP and ‘not too lat’ ?

    \item We could continue the list of what is X is $\mathcal{M}(A,B,X)$: beliefs,
        emotions, history, status, …

\end{itemize}

We have different levels of MM. If A states "B thinks I am god in maths", we
are the second level of MM:  $Model(A, B, Model(B,A, knowledge="good in
math"))$. There is an infinite regress of nested models: If A says "B knows that
I don't expect him to solve this statistics problem" corresponds to $Model(A, B,
Model (B, A (Model (A,B, statistic-skills)))$.

A partner model is probably not a "box", \ie not a monolithical representation
but rather a mosaic of information fragments about the partner, with various
grain sizes and various life cycles. This mosaic is elaborated through various
mechanisms, first for building an initial model of the partner and then for
updating this model.  As two students meet for the first time, mutual models are
initialized by the assumptions they make upon each other from cues such as
his/her membership to large categories (age, culture, profession, ...) include
stereotypes (sportsmen, junkie, business women, Swiss,…) as well as physical
appearance. Scholars studied how initial modeling impacts communication. In
their experiments on initial MM, Slugoski, Lalljee, Lamb \& Ginsburg (1993)
pretended to their subjects that their (fake) partner had or had not received
the same information. They observed that the subjects adapted their dialogue by
focusing the explanation on the items that (s)he was supposed to ignore. Brennan
(1991) showed that the subjects used different initial strategies in forming
queries depending on who they were told their partner was.

Initial common grounds are also initiated by co-presence: they include events to
which A and B attended together (Clark \& Marshall, 1981) in the physical space
or in our cultural space (\eg ‘09-11’). Co-presence means that we can refer to
a shared objects and event,-s it does of course not imply that we give them the
same meaning. Namely, a shared screen does not mean a shared understanding
(Dillenbourg \& Traum, 2006).

After initialization, mutual models are updated along the collaborative work
through verbal and non-verbal interactions. A default inference rule is that "my
partner agrees with me unless he disagrees", which reject the critiques that
mutual modeling generates an unbearable cognitive load. This default rule is
superseded by the several mechanisms for monitoring and repairing one's partner
understanding: acknowledgement, continuous attention, relevance of next turns,
facial expressions including gaze signals, etc.

Mutual modeling does not occur in a vacuum but it is highly contextualized. REF
reviews the features of the collaborative situation, namely the media
(cotemporality… ), may facilitate or hamper mutual modeling. Hutchins (SILENCE)
reported a study in which a short silence between two pilots was perfectly
interpreted because it occurred in a highly constrained communication context.
Some environments are more productive than others in helping peers to detect
their misunderstandings. Roschelle (REF) reformulate the design of CSCL
interfaces as providing ways for peer to detect and repairs their
misunderstanding. In CSCW, researchers have explored various so-called
"awareness tools", \ie functionalities that inform A is about B's actions that
A could not directly perceived because B was working on a different subset of
the virtual space. Different awareness tools will be addressed in this
contribution since they allowed us to manipulate experimentally the MM activity.

\section{Research Plan}

\subsection{State of the Art}

\subsection{Related research of the requirers}

\subsection{Detailed research plan}

\subsubsection{Simple model of mutual modeling}

To report experiments on MM, we use a simple notational system: "A knows that B
knows X", noted $\mathcal{M}(A,B,X)$ is a simplistic but hopefully useful reduction to
reason on mutual modeling. This notation does not mean A as an explicit,
monolithic representation of B: It is an abstraction referring to complex
socio-cognitive processes.

We refer to the degree of accuracy of the mutual model as
$\mathcal{M}^{\circ}(A,B,X)$. To estimate, the MM effort, we need 3 variables.

\begin{enumerate}

    \item Tasks vary a lot with respect to how much they require.  The grounding
        criterion – denoted $\mathcal{M}^{\circ}_{min}$  - refers to how
        important it is to mutually share a piece of information X to succeed
        the task T.. It can be computed as the probability to succeed T despite
        the fact X is not grounded. In the next studies, we estimate
        $\mathcal{M}^{\circ}_{min}(A,B,X)$ as the correlation between
        $\mathcal{M}^{\circ}(A,B,X)$ and task performance.

    \item Before any specific grounding act, there is rarely a null probability
        that X is mutually understood by A and B: because X is part of A's and
        B's cultures, because it is manifest to co-present subjects or simply
        because there is not much space for misunderstanding or disagreement
        about X. We could not simply collaborate without a high level of initial
        grounds. We note the theoretical accuracy of initial grounds:
        $\mathcal{M}^{\circ}_{t1}(A,B,X)$.

    \item The cost of grounding X refers to the physical and cognitive effort
        necessary to perform a grounding act α: a verbal repair (\eg
        rephrasing), a deictic gesture, a physical move to adopt one partner's
        viewpoint, etc. This cost varies according to media features ( Clark \&
        Brennan (REF….).

\end{enumerate}

Based on these 3 parameters, the probability of making an action $\alpha_{t1}$ about
content X at time $t1$ during task T for increasing $\mathcal{M}^{\circ}(A,B,X)$
at time $t2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
costs (3)(Traum \& Dillenbourg, 1996):

$p(\alpha_{t1}(X,T)   M^{\circ}_{t2}(A,B,X)) ≈ (M^{\circ}_{min}(A,B,X,T) -
M^{\circ}_{t1}(A,B,X))/cost (\alpha)$

This is qualitative summary more that a real equation since several parameters
are hard to quantify (\eg the cost an communication acts depends upon the user
as well) but it clarifies the parameters of the experiments we report hereafter.

\begin{itemize}
    \item $\mathcal{M}^{\circ}(A,B,X)$: Our experiments address different contents that can be
        represented in mutual models:

        \begin{enumerate}
            \item $\mathcal{M}(A,B,emotion)$: how accurately A perceives B's emotional state
                (study 3);

            \item $\mathcal{M}(A,B,actions)$ is about how well A guesses what action B has
                done (study 2) or will do next  (study 1)

            \item $\mathcal{M} (A,B,knowledge)$: how accurately A estimates B's knowledge
                with respect to the material they learn together (study 4 and 5)

        \end{enumerate}

    \item $\mathcal{M}^{\circ\circ}(A,B,X,T)$ Our studies concern various collaborative tasks:
        argumentation (study 3), games (study 1 and 2) and concept mapping
        (study 4 and 5). By varying the tasks, we do actually vary the grounding
        criterion but they all require a "reasonably high" grounding criterion.
        All tasks are about building a solution or a representation together; we
        did not address everyday conversations.

    \item $\mathcal{M}^{\circ}_{t1}(A,B,X)$ Along the same reasoning, the initial degree
        of common grounds should be rather low (and hence the difference between
        initial and required degrees rather high) in order to make mutual
        modelling effort more observable. Studies 1, 4 and 5 have been conducted
        with teams of students who did not know each other. They came
        nonetheless from the same university: they hence had some general common
        grounds.  For studies 2 and 3, students knew each other before for
        reasons explained later on. In study 5, we manipulated the initial MM by
        using a JIGSAW script.

    \item $cost(\alpha)$ In all studies but study 4, the cost of  grounding is
        an independent variable. Study 3 uses media richness as independent
        variable, with the hypothesis that modeling emotions is "cheaper" with a
        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
        use awareness tools which, in principle, reduce the cost of MM, but do
        not eliminate all costs: if the tool provide A with information about
        what B does/knows, this additional information may actually increase
        cognitive load. Awareness tools constitute as a kind of MM prosthesis,
        and, like any prosthesis, they many augment MM (by facilitating it or
        even scaffolding it) or inhibit it (by making it useless).

\end{itemize}

Measuring $\mathcal{M}^{\circ}(A,B,X)$ is methodologically difficult. We
discriminate two steps, first to capture $\mathcal{M}(A,B,X)$ and then to
estimate $\mathcal{M}^{\circ}(A,B,X)$.

\begin{itemize}
    \item Capturing M (A,B,X) The simplest method is to ask A what (s)he
        believes about what B knows, feels, intends to do, etc. This raises
        obvious methodological concerns since such a question triggers a
        modeling process beyond what it would 'naturally' be. To avoid this
        bias, one can estimate mutual modeling after task completion. Then, the
        obvious drawback are memory losses and by post-hoc reconstruction. We
        have no other choice:. The first method was used in study 2 and the
        second one in the other studies.

    \item Estimating $\mathcal{M}^{\circ}(A,B,X)$. The accuracy of
        $\mathcal{M}(A,B,X)$ can be
        estimated in 2 ways.
        \begin{itemize}
            \item Subjective accuracy: In study 3, we compute
                $\mathcal{M}^{\circ}(A,B,emotions)$ by measuring if A describe
                B's emotions in the same way B reports her emotions (
                ($\mathcal{M}(A,B)=\mathcal{M}(B,B)$)

            \item Objective accuracy: In studies 4 and 5, w we compute
                $\mathcal{M}^{\circ}(A,B,knowledge)$ by comparing
                $\mathcal{M}^{\circ}A,B,K)$ to B's actual knowledge as it has
                measured by a test.

        \end{itemize}

\end{itemize}

\subsubsection{Research questions}

The experiments we report here address MM across different tasks, some with
dyads, other with triads. They were conducted along 6 years in two different
institutions by different researchers. They used different independent,
intermediate and dependent variables. Nonetheless, we were able to address a few
questions across these studies. We start with 3 simple hypotheses about
$\mathcal{M}^{\circ}(A,B)$:

\begin{itemize}
    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon A’s ability or effort
        to model B,

    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  B’s ability of
        effort to help A to model him/herself

    \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
        interactions among A and B

\end{itemize}

$\mathcal{H}_{2}$ is also called a second level modeling hypothesis since H
needs to monitor B to see if A understood him/her
$\mathcal{M}(B,(A,\mathcal{M}(A,B)))$

Our main issue is the symmetry questions: what is the relationship (fig.1)
between $\mathcal{M}^{\circ}(A,B,X)$ and $\mathcal{M}^{\circ}(B,A,X)$? A low
symmetry would mean that mutual modeling
is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high
correlation might support $\mathcal{H}_{2}$ since there is a low probability
that randomly formed pairs integrate peers with the same level of MM skills.

\begin{figure*}[htb]
\centering

\begin{tikzpicture}[scale=0.5]

\draw(0,0) node[anchor=north] (A) {\scriptsize A};
\draw(10,0) node[anchor=north] (B) {\scriptsize B};
\draw(5,2) node[anchor=north] {\scriptsize $\mathcal{M}^{\circ}(A, B, X)$};
\draw(5,-2) node[anchor=north] {\scriptsize $\mathcal{M}^{\circ}(B, A, X)$};
\draw[dashed, <->] (5,1) -- (5,-1.9) node[midway, sloped, above] {$\Delta_1$};
\draw[->] (A) to[bend left] (B);
\draw[->] (B) to[bend left] (A);

\end{tikzpicture}

\caption{The MM symmetry question $\Delta_1 =  \Delta(\mathcal{M}^{\circ} (A,B,X),
\mathcal{M}^{\circ} (B,A,X))$}

\label{mm_symetry}
\end{figure*}

With triads, we may compute the accuracy of 6 models:
$\mathcal{M}^{\circ}(A,B)$, $\mathcal{M}^{\circ}(B,A)$,
$\mathcal{M}^{\circ}(A,C)$, $\mathcal{M}^{\circ}(C,A)$,
$\mathcal{M}^{\circ}(C,B)$ and $\mathcal{M}^{\circ}(B,C)$. This enables two
triangle questions (see figure 2):

Do A and B have the same accuracy when modeling C? $\Delta2 =
\Delta(\mathcal{M}(A,C,X), \mathcal{M}(B,C,X))$ If it is the same,
$\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain over $\mathcal{H}_{1}$

Does C model more accurately A than B? $\Delta3= \Delta(\mathcal{M}(C,A,X),
\mathcal{M}(C,B,X))$ A positive answer would support $\mathcal{H}_{3}$.

\begin{figure}[htb]
\centering
\subcaptionbox{}{ 
    \begin{tikzpicture}[scale=0.5]

    \draw(0,2) node (A) {\scriptsize A};
    \draw(0,-2) node (B) {\scriptsize B};
    \draw(10,0) node (C) {\scriptsize C};

    \draw(5,2) node {\scriptsize $\mathcal{M}^{\circ}(A, C, X)$};
    \draw(5,-2) node {\scriptsize $\mathcal{M}^{\circ}(B, C, X)$};
    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above] {$\Delta_2$};
    \draw[->] (A) to (C);
    \draw[->] (B) to (C);

    \end{tikzpicture}
}
\subcaptionbox{}{ 
    \begin{tikzpicture}[scale=0.5]

    \draw(0,2) node (A) {\scriptsize A};
    \draw(0,-2) node (B) {\scriptsize B};
    \draw(10,0) node (C) {\scriptsize C};

    \draw(5,2) node {\scriptsize $\mathcal{M}^{\circ}(C, A, X)$};
    \draw(5,-2) node {\scriptsize $\mathcal{M}^{\circ}(C, B, X)$};
    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above]
    {$\Delta_3$};
    \draw[<-] (A) to (C);
    \draw[<-] (B) to (C);

    \end{tikzpicture}
}
\caption{The MM triangle questions}

\label{mm_triangle}
\end{figure}

In addition, the comparison between ($\Delta2$) and ($\Delta3$) could tell us
whether the accuracy of mutual modelling depends more upon the modeller's effort
or the modellee's behaviour.

The rectangle questions

We could go further by comparing self- versus other modeling ($\Delta4$ in Fig.
3) as an indication of metacognitive skills. We could also see if modeling
skills depend upon what aspects are being modeled (X or Y), which would explain
vertical differences ($\Delta5$ in Fig. 3). We do not further develop these
questions because we don’t have the necessary data in our studies.

\begin{figure*}[htb]
\centering

\begin{tikzpicture}[scale=0.5]

    \draw(0,0) node (a) {\scriptsize $\mathcal{M}^{\circ}(A, B, X)$};
    \draw(10,0) node (b) {\scriptsize $\mathcal{M}^{\circ}(A, A, X)$};
    \draw(10,-4) node (c) {\scriptsize $\mathcal{M}^{\circ}(A, A, Y)$};
    \draw(0,-4) node (d) {\scriptsize $\mathcal{M}^{\circ}(A, B, Y)$};
    \draw[<->] (a) -- (b) node[midway, below] {$\Delta_4$};
    \draw[<->] (b) -- (c);
    \draw[<->] (c) -- (d);
    \draw[<->] (d) -- (a) node[midway, sloped, above] {$\Delta_5$};

\end{tikzpicture}

\caption{The rectangle questions}

\label{mm_rectangle}
\end{figure*}

This simple notation does not pretend to provide a mathematical account of
mutual modeling but to be more systematic in describing the following
experiments.

\subsection{Research agenda}

\subsection{Importance of the work}

\subsection{References}

\section{International collaboration}

\end{document}

### 2014-03-20 project-mutual-modeling.tex
Using tex macro to typeset models

@@
 \newcommand{\al}{{\textit{et al.\ }}}
+
+\newcommand{\M}[3]{{\mathcal{M}(#1, #2, #3)}}
+\newcommand{\model}[3]{{$\mathcal{M}(#1, #2, #3)$}}
+\newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}
+
+\newcommand{\concept}[1]{{\small \texttt{#1}}}
+\newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
 
@@
 
-    \item We could continue the list of what is X is $\mathcal{M}(A,B,X)$: beliefs,
+    \item We could continue the list of what is X is \model{A}{B}{X}: beliefs,
         emotions, history, status, …
@@
 his/her membership to large categories (age, culture, profession, ...) include
-stereotypes (sportsmen, junkie, business women, Swiss,…) as well as physical
+stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
 appearance. Scholars studied how initial modeling impacts communication. In
-their experiments on initial MM, Slugoski, Lalljee, Lamb \& Ginsburg (1993)
-pretended to their subjects that their (fake) partner had or had not received
-the same information. They observed that the subjects adapted their dialogue by
-focusing the explanation on the items that (s)he was supposed to ignore. Brennan
-(1991) showed that the subjects used different initial strategies in forming
-queries depending on who they were told their partner was.
+their experiments on initial MM, Slugoski, Lalljee, Lamb \&
+Ginsburg~\cite{slugoski1993attribution} pretended to their subjects that their
+(fake) partner had or had not received the same information. They observed that
+the subjects adapted their dialogue by focusing the explanation on the items
+that (s)he was supposed to ignore. Brennan (1991) showed that the subjects used
+different initial strategies in forming queries depending on who they were told
+their partner was.
 
 Initial common grounds are also initiated by co-presence: they include events to
-which A and B attended together (Clark \& Marshall, 1981) in the physical space
+which A and B attended together~\cite{clark2002definite} in the physical space
 or in our cultural space (\eg ‘09-11’). Co-presence means that we can refer to
 a shared objects and event,-s it does of course not imply that we give them the
-same meaning. Namely, a shared screen does not mean a shared understanding
-(Dillenbourg \& Traum, 2006).
+same meaning. Namely, a shared screen does not mean a shared
+understanding~\cite{dillenbourg2006sharing}.
 
@@
 Some environments are more productive than others in helping peers to detect
-their misunderstandings. Roschelle (REF) reformulate the design of CSCL
-interfaces as providing ways for peer to detect and repairs their
-misunderstanding. In CSCW, researchers have explored various so-called
-"awareness tools", \ie functionalities that inform A is about B's actions that
-A could not directly perceived because B was working on a different subset of
-the virtual space. Different awareness tools will be addressed in this
-contribution since they allowed us to manipulate experimentally the MM activity.
+their misunderstandings. Roschelle~\cite{roschelle1995construction} reformulate
+the design of CSCL interfaces as providing ways for peer to detect and repairs
+their misunderstanding. In CSCW, researchers have explored various so-called
+"awareness tools", \ie functionalities that inform A is about B's actions that A
+could not directly perceived because B was working on a different subset of the
+virtual space. Different awareness tools will be addressed in this contribution
+since they allowed us to manipulate experimentally the MM activity.
 
@@
 
+\paragraph{Building a model of a human during the interaction} Perspective
+Taking is a human ability which allows one to put him/herself in another
+person's point of view. Studied in psychology
+literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
+interacting with people by allowing one to reason on others' understanding of
+the world in terms of visual perception, spatial descriptions, affordances and
+beliefs, etc.  In the last years these notions have been gradually employed in
+HRI.~\cite{Breazeal2006} presents a learning algorithm that takes into account
+information about a teacher's visual perspective in order to learn a task.
+~\cite{Johnson2005} apply visual perspective taking for action recognition
+between two robots.~\cite{Trafton2005} use both visual and spatial perspective
+taking to find out the referent indicated by a human partner.
+
+The cognitive model that the robot builds for the agent it interacts with is
+still simple and mostly focused on geometric features (who sees what? What are
+our relative positions? etc.). Extending this knowledge with more subtle
+perceptions (emotional state for instance) remains to be done.
+
+\paragraph{Joint action}
+Several theories dealing with collaboration~\cite{Cohen1991,Grosz1996,Clark1996}
+emphasize that collaborative tasks have specific requirements compared to
+individual ones, \eg, since the robot and the person share a common goal, they
+have to agree on the manner to realize it, they must show their commitment to
+the goal during execution, etc. Several robotic systems have already been built
+based on these theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they all
+have shown benefits of this approach. They have also shown how difficult it is
+to manage turn-taking between communication partners and to interleave task
+realization and communication in a generic way. Finally, today only few
+systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
+levels.
+
 \subsection{Related research of the requirers}
+
+Our knowledge base implements such a mechanism: when the robots infers a new
+agent has been introduced in the knowledge base, it initializes a new,
+independent, ontology for this agent. All the ontologies that are created share
+the same common-sense knowledge, but rely on each agent's perspective for the
+actual instantiation: the robot (geometrically) computes that the book is in
+its own field of view, but not in the human one. The robot knowledge contains
+the fact \stmt{book isVisible true} while the human model contains \stmt{book
+isVisible false}.
+
+Using separate knowledge models in the knowledge base, we have been able to
+replicate the \emph{Sally and Ann} false-belief experiment with our
+robots~\cite{warnier2012when}.
+
+In our implementation, perspective taking is tightly connected to the symbolic
+knowledge models, and since our knowledge base allows for storage of one
+knowledge model per agent, we have been able to endow the robot with a simple
+theory of mind: we explicitly model
+what the robot knows about its partners in a symbolic way. This knowledge is
+then re-used in different places, to correctly interpret what the human says,
+or to plan tasks that are actually doable for the human.
 
@@
 To report experiments on MM, we use a simple notational system: "A knows that B
-knows X", noted $\mathcal{M}(A,B,X)$ is a simplistic but hopefully useful reduction to
+knows X", noted \model{A}{B}{X} is a simplistic but hopefully useful reduction to
 reason on mutual modeling. This notation does not mean A as an explicit,
@@
 We refer to the degree of accuracy of the mutual model as
-$\mathcal{M}^{\circ}(A,B,X)$. To estimate, the MM effort, we need 3 variables.
+\Model{A}{B}{X}. To estimate, the MM effort, we need 3 variables.
 
@@
         $\mathcal{M}^{\circ}_{min}(A,B,X)$ as the correlation between
-        $\mathcal{M}^{\circ}(A,B,X)$ and task performance.
+        \Model{A}{B}{X} and task performance.
 
@@
         viewpoint, etc. This cost varies according to media features ( Clark \&
-        Brennan (REF….).
+        Brennan (REF)).
 
@@
 Based on these 3 parameters, the probability of making an action $\alpha_{t1}$ about
-content X at time $t1$ during task T for increasing $\mathcal{M}^{\circ}(A,B,X)$
+content X at time $t1$ during task T for increasing \Model{A}{B}{X}
 at time $t2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
@@
 \begin{itemize}
-    \item $\mathcal{M}^{\circ}(A,B,X)$: Our experiments address different contents that can be
+    \item \Model{A}{B}{X}: Our experiments address different contents that can be
         represented in mutual models:
@@
         \begin{enumerate}
-            \item $\mathcal{M}(A,B,emotion)$: how accurately A perceives B's emotional state
+            \item \model{A}{B}{emotion}: how accurately A perceives B's emotional state
                 (study 3);
 
-            \item $\mathcal{M}(A,B,actions)$ is about how well A guesses what action B has
+            \item \model{A}{B}{actions} is about how well A guesses what action B has
                 done (study 2) or will do next  (study 1)
 
-            \item $\mathcal{M} (A,B,knowledge)$: how accurately A estimates B's knowledge
+            \item \model{A}{B}{knowledge}: how accurately A estimates B's knowledge
                 with respect to the material they learn together (study 4 and 5)
@@
 
-    \item $\mathcal{M}^{\circ\circ}(A,B,X,T)$ Our studies concern various collaborative tasks:
+    \item \Model{A,B}{X}{T} Our studies concern various collaborative tasks:
         argumentation (study 3), games (study 1 and 2) and concept mapping
@@
 
-Measuring $\mathcal{M}^{\circ}(A,B,X)$ is methodologically difficult. We
-discriminate two steps, first to capture $\mathcal{M}(A,B,X)$ and then to
-estimate $\mathcal{M}^{\circ}(A,B,X)$.
+Measuring \Model{A}{B}{X} is methodologically difficult. We
+discriminate two steps, first to capture \model{A}{B}{X} and then to
+estimate \Model{A}{B}{X}.
 
@@
 
-    \item Estimating $\mathcal{M}^{\circ}(A,B,X)$. The accuracy of
-        $\mathcal{M}(A,B,X)$ can be
+    \item Estimating \Model{A}{B}{X}. The accuracy of
+        \model{A}{B}{X} can be
         estimated in 2 ways.
@@
             \item Subjective accuracy: In study 3, we compute
-                $\mathcal{M}^{\circ}(A,B,emotions)$ by measuring if A describe
-                B's emotions in the same way B reports her emotions (
+                \Model{A}{B}{emotions} by measuring if A describe
+                B's emotions in the same way B reports her emotions 
                 ($\mathcal{M}(A,B)=\mathcal{M}(B,B)$)
@@
             \item Objective accuracy: In studies 4 and 5, w we compute
-                $\mathcal{M}^{\circ}(A,B,knowledge)$ by comparing
+                \Model{A}{B}{knowledge} by comparing
                 $\mathcal{M}^{\circ}A,B,K)$ to B's actual knowledge as it has
@@
 
-Our main issue is the symmetry questions: what is the relationship (fig.1)
-between $\mathcal{M}^{\circ}(A,B,X)$ and $\mathcal{M}^{\circ}(B,A,X)$? A low
-symmetry would mean that mutual modeling
-is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high
-correlation might support $\mathcal{H}_{2}$ since there is a low probability
-that randomly formed pairs integrate peers with the same level of MM skills.
+Our main issue is the symmetry questions: what is the relationship
+(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and
+\Model{B}{A}{X}? A low symmetry would mean that mutual modeling is
+mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
+might support $\mathcal{H}_{2}$ since there is a low probability that randomly
+formed pairs integrate peers with the same level of MM skills.
 
@@
 \draw(10,0) node[anchor=north] (B) {\scriptsize B};
-\draw(5,2) node[anchor=north] {\scriptsize $\mathcal{M}^{\circ}(A, B, X)$};
-\draw(5,-2) node[anchor=north] {\scriptsize $\mathcal{M}^{\circ}(B, A, X)$};
+\draw(5,2) node[anchor=north] {\scriptsize \Model{A}{ B}{ X}};
+\draw(5,-2) node[anchor=north] {\scriptsize \Model{B}{ A}{ X}};
 \draw[dashed, <->] (5,1) -- (5,-1.9) node[midway, sloped, above] {$\Delta_1$};
@@
 
-\label{mm_symetry}
+\label{mm_symmetry}
 \end{figure*}
@@
 With triads, we may compute the accuracy of 6 models:
-$\mathcal{M}^{\circ}(A,B)$, $\mathcal{M}^{\circ}(B,A)$,
-$\mathcal{M}^{\circ}(A,C)$, $\mathcal{M}^{\circ}(C,A)$,
-$\mathcal{M}^{\circ}(C,B)$ and $\mathcal{M}^{\circ}(B,C)$. This enables two
-triangle questions (see figure 2):
+\Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
+\Model{C}{B}{X} and \Model{B}{C}{X}. This enables two
+triangle questions (see Fig.~\ref{mm_triangles}):
 
-Do A and B have the same accuracy when modeling C? $\Delta2 =
-\Delta(\mathcal{M}(A,C,X), \mathcal{M}(B,C,X))$ If it is the same,
+Do A and B have the same accuracy when modeling C? $\Delta_2 =
+\Delta(\M{A}{C}{X}, \M{B}{C}{X})$ If it is the same,
 $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain over $\mathcal{H}_{1}$
 
-Does C model more accurately A than B? $\Delta3= \Delta(\mathcal{M}(C,A,X),
-\mathcal{M}(C,B,X))$ A positive answer would support $\mathcal{H}_{3}$.
+Does C model more accurately A than B? $\Delta_3= \Delta(\M{C}{A}{X},
+\M{C}{B}{X})$ A positive answer would support $\mathcal{H}_{3}$.
 
@@
 
-    \draw(5,2) node {\scriptsize $\mathcal{M}^{\circ}(A, C, X)$};
-    \draw(5,-2) node {\scriptsize $\mathcal{M}^{\circ}(B, C, X)$};
+    \draw(5,2) node {\scriptsize \Model{A}{ C}{ X}};
+    \draw(5,-2) node {\scriptsize \Model{B}{ C}{ X}};
     \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above] {$\Delta_2$};
@@
 
-    \draw(5,2) node {\scriptsize $\mathcal{M}^{\circ}(C, A, X)$};
-    \draw(5,-2) node {\scriptsize $\mathcal{M}^{\circ}(C, B, X)$};
+    \draw(5,2) node {\scriptsize \Model{C}{ A}{ X}};
+    \draw(5,-2) node {\scriptsize \Model{C}{ B}{ X}};
     \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above]
@@
 
-\label{mm_triangle}
+\label{mm_triangles}
 \end{figure}
 
-In addition, the comparison between ($\Delta2$) and ($\Delta3$) could tell us
+In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
 whether the accuracy of mutual modelling depends more upon the modeller's effort
@@
 
-The rectangle questions
+\paragraph{The rectangle questions}
 
-We could go further by comparing self- versus other modeling ($\Delta4$ in Fig.
-3) as an indication of metacognitive skills. We could also see if modeling
-skills depend upon what aspects are being modeled (X or Y), which would explain
-vertical differences ($\Delta5$ in Fig. 3). We do not further develop these
-questions because we don’t have the necessary data in our studies.
+We could go further by comparing self- versus other modeling $\Delta_4$ in
+Fig.~\ref{mm_rectangle}) as an indication of metacognitive skills. We could also
+see if modeling skills depend upon what aspects are being modeled (X or Y),
+which would explain vertical differences ($\Delta_5$ in
+Fig.~\ref{mm_rectangle}). We do not further develop these questions because we
+don’t have the necessary data in our studies.
 
@@
 
-    \draw(0,0) node (a) {\scriptsize $\mathcal{M}^{\circ}(A, B, X)$};
-    \draw(10,0) node (b) {\scriptsize $\mathcal{M}^{\circ}(A, A, X)$};
-    \draw(10,-4) node (c) {\scriptsize $\mathcal{M}^{\circ}(A, A, Y)$};
-    \draw(0,-4) node (d) {\scriptsize $\mathcal{M}^{\circ}(A, B, Y)$};
+    \draw(0,0) node (a) {\scriptsize \Model{A}{ B}{ X}};
+    \draw(10,0) node (b) {\scriptsize \Model{A}{ A}{ X}};
+    \draw(10,-4) node (c) {\scriptsize \Model{A}{ A}{ Y}};
+    \draw(0,-4) node (d) {\scriptsize \Model{A}{ B}{ Y}};
     \draw[<->] (a) -- (b) node[midway, below] {$\Delta_4$};

### 2014-03-20 project-mutual-modeling.tex
Added the biblio
Added sections on robotic + theory of mind, false belief

@@
 
+Human-robot interaction is a challenge for artificial intelligence. This field
+indeed lays at the crossroad of several other domains of AI and requires to
+tackle them in a holistic manner: Modeling humans and human cognition;
+acquiring, representing, manipulating in a tractable way abstract knowledge;
+reasoning on this knowledge to make decisions; and eventually instantiating
+those decisions into physical actions in coordination with humans. Many AI
+techniques are invited, from visual processing to symbolic reasoning, from task
+planning to \emph{theory of mind} building, from reactive control to action
+recognition and learning.
+
+\paragraph{Theory of Mind}
+
+Theory of Mind (originally defined in~\cite{premack1978does}) is the cognitive
+ability that a subject possesses to represent the mental state of another
+agent, possibly including knowledge that contradicts the subject's own model: for
+example, a book can be at the same time \emph{visible} for myself, and \emph{not
+visible} for you.
+
+Children develop this skill, which is essential to understand others'
+perspectives during interactions, around the age of three. It supposes the
+ability to build, store and retrieve separate models of the knowledge of the
+interactors.
+
+One classical application of this cognitive skill is the so-called
+\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
+experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
+people, A and B, manipulate objects. Then A leaves and B hides away one
+object. When A comes back, we ask the child ``where do you think A will
+look for the object?''. Before acquiring a theory of mind, children are not
+able to separate their own (true) model of the world (where they know that
+the object was hidden) from the model of A, which contains \emph{false
+beliefs} on the world (A still thinks the object is at its original
+position since he did not see B hiding it).
+
 \paragraph{Mutual Modelling}
@@
 term knowledge, skills or traits. It is thus closely related to the notion of
-transactive memory (Wegner, 1987; Moreland, 2000). 'Situational' aspects refer
-to A's representation of B's knowledge, behavior or intentions specifically
-activated in the situation in which A and B are collaborating, some of them
-being valid for 2 seconds, other ones for 2 hours. Here are examples of
-fragments that constitute A's model of B regarding to aspects X, noted
-$Model(A,B,X)$:
+transactive memory~\cite{wegner1987transactive, moreland1999transactive}.
+'Situational' aspects refer to A's representation of B's knowledge, behavior or
+intentions specifically activated in the situation in which A and B are
+collaborating, some of them being valid for 2 seconds, other ones for 2 hours.
+Here are examples of fragments that constitute A's model of B regarding to
+aspects X, noted $Model(A,B,X)$:
 
@@
 
-\subsubsection{Simple model of mutual modeling}
+\subsubsection{Research questions}
 
-To report experiments on MM, we use a simple notational system: "A knows that B
-knows X", noted \model{A}{B}{X} is a simplistic but hopefully useful reduction to
-reason on mutual modeling. This notation does not mean A as an explicit,
-monolithic representation of B: It is an abstraction referring to complex
+The core research question we want to address in this project is
+$(\mathcal{Q}_1)$ how mutual modeling impacts the quality of the human-robot
+interaction?  \emph{Quality}, in our context, is measured as the performance of
+the human-robot pair in selected joint tasks.
+
+To actually instigate a mutual modeling situation, three other research
+questions must be addressed. $(\mathcal{Q}_2)$ how can the robot acquire and
+represent a mental model of its partners? $(\mathcal{Q}_3)$ conversely, how do
+the human partners model the robot and what are the robot's design/behavioural
+factors that influences it? $(\mathcal{Q}_4)$ once the robot detects
+misunderstandings, what should be the interaction strategy?
+\subsubsection{A theoretical model of mutual modeling}
+
+To report experiments on MM, we use a simple notational system: ``A knows that B
+knows X", noted \model{A}{B}{X}. This is a simple but hopefully useful reduction to
+reason on mutual modeling. This notation does not mean A has an explicit,
+monolithic representation of B: it must be understood as an abstraction referring to complex
 socio-cognitive processes.
@@
 We refer to the degree of accuracy of the mutual model as
-\Model{A}{B}{X}. To estimate, the MM effort, we need 3 variables.
+\Model{A}{B}{X}. To estimate the MM effort, we need 3 variables.
 
@@
 
-\subsubsection{Research questions}
+\subsubsection{Studies}
 
@@
 \begin{itemize}
-    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon A’s ability or effort
+    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon A's ability or effort
         to model B,
 
-    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  B’s ability of
+    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  B's ability of
         effort to help A to model him/herself
@@
 
+\bibliographystyle{abbrv}
+\bibliography{biblio}
+
 \section{International collaboration}

### 2014-03-20 project-mutual-modeling.tex
Mostly better latex formatting

@@
 In order to build a shared understanding, do partners have to build a
-representation of each other's knowledge? We refer to \emph{mutual modeling} (MM) as
+representation of each other's knowledge? We refer to \emph{mutual modeling} as
 the process of inferring one's partner mental states. Any claim that students
@@
 incorrect… but it can not remain empty. Dialogues include many instances of
-utterances such as "I thought he would do that" (first level of MM) or even "He
-thought I would do that but I intended something else." (second level of MM).
+utterances such as "I thought he would do that" (first level of mutual
+modelling) or even "He thought I would do that but I intended something else."
+(second level of mutual modelling).
 
@@
 
-We have different levels of MM. If A states "B thinks I am god in maths", we
-are the second level of MM:  $Model(A, B, Model(B,A, knowledge="good in
+We have different levels of mutual modelling. If A states "B thinks I am god in maths", we
+are the second level of mutual modelling:  $Model(A, B, Model(B,A, knowledge="good in
 math"))$. There is an infinite regress of nested models: If A says "B knows that
@@
 appearance. Scholars studied how initial modeling impacts communication. In
-their experiments on initial MM, Slugoski, Lalljee, Lamb \&
+their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
 Ginsburg~\cite{slugoski1993attribution} pretended to their subjects that their
@@
 virtual space. Different awareness tools will be addressed in this contribution
-since they allowed us to manipulate experimentally the MM activity.
+since they allowed us to manipulate experimentally the mutual modelling activity.
 
@@
 
-\subsection{Related research of the requirers}
+\subsection{Related research of the applicants}
 
-Our knowledge base implements such a mechanism: when the robots infers a new
-agent has been introduced in the knowledge base, it initializes a new,
-independent, ontology for this agent. All the ontologies that are created share
-the same common-sense knowledge, but rely on each agent's perspective for the
-actual instantiation: the robot (geometrically) computes that the book is in
-its own field of view, but not in the human one. The robot knowledge contains
-the fact \stmt{book isVisible true} while the human model contains \stmt{book
-isVisible false}.
+Pierre Dillenbourg has been working since several years on the role and importance
+of mutual modelling in human-human interaction, specifically in the context of
+learning sciences.
 
-Using separate knowledge models in the knowledge base, we have been able to
-replicate the \emph{Sally and Ann} false-belief experiment with our
-robots~\cite{warnier2012when}.
+He already conducted several experiments~\cite{sangin2007partner,
+nova2007collaboration} that supported the idea that modelling partners is not only
+an individual skill. When engaged in a social task that could benefit from
+mutual modelling, the accuracy of the models build by each of the participant on
+the other participants increase.
 
-In our implementation, perspective taking is tightly connected to the symbolic
-knowledge models, and since our knowledge base allows for storage of one
-knowledge model per agent, we have been able to endow the robot with a simple
-theory of mind: we explicitly model
-what the robot knows about its partners in a symbolic way. This knowledge is
-then re-used in different places, to correctly interpret what the human says,
-or to plan tasks that are actually doable for the human.
+In the frame of this work, Pierre Dillenbourg has proposed a simple formalism
+that we describe below, and that we will use subsequently to describe the mutual
+modelling situations.
 
-\subsection{Detailed research plan}
+\paragraph{A simple notational system for mutual modeling} We suggest
+the notation \model{A}{B}{X} to denote ``A knows that B knows X". This is a
+simple but hopefully useful reduction to reason on mutual modeling. This
+notation does not mean A has an explicit, monolithic representation of B: it
+must be understood as an abstraction referring to complex socio-cognitive
+processes. Besides, we refer to the \emph{degree of accuracy} of the model as
+\Model{A}{B}{X}.
 
-\subsubsection{Research questions}
-
-The core research question we want to address in this project is
-$(\mathcal{Q}_1)$ how mutual modeling impacts the quality of the human-robot
-interaction?  \emph{Quality}, in our context, is measured as the performance of
-the human-robot pair in selected joint tasks.
-
-To actually instigate a mutual modeling situation, three other research
-questions must be addressed. $(\mathcal{Q}_2)$ how can the robot acquire and
-represent a mental model of its partners? $(\mathcal{Q}_3)$ conversely, how do
-the human partners model the robot and what are the robot's design/behavioural
-factors that influences it? $(\mathcal{Q}_4)$ once the robot detects
-misunderstandings, what should be the interaction strategy?
-\subsubsection{A theoretical model of mutual modeling}
-
-To report experiments on MM, we use a simple notational system: ``A knows that B
-knows X", noted \model{A}{B}{X}. This is a simple but hopefully useful reduction to
-reason on mutual modeling. This notation does not mean A has an explicit,
-monolithic representation of B: it must be understood as an abstraction referring to complex
-socio-cognitive processes.
-
-We refer to the degree of accuracy of the mutual model as
-\Model{A}{B}{X}. To estimate the MM effort, we need 3 variables.
+To estimate the mutual modelling effort, we need 3 variables:
 
@@
     \item Tasks vary a lot with respect to how much they require.  The grounding
-        criterion – denoted $\mathcal{M}^{\circ}_{min}$  - refers to how
+        criterion -- denoted $\mathcal{M}^{\circ}_{min}$ -- refers to how
         important it is to mutually share a piece of information X to succeed
-        the task T.. It can be computed as the probability to succeed T despite
-        the fact X is not grounded. In the next studies, we estimate
-        $\mathcal{M}^{\circ}_{min}(A,B,X)$ as the correlation between
-        \Model{A}{B}{X} and task performance.
+        the task T. It can be computed as the probability to succeed T despite
+        the fact X is not grounded. $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be
+        estimated from the correlation between \Model{A}{B}{X} and task
+        performance.
 
-    \item Before any specific grounding act, there is rarely a null probability
+    \item Before any specific grounding act, there is usually a non-null probability
         that X is mutually understood by A and B: because X is part of A's and
@@
         grounds. We note the theoretical accuracy of initial grounds:
-        $\mathcal{M}^{\circ}_{t1}(A,B,X)$.
+        $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
 
     \item The cost of grounding X refers to the physical and cognitive effort
-        necessary to perform a grounding act α: a verbal repair (\eg
+        necessary to perform a grounding act $\alpha$: a verbal repair (\eg
         rephrasing), a deictic gesture, a physical move to adopt one partner's
@@
 
-Based on these 3 parameters, the probability of making an action $\alpha_{t1}$ about
-content X at time $t1$ during task T for increasing \Model{A}{B}{X}
-at time $t2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
-costs (3)(Traum \& Dillenbourg, 1996):
+Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
+content X at time $t_1$ during task T in order to increase \Model{A}{B}{X}
+at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
+costs (3)~\cite{traum1996miscommunication}:
 
-$p(\alpha_{t1}(X,T)   M^{\circ}_{t2}(A,B,X)) ≈ (M^{\circ}_{min}(A,B,X,T) -
-M^{\circ}_{t1}(A,B,X))/cost (\alpha)$
+\begin{equation}
+p(\alpha_{t_1}(X,T) / M^{\circ}_{t_2}(A,B,X) \nearrow) \approx \frac{(M^{\circ}_{min}(A,B,X,T) -
+M^{\circ}_{t_1}(A,B,X))}{cost (\alpha)}
+\end{equation}
 
@@
 are hard to quantify (\eg the cost an communication acts depends upon the user
-as well) but it clarifies the parameters of the experiments we report hereafter.
+as well) but it clarifies the parameters of the experiments we propose hereafter.
 
@@
         grounds.  For studies 2 and 3, students knew each other before for
-        reasons explained later on. In study 5, we manipulated the initial MM by
+        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
         using a JIGSAW script.
@@
         richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
-        use awareness tools which, in principle, reduce the cost of MM, but do
+        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
         not eliminate all costs: if the tool provide A with information about
         what B does/knows, this additional information may actually increase
-        cognitive load. Awareness tools constitute as a kind of MM prosthesis,
-        and, like any prosthesis, they many augment MM (by facilitating it or
+        cognitive load. Awareness tools constitute as a kind of mutual modelling prosthesis,
+        and, like any prosthesis, they many augment mutual modelling (by facilitating it or
         even scaffolding it) or inhibit it (by making it useless).
@@
 might support $\mathcal{H}_{2}$ since there is a low probability that randomly
-formed pairs integrate peers with the same level of MM skills.
+formed pairs integrate peers with the same level of mutual modelling skills.
 
@@
 
-\caption{The MM symmetry question $\Delta_1 =  \Delta(\mathcal{M}^{\circ} (A,B,X),
+\caption{The mutual modelling symmetry question $\Delta_1 =  \Delta(\mathcal{M}^{\circ} (A,B,X),
 \mathcal{M}^{\circ} (B,A,X))$}
@@
 }
-\caption{The MM triangle questions}
+\caption{The mutual modelling triangle questions}
 

### 2014-03-20 project-mutual-modeling.tex
Formatting, syntax

@@
-\documentclass{report}
+\documentclass{article}
 
@@
 \usepackage{subcaption}
+
+\usepackage[draft]{fixme}
 
@@
 
+\newcommand{\groundingcriterion}{{$\mathcal{M}^{\circ}_{min}$}}
+\newcommand{\inigrounding}{{$\mathcal{M}^{\circ}_{t_0}$}}
+
 \newcommand{\concept}[1]{{\small \texttt{#1}}}
@@
 
+\title{Is Partner Modeling Mutual?}
+
 \begin{document}
+\maketitle
 
-\title{Mutual Modeling in Human-Robot Interaction - Scientific part}
+\begin{abstract}
+    TDB
+\end{abstract}
 
-\chapter{Scientific Part}
-\section{Summary}
+\section{Introduction}
 
-Human-robot interaction is a challenge for artificial intelligence. This field
-indeed lays at the crossroad of several other domains of AI and requires to
-tackle them in a holistic manner: Modeling humans and human cognition;
-acquiring, representing, manipulating in a tractable way abstract knowledge;
-reasoning on this knowledge to make decisions; and eventually instantiating
-those decisions into physical actions in coordination with humans. Many AI
-techniques are invited, from visual processing to symbolic reasoning, from task
-planning to \emph{theory of mind} building, from reactive control to action
-recognition and learning.
+From its very beginning, CSCL research has been following Roschelle \& Teasley's
+(1986) suggestion that collaborative learning has something to do with the
+process of constructing and maintaining a shared understanding of the task at
+hand. Building a shared/mutual understanding refers to the upper class of
+collaborative learning situations, those in which students should build upon
+each other's understanding to refine their own understanding. What is expected
+to produce learning is not the mere fact that two students build the same
+understanding but the cognitive effort they have to engage to build this shared
+understanding (Schwartz, 1995). This effort can be observed by the frequency of
+rich interactions, i.e. interactions whose occurrence has been related to
+learning: (self-) explanations in cognitive science (Chi ; Webb), conflict
+resolution in socio-cognitive theories (Doise), mutual regulation (Blaye,) in a
+Vygostkian perspective, etc. The construction of a shared understanding has been
+investigated for several years in psycholinguistics, under the  notion of
+“grounding” (Clark \&Wilkes-Gibbs, 1986). However, the relevance of grounding
+mechanisms for explaining learning outcomes has been questioned. The monitoring
+and repair of mis-understanding explains for instance referential failures in
+short dialogue episodes but does hardly predict conceptual change over longer
+sessions (Dillenbourg \& Traum, 2006). The cumulative effect of grounding
+episodes can probably be better understood from a socio-cultural perspective:
+"collaborative learning is associated with the increased cognitive-interactional
+effort involved in the transition from learning to understand each other to
+learning to understand the meanings of the semiotic tools that constitute the
+mediators of interpersonal interaction" (Baker, Hansen, Joiner \& Traum, 1999).
+Several scholars suggest that CSCL research should go deeper in the
+understanding of how partners engage into shared meaning making (Stahl, 2006) or
+'intersubjective' meaning making (Suthers (2006).
 
-\paragraph{Theory of Mind}
+Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
+change viewpoint, it is criticized at the same time as being too cognitivist by
+some psycholinguists, i.e. as overestimating the amount shared knowledge and
+mutual representations actually necessary to conduct a dialogue. The fundamental
+issue, as old as philosophy, is the degree of coupling between the different
+levels of dialogue, mostly between the lexical / syntactical level and the
+deeper semantic levels. Pickering \& Garrod (2004) argue that the mutual
+understanding starts mostly with a 'local alignment' at the level of the
+linguistic representations, due to priming mechanisms, and that this superficial
+alignment may – in some cases- lead to a 'global alignment' of the semantic
+level.  For these authors, the convergence in dialogue, and even the repair of
+some misunderstandings, is explained by this mimetic behavior more than by a
+monitoring of each other's knowledge: "…interlocutors do not need to monitor and
+develop full common ground as a regular, constant part of routine conversation,
+as it would be unnecessary and far too costly. Establishment of full common
+ground is, we argue, a specialized and non-automatic process that is used
+primarily in times of difficulty (when radical misalignment becomes apparent)."
+(p. 179). This view is actually not incompatible with Clark's grounding
+criterion (REF): the degree of shared understanding that peers need to reach
+depends upon the task they perform. For instance, a dialogue between two
+surgeons might rely on superficial alignment if they talk about their friends
+but has to guarantee accurate common grounds when talking about which
+intervention will be conducted in which way on which patient.  Some CSCL
+scholars stressed the 'illusion of shared understanding' (JACCO?), i.e. the
+potential decoupling between linguistic alignment and actual shared meanings.
 
-Theory of Mind (originally defined in~\cite{premack1978does}) is the cognitive
-ability that a subject possesses to represent the mental state of another
-agent, possibly including knowledge that contradicts the subject's own model: for
-example, a book can be at the same time \emph{visible} for myself, and \emph{not
-visible} for you.
+This interesting cognitive science debate mostly occurred outside the field of
+learning. In education, the question is to relate these mechanisms to learning
+outcomes. Is linguistic alignment sufficient to trigger conceptual change? Does
+negotiation of meaning only occurs when partners monitor and diagnose each
+other's knowledge. If the ratio between shallow alignment and deep grounding
+depends upon the task, and if deep grounding is a condition for learning, then
+the pedagogical challenge is to design tasks that require deep grounding. Most
+empirical studies on grounding and alignment are conducted with tasks, which
+despite being qualified as 'ecologically valid' by their authors, are mere
+referencing tasks such as asking the way to the train station or helping the
+peer to choose a picture among many. In this contribution, we explore several
+richer tasks such as arguing on a sensitive issue or building a concept map.
 
-Children develop this skill, which is essential to understand others'
-perspectives during interactions, around the age of three. It supposes the
-ability to build, store and retrieve separate models of the knowledge of the
-interactors.
+Deep grounding or shared meaning making requires some cognitive load. For Clark
+\& Wilkes-Gibbs (1986), what is important is not the individual effort made by
+the receiver of a communicative act, but the overall least collaborative effort.
+The cost of producing a perfect utterance may be higher than the cost of
+repairing the problems that may arise through misunderstandings. For instance,
+subjects are less careful about adapting their utterances to their partner when
+they know they can provide feedback on his/her understanding (Schober, 1993). We
+introduced the notion of ‘optimal collaborative effort’ (Dillenbourg et al,
+1996) to stress that misunderstanding should not be viewed as something to be
+avoided (if this was possible), but as an opportunity to engage into
+verbalization, explanation, negotiation, and so forth. This issue is related to
+the global argument regarding cognitive load in learning activities, especially
+in discovery learning environments: there is no learning without some cognitive
+load, but overload may hinder learning (Paas, Renkl \& Sweller, 2003). In the
+context of collaborative learning, we understand the cognitive load induced by
+mutual modeling as part of Schwartz’s (1995) notion of effort towards a shared
+understanding. For instance, CSCL researchers expanded the use of 'collaboration
+scripts' (Aronson,?). A script is a pedagogical method that collaborative
+learning activities in order to foster the emergence of productive interactions
+such as argumentation, explanation or conflict. Conflict-resolution scripts such
+as the ArgueGraph (Dillenbourg \& Hong, 2008) forms pair of students with
+opposite opinions, which increases the difficulty of consensus building,
+requiring more justifications, more negotiation, more load. Similarly, JIGSAW
+scripts provide peers with different but complementary knowledge for augmenting
+(reasonably) the effort group members have to engage to reach a shared solution.
 
-One classical application of this cognitive skill is the so-called
-\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
-experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
-people, A and B, manipulate objects. Then A leaves and B hides away one
-object. When A comes back, we ask the child ``where do you think A will
-look for the object?''. Before acquiring a theory of mind, children are not
-able to separate their own (true) model of the world (where they know that
-the object was hidden) from the model of A, which contains \emph{false
-beliefs} on the world (A still thinks the object is at its original
-position since he did not see B hiding it).
+%\paragraph{Theory of Mind}
+%
+%Theory of Mind (originally defined in~\cite{premack1978does}) is the cognitive
+%ability that a subject possesses to represent the mental state of another
+%agent, possibly including knowledge that contradicts the subject's own model: for
+%example, a book can be at the same time \emph{visible} for myself, and \emph{not
+%visible} for you.
+%
+%Children develop this skill, which is essential to understand others'
+%perspectives during interactions, around the age of three. It supposes the
+%ability to build, store and retrieve separate models of the knowledge of the
+%interactors.
+%
+%One classical application of this cognitive skill is the so-called
+%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
+%experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
+%people, A and B, manipulate objects. Then A leaves and B hides away one
+%object. When A comes back, we ask the child ``where do you think A will
+%look for the object?''. Before acquiring a theory of mind, children are not
+%able to separate their own (true) model of the world (where they know that
+%the object was hidden) from the model of A, which contains \emph{false
+%beliefs} on the world (A still thinks the object is at its original
+%position since he did not see B hiding it).
 
-\paragraph{Mutual Modelling}
+\section{Mutual Modeling}
 
@@
 the subjects adapted their dialogue by focusing the explanation on the items
-that (s)he was supposed to ignore. Brennan (1991) showed that the subjects used
+that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
 different initial strategies in forming queries depending on who they were told
@@
 
-\section{Research Plan}
+%\subsection{State of the Art}
+%
+%\paragraph{Building a model of a human during the interaction} Perspective
+%Taking is a human ability which allows one to put him/herself in another
+%person's point of view. Studied in psychology
+%literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
+%interacting with people by allowing one to reason on others' understanding of
+%the world in terms of visual perception, spatial descriptions, affordances and
+%beliefs, etc.  In the last years these notions have been gradually employed in
+%HRI.~\cite{Breazeal2006} presents a learning algorithm that takes into account
+%information about a teacher's visual perspective in order to learn a task.
+%~\cite{Johnson2005} apply visual perspective taking for action recognition
+%between two robots.~\cite{Trafton2005} use both visual and spatial perspective
+%taking to find out the referent indicated by a human partner.
+%
+%The cognitive model that the robot builds for the agent it interacts with is
+%still simple and mostly focused on geometric features (who sees what? What are
+%our relative positions? etc.). Extending this knowledge with more subtle
+%perceptions (emotional state for instance) remains to be done.
+%
+%\paragraph{Joint action}
+%Several theories dealing with collaboration~\cite{Cohen1991,Grosz1996,Clark1996}
+%emphasize that collaborative tasks have specific requirements compared to
+%individual ones, \eg, since the robot and the person share a common goal, they
+%have to agree on the manner to realize it, they must show their commitment to
+%the goal during execution, etc. Several robotic systems have already been built
+%based on these theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they all
+%have shown benefits of this approach. They have also shown how difficult it is
+%to manage turn-taking between communication partners and to interleave task
+%realization and communication in a generic way. Finally, today only few
+%systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
+%levels.
+%
+\section{A model for mutual modelling evaluation}
 
-\subsection{State of the Art}
-
-\paragraph{Building a model of a human during the interaction} Perspective
-Taking is a human ability which allows one to put him/herself in another
-person's point of view. Studied in psychology
-literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
-interacting with people by allowing one to reason on others' understanding of
-the world in terms of visual perception, spatial descriptions, affordances and
-beliefs, etc.  In the last years these notions have been gradually employed in
-HRI.~\cite{Breazeal2006} presents a learning algorithm that takes into account
-information about a teacher's visual perspective in order to learn a task.
-~\cite{Johnson2005} apply visual perspective taking for action recognition
-between two robots.~\cite{Trafton2005} use both visual and spatial perspective
-taking to find out the referent indicated by a human partner.
-
-The cognitive model that the robot builds for the agent it interacts with is
-still simple and mostly focused on geometric features (who sees what? What are
-our relative positions? etc.). Extending this knowledge with more subtle
-perceptions (emotional state for instance) remains to be done.
-
-\paragraph{Joint action}
-Several theories dealing with collaboration~\cite{Cohen1991,Grosz1996,Clark1996}
-emphasize that collaborative tasks have specific requirements compared to
-individual ones, \eg, since the robot and the person share a common goal, they
-have to agree on the manner to realize it, they must show their commitment to
-the goal during execution, etc. Several robotic systems have already been built
-based on these theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they all
-have shown benefits of this approach. They have also shown how difficult it is
-to manage turn-taking between communication partners and to interleave task
-realization and communication in a generic way. Finally, today only few
-systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
-levels.
-
-\subsection{Related research of the applicants}
-
-Pierre Dillenbourg has been working since several years on the role and importance
-of mutual modelling in human-human interaction, specifically in the context of
-learning sciences.
-
-He already conducted several experiments~\cite{sangin2007partner,
-nova2007collaboration} that supported the idea that modelling partners is not only
-an individual skill. When engaged in a social task that could benefit from
-mutual modelling, the accuracy of the models build by each of the participant on
-the other participants increase.
-
-In the frame of this work, Pierre Dillenbourg has proposed a simple formalism
-that we describe below, and that we will use subsequently to describe the mutual
-modelling situations.
-
-\paragraph{A simple notational system for mutual modeling} We suggest
+To report experiments on mutual modeling, we suggest
 the notation \model{A}{B}{X} to denote ``A knows that B knows X". This is a
-simple but hopefully useful reduction to reason on mutual modeling. This
+simple but useful reduction to reason on mutual modelling. This
 notation does not mean A has an explicit, monolithic representation of B: it
@@
 
-To estimate the mutual modelling effort, we need 3 variables:
+We parametrize and assess the mutual modelling effort through 3 variables:
 
@@
     \item Tasks vary a lot with respect to how much they require.  The grounding
-        criterion -- denoted $\mathcal{M}^{\circ}_{min}$ -- refers to how
+        criterion -- denoted \groundingcriterion -- refers to how
         important it is to mutually share a piece of information X to succeed
@@
 
-The experiments we report here address MM across different tasks, some with
+The experiments we report here address mutual modelling across different tasks, some with
 dyads, other with triads. They were conducted along 6 years in two different
@@
 
-\subsection{Research agenda}
-
-\subsection{Importance of the work}
-
-\subsection{References}
-
 \bibliographystyle{abbrv}
@@
 
-\section{International collaboration}
-
 \end{document}

### 2014-03-24 project-mutual-modeling.tex
Formatting, syntax

@@
 \newcommand{\model}[3]{{$\mathcal{M}(#1, #2, #3)$}}
+\newcommand{\refmodel}[2]{{$\mathcal{M}(#1, #2)$}}
 \newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}
@@
 
-\begin{equation}
+\begin{equation} \label{eq:probrepair}
 p(\alpha_{t_1}(X,T) / M^{\circ}_{t_2}(A,B,X) \nearrow) \approx \frac{(M^{\circ}_{min}(A,B,X,T) -
@@
 
-Measuring \Model{A}{B}{X} is methodologically difficult. We
+Each of the studies that we present in the next section have been design in such a way that, while tractable on a
+robot, mutual modelling plays a key role for the task performance: the tasks
+require a ``reasonably high" grounding criterion \groundingcriterion,
+and all tasks are about building a solution or a representation together.
+
+Along the same reasoning, the initial degree of common grounds \inigrounding
+should be rather low (and hence the difference between initial and required
+degrees rather high) in order to make mutual modelling effort more observable.
+While we can exactly control the actual initial knowledge available to the
+robot, estimating \inigrounding still requires special care during the design of
+the experiments: as we have previously shown~\cite{lemaignan2014dynamics,
+lemaignan2014cognitive} and because the interaction takes place between a human and
+a robot, the initial human model of the robot can widely vary depending on
+several factors like the user familiarity with technology, the effect of
+novelty, the user expectations or the anthropomorphic features of the robot.
+
+More generally, while we have explicit access to \model{R}{H}{X} (from the facts
+stored in the robot belief state, as illustrated in
+Table~\ref{table|beliefsfig7}), estimating \Model{R}{H}{X} also requires
+\model{H}{H}{X}, which depends on the meta-cognitive competencies of the human
+partner.
+
+Conversely, measuring \Model{A}{B}{X} is methodologically difficult. We
 discriminate two steps, first to capture \model{A}{B}{X} and then to
@@
 \begin{itemize}
-    \item Capturing M (A,B,X) The simplest method is to ask A what (s)he
-        believes about what B knows, feels, intends to do, etc. This raises
+    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what (s)he
+        believes about what $B$ knows, feels, intends to do, etc. This raises
         obvious methodological concerns since such a question triggers a
@@
         bias, one can estimate mutual modeling after task completion. Then, the
-        obvious drawback are memory losses and by post-hoc reconstruction. We
-        have no other choice:. The first method was used in study 2 and the
-        second one in the other studies.
+        obvious drawback are memory losses and by post-hoc reconstruction.
+        The first method was used in study 2 and the
+        second one in the other studies. Another option would be to rely on external behavioural metrics like
+        eye-tracking: we hypothesise for instance that the fixation time
+        reflects the efforts engaged by the human to understand, hence, model,
+        the others. Such an approach has however not yet been investigated in the presented studies.
 
-    \item Estimating \Model{A}{B}{X}. The accuracy of
-        \model{A}{B}{X} can be
-        estimated in 2 ways.
+    \item {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
+        need to access the reference model \refmodel{B}{X} to estimate its
+        \emph{accuracy}. Since we can only indirectly access it via what $B$
+        reports (\ie \model{B}{B}{X}), accuracy can be estimated in 2 ways:
+
         \begin{itemize}
+
             \item Subjective accuracy: In study 3, we compute
-                \Model{A}{B}{emotions} by measuring if A describe
-                B's emotions in the same way B reports her emotions 
-                ($\mathcal{M}(A,B)=\mathcal{M}(B,B)$)
+                \Model{A}{B}{X} by measuring if $A$ describes
+                $B$'s emotions in the same way $B$ reports its emotions 
+                ($\M{A}{B}{X} = \M{B}{B}{X}$).
 
-            \item Objective accuracy: In studies 4 and 5, w we compute
-                \Model{A}{B}{knowledge} by comparing
-                $\mathcal{M}^{\circ}A,B,K)$ to B's actual knowledge as it has
-                measured by a test.
+            \item Objective accuracy: In studies 4 and 5, we compute
+                \Model{A}{B}{X} by comparing \Model{A}{B}{K} to $B$'s
+                actual knowledge as it has measured by a test.
 

### 2014-08-26 project-mutual-modeling.tex
Turn it into a paper

--- 2014-03-20 project-mutual-modeling.tex
@@
 \newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}
-
-\newcommand{\groundingcriterion}{{$\mathcal{M}^{\circ}_{min}$}}
-\newcommand{\inigrounding}{{$\mathcal{M}^{\circ}_{t_0}$}}
 
@@
 %
-\section{A model for mutual modelling evaluation}
+\section{A simple notation system}
 
@@
 the notation \model{A}{B}{X} to denote ``A knows that B knows X". This is a
-simple but useful reduction to reason on mutual modelling. This
+simple but hopefully useful reduction to reason on mutual modeling. This
 notation does not mean A has an explicit, monolithic representation of B: it
@@
 
-We parametrize and assess the mutual modelling effort through 3 variables:
+To estimate the mutual modelling effort, we need 3 variables:
 
@@
     \item Tasks vary a lot with respect to how much they require.  The grounding
-        criterion -- denoted \groundingcriterion -- refers to how
+        criterion -- denoted $\mathcal{M}^{\circ}_{min}$ -- refers to how
         important it is to mutually share a piece of information X to succeed
@@
 
-The experiments we report here address mutual modelling across different tasks, some with
+The experiments we report here address MM across different tasks, some with
 dyads, other with triads. They were conducted along 6 years in two different

### 2014-08-26 project-mutual-modeling.tex
Imported Study 1 from original paper

--- 2014-03-24 project-mutual-modeling.tex
@@
 \documentclass{article}
+
+\usepackage[english]{babel}
+
+\usepackage{graphicx}
+\graphicspath{{figs/}}
 
@@
 
+\section{Studies}
+
+\subsection{Study 1: Effect of an awareness tool on \Model{A}{B} in a virtual
+game}
+
+\paragraph{Questions}
+
+We studied the impact of an awareness tool on group performance and mutual
+modeling. (Nova, Wehrle, Goslin, Bourquin \& Dillenbourg (2006). The availability
+of an awareness tool was our independent variable. In previous studies, we
+replayed a video of the game to subjects who surprised us by their ability to
+remember former states of their mutual model: "I did that because I thought that
+you would do that". Hence, this experiment focused the representation of each
+other's action plans. During the game, we asked them to anticipate the next
+action of their partner as well as to announce their own action. It is clear
+asking them a question about their partner increases their attention for the
+rest of the game.
+
+\paragraph{Experimental setting}
+
+SpaceMiners is a 3D game that involves two players collecting minerals located
+in asteroids. To do so, they shoot drones through the space after choosing their
+initial direction and speed. Once launched, the trajectory of drones is only
+influenced by the gravity of planets and by trajectory modification tools.
+During the experiment, the teams were confronted with three increasingly complex
+situations. The experiment was 2 hours long: a 30 minutes tutorial and 3 levels
+of 30 minutes. Players uses using a regular joystick and communicated with each
+other through an audio channel.
+
+The independent variable was the availability of an awareness tool that shows to
+player A  the location and gaze direction of player B. In the awareness
+condition, players could switch to the 'scout mode' where they could view what
+their partner was looking at. We hypothesize that this would enable subjects to
+more accurately infer their teammate's intentions. Each player sat in front of a
+distinct computer located in different rooms.
+
+\paragraph{Subjects}
+
+Thirty-six persons participated in this study, all native French speakers. We
+constituted 18 pairs of participants (N = 18) who were not familiar with each
+other. The pairs were assigned randomly to either the control condition (without
+the awareness tool) or the awareness condition (with the awareness tool).
+
+\begin{figure}
+        \centering
+        \includegraphics[width=0.9\columnwidth]{image4.png}
+        \caption{Snapshot of the SpaceMiner Game}
+        \label{spaceminer}
+\end{figure}
+
+\paragraph{Variables}
+
+Task performance was measured by the score reached by the two subjects after
+three situations. The effort of mutual modeling was measured as the ratio of
+time that players would spend in the scout mode (divided by total time), which
+is the time during which players are not performing their own actions but
+monitoring their partner's actions.
+
+In order to evaluate m°(A,B) during the task, we used two questionnaires (Figure
+4) that were displayed during each of the three games, as a transparent layer
+appearing over the game display. The first questionnaire concerned the player’s
+intended actions. The second questionnaire asked each player about what he
+thought her or his partner was intending to do. Some answers were identical in
+both questionnaires (like “adjusting a shot”) while others were reversed ("to
+guide my partner" versus "to guide me"). We chose this subjective measure of
+accuracy (∆ M(A,B), M(B,B)) rather than an objective measure (i.e. model (A,B)
+is compared to B's next action) because some the activities proposed by the
+questionnaire were not observable by the environment (e.g. establishing a
+strategy). We counted M°(A,B,activity,gamei) as the number of common answers
+between questionnaires M(A,B) and M(B,B) in each game and computed the average
+value a across the 3 games.
+
+Figure 4: M(A,A) and M(A,B) questionnaires in SpaceMiners (translated from French)
+
+\paragraph{Results}
+
+Grounding criterion. The grounding criterion was high: the correlation between
+M°(A,B) and task performance was 0.42, p = 0.05. Pairs with an accurate mutual
+model reached higher scores. A regression analysis confirmed the positive and
+significant relation between group performance and mutual modelling accuracy (β
+=54, p = .02).
+
+Study-specific questions. The awareness tool permitted higher group performance,
+but it did not improve the accuracy of the mutual model. Since teams were free
+to use the awareness tool or not (the 'scout' mode), we performed a post-hoc
+split of players depending on how much time they used it. The split point was
+the mean of time spent in the scout mode and it led to the constitution of two
+groups made up of 12 individuals “short time in scout mode” and 24 individuals
+“long time in scout mode”. A two-way analysis of variance conducted on these
+contrasted groups revealed that pairs in the awareness condition who spent more
+time in the scout mode reached higher levels of M°(A,B)(F = 8.02, p = 0.015). Of
+course, a post-hoc split does not support a causal direction. An alternative
+explanation could be that good modellers are more social and hence appreciate
+the awareness tool.
+
+Symmetry question. We computed intra-class correlation as described by Kenny et
+al. (1998) from the answers to the cross-questionnaires. Considering all pairs
+in both conditions, we found a positive and significant correlation (r = .38, p
+< .05) between M°(A,B) and M°(B,A). Interestingly, this was higher in the
+control group (r = 0.44) than in the experimental group (r = 0.24). Actually, ∆
+(M°(A,B),M°(B,A)), i.e. the absolute differences between the models accuracy,
+was not significantly different with or without the awareness tool ( F [1,13]=
+0.144, p > 0.5). This result could be explained by the fact that the players
+without awareness tools communicated more.
+
+We cannot answer the triangle questions since this study was conducted with
+pairs-
+
+\paragraph{Discussion}
+
+How to interpret a correlation of 0.38 between M°(A,B) and M°(B,A)? If it was
+null, modelling would be interpreted as an individual process that varies
+according to personality traits or cognitive skills: some persons tend to model
+more accurately their peer while others don't care or can simply not. If the
+correlation was close to 1, we could conclude that the modelling accuracy is an
+emergent property of teams since there would be low probability that, by chance,
+low accurate modellers are paired with accurate modellers and conversely. Our
+correlation is between these two extremes but the very notion of significant
+intra-class correlation tend to indicate that one may not consider peers as
+independent subjects, which is an argument for the 'interaction' hypothesis
+(H3).
+
 \bibliographystyle{abbrv}

### 2014-08-27 project-mutual-modeling.tex
Two columns, figure layout
Raw import of studies 2->5 from Word

@@
-\documentclass{article}
+\documentclass[twocolumn]{article}
 
@@
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
 \centering
@@
 \label{mm_symmetry}
-\end{figure*}
+\end{figure}
 
@@
 \label{mm_rectangle}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection{Study 1: Effect of an awareness tool on \Model{A}{B} in a virtual
+\subsection{Study 1: Effect of an awareness tool on \Model{A}{B}{x} in a virtual
 game}
@@
 
-\begin{figure}
+\begin{figure*}
         \centering
-        \includegraphics[width=0.9\columnwidth]{image4.png}
+        \includegraphics[width=0.9\textwidth]{image4.png}
         \caption{Snapshot of the SpaceMiner Game}
         \label{spaceminer}
-\end{figure}
+\end{figure*}
 
@@
 
-Figure 4: M(A,A) and M(A,B) questionnaires in SpaceMiners (translated from French)
+\begin{figure}[ht!]
+        \centering
+        \includegraphics[width=\columnwidth]{image5.png}
+        \caption{M(A,A) and M(A,B) questionnaires in SpaceMiners (translated
+        from French)}
+        \label{study1:questionnaires}
+\end{figure}
 
@@
 
+\subsection{Study 2: Effect of an awareness tool on M°(A,B)  in a pervasive
+game}
+
+\paragraph{Questions}
+
+This study concerns a collaborative game that occurred in the physical space
+(Nova et al , 200?). We studied whether players build an accurate model of the
+path followed by their partners because this path reflected their problem
+solving strategy. We used an objective measure of M°(A,B): the distance between
+where A believes B has been walking and where B actually went. The main
+hypothesis concerned the effect of awareness tools on group performance and on
+M°(A,B).
+
+\paragraph{Experimental setting}
+
+CatchBob is a mobile game in which groups of 3 players have to solve a joint
+task. The game was played on EPFL campus. Participants had to find a virtual
+object ('Bob') and "to catch it" by forming a triangle around it. Players used a
+Tablet PC that displayed a map of the campus and an indication of their personal
+distance from Bob. Their annotations on the map were shared with the two other
+players (A could see what B and C wrote). These annotations faded out after a
+few minutes to avoid covering the full display. The awareness tool displayed the
+location of the two other players on the map. The availability of location
+awareness tool constituted our independent variable.
+
+\begin{figure*}[h!t]
+        \centering
+        \begin{subfigure}{.3\textwidth}
+            \includegraphics[width=\linewidth]{image6.png}
+            \caption{A's drawing A of B's path.}
+        \end{subfigure}
+        \begin{subfigure}{.3\textwidth}
+            \includegraphics[width=\linewidth]{image7.png}
+            \caption{Actual path followed by B.}
+        \end{subfigure}
+        \caption{Bla bla}
+        \label{study2:paths}
+\end{figure*}
+
+\paragraph{Subjects}
+
+Ninety students participated in this experiment. We chosen only students from
+our University because knowing the campus geography had an impact both on group
+performance but also on mutual modeling: it is hard to represent the path of
+someone across some space without having a mental map of this space. We selected
+groups of students who knew each other. We assigned 10 triads to each of our
+three experimental conditions. Each condition was made up of approximately 25%
+of women, but we did not control gender repartition within each triad.
+
+\paragraph{Variables}
+
+The independent variable was the awareness tool: the control condition (without
+tool) and two experimental conditions: synchronous awareness (display current
+position of each player) and asynchronous awareness (display current position of
+each player and their recent spatial trace).  As a dependent variable, we had
+the task performance which was the distance covered by the team to catch Bob and
+M°(A,B). To estimate M°(A,B), we asked players to draw on paper their own path
+and the path of each of their partners after the game. This enabled us to
+calculate the number of errors players made while drawing the path of their
+partners. We compared the path that player A attributed to B with B's real paths
+recorded by the system and the same for A\&C or B\&C as depicted on Figure 5.
+
+M°(A,B) is the sum of errors made by A about B’s paths. An error was either
+drawing a place where the partner had not been or not drawing a place where
+he/she had gone. One could argue that M°(A,B) is biased by the subjects' ability
+to translate their trajectories souvenir into a map drawing. However, 85% of
+subjects made no mistake at all when drawing their own path. We therefore
+consider mistakes in their partners' path as being due to a lack of mutual
+modeling accuracy instead of being due to spatial reasoning skills.
+
+\paragraph{Results}
+
+Grounding criterion. The correlation between M°(A,B) and the task performance
+(path lengths) was low: 0.15. Using a post-hoc split on M°(A,B), there is no
+significant difference between the performance of the groups with high and low
+M°(A,B)  (F = 1.45, p = 0.24). Conversely, a post-hoc split of the groups
+according to their performance did not show any significant differences on
+M°(A,B) (F = 1.16, p = 0.29).
+
+Study-specific questions. There was no significant difference regarding the task
+performance. However, our surprise was that the absence of the awareness tool
+was related to a higher M°(A,B): players better remembered their partners’ path
+if they did not see their position permanently! We will not enter into the
+details of these results but, to be short, teams without awareness tool made
+more annotations on the map while permanent MLA has an underwhelming effect
+(Nova et al., 2005).
+
+Symmetry question. The correlation between M°(A,B)  and M°(B,A)  is positive (r
+= .41) and significant (p < .01);  the more A makes errors about B, the more B
+does the same.
+
+Triangle questions. About ∆2, the correlation between M°(A,B) and M°(A,C)is
+significant: r=0.30, p <.01. About ∆3, the correlation between M°(A,C) and
+M°(B,C) is significant: r=0.43, p <.001.
+
+\paragraph{Discussion}
+
+This positive result regarding to the symmetry question is even more surprising
+than the previous one: despite the high heterogeneity of spatial skills among
+adults (Liben et al, 1981), a significant correlation could indicate that mutual
+modeling relies more on group processes than on individual processes. The
+significant results on the triangle questions probably go in the same direction:
+either the modeler or the modeled both contribute to accurate modeling, or that
+the quality of teamwork has a global effect. In this experiment, since team
+members did not interact intensively during the task, the factor that
+facilitates accurate mutual modeling is probably less the quality of verbal
+interactions that the group strategy: a clear strategy facilitates memorizing
+one’s partner path.
+
+\subsection{Study 3:  Effect of media richness on M°(A,B) in argumentation}
+
+\paragraph{Questions}
+
+The aim of this unpublished study was to evaluate the effect of media richness
+on M°(A,B,emotions). The hypothesis was that video communication would lead to a
+better M°(A,B) than audio only since emotions often impact on facial
+expressions.
+
+\begin{figure*}[ht!]
+        \centering
+        \includegraphics[width=0.9\textwidth]{image8.jpg}
+        \caption{Example of argumentation graph}
+        \label{study3:argumentation_graph}
+\end{figure*}
+
+\paragraph{Experimental settings}
+
+Triads had to address an emotional societal debate: authorizing or not adoption
+by homosexual couples). They worked on-line and had to structure their
+argumentation with a shared concept map tool (TeamWave) as illustrated in Fig.
+6. Ten groups had only an audio connection while ten groups had audio plus
+video. The video communication was provided by a web cam and the software
+IVisit. For the audio link, we used microphones, headsets and a software called
+BattleCom. In the audio+video condition, the screen was divided in three
+sections. The main part was devoted to the concept map window (24 X 24 cm). In
+the remaining part appeared the image of the two peers (8.5 X 8.5 cm each). In
+the audio condition, this video zone was left empty so that the size of the
+concept map was equal in each condition. The subjects were located in the same
+room, separated by mobile walls. Despite their headsets, it is not impossible
+that they could hear non-verbal audio cues (e.g. tapping the sole with feet).
+The task last in average 61 minutes.
+
+\paragraph{Subjects}
+
+Twenty triads from the University of Geneva participated to this experiment. We
+formed groups of subjects who knew each other, i.e. they had ab initio mutual
+models: the task required the discussion of sensitive issues which required to
+feel quite comfortable with peers. There were 36 women and 24 men, but, since
+groups were formed a priori, we could not control the balance of men and women
+in each condition.
+
+\paragraph{Variables}
+
+The independent variable was the presence of not of video link. The dependent
+variable M°(A,B) was subjectively by using 3 questionnaires: in the first one, A
+described his or her own emotions M(A,A) , while in the two other questionnaire,
+A describes B’s and C’s emotions.   The questionnaire included 18 items (7 steps
+Likert Scale) describing emotions labeled as adjectives: anxious, enthusiastic,
+agitated, proud, excited, quiet, calm, stressed, bored, upset, relaxed,
+irritated, determined, hostile, active, etc. This list of emotions was
+elaborated from (REF?). The proposed emotions were . M(A,B) is a vector of 18
+numerical values {x1 … xi} corresponding to their answers on each questionnaire.
+M°(A,B) has been computed as the correlation between the two vectors :
+
+M°(A,B,emotions) = ∆M(A,B,{x1 … xi}, M(B,B,{y1, … ,yi}) = r ({x1 … xi}, {y1, …
+,yi}))
+
+\paragraph{Results}
+
+\begin{figure*}[ht!]
+        \centering
+        \includegraphics[width=0.9\textwidth]{image9.pdf}
+        \caption{Average values of M°(A,B,X) where X is one of the proposed
+        emotions (max = 7).}
+        \label{study3:deg_m_values}
+\end{figure*}
+
+Grounding criterion. The maps produced by teams were ranked. We use the average
+rank as a estimation of the team performance and correlated with the average of
+the 6 values of M°(A,B) per team (M°(A,B), M°(A,C), M°(BA),…): The correlation
+is -0.22: teams with a good M°(A,B) team to be better ranked.
+
+Study-specific questions. Our hypothesis about media richness is rejected: the
+average M°(A,B, emotions) was .62 (SD .28) with the audio+video condition and
+.71 (SD .19) in the audio condition (F=3.4, p: 0.065). Our interpretation is
+that viewing what one's partner sees (shared graphical editor) was, in this
+task, more important than seeing each other (Gaver et al 1993; Anderson et al.,
+2000).
+
+Symmetry questions. Regarding the main items, the mutual modeling of emotions,
+the correlation is .25 or .36  a reprendre avec - Patrick
+
+Triangle questions. Check (∆2) . For (∆3) If we compute for each subject A, the
+correlation between the vectors of emotions in M(B,A) and the vector in M(C,A),
+and take the average across all 60 subjects, we obtain a very impressive
+correlation of .66. This average is about the same in each condition.
+
+Rectangle questions We cannot address the relationship (∆4) between self-and
+social accuracy here because we don't have an estimation of self-accuracy:
+subjects describe their emotions but we have no way to check if there are
+correct. However, by measuring M°(A,B) on  18 emotional labels, we can however
+have a glimpse about ∆5: how M°(A,B,X) varies according to X. Figure 7 shows the
+range of modelling errors: the difference between M(A,B) and M(B,A), on a scale
+of 7, is in average of 0.3 for the emotion "discasted" up to 1.9 for the emotion
+"calm". This is probably specific to the variety of scales  (SD= 0-5 for
+"discated" versus .7 for "calm"), our point is not to interpret this too far,
+but to show that there are large varations even with one area  (perceiving
+emotions).
+
+\paragraph{Discussion}
+
+Chek si il faut modifier les resultats de la symmetries
+
+\subsection{Study 4:  Effects of a script on M°(A,B)  in concept mapping}
+
+\paragraph{Questions}
+
+This study investigated the effect of a collaboration script on collaborative
+learning (REF).. The script chosen is a JIGSAW: two students receive different
+but complementary subsets of the knowledge (texts) which have to be integrated
+to build a shared concept map.  This script increases the cognitive effort to
+build the map, not only to recon ciliate the viewpoints of each team members
+but, before that, to find out what the other knows.
+
+\paragraph{Experimental settings}
+
+The instructional material consisted of an explanatory text about the
+neurophysiologic phenomenon of “action potential”. The text was divided into 3
+chapters. The peers were located in two rooms equipped with the same computer.
+The experimental session lasted around 90 minutes and consisted of 6 phases:
+Participants used two software components, CmapTools (http://cmap.ihmc.us/) and
+TeamSpeak (http://www.goteamspeak.com/).
+
+1. As a pre-test, participants were asked to write down all they knew about the
+neuron and its functioning (5 minutes).
+
+2. Participants were instructed to read a text (12 minutes).
+
+3. Participants were asked to build individually a concept map in order to
+graphically represent what they learnt from the text (10 minutes).
+
+4. Dyads had to built a concept map during 20 minutes, communicating by audio.
+The screen layout was structured into three areas (see Figure 8)
+
+5. Participants were invited to individually complete a knowledge test (15-20
+minutes).
+
+6. Participants where asked to estimate their own- and their partner’s final
+knowledge in a questionnaire.
+
+Figure ?...
+
+\begin{figure*}
+    \centering
+    \begin{tikzpicture}[scale=0.8]
+        \node[anchor=south west,inner sep=0] at (0,0)
+            {\includegraphics[width=0.65\textwidth]{image10.png}};
+
+            \node[rectangle,draw, thick] at (11,6.2) {Self map};
+            \node[rectangle,draw, thick] at (11,3) {Partner map};
+            \node[rectangle,draw, thick] at (3,2) {Collaborative map};
+    \end{tikzpicture}
+    \caption{The group concept map and the individual concept maps in study
+    4}
+    \label{study4:concept_map}
+\end{figure*}
+
+\paragraph{Subjects}
+
+Fifty-eight first year students from EPFL, 47 men and 11 women, with a mean age
+of 20.46 were remunerated for participation. Dyads were randomly assigned to one
+of the two experimental conditions. The men/women ratio was equivalent in each
+condition. Participants did not know each other before the experiment. Students
+from Life Sciences were not recruited since they had too high background
+knowledge on the learning domain.
+
+\paragraph{Variables}
+
+The independent variable, script versus no-script, was implemented by the
+difference of texts that individuals had to read. In the “same information” (SI)
+condition, the same text was given to both partners. In the “complementary
+information” (CI) condition. it was divided into two sub-texts, one about the
+electrical processes of the neuron while the second one about the chemical
+processes. These two versions were equivalent in terms of number of information
+pieces.
+
+The dependent variables were the post-test scores, but this paper is mostly
+concerned by the M(A,B, knowledge).  In phase 6, participants were ask to
+estimate (7-point Likert scale) their own- and their partner’s outcome knowledge
+with respect to each chapter of the zexz. The order of questions about oneself
+and about the other was counterbalanced across participants. We measure M°(A,B)
+in an objective (1 \& 2) as well as in a subjective way (3).
+
+(1) M°(A,B, knowledge) = Σi=1to3  | M(A,B, chapteri) – score (B, chapteri) |
+
+(2) M°(A,A, knowledge) = Σi=1to3  | M(A,A, chapteri) – score (A, chapteri) |
+
+(3) m°(A,B, knowledge) = Σi=1to3  | M(A,B, chapteri) – M(B,A, chapteri) |
+
+\paragraph{Results}
+
+Grounding criterion. In this task, the grounding criterion was low, but we did
+not evaluate task performance (e.g. the quality of the jointly produced concept
+map) but the learning gains. The correlation between M°(A,B) and A's learning
+gains is not significant (β =.08, ,NS, N = 60). It is also not significant
+within each condition.
+
+Study-specific questions. We performed a non-parametric Mann-Whitney test on
+post-test scores for electric questions and a one-way ANOVA on scores for ionic
+questions (Levene tests for homogeneity of variances: p = .02 and ns,
+respectively. Results  did not show any significant difference between the “same
+information” (SI) condition and the “complementary information” (CI) condition,
+neither for electric questions [U = 388.50, z = -0.88, ns] nor for ionic
+questions [F(1, 58) = 0.17, ns].  The effect of scripts on M°(A,B) was not
+significant (F(1, 58) = 0.78, ns) when considering the absolute difference
+between M°(A,B) and B's post-test score. However, A tended underestimate B's
+score in the SI condition (M = -2.06) and to overestimate it in the CI condition
+(M = 1.21). (F(1, 58) = 6.44, p<.01). Regarding to M°(A,A), there was no
+significant difference between conditions.
+
+Symmetry question. The interclass correlation between M°(A,B) and M° (B,A) is
+not significant, but almost (r=.26 (F(1,29)=1.71, p= 0.075)). Actually, it is
+significant when students read the same text (SI condition: r= 0.43,
+F(1,15)=2.53, p<.05) but not when they read different texts (r=.13, F(1,12)=1.3,
+NS).
+
+Triangle questions do not apply here since we worked with dyads.
+
+Rectangle question ∆4: the correlation between M(A,A) and M(A,B) is globally not
+significant (r=0.05) as well as not significant in each condition: someone good
+at self-modeling is not necessarily good at modeling someone else and
+vice-versa.  Discussion
+
+In this experiment, the symmetry of mutual modelling is only found when subjects
+are in conditions of symmetry of information. This consolidates the hypothesis
+of M°(A,B) being a property of interactions in context rather than A's cognitive
+attitude.
+
+\subsection{Study 5:  M°(A,B, knowledge)  in concept mapping}
+
+\paragraph{Questions}
+
+This study investigates if M°(A,B, knowledge) is related to learning outcomes by
+comparing teams with or without a knowledge awareness tool (KAT), i.e. a tool
+that informs A about B's knowledge as measured through a pre-test.
+
+\paragraph{Experimental setting}
+
+The peers were located into two different rooms. A complete description of the
+study is provided in Sangin et al (2008).The experiment lasted 90 minutes. It
+started with the same two steps as in study 4.
+
+3. Subjects passed a pre-test, with ten questions per chapter.
+
+4. Participants had 20 minutes to draw a collaborative concept map reporting the
+content of the texts. They were able to communicate orally through headsets.  We
+used Tobbii eye tracking devices to record their gazes.
+
+5. The post-test included the same items than the pre-test but in a different
+order.
+
+6. Finally, participants were asked to estimate their partner’s knowledge at the
+post-test for each of the three chapters on a 7-point likert-like survey.
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.9\textwidth]{image11.png}
+        \caption{Screenshot of the KAT condition during the concept-map building
+        phase.}
+        \label{study5:kat}
+\end{figure*}
+
+\paragraph{Subjects}
+
+Sixty-four first year EPFL students (18 women, mean age = 21.2 years)
+participated to the study. They were involved in curricula that do not imply
+advanced neurophysiologic notions. They were randomly assigned to conditions and
+did not know each other before.
+
+\paragraph{Variables}
+
+The participants of the experimental condition group were provided with the KAT
+on the bottom part of the screen (see Figure 9): each line represents the score
+obtained by the partner at the pre-test for a chapter. Participants did not see
+they own score, but they usually started their discussion by exchanging this
+information.
+
+\paragraph{Results}
+
+Grounding criterion. A linear regression revealed a positive relation between M°
+(A,B) and the collaborative learning gain of the pair (AB): β = .401, p < .001,
+r2adj. =.15, large effect. This relationship was not significant in the previous
+experiment which as been conducted with the same task: using the same task
+should produce the same grounding criterion This probably due to the fact that
+M° (A,B) was influence by the KAT.
+
+Study-specific questions. The t-test reported a significant difference between
+the KAT condition participants (M = 13.4\%) and the control group [M = 3.6%, t(1,
+60) = 2.73, p < .01,Cohen’s d = 0.7, medium to large effect]. Providing learners
+with cues about the prior-knowledge of their partner enhances their
+collaborative learning. As a treatment check, we found a positive and
+significant correlation between the amount of gazes on KAT (using eye tracking
+devices) and the learning gains (r(22) = .54, p = .01). A detailed analysis
+revealed that subjects look at KAT to assess their peer's credibility when (s)he
+provided new information.
+
+The KAT has a significant effect on M°(A,B): peers more accurately estimated
+their partners knowledge (M = 1.11) than those in the control condition [M =
+.98, t(1, 60) = 3.19, p < .01, Cohen’s d = 0.83, large effect]. This is a
+trivial result since the KAT provided them with an initial M(A,B). However,
+subjects have to predict the post-test score while the KAT informed them about
+the pre-test score. Actually, pairs in the KAT condition produced significantly
+more instances of 3 specific categories of interactions: (1) elaborated
+utterances with rich contents,  (2) utterances asking about the other's
+knowledge such as "Did you understand how transmission works?" (3) utterances
+describing one's own knowledge (M(A,A)) such as "I don’t remember the Ranvier’s
+thing…" . These 3 categories provide different account of M°(A,B): (1) as an
+effect of the quality of interaction, (2) as an effect of A's effort to model B
+or (3) as B's effort to give cues to A about his own knowledge.
+
+We examined M° (A,B) as potentially mediating the effect of the KAT factor on
+the relative learning gain. A linear regression confirmed that M° (A,B) was
+significantly related to the KAT-factor (β = .381, p < .01, r2 = .15). The
+KAT-factor was also significantly and positively related to the RLG (β = .332, p
+< .01, r2 = .11). We then tested the relation between the independent variable
+(KAT) and the dependent variable (gains) when controlling for the mediating
+variable (M° (A,B)). A multiple regression showed that the KAT-factor was no
+longer a significant predictor (β = .210; p = ns) whereas the M° (A,B) was still
+a significant predictor (β = .32, p < .01). Thus, it can be concluded that M°
+(A,B) mediated the KAT-factor’s effect on the learners’ RLG. The Sobel (1982)
+significance test for indirect effects was significant [z = 1.99, p < .05].
+
+Symmetry question.  We did not find an intra-pair correlation (ICC = 0.05, NS).
+Rectangle question ∆4: the correlation between M(A,A) and M(A,B) is \fixme{???
+MIRWEIS A FAIRE}.
+
+\paragraph{Discussion}
+
+Despite the fact that the learning task was the same as in study 4, the
+conditions of collaboration (viewing multiple maps or not) and the conditions
+(scripted or not, awareness tool or not) probably explain differences in terms
+of mutual modeling
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{abbrv}

### 2014-08-27 mutual-modeling.tex
Texifying...
While here:
{project- ->}mutual-modeling.tex

--- 2014-08-26 project-mutual-modeling.tex
@@
 
+\usepackage{amsmath}
+
 \usepackage{tikz}
@@
 
+\usepackage{hyperref}
 \usepackage{subcaption}
 
-\usepackage[draft]{fixme}
+\usepackage[draft, footnote, margin=false]{fixme}
 
@@
 \newcommand{\model}[3]{{$\mathcal{M}(#1, #2, #3)$}}
+% generic mutual model (gmodel) 
+\newcommand{\gmodel}[2]{{$\mathcal{M}(#1, #2)$}}
 \newcommand{\refmodel}[2]{{$\mathcal{M}(#1, #2)$}}
 \newcommand{\Model}[3]{{$\mathcal{M}^{\circ}(#1, #2, #3)$}}
+\newcommand{\gModel}[2]{{$\mathcal{M}^{\circ}(#1, #2)$}}
+\newcommand{\Mdeg}[3]{{\mathcal{M}^{\circ}(#1, #2, #3)}}
+\newcommand{\gMdeg}[2]{{\mathcal{M}^{\circ}(#1, #2)}}
 
@@
 understanding (Schwartz, 1995). This effort can be observed by the frequency of
-rich interactions, i.e. interactions whose occurrence has been related to
+rich interactions, \ie interactions whose occurrence has been related to
 learning: (self-) explanations in cognitive science (Chi ; Webb), conflict
@@
 change viewpoint, it is criticized at the same time as being too cognitivist by
-some psycholinguists, i.e. as overestimating the amount shared knowledge and
+some psycholinguists, \ie as overestimating the amount shared knowledge and
 mutual representations actually necessary to conduct a dialogue. The fundamental
@@
 intervention will be conducted in which way on which patient.  Some CSCL
-scholars stressed the 'illusion of shared understanding' (JACCO?), i.e. the
+scholars stressed the 'illusion of shared understanding' (JACCO?), \ie the
 potential decoupling between linguistic alignment and actual shared meanings.
@@
 they know they can provide feedback on his/her understanding (Schober, 1993). We
-introduced the notion of ‘optimal collaborative effort’ (Dillenbourg et al,
+introduced the notion of ‘optimal collaborative effort' (Dillenbourg et al,
 1996) to stress that misunderstanding should not be viewed as something to be
@@
 context of collaborative learning, we understand the cognitive load induced by
-mutual modeling as part of Schwartz’s (1995) notion of effort towards a shared
+mutual modeling as part of Schwartz's (1995) notion of effort towards a shared
 understanding. For instance, CSCL researchers expanded the use of 'collaboration
@@
     \item $Model(A,B, "urgent")$: What does A about know B's understanding of A's
-        last utterance: does ‘urgent’ means now, ASAP and ‘not too lat’ ?
+        last utterance: does ‘urgent' means now, ASAP and ‘not too lat' ?
 
@@
 which A and B attended together~\cite{clark2002definite} in the physical space
-or in our cultural space (\eg ‘09-11’). Co-presence means that we can refer to
+or in our cultural space (\eg ‘09-11'). Co-presence means that we can refer to
 a shared objects and event,-s it does of course not imply that we give them the
@@
 
-\begin{equation} \label{eq:probrepair}
-p(\alpha_{t_1}(X,T) / M^{\circ}_{t_2}(A,B,X) \nearrow) \approx \frac{(M^{\circ}_{min}(A,B,X,T) -
-M^{\circ}_{t_1}(A,B,X))}{cost (\alpha)}
-\end{equation}
+\begin{multline} \label{eq:probrepair}
+p(\alpha_{t_1}(X,T) / M^{\circ}_{t_2}(A,B,X) \nearrow) \approx \\
+\frac{(M^{\circ}_{min}(A,B,X,T) - M^{\circ}_{t_1}(A,B,X))}{cost (\alpha)}
+\end{multline}
 
@@
 
-\subsubsection{Studies}
+\subsubsection*{Studies}
 
@@
 Fig.~\ref{mm_rectangle}). We do not further develop these questions because we
-don’t have the necessary data in our studies.
+don't have the necessary data in our studies.
 
-\begin{figure*}[htb]
+\begin{figure}[htb]
 \centering
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Studies}
 
-\subsection{Study 1: Effect of an awareness tool on \Model{A}{B} in a virtual
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Study 1: Effect of an awareness tool on \gModel{A}{B} in a virtual
 game}
 
-\paragraph{Questions}
+\subsubsection*{Questions}
 
 We studied the impact of an awareness tool on group performance and mutual
-modeling. (Nova, Wehrle, Goslin, Bourquin \& Dillenbourg (2006). The availability
+modeling~\cite{nova2007collaboration}. The availability
 of an awareness tool was our independent variable. In previous studies, we
 replayed a video of the game to subjects who surprised us by their ability to
-remember former states of their mutual model: "I did that because I thought that
-you would do that". Hence, this experiment focused the representation of each
+remember former states of their mutual model: ``I did that because I thought that
+you would do that''. Hence, this experiment focused the representation of each
 other's action plans. During the game, we asked them to anticipate the next
@@
 
-\paragraph{Experimental setting}
+\subsubsection*{Experimental setting}
 
 SpaceMiners is a 3D game that involves two players collecting minerals located
-in asteroids. To do so, they shoot drones through the space after choosing their
+in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they shoot drones through the space after choosing their
 initial direction and speed. Once launched, the trajectory of drones is only
@@
 player A  the location and gaze direction of player B. In the awareness
-condition, players could switch to the 'scout mode' where they could view what
+condition, players could switch to the \emph{scout mode} where they could view what
 their partner was looking at. We hypothesize that this would enable subjects to
@@
 
-\paragraph{Subjects}
+\subsubsection*{Subjects}
 
@@
 
-\begin{figure}
+\begin{figure*}
         \centering
-        \includegraphics[width=0.9\columnwidth]{image4.png}
+        \includegraphics[width=0.9\textwidth]{image4.png}
         \caption{Snapshot of the SpaceMiner Game}
-        \label{spaceminer}
-\end{figure}
+        \label{study1:spaceminer}
+\end{figure*}
 
-\paragraph{Variables}
+\subsubsection*{Variables}
 
@@
 
-In order to evaluate m°(A,B) during the task, we used two questionnaires (Figure
-4) that were displayed during each of the three games, as a transparent layer
-appearing over the game display. The first questionnaire concerned the player’s
-intended actions. The second questionnaire asked each player about what he
-thought her or his partner was intending to do. Some answers were identical in
-both questionnaires (like “adjusting a shot”) while others were reversed ("to
-guide my partner" versus "to guide me"). We chose this subjective measure of
-accuracy (∆ M(A,B), M(B,B)) rather than an objective measure (i.e. model (A,B)
-is compared to B's next action) because some the activities proposed by the
-questionnaire were not observable by the environment (e.g. establishing a
-strategy). We counted M°(A,B,activity,gamei) as the number of common answers
-between questionnaires M(A,B) and M(B,B) in each game and computed the average
-value a across the 3 games.
+In order to evaluate \gModel{A}{B} during the task, we used two questionnaires
+(Fig.~\ref{study1:questionnaires}) that were displayed during each of the three
+games, as a transparent layer appearing over the game display. The first
+questionnaire concerned the player's intended actions. The second questionnaire
+asked each player about what he thought her or his partner was intending to do.
+Some answers were identical in both questionnaires (like ``adjusting a shot'')
+while others were reversed (``to guide my partner'' versus ``to guide me''). We
+chose this subjective measure of accuracy ($\Delta(\M{A}{B}{x}, \M{B}{B}{x})$)
+rather than an objective measure (\ie the model \gmodel{A}{B} is compared to B's
+next action) because some the activities proposed by the questionnaire were not
+observable by the environment (\eg establishing a strategy). We counted
+\Model{A}{B}{activity_{game_{i}}}) as the number of common answers between
+questionnaires \gmodel{A}{B} and \gmodel{B}{B} in each game and computed the
+average value a across the 3 games.
 
-Figure 4: M(A,A) and M(A,B) questionnaires in SpaceMiners (translated from French)
+\begin{figure}[ht!]
+        \centering
+        \includegraphics[width=\columnwidth]{image5.png}
+        \caption{\gmodel{A}{A} and \gmodel{A}{B} questionnaires in SpaceMiners
+        (translated from French).}
 
-\paragraph{Results}
+        \label{study1:questionnaires}
+\end{figure}
 
-Grounding criterion. The grounding criterion was high: the correlation between
-M°(A,B) and task performance was 0.42, p = 0.05. Pairs with an accurate mutual
-model reached higher scores. A regression analysis confirmed the positive and
-significant relation between group performance and mutual modelling accuracy (β
-=54, p = .02).
+\subsubsection*{Results}
 
-Study-specific questions. The awareness tool permitted higher group performance,
-but it did not improve the accuracy of the mutual model. Since teams were free
-to use the awareness tool or not (the 'scout' mode), we performed a post-hoc
-split of players depending on how much time they used it. The split point was
-the mean of time spent in the scout mode and it led to the constitution of two
-groups made up of 12 individuals “short time in scout mode” and 24 individuals
-“long time in scout mode”. A two-way analysis of variance conducted on these
-contrasted groups revealed that pairs in the awareness condition who spent more
-time in the scout mode reached higher levels of M°(A,B)(F = 8.02, p = 0.015). Of
-course, a post-hoc split does not support a causal direction. An alternative
-explanation could be that good modellers are more social and hence appreciate
-the awareness tool.
+\paragraph{Grounding criterion} The grounding criterion was high: the
+correlation between \gModel{A}{B} and task performance was 0.42, $p = 0.05$.
+Pairs with an accurate mutual model reached higher scores. A regression analysis
+confirmed the positive and significant relation between group performance and
+mutual modelling accuracy ($\beta=54, p = .02$).
 
-Symmetry question. We computed intra-class correlation as described by Kenny et
-al. (1998) from the answers to the cross-questionnaires. Considering all pairs
-in both conditions, we found a positive and significant correlation (r = .38, p
-< .05) between M°(A,B) and M°(B,A). Interestingly, this was higher in the
-control group (r = 0.44) than in the experimental group (r = 0.24). Actually, ∆
-(M°(A,B),M°(B,A)), i.e. the absolute differences between the models accuracy,
-was not significantly different with or without the awareness tool ( F [1,13]=
-0.144, p > 0.5). This result could be explained by the fact that the players
-without awareness tools communicated more.
+\paragraph{Study-specific questions} The awareness tool permitted higher group
+performance, but it did not improve the accuracy of the mutual model. Since
+teams were free to use the awareness tool or not (the \emph{scout} mode), we
+performed a post-hoc split of players depending on how much time they used it.
+The split point was the mean of time spent in the scout mode and it led to the
+constitution of two groups made up of 12 individuals ``short time in scout
+mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
+variance conducted on these contrasted groups revealed that pairs in the
+awareness condition who spent more time in the scout mode reached higher levels
+of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
+does not support a causal direction. An alternative explanation could be that
+good modellers are more social and hence appreciate the awareness tool.
 
-We cannot answer the triangle questions since this study was conducted with
-pairs-
+\paragraph{Symmetry question} We computed intra-class correlation as described
+by~\cite{kenny1998data} from the answers to the cross-questionnaires.
+Considering all pairs in both conditions, we found a positive and significant
+correlation ($r = .38, p < .05$) between \gModel{A}{B} and \gModel{B}{A}.
+Interestingly, this was higher in the control group ($r = 0.44$) than in the
+experimental group ($r = 0.24$). Actually, $\Delta(\gMdeg{A}{B},\gMdeg{B}{A})$,
+\ie the absolute differences between the models accuracy, was not significantly
+different with or without the awareness tool ($F [1,13]= 0.144, p > 0.5$). This
+result could be explained by the fact that the players without awareness tools
+communicated more.
 
-\paragraph{Discussion}
+\paragraph{Triangle questions} We cannot answer the triangle questions since
+this study was conducted with dyads.
 
-How to interpret a correlation of 0.38 between M°(A,B) and M°(B,A)? If it was
-null, modelling would be interpreted as an individual process that varies
-according to personality traits or cognitive skills: some persons tend to model
-more accurately their peer while others don't care or can simply not. If the
-correlation was close to 1, we could conclude that the modelling accuracy is an
-emergent property of teams since there would be low probability that, by chance,
-low accurate modellers are paired with accurate modellers and conversely. Our
-correlation is between these two extremes but the very notion of significant
-intra-class correlation tend to indicate that one may not consider peers as
-independent subjects, which is an argument for the 'interaction' hypothesis
-(H3).
+\subsubsection*{Discussion}
 
+How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
+If it was null, modelling would be interpreted as an individual process that
+varies according to personality traits or cognitive skills: some persons tend to
+model more accurately their peer while others don't care or can simply not. If
+the correlation was close to 1, we could conclude that the modelling accuracy is
+an emergent property of teams since there would be low probability that, by
+chance, low accurate modellers are paired with accurate modellers and
+conversely. Our correlation is between these two extremes but the very notion of
+significant intra-class correlation tend to indicate that one may not consider
+peers as independent subjects, which is an argument for the \emph{interaction}
+hypothesis $\mathcal{H}_{3}$.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Study 2: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
+game}
+
+\subsubsection*{Questions}
+
+This study concerns a collaborative game that occurred in the physical
+space~\cite{nova2006underwhelming}. We studied whether players build an accurate
+model of the path followed by their partners because this path reflected their
+problem solving strategy. We used an objective measure of \gModel{A}{B}: the
+distance between where A believes B has been walking and where B actually went.
+The main hypothesis concerned the effect of awareness tools on group performance
+and on \gModel{A}{B}.
+
+\subsubsection*{Experimental setting}
+
+{\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
+joint task. The game was played on EPFL campus. Participants had to find a
+virtual object (\emph{Bob}) and ``to catch it'' by forming a triangle around it.
+Players used a Tablet PC that displayed a map of the campus and an indication of
+their personal distance to Bob. Their annotations on the map were shared with
+the two other players (A could see what B and C wrote). These annotations faded
+out after a few minutes to avoid covering the full display. The awareness tool
+displayed as well the location of the two other players on the map. The
+availability of location awareness tool constituted our independent variable.
+
+\begin{figure*}[h!t]
+        \centering
+        \begin{subfigure}{.3\textwidth}
+            \includegraphics[width=\linewidth]{image6.png}
+            \caption{A's drawing of B's path.}
+        \end{subfigure}
+        \begin{subfigure}{.3\textwidth}
+            \includegraphics[width=\linewidth]{image7.png}
+            \caption{Actual path followed by B.}
+        \end{subfigure}
+        \caption{Reported and actual path of one of the player, during the {\sc
+        CatchBob} game.}
+        \label{study2:paths}
+\end{figure*}
+
+\subsubsection*{Subjects}
+
+Ninety students participated in this experiment. We chosen only students from
+our university because knowing the campus geography had an impact both on group
+performance but also on mutual modeling: it is hard to represent the path of
+someone across some space without having a mental map of this space. We selected
+groups of students who knew each other. We assigned 10 triads to each of our
+three experimental conditions. Each condition was made up of approximately 25\%
+of women, but we did not control gender repartition within each triad.
+
+\subsubsection*{Variables}
+
+The independent variable was the awareness tool: the control condition (without
+tool) and two experimental conditions: synchronous awareness (display current
+position of each player) and asynchronous awareness (display current position of
+each player and their recent spatial trace).  As a dependent variable, we had
+the task performance which was the distance covered by the team to catch Bob and
+\gModel{A}{B}. To estimate \gModel{A}{B}, we asked players to draw on paper
+their own path and the path of each of their partners after the game. This
+enabled us to calculate the number of errors players made while drawing the path
+of their partners. We compared the path that player A attributed to B with B's
+real path recorded by the system and the same for A \& C and B \& C as depicted
+on Fig.~\ref{study2:paths}.
+
+\gModel{A}{B} is the sum of errors made by A about B's paths. An error was
+either drawing a place where the partner had not been or not drawing a place
+where he/she had gone. One could argue that \gModel{A}{B} is biased by the
+subjects' ability to translate their trajectories souvenir into a map drawing.
+However, 85\% of subjects made no mistake at all when drawing their own path. We
+therefore consider mistakes in their partners' path as being due to a lack of
+mutual modeling accuracy instead of being due to spatial reasoning skills.
+
+\subsubsection*{Results}
+
+\paragraph{Grounding criterion} The correlation between \gModel{A}{B} and the
+task performance (path lengths) was low: 0.15. Using a \emph{post-hoc} split on
+\gModel{A}{B}, there is no significant difference between the performance of the
+groups with high and low \gModel{A}{B}  ($F = 1.45, p = 0.24$). Conversely, a
+\emph{post-hoc} split of the groups according to their performance did not show
+any significant differences on \gModel{A}{B} ($F = 1.16, p = 0.29$).
+
+\paragraph{Study-specific questions} There was no significant difference
+regarding the task performance. However, our surprise was that the absence of
+the awareness tool was related to a higher \gModel{A}{B}: players better
+remembered their partners' path if they did not see their position permanently!
+We will not enter into the details of these results but, to be short, teams
+without awareness tool made more annotations on the map while permanent MLA has
+an underwhelming effect~\cite{nova2005location}.
+
+\paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
+\gModel{B}{A}  is positive ($r = .41$) and significant ($p < .01$);  the more A
+makes errors about B, the more B does the same.
+
+\paragraph{Triangle questions} Regarding $\Delta_2$, the correlation between
+\gModel{A}{B} and \gModel{A}{C} is significant: $r=0.30, p <.01$. Concerning
+$\Delta_3$, the correlation between \gModel{A}{C} and \gModel{B}{C} is
+significant as well: $r=0.43, p <.001$.
+
+\subsubsection*{Discussion}
+
+This positive result regarding to the symmetry question is even more surprising
+than the previous one: despite the high heterogeneity of spatial skills among
+adults~\cite{liben1981spatial}, a significant correlation could indicate that
+mutual modeling relies more on group processes than on individual processes. The
+significant results on the triangle questions probably go in the same direction:
+either the modeler or the modeled both contribute to accurate modeling, or that
+the quality of teamwork has a global effect. In this experiment, since team
+members did not interact intensively during the task, the factor that
+facilitates accurate mutual modeling is probably less the quality of verbal
+interactions that the group strategy: a clear strategy facilitates memorizing
+one's partner path.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Study 3:  Effect of media richness on \gModel{A}{B} in argumentation}
+
+\subsubsection*{Questions}
+
+The aim of this unpublished study was to evaluate the effect of media richness
+on \Model{A}{B}{emotions}. The hypothesis was that video communication would
+lead to a better \gModel{A}{B} than audio only since emotions often impact on
+facial expressions.
+
+\begin{figure*}[ht!]
+        \centering
+        \includegraphics[width=0.9\textwidth]{image8.jpg}
+        \caption{Example of argumentation graph}
+        \label{study3:argumentation_graph}
+\end{figure*}
+
+\subsubsection*{Experimental settings}
+
+Triads had to address an emotional societal debate: authorizing or not adoption
+by homosexual couples. They worked on-line and had to structure their
+argumentation with a shared concept map tool {\sc TeamWave} as illustrated in
+Fig.~\ref{study3:argumentation_graph}. Ten groups had only an audio connection
+while ten groups had audio plus video. The video communication was provided by a
+web cam and the software {\sc IVisit}. For the audio link, we used microphones,
+headsets and a software called {\sc BattleCom}. In the audio+video condition,
+the screen was divided in three sections. The main part was devoted to the
+concept map window (24x24cm). In the remaining part appeared the image of the
+two peers (8.5x8.5cm each). In the audio condition, this video zone was left
+empty so that the size of the concept map was equal in each condition. The
+subjects were located in the same room, separated by mobile walls. Despite their
+headsets, it is not impossible that they could hear non-verbal audio cues (\eg
+tapping the sole with feet).  The task last in average 61 minutes.
+
+\subsubsection*{Subjects}
+
+Twenty triads from the University of Geneva participated to this experiment. We
+formed groups of subjects who knew each other, \ie they had \emph{ab initio}
+mutual models: the task required the discussion of sensitive issues which
+required to feel quite comfortable with peers. There were 36 women and 24 men,
+but, since groups were formed a priori, we could not control the balance of men
+and women in each condition.
+
+\subsubsection*{Variables}
+
+The independent variable was the presence of not of video link. The dependent
+variable \gModel{A}{B} was subjectively by using 3 questionnaires: in the first
+one, A described his or her own emotions \gmodel{A}{A} , while in the two other
+questionnaire, A describes B's and C's emotions. The questionnaire included 18
+items (7-point Likert Scale) describing emotions labeled as adjectives:
+\emph{anxious}, \emph{enthusiastic}, \emph{agitated}, \emph{proud},
+\emph{excited}, \emph{quiet}, \emph{calm}, \emph{stressed}, \emph{bored},
+\emph{upset}, \emph{relaxed}, \emph{irritated}, \emph{determined},
+\emph{hostile}, \emph{active}, etc. This list of emotions was
+elaborated from \fixme{REF?}. The proposed emotions were \fixme{?}. \gmodel{A}{B} is a vector
+of 18 numerical values ${x_1 \dots x_i}$ corresponding to their answers on each
+questionnaire. \gModel{A}{B} has been computed as the correlation between the
+two vectors (Eq.~\ref{eq:study3}).
+
+\begin{multline} \label{eq:study3}
+    \Mdeg{A}{B}{emotions}\\
+    = \Delta(\M{A}{B}{\{x_1 \dots x_i\}}, \M{B}{B}{\{y_1 \dots y_i\}}) \\ 
+    = r(\{x_1 \dots x_i\}, \{y_1 \dots y_i\})
+\end{multline}
+
+\subsubsection*{Results}
+
+\begin{figure*}[ht!]
+        \centering
+        \includegraphics[width=0.9\textwidth]{image9.pdf}
+        \caption{Average values of \Model{A}{B}{X} where X is one of the proposed
+        emotions (max = 7).}
+        \label{study3:deg_m_values}
+\end{figure*}
+
+\paragraph{Grounding criterion} The maps produced by teams were ranked. We use
+the average rank as a estimation of the team performance and correlated with the
+average of the 6 values of \gModel{A}{B} per team (\gModel{A}{B}, \gModel{A}{C},
+\gModel{B}{A},...): The correlation is -0.22: teams with a good \gModel{A}{B}
+team to be better ranked.
+
+\paragraph{Study-specific questions} Our hypothesis about media richness is
+rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
+audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
+0.065$). Our interpretation is that viewing what one's partner sees (shared
+graphical editor) was, in this task, more important than seeing each
+other~\cite{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
+Anderson et al., 2000}.
+
+\paragraph{Symmetry question} Regarding the main items, the mutual modeling of
+emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}
+
+\paragraph{Triangle questions} \fixme{Check ($\Delta_2$)} For $\Delta_3$, if we
+compute for each subject A, the correlation between the vectors of emotions in
+\gmodel{B}{A} and the vector in \gmodel{C}{A}, and take the average across all
+60 subjects, we obtain a very impressive correlation of .66. This average is
+about the same in each condition.
+
+\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
+between self-and social accuracy here because we don't have an estimation of
+self-accuracy: subjects describe their emotions but we have no way to check if
+there are correct. However, by measuring \gModel{A}{B} on  18 emotional labels,
+we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
+according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
+errors: the difference between \gmodel{A}{B} and \gmodel{B}{A}, on a scale of 7,
+is in average of 0.3 for the emotion \emph{discasted}, and up to 1.9 for the
+emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
+0-5$ for \emph{discated} versus $SD=.7$ for \emph{calm}), our point is not to
+interpret this too far, but to show that there are large varations even with one
+area  (perceiving emotions).
+
+\subsubsection*{Discussion}
+
+\fixme{TBD}
+\fixme{Chek si il faut modifier les resultats de la symmetries}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Study 4:  Effects of a script on \gModel{A}{B}  in concept mapping}
+
+\subsubsection*{Questions}
+
+This study investigated the effect of a collaboration script on collaborative
+learning~\cite{molinari2008effects}. The script chosen is a JIGSAW: two students
+receive different but complementary subsets of the knowledge (texts) which have
+to be integrated to build a shared concept map.  This script increases the
+cognitive effort to build the map, not only to recon ciliate the viewpoints of
+each team members but, before that, to find out what the other knows.
+
+\subsubsection*{Experimental settings}
+
+The instructional material consisted of an explanatory text about the
+neurophysiologic phenomenon of \emph{action potential}. The text was divided
+into 3 chapters. The peers were located in two rooms equipped with the same
+computer.  The experimental session lasted around 90 minutes and consisted of 6
+phases: Participants used two software components, {\sc
+CmapTools}\footnote{\url{http://cmap.ihmc.us/}} and {\sc
+TeamSpeak}\footnote{\url{http://www.goteamspeak.com/}}.
+
+\begin{enumerate}
+
+    \item As a pre-test, participants were asked to write down all they knew
+        about the neuron and its functioning (5 minutes).
+
+    \item Participants were instructed to read a text (12 minutes).
+
+    \item Participants were asked to build individually a concept map in order to
+        graphically represent what they learnt from the text (10 minutes).
+
+    \item Dyads had to built a concept map during 20 minutes, communicating by
+        audio.  The screen layout was structured into three areas (see Figure 8)
+
+    \item Participants were invited to individually complete a knowledge test
+        (15-20 minutes).
+
+    \item Participants where asked to estimate their own- and their partner's
+        final knowledge in a questionnaire.
+
+\end{enumerate}
+
+\begin{figure*}
+    \centering
+    \begin{tikzpicture}[scale=0.8]
+        \node[anchor=south west,inner sep=0] at (0,0)
+            {\includegraphics[width=0.65\textwidth]{image10.png}};
+
+            \node[rectangle,draw, thick] at (11,6.2) {Self map};
+            \node[rectangle,draw, thick] at (11,3) {Partner map};
+            \node[rectangle,draw, thick] at (3,2) {Collaborative map};
+    \end{tikzpicture}
+    \caption{The group concept map and the individual concept maps in study
+    4}
+    \label{study4:concept_map}
+\end{figure*}
+
+\subsubsection*{Subjects}
+
+Fifty-eight first year students from EPFL, 47 men and 11 women, with a mean age
+of 20.46 were remunerated for participation. Dyads were randomly assigned to one
+of the two experimental conditions. The men/women ratio was equivalent in each
+condition. Participants did not know each other before the experiment. Students
+from Life Sciences were not recruited since they had too high background
+knowledge on the learning domain.
+
+\subsubsection*{Variables}
+
+The independent variable, script versus no-script, was implemented by the
+difference of texts that individuals had to read. In the \emph{same information}
+(SI) condition, the same text was given to both partners. In the
+\emph{complementary information} (CI) condition, it was divided into two
+sub-texts, one about the electrical processes of the neuron while the second one
+about the chemical processes. These two versions were equivalent in terms of
+number of information pieces.
+
+The dependent variables were the post-test scores, but this paper is mostly
+concerned by the \Model{A}{B}{knowledge}. In phase 6, participants were ask to
+estimate (7-point Likert scale) their own and their partner's outcome knowledge
+with respect to each chapter of the \fixme{zexz}. The order of questions about
+oneself and about the other was counterbalanced across participants. We measure
+\gModel{A}{B} in an objective (Eq.~\ref{eq:study4.1} and \ref{eq:study4.2}) as
+well as in a subjective way (Eq.~\ref{eq:study4.3}).\fixme{check  subjective
+assess: 2nd equation?}
+
+\begin{multline} \label{eq:study4.1}
+    \Mdeg{A}{B}{knowledge} = \\
+        \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
+\end{multline}
+
+\begin{multline} \label{eq:study4.2}
+    \Mdeg{A}{A}{knowledge} = \\
+        \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
+\end{multline}
+
+\begin{multline} \label{eq:study4.3}
+    \Mdeg{A}{B}{knowledge} = \\
+        \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
+\end{multline}
+
+\subsubsection*{Results}
+
+\paragraph{Grounding criterion} In this task, the grounding criterion was low,
+but we did not evaluate task performance (\eg the quality of the jointly
+produced concept map) but the learning gains. The correlation between
+\gModel{A}{B} and A's learning gains is not significant ($\beta =.08, ,NS, N =
+60$). It is also not significant within each condition.
+
+\paragraph{Study-specific questions} We performed a non-parametric Mann-Whitney
+test on post-test scores for electric questions and a one-way ANOVA on scores
+for ionic questions (Levene tests for homogeneity of variances: $p = .02$ and
+$ns$, respectively. Results  did not show any significant difference between the
+\emph{same information} (SI) condition and the \emph{complementary information}
+(CI) condition, neither for electric questions [$U = 388.50, z = -0.88, ns$] nor
+for ionic questions [$F(1, 58) = 0.17, ns$].  The effect of scripts on
+\gModel{A}{B} was not significant ($F(1, 58) = 0.78, ns$) when considering the
+absolute difference between \gModel{A}{B} and B's post-test score. However, A
+tended underestimate B's score in the SI condition ($M = -2.06$) and to
+overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<.01$).
+Regarding to \gModel{A}{A}, there was no significant difference between
+conditions.
+
+\paragraph{Symmetry question} The interclass correlation between \gModel{A}{B}
+and \Model{B}{A} is not significant, but almost ($r=.26, F(1,29)=1.71, p=
+0.075$).  Actually, it is significant when students read the same text (SI
+condition: $r=0.43, F(1,15)=2.53, p<.05$) but not when they read different texts
+($r=.13, F(1,12)=1.3, NS$).
+
+\paragraph{Triangle questions} Triangle questions do not apply here since we
+worked with dyads.
+
+\paragraph{Rectangle question}$\Delta_4$: the correlation between \gmodel{A}{A}
+and \gmodel{A}{B} is globally not significant ($r=0.05$) as well as not
+significant in each condition: someone good at self-modeling is not necessarily
+good at modeling someone else and vice-versa.
+
+\subsubsection*{Discussion}
+
+In this experiment, the symmetry of mutual modelling is only found when subjects
+are in conditions of symmetry of information. This consolidates the hypothesis
+of \gModel{A}{B} being a property of interactions in context rather than A's
+cognitive attitude.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Study 5: \Model{A}{B}{knowledge} in concept mapping}
+
+\subsubsection*{Questions}
+
+This study investigates if \Model{A}{B}{knowledge} is related to learning
+outcomes by comparing teams with or without a knowledge awareness tool (KAT),
+\ie a tool that informs A about B's knowledge as measured through a pre-test.
+
+\subsubsection*{Experimental setting}
+
+The peers were located into two different rooms. A complete description of the
+study is provided in~\cite{sangin2008learners}. The experiment lasted 90 minutes.
+
+It started with the same two steps as in study 4, followed by:
+
+\begin{enumerate}
+    \setcounter{enumi}{2}
+
+    \item Subjects passed a pre-test, with ten questions per chapter.
+
+    \item Participants had 20 minutes to draw a collaborative concept map
+        reporting the content of the texts. They were able to communicate
+        orally through headsets.  We used Tobbii eye tracking devices to
+        record their gazes.
+
+    \item The post-test included the same items than the pre-test but in a
+        different order.
+
+    \item  Finally, participants were asked to estimate their partner's
+        knowledge at the post-test for each of the three chapters on a 7-point
+        Likert-like survey.
+
+\end{enumerate}
+
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.9\textwidth]{image11.png}
+        \caption{Screenshot of the KAT condition during the concept-map building
+        phase.}
+        \label{study5:kat}
+\end{figure*}
+
+\subsubsection*{Subjects}
+
+Sixty-four first year EPFL students (18 women, mean age = 21.2 years)
+participated to the study. They were involved in curricula that do not imply
+advanced neurophysiologic notions. They were randomly assigned to conditions and
+did not know each other before.
+
+\subsubsection*{Variables}
+
+The participants of the experimental condition group were provided with the KAT
+on the bottom part of the screen (Fig.~\ref{study5:kat}): each line represents
+the score obtained by the partner at the pre-test for a chapter. Participants
+did not see they own score, but they usually started their discussion by
+exchanging this information.
+
+\subsubsection*{Results}
+
+\paragraph{Grounding criterion} A linear regression revealed a positive relation
+between \gModel{A}{B} and the collaborative learning gain of the pair $\{A,
+B\}$: $\beta= .401, p < .001, r2adj. =.15$, large effect. This relationship was
+not significant in the previous experiment which as been conducted with the same
+task: using the same task should produce the same grounding criterion. This
+probably due to the fact that \gModel{A}{B} was influenced by the KAT.
+
+\paragraph{Study-specific questions} The $t$-test reported a significant
+difference between the KAT condition participants ($M = 13.4\%$) and the control
+group [$M = 3.6\%, t(1, 60) = 2.73, p < .01$, Cohen's $d = 0.7$, medium to large
+effect]. Providing learners with cues about the prior-knowledge of their partner
+enhances their collaborative learning. As a treatment check, we found a positive
+and significant correlation between the amount of gazes on KAT (using eye
+tracking devices) and the learning gains ($r(22) = .54, p = .01$). A detailed
+analysis revealed that subjects look at KAT to assess their peer's credibility
+when (s)he provided new information.
+
+The KAT has a significant effect on \gModel{A}{B}: peers more accurately
+estimated their partners knowledge ($M = 1.11$) than those in the control
+condition [$M = .98, t(1, 60) = 3.19, p < .01$, Cohen's $d = 0.83$, large
+effect]. This is a trivial result since the KAT provided them with an initial
+\gmodel{A}{B}. However, subjects have to predict the post-test score while the
+KAT informed them about the pre-test score. Actually, pairs in the KAT condition
+produced significantly more instances of 3 specific categories of interactions:
+(1) elaborated utterances with rich contents,  (2) utterances asking about the
+other's knowledge such as ``Did you understand how transmission works?'' (3)
+utterances describing one's own knowledge (\gmodel{A}{A}) such as ``I don't
+remember the Ranvier's thing...''. These 3 categories provide different account
+of \gModel{A}{B}: (1) as an effect of the quality of interaction, (2) as an
+effect of A's effort to model B or (3) as B's effort to give cues to A about his
+own knowledge.
+
+We examined \gModel{A}{B} as potentially mediating the effect of the KAT factor
+on the relative learning gain. A linear regression confirmed that \gModel{A}{B}
+was significantly related to the KAT-factor ($\beta= .381, p < .01, r2 = .15$).
+The KAT-factor was also significantly and positively related to the RLG ($\beta=
+.332, p < .01, r2 = .11$). We then tested the relation between the independent
+variable (KAT) and the dependent variable (gains) when controlling for the
+mediating variable (\gModel{A}{B}). A multiple regression showed that the
+KAT-factor was no longer a significant predictor ($\beta= .210; p = ns$) whereas
+the \gModel{A}{B} was still a significant predictor ($\beta= .32, p < .01$).
+Thus, it can be concluded that \gModel{A}{B} mediated the KAT-factor's effect on
+the learners' RLG. The Sobel significance test for indirect effects was
+significant [$z = 1.99, p < .05$].
+
+\paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
+0.05, NS$).
+
+\paragraph{Rectangle question} $\Delta_4$: the correlation between \gmodel{A}{A} and \gmodel{A}{B} is
+\fixme{???  MIRWEIS A FAIRE}.
+
+\subsubsection*{Discussion}
+
+Despite the fact that the learning task was the same as in study 4, the
+conditions of collaboration (viewing multiple maps or not) and the conditions
+(scripted or not, awareness tool or not) probably explain differences in terms
+of mutual modeling.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{abbrv}

### 2014-08-27 mutual-modeling.tex
Import of original Word doc is done

@@
 robot, mutual modelling plays a key role for the task performance: the tasks
-require a ``reasonably high" grounding criterion \groundingcriterion,
+require a ``reasonably high'' grounding criterion \groundingcriterion,
 and all tasks are about building a solution or a representation together.
@@
 
+\section{Synthesis}
+
+Out of 5 studies, 3 revealed a correlation between the model peers built about
+each other: when A builds an accurate model of B, B also tends to build an
+accurate model of A.  In one study, this was the case only in one condition.
+The conclusion is hence that modeling is mutual but only in certain conditions,
+summarized in Table~\ref{synthesis_table}.
+
+\begin{table*}[h!t]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{p{3cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}}
+\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
+\textit{Task}                 & Game in virtual Space            & Game in physical Space          & Building an argument map  & Building a concept map    & Building a concept map       \\
+\textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
+\textit{Shared Editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
+\textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
+\textit{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
+\textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
+\textit{Dependent variable}   & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
+\textit{Independent variable} & Awareness tool vs not.           & Awareness tool vs not.          & Audio only vs Audio+Video & Script vs  not            & Awareness tool vs not.       \\
+\textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
+\textit{$\Delta_1$}                   & +0.38 (Sig)                      & +0.41 (Sig)                     & ???                       & +0.26 (near SIG)          & +0.05 (non SIG)              \\
+\textit{$\Delta_2$}                   &                                  & +0.30                           & +0.43                     &                           &                              \\
+\textit{$\Delta_3$}                   &                                  & ???                             & +0.66                     &                           &                              \\
+\textit{$\Delta_4$}                   &                                  &                                 &                           & +0.05                     &                             
+\end{tabular}
+}
+\caption{Comparison of the studies}
+\label{synthesis_table}
+\end{table*}
+
+Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
+and 3 reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
+\emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
+modeling one's partner is not simply the result of an individual cognitive
+effort but overall the side-effect of the quality of interactions. The
+\emph{second-order-modeling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
+accurate in modeling B, A should notice when B misunderstands A and should be
+able to repair. A good modeler would somehow drive the modeling of his or her
+partner.  This \emph{drive} will however occur through verbal cues and repairs
+and is hence not easy to differentiate from $\mathcal{H}_{3}$.  The
+contradictory $\Delta_1$ results in studies 4 (to some extent) and 5 weaken the
+hypothesis.
+
+The triangular questions lead to similar contradictions. The $\Delta_2$
+correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
+support $\mathcal{H}_{2}$ if modeling skills were purely individual, there would
+be no reason for modeling skills of A and B to be correlated.  The role of C in
+\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
+the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (studies
+3 and 2?) question $\mathcal{H}_{3}$: if the role of the modellee was so
+important, it would lead be surprising to have by chance a correlation between
+B's and C' ability to drive A's modeling.
+
+In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
+($\Delta_2$ in study 4 and study 5) could also contributes to the idea that A's
+modeling skills is not the key factor. However, between modeling someone and
+modeling oneself, \ie  metacognition, they are so many differences, that we
+would hardly conclude from that,
+
+Among these complex results, it is important to stress that, for $\Delta_2$, the
+significant correlations are all below .50, not around .90.  So even if all
+$\Delta_2$ correlations were significant in all experiments, we would still not
+conclude that modelling purely reflects the quality of social interactions.
+There is obviously a large part of individual variance; some of us are much
+better in modelling peers. This is not new. The contribution of these studies is
+rather to establish the importance of team processes on something that a priori
+seemed to be more of a personal cognitive skill.  In other words, we stress that
+the bottle is half-full.
+
+These conclusions must be presented with multiple disclaimers. First, they
+heavily rely on correlations; hence we cannot identify causal links. Second, we
+faced hard methodological issues. Providing learners with on-task questionnaires
+introduce a bias: they will pay more attention to their partners in the
+remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
+and rationalization biases. The nature of mutual modeling implies methodological
+challenges that call for new  measurement methods. We have promising results for
+using eye tracking methods to address this challenge.
+
+Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
+imply we have a mechanical view of modeling. It's just a language for referring
+to different components of modeling each other in teams.  This simple notation
+emerges from the need to compare five mutual modelling studies; it did not
+generate a consistent program of consistent studies.  Our results are not very
+robust because they emerge from a \emph{post-hoc} reinterpretation of studies
+that addressed different research questions (media richness, awareness tools,
+scripts,...). This diversity makes our results difficult to integrate, partly
+contradictory. Nonetheless, this diversity gives some generalizability: mutual
+modelling has been investigated in different contexts (virtual space versus real
+space), with different groups sizes (pairs and triads) and different tasks.
+Moreover, these results have been found using different methods for measuring
+modeling: on-task in study 1 versus off-task in study 2, subjective validation
+in study 1 versus objective validation in study 2 (comparing A's model with B's
+behavior).
+
+\section*{Acknowledgments}
+
+The experiments were developed with the help of Patrick Jermann,  Fabien
+Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan Bourquin and  Jeremy Goslin.
+The main funding has been provided from a NSF Grant grant \#102511-106940.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-08-27 mutual-modeling.tex
Completed many of the missing ref, switched to Harvard-style citations

@@
 
+% UTF8 support
+\usepackage[utf8x]{inputenc}
+
 \usepackage[english]{babel}
+
+\usepackage[round]{natbib}
 
@@
 
+\author{
+    Pierre Dillenbourg, Séverin Lemaignan \\
+    \small Computer-Human Interaction in Learning and Instruction (CHILI) \\
+    \small Ecole Polytechnique Fédérale de Lausanne (EPFL) \\
+    \small CH-1015 Lausanne, Switzerland \\
+    {\tt firstname.lastname@epfl.ch}
+    \and
+    Mirweis Sangin \\
+    \and
+    Nicolas Nova \\
+    \small Haute Ecole Arts \& Design de Genève\\
+    {\tt nicolas.nova@hesge.ch}
+    \and
+    René Glaus
+    \and
+    Gaëlle Molinari \\
+    \small Fernuni Schweiz \\
+    {\tt gaelle.molinari@fernuni.ch}
+}
 \begin{document}
@@
 
-From its very beginning, CSCL research has been following Roschelle \& Teasley's
-(1986) suggestion that collaborative learning has something to do with the
-process of constructing and maintaining a shared understanding of the task at
-hand. Building a shared/mutual understanding refers to the upper class of
-collaborative learning situations, those in which students should build upon
-each other's understanding to refine their own understanding. What is expected
-to produce learning is not the mere fact that two students build the same
-understanding but the cognitive effort they have to engage to build this shared
-understanding (Schwartz, 1995). This effort can be observed by the frequency of
-rich interactions, \ie interactions whose occurrence has been related to
-learning: (self-) explanations in cognitive science (Chi ; Webb), conflict
-resolution in socio-cognitive theories (Doise), mutual regulation (Blaye,) in a
-Vygostkian perspective, etc. The construction of a shared understanding has been
-investigated for several years in psycholinguistics, under the  notion of
-“grounding” (Clark \&Wilkes-Gibbs, 1986). However, the relevance of grounding
+From its very beginning, CSCL research has been following
+\citet{roschelle1995construction} suggestion that collaborative learning has
+something to do with the process of constructing and maintaining a shared
+understanding of the task at hand. Building a shared/mutual understanding refers
+to the upper class of collaborative learning situations, those in which students
+should build upon each other's understanding to refine their own understanding.
+What is expected to produce learning is not the mere fact that two students
+build the same understanding but the cognitive effort they have to engage to
+build this shared understanding~\citep{schwartz1995emergence}. This effort can
+be observed by the frequency of rich interactions, \ie interactions whose
+occurrence has been related to learning: (self-) explanations in cognitive
+science (Chi ; Webb), conflict resolution in socio-cognitive theories (Doise),
+mutual regulation (Blaye,) in a Vygostkian perspective, etc. The construction of
+a shared understanding has been investigated for several years in
+psycholinguistics, under the  notion of
+``grounding''~\citep{clark1986referring}.  However, the relevance of grounding
 mechanisms for explaining learning outcomes has been questioned. The monitoring
@@
 short dialogue episodes but does hardly predict conceptual change over longer
-sessions (Dillenbourg \& Traum, 2006). The cumulative effect of grounding
+sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
 episodes can probably be better understood from a socio-cultural perspective:
-"collaborative learning is associated with the increased cognitive-interactional
-effort involved in the transition from learning to understand each other to
-learning to understand the meanings of the semiotic tools that constitute the
-mediators of interpersonal interaction" (Baker, Hansen, Joiner \& Traum, 1999).
-Several scholars suggest that CSCL research should go deeper in the
-understanding of how partners engage into shared meaning making (Stahl, 2006) or
-'intersubjective' meaning making (Suthers (2006).
+``collaborative learning is associated with the increased
+cognitive-interactional effort involved in the transition from learning to
+understand each other to learning to understand the meanings of the semiotic
+tools that constitute the mediators of interpersonal
+interaction''~\citep{baker1999role}.  Several scholars suggest that CSCL
+research should go deeper in the understanding of how partners engage into
+shared meaning making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
+making~\citep{suthers2006technology}.
 
@@
 levels of dialogue, mostly between the lexical / syntactical level and the
-deeper semantic levels. Pickering \& Garrod (2004) argue that the mutual
-understanding starts mostly with a 'local alignment' at the level of the
+deeper semantic levels. \citet{pickering2006alignment} argue that the mutual
+understanding starts mostly with a \emph{local alignment} at the level of the
 linguistic representations, due to priming mechanisms, and that this superficial
-alignment may – in some cases- lead to a 'global alignment' of the semantic
-level.  For these authors, the convergence in dialogue, and even the repair of
-some misunderstandings, is explained by this mimetic behavior more than by a
-monitoring of each other's knowledge: "…interlocutors do not need to monitor and
-develop full common ground as a regular, constant part of routine conversation,
-as it would be unnecessary and far too costly. Establishment of full common
-ground is, we argue, a specialized and non-automatic process that is used
-primarily in times of difficulty (when radical misalignment becomes apparent)."
-(p. 179). This view is actually not incompatible with Clark's grounding
-criterion (REF): the degree of shared understanding that peers need to reach
-depends upon the task they perform. For instance, a dialogue between two
-surgeons might rely on superficial alignment if they talk about their friends
-but has to guarantee accurate common grounds when talking about which
-intervention will be conducted in which way on which patient.  Some CSCL
-scholars stressed the 'illusion of shared understanding' (JACCO?), \ie the
-potential decoupling between linguistic alignment and actual shared meanings.
+alignment may -- in some cases -- lead to a \emph{global alignment} of the
+semantic level.  For these authors, the convergence in dialogue, and even the
+repair of some misunderstandings, is explained by this mimetic behavior more
+than by a monitoring of each other's knowledge: ``...interlocutors do not need
+to monitor and develop full common ground as a regular, constant part of routine
+conversation, as it would be unnecessary and far too costly. Establishment of
+full common ground is, we argue, a specialized and non-automatic process that is
+used primarily in times of difficulty (when radical misalignment becomes
+apparent).'' (p. 179). This view is actually not incompatible with Clark's
+grounding criterion \fixme{(REF)}: the degree of shared understanding that peers
+need to reach depends upon the task they perform. For instance, a dialogue
+between two surgeons might rely on superficial alignment if they talk about
+their friends but has to guarantee accurate common grounds when talking about
+which intervention will be conducted in which way on which patient.  Some CSCL
+scholars stressed the ``illusion of shared understanding'' \fixme{REF:(JACCO?)},
+\ie the potential decoupling between linguistic alignment and actual shared
+meanings.
 
@@
 empirical studies on grounding and alignment are conducted with tasks, which
-despite being qualified as 'ecologically valid' by their authors, are mere
+despite being qualified as ``ecologically valid'' by their authors, are mere
 referencing tasks such as asking the way to the train station or helping the
@@
 
-Deep grounding or shared meaning making requires some cognitive load. For Clark
-\& Wilkes-Gibbs (1986), what is important is not the individual effort made by
-the receiver of a communicative act, but the overall least collaborative effort.
-The cost of producing a perfect utterance may be higher than the cost of
-repairing the problems that may arise through misunderstandings. For instance,
-subjects are less careful about adapting their utterances to their partner when
-they know they can provide feedback on his/her understanding (Schober, 1993). We
-introduced the notion of ‘optimal collaborative effort' (Dillenbourg et al,
-1996) to stress that misunderstanding should not be viewed as something to be
-avoided (if this was possible), but as an opportunity to engage into
-verbalization, explanation, negotiation, and so forth. This issue is related to
-the global argument regarding cognitive load in learning activities, especially
-in discovery learning environments: there is no learning without some cognitive
-load, but overload may hinder learning (Paas, Renkl \& Sweller, 2003). In the
-context of collaborative learning, we understand the cognitive load induced by
-mutual modeling as part of Schwartz's (1995) notion of effort towards a shared
-understanding. For instance, CSCL researchers expanded the use of 'collaboration
-scripts' (Aronson,?). A script is a pedagogical method that collaborative
-learning activities in order to foster the emergence of productive interactions
-such as argumentation, explanation or conflict. Conflict-resolution scripts such
-as the ArgueGraph (Dillenbourg \& Hong, 2008) forms pair of students with
+Deep grounding or shared meaning making requires some cognitive load. For
+\citet{clark1986referring}, what is important is not the individual effort made
+by the receiver of a communicative act, but the overall least collaborative
+effort.  The cost of producing a perfect utterance may be higher than the cost
+of repairing the problems that may arise through misunderstandings. For
+instance, subjects are less careful about adapting their utterances to their
+partner when they know they can provide feedback on his/her
+understanding~\citep{schober1993spatial}. We introduced the notion of
+\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
+that misunderstanding should not be viewed as something to be avoided (if this
+was possible), but as an opportunity to engage into verbalization, explanation,
+negotiation, and so forth. This issue is related to the global argument
+regarding cognitive load in learning activities, especially in discovery
+learning environments: there is no learning without some cognitive load, but
+overload may hinder learning~\citep{paas2003cognitive}. In the context of
+collaborative learning, we understand the cognitive load induced by mutual
+modeling as part of~\citet{schwartz1995emergence} notion of effort towards a
+shared understanding. For instance, CSCL researchers expanded the use of
+\emph{collaboration scripts} \fixme{REF (Aronson,?)}. A script is a pedagogical
+method that collaborative learning activities in order to foster the emergence
+of productive interactions such as argumentation, explanation or conflict.
+Conflict-resolution scripts such as the {\sc
+ArgueGraph}~\citep{dillenbourg2008mechanics} forms pair of students with
 opposite opinions, which increases the difficulty of consensus building,
@@
 %
-%Theory of Mind (originally defined in~\cite{premack1978does}) is the cognitive
+%Theory of Mind (originally defined in~\citep{premack1978does}) is the cognitive
 %ability that a subject possesses to represent the mental state of another
@@
 %\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
-%experiment)~\cite{Leslie2000}: a child is asked to watch a scene where two
+%experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
 %people, A and B, manipulate objects. Then A leaves and B hides away one
@@
 term knowledge, skills or traits. It is thus closely related to the notion of
-transactive memory~\cite{wegner1987transactive, moreland1999transactive}.
+transactive memory~\citep{wegner1987transactive, moreland1999transactive}.
 'Situational' aspects refer to A's representation of B's knowledge, behavior or
@@
 their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
-Ginsburg~\cite{slugoski1993attribution} pretended to their subjects that their
+Ginsburg~\citep{slugoski1993attribution} pretended to their subjects that their
 (fake) partner had or had not received the same information. They observed that
@@
 Initial common grounds are also initiated by co-presence: they include events to
-which A and B attended together~\cite{clark2002definite} in the physical space
+which A and B attended together~\citep{clark2002definite} in the physical space
 or in our cultural space (\eg ‘09-11'). Co-presence means that we can refer to
@@
 same meaning. Namely, a shared screen does not mean a shared
-understanding~\cite{dillenbourg2006sharing}.
+understanding~\citep{dillenbourg2006sharing}.
 
@@
 Some environments are more productive than others in helping peers to detect
-their misunderstandings. Roschelle~\cite{roschelle1995construction} reformulate
+their misunderstandings. Roschelle~\citep{roschelle1995construction} reformulate
 the design of CSCL interfaces as providing ways for peer to detect and repairs
@@
 %person's point of view. Studied in psychology
-%literature~\cite{Flavell1992,Tversky1999}, this ability is crucial when
+%literature~\citep{Flavell1992,Tversky1999}, this ability is crucial when
 %interacting with people by allowing one to reason on others' understanding of
@@
 %beliefs, etc.  In the last years these notions have been gradually employed in
-%HRI.~\cite{Breazeal2006} presents a learning algorithm that takes into account
+%HRI.~\citep{Breazeal2006} presents a learning algorithm that takes into account
 %information about a teacher's visual perspective in order to learn a task.
-%~\cite{Johnson2005} apply visual perspective taking for action recognition
-%between two robots.~\cite{Trafton2005} use both visual and spatial perspective
+%~\citep{Johnson2005} apply visual perspective taking for action recognition
+%between two robots.~\citep{Trafton2005} use both visual and spatial perspective
 %taking to find out the referent indicated by a human partner.
@@
 %\paragraph{Joint action}
-%Several theories dealing with collaboration~\cite{Cohen1991,Grosz1996,Clark1996}
+%Several theories dealing with collaboration~\citep{Cohen1991,Grosz1996,Clark1996}
 %emphasize that collaborative tasks have specific requirements compared to
@@
 %the goal during execution, etc. Several robotic systems have already been built
-%based on these theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they all
+%based on these theories~\citep{Rich1997,Sidner2005,Breazeal2003} and they all
 %have shown benefits of this approach. They have also shown how difficult it is
@@
 %realization and communication in a generic way. Finally, today only few
-%systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
+%systems~\citep{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
 %levels.
@@
 at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
-costs (3)~\cite{traum1996miscommunication}:
+costs (3)~\citep{traum1996miscommunication}:
 
@@
 robot, estimating \inigrounding still requires special care during the design of
-the experiments: as we have previously shown~\cite{lemaignan2014dynamics,
+the experiments: as we have previously shown~\citep{lemaignan2014dynamics,
 lemaignan2014cognitive} and because the interaction takes place between a human and
@@
 We studied the impact of an awareness tool on group performance and mutual
-modeling~\cite{nova2007collaboration}. The availability
+modeling~\citep{nova2007collaboration}. The availability
 of an awareness tool was our independent variable. In previous studies, we
@@
 \paragraph{Symmetry question} We computed intra-class correlation as described
-by~\cite{kenny1998data} from the answers to the cross-questionnaires.
+by~\citep{kenny1998data} from the answers to the cross-questionnaires.
 Considering all pairs in both conditions, we found a positive and significant
@@
 This study concerns a collaborative game that occurred in the physical
-space~\cite{nova2006underwhelming}. We studied whether players build an accurate
+space~\citep{nova2006underwhelming}. We studied whether players build an accurate
 model of the path followed by their partners because this path reflected their
@@
 without awareness tool made more annotations on the map while permanent MLA has
-an underwhelming effect~\cite{nova2005location}.
+an underwhelming effect~\citep{nova2005location}.
 
@@
 than the previous one: despite the high heterogeneity of spatial skills among
-adults~\cite{liben1981spatial}, a significant correlation could indicate that
+adults~\citep{liben1981spatial}, a significant correlation could indicate that
 mutual modeling relies more on group processes than on individual processes. The
@@
 graphical editor) was, in this task, more important than seeing each
-other~\cite{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
+other~\citep{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
 Anderson et al., 2000}.
@@
 This study investigated the effect of a collaboration script on collaborative
-learning~\cite{molinari2008effects}. The script chosen is a JIGSAW: two students
+learning~\citep{molinari2008effects}. The script chosen is a JIGSAW: two students
 receive different but complementary subsets of the knowledge (texts) which have
@@
 The peers were located into two different rooms. A complete description of the
-study is provided in~\cite{sangin2008learners}. The experiment lasted 90 minutes.
+study is provided in~\citep{sangin2008learners}. The experiment lasted 90 minutes.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{abbrv}
+\bibliographystyle{plainnat}
 \bibliography{biblio}

### 2015-04-01 mutual-modeling.tex
Merged modifs from HRI paper + a few automated fixes like double quotes

@@
 \citet{roschelle1995construction} suggestion that collaborative learning has
-something to do with the process of constructing and maintaining a shared
-understanding of the task at hand. Building a shared/mutual understanding refers
-to the upper class of collaborative learning situations, those in which students
-should build upon each other's understanding to refine their own understanding.
-What is expected to produce learning is not the mere fact that two students
-build the same understanding but the cognitive effort they have to engage to
-build this shared understanding~\citep{schwartz1995emergence}. This effort can
-be observed by the frequency of rich interactions, \ie interactions whose
-occurrence has been related to learning: (self-) explanations in cognitive
+something to do with the process of constructing and maintaining a \emph{shared
+understanding} of the task at hand. Building a shared/mutual understanding
+refers to the upper class of collaborative learning situations, those in which
+students should build upon each other's understanding to refine their own
+understanding.  What is expected to produce learning is not the mere fact that
+two students build the same understanding but the cognitive effort they have to
+engage to build this shared understanding~\citep{schwartz1995emergence}. This
+effort can be observed by the frequency of rich interactions, \ie interactions
+whose occurrence has been related to learning: (self-) explanations in cognitive
 science (Chi ; Webb), conflict resolution in socio-cognitive theories (Doise),
@@
 ``grounding''~\citep{clark1986referring}.  However, the relevance of grounding
-mechanisms for explaining learning outcomes has been questioned. The monitoring
-and repair of mis-understanding explains for instance referential failures in
-short dialogue episodes but does hardly predict conceptual change over longer
-sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
+mechanisms for explaining learning outcomes has been questioned in learning
+sciences. The monitoring and repair of misunderstanding explains for instance
+referential failures in short dialogue episodes but does hardly predict
+\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
+new belief into one's mental model\fxnote{check with Pierre}) over longer
+sessions~\cite{dillenbourg2006sharing}. The cumulative effect of grounding
 episodes can probably be better understood from a socio-cultural perspective:
-``collaborative learning is associated with the increased
-cognitive-interactional effort involved in the transition from learning to
-understand each other to learning to understand the meanings of the semiotic
+
+\begin{quote}
+Collaborative learning is associated with the increased
+cognitive-interactional effort involved in the transition from \emph{learning to
+understand each other} to \emph{learning to understand the meanings of the semiotic
 tools that constitute the mediators of interpersonal
-interaction''~\citep{baker1999role}.  Several scholars suggest that CSCL
-research should go deeper in the understanding of how partners engage into
-shared meaning making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
-making~\citep{suthers2006technology}.
+interaction}~\citep{baker1999role}
+\end{quote}
+
+Along this line, several scholars suggest that CSCL research should go deeper in
+the understanding of how partners engage into shared meaning
+making~\cite{stahl2007meaning} or \emph{intersubjective} meaning
+making~\cite{suthers2006technology}.
 
@@
 change viewpoint, it is criticized at the same time as being too cognitivist by
-some psycholinguists, \ie as overestimating the amount shared knowledge and
+some psycholinguists, \ie as overestimating the amount of shared knowledge and
 mutual representations actually necessary to conduct a dialogue. The fundamental
 issue, as old as philosophy, is the degree of coupling between the different
-levels of dialogue, mostly between the lexical / syntactical level and the
-deeper semantic levels. \citet{pickering2006alignment} argue that the mutual
-understanding starts mostly with a \emph{local alignment} at the level of the
-linguistic representations, due to priming mechanisms, and that this superficial
+levels of dialogue, mostly between the lexical/syntactical level and the deeper
+semantic levels. \citet{pickering2006alignment} argue that the mutual
+understanding starts mostly with a \emph{superficial alignment} at the level of
+the linguistic representations, due to priming mechanisms, and that this local
 alignment may -- in some cases -- lead to a \emph{global alignment} of the
-semantic level.  For these authors, the convergence in dialogue, and even the
-repair of some misunderstandings, is explained by this mimetic behavior more
-than by a monitoring of each other's knowledge: ``...interlocutors do not need
-to monitor and develop full common ground as a regular, constant part of routine
-conversation, as it would be unnecessary and far too costly. Establishment of
-full common ground is, we argue, a specialized and non-automatic process that is
-used primarily in times of difficulty (when radical misalignment becomes
-apparent).'' (p. 179). This view is actually not incompatible with Clark's
-grounding criterion \fixme{(REF)}: the degree of shared understanding that peers
-need to reach depends upon the task they perform. For instance, a dialogue
+semantic level (\emph{deep grounding}).  For these authors, the convergence in
+dialogue, and even the repair of some misunderstandings, is explained by this
+mimetic behavior more than by a monitoring of each other's knowledge:
+\emph{``...interlocutors do not need to monitor and develop full common ground
+as a regular, constant part of routine conversation, as it would be
+unnecessary and far too costly. Establishment of full common ground is, we
+argue, a specialized and non-automatic process that is used primarily in
+times of difficulty (when radical misalignment becomes apparent).''} (p. 179).
+This view is actually not incompatible with Clark's \emph{grounding
+criterion}~\cite{clark1989contributing}: the degree of shared understanding that
+peers need to reach depends upon the task they perform. For instance, a dialogue
 between two surgeons might rely on superficial alignment if they talk about
@@
 \citet{clark1986referring}, what is important is not the individual effort made
-by the receiver of a communicative act, but the overall least collaborative
-effort.  The cost of producing a perfect utterance may be higher than the cost
+by the receiver of a communicative act, but the overall \emph{least collaborative
+effort}.  The cost of producing a perfect utterance may be higher than the cost
 of repairing the problems that may arise through misunderstandings. For
@@
 incorrect… but it can not remain empty. Dialogues include many instances of
-utterances such as "I thought he would do that" (first level of mutual
-modelling) or even "He thought I would do that but I intended something else."
+utterances such as ``I thought he would do that'' (first level of mutual
+modelling) or even ``He thought I would do that but I intended something else.''
 (second level of mutual modelling).
@@
 
-    \item $Model(A,B, "urgent")$: What does A about know B's understanding of A's
+    \item $Model(A,B, ``urgent'')$: What does A about know B's understanding of A's
         last utterance: does ‘urgent' means now, ASAP and ‘not too lat' ?
@@
 
-We have different levels of mutual modelling. If A states "B thinks I am god in maths", we
-are the second level of mutual modelling:  $Model(A, B, Model(B,A, knowledge="good in
-math"))$. There is an infinite regress of nested models: If A says "B knows that
-I don't expect him to solve this statistics problem" corresponds to $Model(A, B,
-Model (B, A (Model (A,B, statistic-skills)))$.
-
-A partner model is probably not a "box", \ie not a monolithical representation
+We have different levels of mutual modelling. If A states ``B thinks I am good in
+maths'', we are the second level of mutual modelling:  $Model(A, B, Model(B,A,
+knowledge=``good in math''))$. There is an infinite regress of nested models: If A
+says ``B knows that I don't expect him to solve this statistics problem''
+corresponds to $Model(A, B, Model (B, A (Model (A,B, statistic-skills)))$.
+
+A partner model is probably not a ``box'', \ie not a monolithical representation
 but rather a mosaic of information fragments about the partner, with various
@@
 After initialization, mutual models are updated along the collaborative work
-through verbal and non-verbal interactions. A default inference rule is that "my
-partner agrees with me unless he disagrees", which reject the critiques that
+through verbal and non-verbal interactions. A default inference rule is that ``my
+partner agrees with me unless he disagrees'', which reject the critiques that
 mutual modeling generates an unbearable cognitive load. This default rule is
@@
 their misunderstanding. In CSCW, researchers have explored various so-called
-"awareness tools", \ie functionalities that inform A is about B's actions that A
+``awareness tools'', \ie functionalities that inform A is about B's actions that A
 could not directly perceived because B was working on a different subset of the
@@
 To report experiments on mutual modeling, we suggest
-the notation \model{A}{B}{X} to denote ``A knows that B knows X". This is a
+the notation \model{A}{B}{X} to denote ``A knows that B knows X''. This is a
 simple but useful reduction to reason on mutual modelling. This
@@
 
-    \item Tasks vary a lot with respect to how much they require.  The grounding
+    \item Tasks vary a lot with respect to how much they require mutual
+        understanding.  The grounding
         criterion -- denoted \groundingcriterion -- refers to how
-        important it is to mutually share a piece of information X to succeed
-        the task T. It can be computed as the probability to succeed T despite
-        the fact X is not grounded. $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be
-        estimated from the correlation between \Model{A}{B}{X} and task
+        important it is to mutually share a piece of information $X$ to succeed
+        the task $T$. It can be computed as the probability to succeed T despite
+        the fact $X$ is not grounded. $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be
+        estimated from the correlation between \Model{A}{B}{X} and the task
         performance.
 
-    \item Before any specific grounding act, there is usually a non-null probability
-        that X is mutually understood by A and B: because X is part of A's and
-        B's cultures, because it is manifest to co-present subjects or simply
-        because there is not much space for misunderstanding or disagreement
-        about X. We could not simply collaborate without a high level of initial
-        grounds. We note the theoretical accuracy of initial grounds:
+    \item Before any specific grounding action, there is usually a non-null probability
+        that $X$ is mutually understood by $A$ and $B$ (\eg $X$
+        is part of $A$'s and $B$'s cultures, it is manifest to co-present
+        subjects or simply there is not much space for misunderstanding
+        or disagreement about $X$). We simply could not collaborate without a high level of initial
+        grounds. We note the theoretical accuracy of \emph{initial grounds}
         $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
 
-    \item The cost of grounding X refers to the physical and cognitive effort
-        necessary to perform a grounding act $\alpha$: a verbal repair (\eg
+    \item The cost of grounding $X$ refers to the physical and cognitive effort
+        required to perform a grounding act $\alpha$: a verbal repair (\eg
         rephrasing), a deictic gesture, a physical move to adopt one partner's
-        viewpoint, etc. This cost varies according to media features ( Clark \&
-        Brennan (REF)).
+        viewpoint, etc. This cost varies according to media
+        features~\cite{clark1991grounding}.
 
@@
         (study 4 and 5). By varying the tasks, we do actually vary the grounding
-        criterion but they all require a "reasonably high" grounding criterion.
+        criterion but they all require a ``reasonably high'' grounding criterion.
         All tasks are about building a solution or a representation together; we
@@
         an independent variable. Study 3 uses media richness as independent
-        variable, with the hypothesis that modeling emotions is "cheaper" with a
+        variable, with the hypothesis that modeling emotions is ``cheaper'' with a
         richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
@@
 
-Each of the studies that we present in the next section have been design in such a way that, while tractable on a
-robot, mutual modelling plays a key role for the task performance: the tasks
-require a ``reasonably high'' grounding criterion \groundingcriterion,
-and all tasks are about building a solution or a representation together.
+Each of the studies that we present in the next section have been design in such
+a way that, while tractable on a robot, mutual modelling plays a key role for
+the task performance: the tasks require a ``reasonably high'' grounding
+criterion \groundingcriterion, and all tasks are about building a solution or a
+representation together.
 
@@
         bias, one can estimate mutual modeling after task completion. Then, the
-        obvious drawback are memory losses and by post-hoc reconstruction.
-        The first method was used in study 2 and the
-        second one in the other studies. Another option would be to rely on external behavioural metrics like
-        eye-tracking: we hypothesise for instance that the fixation time
+        obvious drawback are memory losses and by post-hoc reconstruction.  The
+        first method was used in study 2 and the second one in the other
+        studies. Another option would be to rely on external behavioural metrics
+        like eye-tracking: we hypothesise for instance that the fixation time
         reflects the efforts engaged by the human to understand, hence, model,
-        the others. Such an approach has however not yet been investigated in the presented studies.
+        the others. Such an approach has however not yet been investigated in
+        the presented studies.
 
@@
 
-The experiments we report here address mutual modelling across different tasks, some with
-dyads, other with triads. They were conducted along 6 years in two different
-institutions by different researchers. They used different independent,
-intermediate and dependent variables. Nonetheless, we were able to address a few
-questions across these studies. We start with 3 simple hypotheses about
-$\mathcal{M}^{\circ}(A,B)$:
+The experiments we report here address mutual modelling across different tasks,
+some with dyads, other with triads. They were conducted along 6 years in two
+different institutions by different researchers. They used different
+independent, intermediate and dependent variables. Nonetheless, we were able to
+address a few questions across these studies. We start with 3 simple hypotheses
+about $\mathcal{M}^{\circ}(A,B)$:
 
@@
 Our main issue is the symmetry questions: what is the relationship
-(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and
-\Model{B}{A}{X}? A low symmetry would mean that mutual modeling is
-mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
-might support $\mathcal{H}_{2}$ since there is a low probability that randomly
-formed pairs integrate peers with the same level of mutual modelling skills.
+(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and \Model{B}{A}{X}? A low
+symmetry would mean that mutual modeling is mainly an individual
+attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might support
+$\mathcal{H}_{2}$ since there is a low probability that randomly formed pairs
+integrate peers with the same level of mutual modelling skills.
 
@@
 
-\begin{tikzpicture}[scale=0.5]
+\begin{tikzpicture}[>=latex, scale=0.5]
 
@@
 
-\caption{The mutual modelling symmetry question $\Delta_1 =  \Delta(\mathcal{M}^{\circ} (A,B,X),
+\caption{\small Mutual modelling in a dyadic interaction, $\Delta_1 =
+    \Delta(\mathcal{M}^{\circ} (A,B,X),
 \mathcal{M}^{\circ} (B,A,X))$}
@@
 \Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
-\Model{C}{B}{X} and \Model{B}{C}{X}. This enables two
-triangle questions (see Fig.~\ref{mm_triangles}):
-
-Do A and B have the same accuracy when modeling C? $\Delta_2 =
-\Delta(\M{A}{C}{X}, \M{B}{C}{X})$ If it is the same,
-$\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain over $\mathcal{H}_{1}$
-
-Does C model more accurately A than B? $\Delta_3= \Delta(\M{C}{A}{X},
+\Model{C}{B}{X} and \Model{B}{C}{X}.
+
+This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do A and
+B have the same accuracy when modeling C? $\Delta_2 = \Delta(\M{A}{C}{X},
+\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain
+over $\mathcal{H}_{1}$
+
+Conversely, does C model more accurately A than B? $\Delta_3= \Delta(\M{C}{A}{X},
 \M{C}{B}{X})$ A positive answer would support $\mathcal{H}_{3}$.
@@
 \subcaptionbox{}{ 
-    \begin{tikzpicture}[scale=0.5]
+    \begin{tikzpicture}[>=latex, scale=0.5]
 
@@
 }
-\caption{The mutual modelling triangle questions}
+\caption{\small Mutual modelling in a triadic interaction}
 
@@
 
-We could go further by comparing self- versus other modeling $\Delta_4$ in
-Fig.~\ref{mm_rectangle}) as an indication of metacognitive skills. We could also
-see if modeling skills depend upon what aspects are being modeled (X or Y),
+We could go further by comparing self- versus other modeling ($\Delta_4$ in
+Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
+question if modelling skills depend upon what aspects are being modeled ($X$ or $Y$),
 which would explain vertical differences ($\Delta_5$ in
-Fig.~\ref{mm_rectangle}). We do not further develop these questions because we
+Fig.~\ref{mm_rectangle}). We do not however further develop these questions because we
 don't have the necessary data in our studies.
@@
 
-\caption{The rectangle questions}
+\caption{\small Meta-cognitive skills and domain-dependent modelling}
 
@@
 SpaceMiners is a 3D game that involves two players collecting minerals located
-in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they shoot drones through the space after choosing their
-initial direction and speed. Once launched, the trajectory of drones is only
-influenced by the gravity of planets and by trajectory modification tools.
-During the experiment, the teams were confronted with three increasingly complex
-situations. The experiment was 2 hours long: a 30 minutes tutorial and 3 levels
-of 30 minutes. Players uses using a regular joystick and communicated with each
-other through an audio channel.
+in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they shoot drones through
+the space after choosing their initial direction and speed. Once launched, the
+trajectory of drones is only influenced by the gravity of planets and by
+trajectory modification tools.  During the experiment, the teams were confronted
+with three increasingly complex situations. The experiment was 2 hours long: a
+30 minutes tutorial and 3 levels of 30 minutes. Players uses using a regular
+joystick and communicated with each other through an audio channel.
 

### 2015-04-01 mutual-modeling.tex
{modeling->modelling} to follow bristish english rules

@@
 
-\title{Is Partner Modeling Mutual?}
+\title{Is Partner Modelling Mutual?}
 
@@
 collaborative learning, we understand the cognitive load induced by mutual
-modeling as part of~\citet{schwartz1995emergence} notion of effort towards a
+modelling as part of~\citet{schwartz1995emergence} notion of effort towards a
 shared understanding. For instance, CSCL researchers expanded the use of
@@
 
-\section{Mutual Modeling}
+\section{Mutual Modelling}
 
 In order to build a shared understanding, do partners have to build a
-representation of each other's knowledge? We refer to \emph{mutual modeling} as
+representation of each other's knowledge? We refer to \emph{mutual modelling} as
 the process of inferring one's partner mental states. Any claim that students
 carry out a detailed monitoring of their peers would be as incorrect as any
-claim that they do not maintain any representation at all. If mutual modeling
+claim that they do not maintain any representation at all. If mutual modelling
 had to be permanently detailed and accurate, subjects would obviously face a
 huge cognitive load. Conversely, peers could not collaborate without some
-minimal amount of mutual modeling. For instance, A cannot disagree with B
+minimal amount of mutual modelling. For instance, A cannot disagree with B
 without knowing that B has a different opinion. The mutual model can be implicit
@@
 stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
-appearance. Scholars studied how initial modeling impacts communication. In
+appearance. Scholars studied how initial modelling impacts communication. In
 their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
@@
 partner agrees with me unless he disagrees'', which reject the critiques that
-mutual modeling generates an unbearable cognitive load. This default rule is
+mutual modelling generates an unbearable cognitive load. This default rule is
 superseded by the several mechanisms for monitoring and repairing one's partner
@@
 
-Mutual modeling does not occur in a vacuum but it is highly contextualized. REF
+Mutual modelling does not occur in a vacuum but it is highly contextualized. REF
 reviews the features of the collaborative situation, namely the media
-(cotemporality… ), may facilitate or hamper mutual modeling. Hutchins (SILENCE)
+(cotemporality… ), may facilitate or hamper mutual modelling. Hutchins (SILENCE)
 reported a study in which a short silence between two pilots was perfectly
@@
 
-To report experiments on mutual modeling, we suggest
+To report experiments on mutual modelling, we suggest
 the notation \model{A}{B}{X} to denote ``A knows that B knows X''. This is a
@@
         an independent variable. Study 3 uses media richness as independent
-        variable, with the hypothesis that modeling emotions is ``cheaper'' with a
+        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
         richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
@@
         obvious methodological concerns since such a question triggers a
-        modeling process beyond what it would 'naturally' be. To avoid this
-        bias, one can estimate mutual modeling after task completion. Then, the
+        modelling process beyond what it would 'naturally' be. To avoid this
+        bias, one can estimate mutual modelling after task completion. Then, the
         obvious drawback are memory losses and by post-hoc reconstruction.  The
@@
 
-$\mathcal{H}_{2}$ is also called a second level modeling hypothesis since H
+$\mathcal{H}_{2}$ is also called a second level modelling hypothesis since H
 needs to monitor B to see if A understood him/her
@@
 (Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and \Model{B}{A}{X}? A low
-symmetry would mean that mutual modeling is mainly an individual
+symmetry would mean that mutual modelling is mainly an individual
 attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might support
@@
 This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do A and
-B have the same accuracy when modeling C? $\Delta_2 = \Delta(\M{A}{C}{X},
+B have the same accuracy when modelling C? $\Delta_2 = \Delta(\M{A}{C}{X},
 \M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain
@@
 
-We could go further by comparing self- versus other modeling ($\Delta_4$ in
+We could go further by comparing self- versus other modelling ($\Delta_4$ in
 Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
@@
 This simple notation does not pretend to provide a mathematical account of
-mutual modeling but to be more systematic in describing the following
+mutual modelling but to be more systematic in describing the following
 experiments.
@@
 We studied the impact of an awareness tool on group performance and mutual
-modeling~\citep{nova2007collaboration}. The availability
+modelling~\citep{nova2007collaboration}. The availability
 of an awareness tool was our independent variable. In previous studies, we
@@
 Task performance was measured by the score reached by the two subjects after
-three situations. The effort of mutual modeling was measured as the ratio of
+three situations. The effort of mutual modelling was measured as the ratio of
 time that players would spend in the scout mode (divided by total time), which
@@
 our university because knowing the campus geography had an impact both on group
-performance but also on mutual modeling: it is hard to represent the path of
+performance but also on mutual modelling: it is hard to represent the path of
 someone across some space without having a mental map of this space. We selected
@@
 therefore consider mistakes in their partners' path as being due to a lack of
-mutual modeling accuracy instead of being due to spatial reasoning skills.
+mutual modelling accuracy instead of being due to spatial reasoning skills.
 
@@
 adults~\citep{liben1981spatial}, a significant correlation could indicate that
-mutual modeling relies more on group processes than on individual processes. The
+mutual modelling relies more on group processes than on individual processes. The
 significant results on the triangle questions probably go in the same direction:
-either the modeler or the modeled both contribute to accurate modeling, or that
+either the modeler or the modeled both contribute to accurate modelling, or that
 the quality of teamwork has a global effect. In this experiment, since team
 members did not interact intensively during the task, the factor that
-facilitates accurate mutual modeling is probably less the quality of verbal
+facilitates accurate mutual modelling is probably less the quality of verbal
 interactions that the group strategy: a clear strategy facilitates memorizing
@@
 
-\paragraph{Symmetry question} Regarding the main items, the mutual modeling of
+\paragraph{Symmetry question} Regarding the main items, the mutual modelling of
 emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}
@@
 and \gmodel{A}{B} is globally not significant ($r=0.05$) as well as not
-significant in each condition: someone good at self-modeling is not necessarily
-good at modeling someone else and vice-versa.
+significant in each condition: someone good at self-modelling is not necessarily
+good at modelling someone else and vice-versa.
 
@@
 (scripted or not, awareness tool or not) probably explain differences in terms
-of mutual modeling.
+of mutual modelling.
 
@@
 accurate model of A.  In one study, this was the case only in one condition.
-The conclusion is hence that modeling is mutual but only in certain conditions,
+The conclusion is hence that modelling is mutual but only in certain conditions,
 summarized in Table~\ref{synthesis_table}.
@@
 \emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
-modeling one's partner is not simply the result of an individual cognitive
+modelling one's partner is not simply the result of an individual cognitive
 effort but overall the side-effect of the quality of interactions. The
-\emph{second-order-modeling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
-accurate in modeling B, A should notice when B misunderstands A and should be
-able to repair. A good modeler would somehow drive the modeling of his or her
+\emph{second-order-modelling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
+accurate in modelling B, A should notice when B misunderstands A and should be
+able to repair. A good modeler would somehow drive the modelling of his or her
 partner.  This \emph{drive} will however occur through verbal cues and repairs
@@
 correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
-support $\mathcal{H}_{2}$ if modeling skills were purely individual, there would
-be no reason for modeling skills of A and B to be correlated.  The role of C in
+support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
+be no reason for modelling skills of A and B to be correlated.  The role of C in
 \gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
@@
 important, it would lead be surprising to have by chance a correlation between
-B's and C' ability to drive A's modeling.
+B's and C' ability to drive A's modelling.
 
@@
 ($\Delta_2$ in study 4 and study 5) could also contributes to the idea that A's
-modeling skills is not the key factor. However, between modeling someone and
-modeling oneself, \ie  metacognition, they are so many differences, that we
+modelling skills is not the key factor. However, between modelling someone and
+modelling oneself, \ie  metacognition, they are so many differences, that we
 would hardly conclude from that,
@@
 remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
-and rationalization biases. The nature of mutual modeling implies methodological
+and rationalization biases. The nature of mutual modelling implies methodological
 challenges that call for new  measurement methods. We have promising results for
@@
 Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
-imply we have a mechanical view of modeling. It's just a language for referring
-to different components of modeling each other in teams.  This simple notation
+imply we have a mechanical view of modelling. It's just a language for referring
+to different components of modelling each other in teams.  This simple notation
 emerges from the need to compare five mutual modelling studies; it did not
@@
 Moreover, these results have been found using different methods for measuring
-modeling: on-task in study 1 versus off-task in study 2, subjective validation
+modelling: on-task in study 1 versus off-task in study 2, subjective validation
 in study 1 versus objective validation in study 2 (comparing A's model with B's

### 2015-04-01 mutual-modeling.tex
First version abstract

@@
 
-\title{Is Partner Modelling Mutual?}
+\title{Do I Need to Model You in order to Learn With You?}
 
@@
 \begin{abstract}
-    TDB
+    Collaborative learning has something to do with building a shared
+    understanding of the situation at hand. Yet, the psycholinguistics mechanisms
+    at work while establishing common grounds are complex and certainly involve
+    modelling what the partner knows/wants/intends at a given time.
+
+    Building on a set of five independent studies that where conducted over
+    several years, we propose in this article a simple model of the mechanisms
+    of mutual modelling. It integrates several concepts related to the classical
+    idea of knowledge grounding, and we apply it as an operational tool to
+    analyse the interplay of learning and mutual modelling in the five studies.
+
+    We come up with mixed experimental results that evidence the dual nature of
+    mutual modelling: a individual meta-cognitive aptitude that co-exists with
+    modelling elicited and fostered by social interactions.
+
 \end{abstract}

### 2015-04-01 mutual-modeling.tex
Fixed a few ref + style + formatting in the first paragraphs of article

@@
 students should build upon each other's understanding to refine their own
-understanding.  What is expected to produce learning is not the mere fact that
+understanding: what is expected to produce learning is not the mere fact that
 two students build the same understanding but the cognitive effort they have to
@@
 whose occurrence has been related to learning: (self-) explanations in cognitive
-science (Chi ; Webb), conflict resolution in socio-cognitive theories (Doise),
-mutual regulation (Blaye,) in a Vygostkian perspective, etc. The construction of
-a shared understanding has been investigated for several years in
-psycholinguistics, under the  notion of
-``grounding''~\citep{clark1986referring}.  However, the relevance of grounding
+science \fixme{REF:Chi} \fixme{REF: Webb}, conflict resolution in
+socio-cognitive theories \fixme{REF:Doise},
+mutual regulation \fixme{REF: Blaye} in a Vygostkian perspective, etc. The
+construction of a shared understanding has been investigated for several years
+in psycholinguistics, under the  notion of
+\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
 mechanisms for explaining learning outcomes has been questioned in learning
@@
 the understanding of how partners engage into shared meaning
-making~\cite{stahl2007meaning} or \emph{intersubjective} meaning
-making~\cite{suthers2006technology}.
+making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
+making~\citep{suthers2006technology}.
 
@@
 This view is actually not incompatible with Clark's \emph{grounding
-criterion}~\cite{clark1989contributing}: the degree of shared understanding that
+criterion}~\citep{clark1989contributing}: the degree of shared understanding that
 peers need to reach depends upon the task they perform. For instance, a dialogue
@@
 their friends but has to guarantee accurate common grounds when talking about
-which intervention will be conducted in which way on which patient.  Some CSCL
-scholars stressed the ``illusion of shared understanding'' \fixme{REF:(JACCO?)},
-\ie the potential decoupling between linguistic alignment and actual shared
-meanings.
-
-This interesting cognitive science debate mostly occurred outside the field of
+which intervention will be conducted in which way on which patient. 
+%Some CSCL
+%scholars stressed the ``illusion of shared understanding'' \fixme{REF: Jacco?},
+%\ie the potential decoupling between linguistic alignment and actual shared
+%meanings.
+
+This interesting cognitive science debate occurred mostly outside the field of
 learning. In education, the question is to relate these mechanisms to learning
-outcomes. Is linguistic alignment sufficient to trigger conceptual change? Does
-negotiation of meaning only occurs when partners monitor and diagnose each
-other's knowledge. If the ratio between shallow alignment and deep grounding
-depends upon the task, and if deep grounding is a condition for learning, then
-the pedagogical challenge is to design tasks that require deep grounding. Most
-empirical studies on grounding and alignment are conducted with tasks, which
-despite being qualified as ``ecologically valid'' by their authors, are mere
-referencing tasks such as asking the way to the train station or helping the
-peer to choose a picture among many. In this contribution, we explore several
-richer tasks such as arguing on a sensitive issue or building a concept map.
+outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
+change?} \textbf{Does negotiation of meaning only occurs when partners monitor
+and diagnose each other's knowledge?} If the ratio between shallow alignment and
+deep grounding depends upon the task, and if deep grounding is a condition for
+learning, then the pedagogical challenge is indeed to design tasks that require
+deep grounding. Most empirical studies on grounding and alignment are conducted
+with tasks, which despite being qualified as ``ecologically valid'' by their
+authors, are mere referencing tasks such as asking the way to the train station
+or helping the peer to choose a picture among many. In this contribution, we
+explore several richer tasks such as arguing on a sensitive issue or building a
+concept map.
 
@@
 effort}.  The cost of producing a perfect utterance may be higher than the cost
-of repairing the problems that may arise through misunderstandings. For
-instance, subjects are less careful about adapting their utterances to their
-partner when they know they can provide feedback on his/her
+of repairing the problems that may arise through misunderstandings, and in fact,
+subjects tend to make less efforts adapting their utterances to a specific
+partner when they know that they can later provide feedback on his/her
 understanding~\citep{schober1993spatial}. We introduced the notion of
@@
 shared understanding. For instance, CSCL researchers expanded the use of
-\emph{collaboration scripts} \fixme{REF (Aronson,?)}. A script is a pedagogical
-method that collaborative learning activities in order to foster the emergence
-of productive interactions such as argumentation, explanation or conflict.
-Conflict-resolution scripts such as the {\sc
-ArgueGraph}~\citep{dillenbourg2008mechanics} forms pair of students with
+\emph{collaboration scripts}~\citep{kobbe2007specifying}. A script is a pedagogical
+method that frames collaborative learning activities in order to foster the
+emergence of productive interactions such as argumentation, explanation or
+conflict. Conflict-resolution scripts such as the {\sc
+ArgueGraph}~\citep{dillenbourg2008mechanics} form pair of students with
 opposite opinions, which increases the difficulty of consensus building,
-requiring more justifications, more negotiation, more load. Similarly, JIGSAW
-scripts provide peers with different but complementary knowledge for augmenting
-(reasonably) the effort group members have to engage to reach a shared solution.
+requiring more justifications, more negotiation, more load. Similarly, {\sc
+jigsaw} scripts~\citep{aronson1978jigsaw} provide peers with different but
+complementary knowledge for augmenting (reasonably) the efforts that group members
+have to engage into to reach a shared solution.
 
@@
 %experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
-%people, A and B, manipulate objects. Then A leaves and B hides away one
-%object. When A comes back, we ask the child ``where do you think A will
+%people, $A$ and $B$, manipulate objects. Then $A$ leaves and $B$ hides away one
+%object. When $A$ comes back, we ask the child ``where do you think $A$ will
 %look for the object?''. Before acquiring a theory of mind, children are not
@@
 %the object was hidden) from the model of A, which contains \emph{false
-%beliefs} on the world (A still thinks the object is at its original
-%position since he did not see B hiding it).
+%beliefs} on the world ($A$ still thinks the object is at its original
+%position since he did not see $B$ hiding it).
+
+\vspace{2em}
+These preliminary remarks frame the main question that this article address:
+in order to engage into shared understanding, which is a desirable learning
+situation, do partners have to build a representation of each other's knowledge?
+if so, what exactly is the nature of this mutual modelling? We adopt hereafter a
+constructive approach, first introducing a simple yet convenient model of mutual
+modelling, then analyzing five independent studies in terms of the interactions
+between learning and mutual modelling.
 
@@
 
-In order to build a shared understanding, do partners have to build a
-representation of each other's knowledge? We refer to \emph{mutual modelling} as
-the process of inferring one's partner mental states. Any claim that students
-carry out a detailed monitoring of their peers would be as incorrect as any
-claim that they do not maintain any representation at all. If mutual modelling
-had to be permanently detailed and accurate, subjects would obviously face a
-huge cognitive load. Conversely, peers could not collaborate without some
-minimal amount of mutual modelling. For instance, A cannot disagree with B
-without knowing that B has a different opinion. The mutual model can be implicit
-(A is not aware of what he knows about B), by-default (I believe that B beliefs
-what I believe unless contrary evidence), opportunistic (A does not model B
-unless the conversation requires it), global (A infers B's beliefs based on
-categories such as age, culture or profession) and, of course, it can be
-incorrect… but it can not remain empty. Dialogues include many instances of
-utterances such as ``I thought he would do that'' (first level of mutual
-modelling) or even ``He thought I would do that but I intended something else.''
-(second level of mutual modelling).
-
-The content of mutual models ranges from 'dispositional' versus 'situational'
-aspects. The 'dispositional' aspects refer to A's representation of B's long
-term knowledge, skills or traits. It is thus closely related to the notion of
-transactive memory~\citep{wegner1987transactive, moreland1999transactive}.
-'Situational' aspects refer to A's representation of B's knowledge, behavior or
-intentions specifically activated in the situation in which A and B are
-collaborating, some of them being valid for 2 seconds, other ones for 2 hours.
-Here are examples of fragments that constitute A's model of B regarding to
-aspects X, noted $Model(A,B,X)$:
+We refer to \emph{mutual modelling} as the process of inferring one's partner
+mental states. Any claim that students carry out a detailed monitoring of their
+peers would be as incorrect as any claim that they do not maintain any
+representation at all. If mutual modelling had to be permanently detailed and
+accurate, subjects would obviously face a huge cognitive load. Conversely, peers
+could not collaborate without some minimal amount of mutual modelling. For
+instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
+opinion. The mutual model can be implicit ($A$ is not aware of what he knows about
+B), by-default (I believe that $B$ beliefs what I believe unless contrary
+evidence), opportunistic ($A$ does not model $B$ unless the conversation requires
+it), global ($A$ infers $B$'s beliefs based on categories such as age, culture or
+profession) and, of course, it can be incorrect… but it can not remain empty.
+Dialogues include many instances of utterances such as ``I thought he would do
+that'' (first level of mutual modelling) or even ``He thought I would do that
+but I intended something else.'' (second level of mutual modelling).
+
+The content of mutual models ranges from \emph{dispositional} versus
+\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
+representation of $B$'s long term knowledge, skills or traits. It is thus closely
+related to the notion of transactive memory~\citep{wegner1987transactive,
+moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
+B's knowledge, behavior or intentions specifically activated in the situation in
+which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
+ones for 2 hours.  Here are examples of fragments that constitute $A$'s model of B
+regarding to aspects X, noted $Model(A,B,X)$:
 
@@
 
-    \item $Model(A,B, knowledge)$: What does A know about B's knowledge with
-        respect to the task at hand or, inversely, about B's knowledge gaps?
-        When can A consider B's statements as reliable?
-
-    \item $Model(A,B, skills)$: What does A know about B's skills with respect to
-        the task at hand? May A expect B to perform well in a specific subtask?
+    \item $Model(A,B, knowledge)$: What does $A$ know about $B$'s knowledge with
+        respect to the task at hand or, inversely, about $B$'s knowledge gaps?
+        When can $A$ consider $B$'s statements as reliable?
+
+    \item $Model(A,B, skills)$: What does $A$ know about $B$'s skills with respect to
+        the task at hand? May $A$ expect $B$ to perform well in a specific subtask?
         The effectiveness of division of labor depends on the quality of this
@@
 
-    \item $Model(A,B, goals)$: What does A know about B's intentions with respect
-        to the project, including B' motivation and commitment? Can A trust B
-        when B promises to deliver?
-
-    \item $Model(A,B, task)$: What does A know about B's representation of the
-        situation and the task: does A knows whether B has the same
+    \item $Model(A,B, goals)$: What does $A$ know about $B$'s intentions with respect
+        to the project, including $B$' motivation and commitment? Can $A$ trust B
+        when $B$ promises to deliver?
+
+    \item $Model(A,B, task)$: What does $A$ know about $B$'s representation of the
+        situation and the task: does $A$ knows whether $B$ has the same
         understanding of the problem at stake?
 
-    \item $Model(A,B, plans)$: What does A know about B's strategy. Does A
-        understand why B did what he did? Is A able to anticipate what B will do
+    \item $Model(A,B, plans)$: What does $A$ know about $B$'s strategy. Does A
+        understand why $B$ did what he did? Is $A$ able to anticipate what $B$ will do
         next?
 
-    \item $Model(A,B, ``urgent'')$: What does A about know B's understanding of A's
+    \item $Model(A,B, ``urgent'')$: What does $A$ about know $B$'s understanding of $A$'s
         last utterance: does ‘urgent' means now, ASAP and ‘not too lat' ?
 
-    \item We could continue the list of what is X is \model{A}{B}{X}: beliefs,
+    \item We could continue the list of what is $X$ is \model{A}{B}{X}: beliefs,
         emotions, history, status, …
@@
 
-We have different levels of mutual modelling. If A states ``B thinks I am good in
-maths'', we are the second level of mutual modelling:  $Model(A, B, Model(B,A,
+We have different levels of mutual modelling. If $A$ states ``B thinks I am good in
+maths'', we are the second level of mutual modelling:  $Model(A, $B$, Model(B,A,
 knowledge=``good in math''))$. There is an infinite regress of nested models: If A
 says ``B knows that I don't expect him to solve this statistics problem''
-corresponds to $Model(A, B, Model (B, A (Model (A,B, statistic-skills)))$.
+corresponds to $Model(A, $B$, Model (B, A (Model (A,B, statistic-skills)))$.
 

### 2015-04-01 mutual-modeling.tex
{A -> $A$}

@@
 Initial common grounds are also initiated by co-presence: they include events to
-which A and B attended together~\citep{clark2002definite} in the physical space
+which $A$ and $B$ attended together~\citep{clark2002definite} in the physical space
 or in our cultural space (\eg ‘09-11'). Co-presence means that we can refer to
@@
 their misunderstanding. In CSCW, researchers have explored various so-called
-``awareness tools'', \ie functionalities that inform A is about B's actions that A
-could not directly perceived because B was working on a different subset of the
+``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions that A
+could not directly perceived because $B$ was working on a different subset of the
 virtual space. Different awareness tools will be addressed in this contribution
@@
 To report experiments on mutual modelling, we suggest
-the notation \model{A}{B}{X} to denote ``A knows that B knows X''. This is a
+the notation \model{A}{B}{X} to denote ``A knows that $B$ knows X''. This is a
 simple but useful reduction to reason on mutual modelling. This
-notation does not mean A has an explicit, monolithic representation of B: it
+notation does not mean $A$ has an explicit, monolithic representation of B: it
 must be understood as an abstraction referring to complex socio-cognitive
@@
 Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
-content X at time $t_1$ during task T in order to increase \Model{A}{B}{X}
+content $X$ at time $t_1$ during task T in order to increase \Model{A}{B}{X}
 at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
@@
     \begin{enumerate}
-        \item \model{A}{B}{emotion}: how accurately A perceives B's emotional state
+        \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
             (study 3);
 
-        \item \model{A}{B}{actions} is about how well A guesses what action B has
+        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
             done (study 2) or will do next  (study 1)
 
-        \item \model{A}{B}{knowledge}: how accurately A estimates B's knowledge
+        \item \model{A}{B}{knowledge}: how accurately $A$ estimates $B$'s knowledge
             with respect to the material they learn together (study 4 and 5)
@@
         use awareness tools which, in principle, reduce the cost of mutual modelling, but do
-        not eliminate all costs: if the tool provide A with information about
-        what B does/knows, this additional information may actually increase
+        not eliminate all costs: if the tool provide $A$ with information about
+        what $B$ does/knows, this additional information may actually increase
         cognitive load. Awareness tools constitute as a kind of mutual modelling prosthesis,
@@
         obvious methodological concerns since such a question triggers a
-        modelling process beyond what it would 'naturally' be. To avoid this
+        modelling process beyond what would naturally occur. To avoid this
         bias, one can estimate mutual modelling after task completion. Then, the
@@
 \begin{itemize}
-    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon A's ability or effort
-        to model B,
-
-    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  B's ability of
-        effort to help A to model him/herself
+    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon $A$'s ability or effort
+        to model $B$,
+
+    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability of
+        effort to help $A$ to model him/herself
 
     \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
-        interactions among A and B
+        interactions among $A$ and B
 
@@
 $\mathcal{H}_{2}$ is also called a second level modelling hypothesis since H
-needs to monitor B to see if A understood him/her
+needs to monitor $B$ to see if $A$ understood him/her
 $\mathcal{M}(B,(A,\mathcal{M}(A,B)))$
@@
 
-This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do A and
+This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$ and
 B have the same accuracy when modelling C? $\Delta_2 = \Delta(\M{A}{C}{X},
@@
 
-Conversely, does C model more accurately A than B? $\Delta_3= \Delta(\M{C}{A}{X},
+Conversely, does $C$ model more accurately $A$ than B? $\Delta_3= \Delta(\M{C}{A}{X},
 \M{C}{B}{X})$ A positive answer would support $\mathcal{H}_{3}$.
@@
 The independent variable was the availability of an awareness tool that shows to
-player A  the location and gaze direction of player B. In the awareness
+player $A$  the location and gaze direction of player $B$. In the awareness
 condition, players could switch to the \emph{scout mode} where they could view what
@@
 chose this subjective measure of accuracy ($\Delta(\M{A}{B}{x}, \M{B}{B}{x})$)
-rather than an objective measure (\ie the model \gmodel{A}{B} is compared to B's
+rather than an objective measure (\ie the model \gmodel{A}{B} is compared to $B$'s
 next action) because some the activities proposed by the questionnaire were not
@@
 problem solving strategy. We used an objective measure of \gModel{A}{B}: the
-distance between where A believes B has been walking and where B actually went.
+distance between where $A$ believes $B$ has been walking and where $B$ actually went.
 The main hypothesis concerned the effect of awareness tools on group performance
@@
 their personal distance to Bob. Their annotations on the map were shared with
-the two other players (A could see what B and C wrote). These annotations faded
+the two other players ($A$ could see what $B$ and $C$ wrote). These annotations faded
 out after a few minutes to avoid covering the full display. The awareness tool
@@
             \includegraphics[width=\linewidth]{image6.png}
-            \caption{A's drawing of B's path.}
+            \caption{A's drawing of $B$'s path.}
         \end{subfigure}
@@
             \includegraphics[width=\linewidth]{image7.png}
-            \caption{Actual path followed by B.}
+            \caption{Actual path followed by $B$.}
         \end{subfigure}
@@
 enabled us to calculate the number of errors players made while drawing the path
-of their partners. We compared the path that player A attributed to B with B's
-real path recorded by the system and the same for A \& C and B \& C as depicted
+of their partners. We compared the path that player $A$ attributed to $B$ with $B$'s
+real path recorded by the system and the same for $A$ \& $C$ and $B$ \& $C$ as depicted
 on Fig.~\ref{study2:paths}.
 
-\gModel{A}{B} is the sum of errors made by A about B's paths. An error was
+\gModel{A}{B} is the sum of errors made by $A$ about $B$'s paths. An error was
 either drawing a place where the partner had not been or not drawing a place
@@
 \gModel{B}{A}  is positive ($r = .41$) and significant ($p < .01$);  the more A
-makes errors about B, the more B does the same.
+makes errors about $B$, the more $B$ does the same.
 
@@
 variable \gModel{A}{B} was subjectively by using 3 questionnaires: in the first
-one, A described his or her own emotions \gmodel{A}{A} , while in the two other
-questionnaire, A describes B's and C's emotions. The questionnaire included 18
+one, $A$ described his or her own emotions \gmodel{A}{A} , while in the two other
+questionnaire, $A$ describes $B$'s and C's emotions. The questionnaire included 18
 items (7-point Likert Scale) describing emotions labeled as adjectives:
@@
         \includegraphics[width=0.9\textwidth]{image9.pdf}
-        \caption{Average values of \Model{A}{B}{X} where X is one of the proposed
+        \caption{Average values of \Model{A}{B}{X} where $X$ is one of the proposed
         emotions (max = 7).}
@@
 produced concept map) but the learning gains. The correlation between
-\gModel{A}{B} and A's learning gains is not significant ($\beta =.08, ,NS, N =
+\gModel{A}{B} and $A$'s learning gains is not significant ($\beta =.08, ,NS, N =
 60$). It is also not significant within each condition.
@@
 \gModel{A}{B} was not significant ($F(1, 58) = 0.78, ns$) when considering the
-absolute difference between \gModel{A}{B} and B's post-test score. However, A
-tended underestimate B's score in the SI condition ($M = -2.06$) and to
+absolute difference between \gModel{A}{B} and $B$'s post-test score. However, A
+tended underestimate $B$'s score in the SI condition ($M = -2.06$) and to
 overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<.01$).
@@
 are in conditions of symmetry of information. This consolidates the hypothesis
-of \gModel{A}{B} being a property of interactions in context rather than A's
+of \gModel{A}{B} being a property of interactions in context rather than $A$'s
 cognitive attitude.
@@
 outcomes by comparing teams with or without a knowledge awareness tool (KAT),
-\ie a tool that informs A about B's knowledge as measured through a pre-test.
+\ie a tool that informs $A$ about $B$'s knowledge as measured through a pre-test.
 
@@
 of \gModel{A}{B}: (1) as an effect of the quality of interaction, (2) as an
-effect of A's effort to model B or (3) as B's effort to give cues to A about his
+effect of $A$'s effort to model $B$ or (3) as $B$'s effort to give cues to $A$ about his
 own knowledge.
@@
 Out of 5 studies, 3 revealed a correlation between the model peers built about
-each other: when A builds an accurate model of B, B also tends to build an
+each other: when $A$ builds an accurate model of $B$, $B$ also tends to build an
 accurate model of A.  In one study, this was the case only in one condition.
@@
 effort but overall the side-effect of the quality of interactions. The
-\emph{second-order-modelling hypothesis} $\mathcal{H}_{2}$ stated that, if A is
-accurate in modelling B, A should notice when B misunderstands A and should be
+\emph{second-order-modelling hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is
+accurate in modelling $B$, $A$ should notice when $B$ misunderstands $A$ and should be
 able to repair. A good modeler would somehow drive the modelling of his or her
@@
 support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
-be no reason for modelling skills of A and B to be correlated.  The role of C in
+be no reason for modelling skills of $A$ and $B$ to be correlated.  The role of $C$ in
 \gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
@@
 important, it would lead be surprising to have by chance a correlation between
-B's and C' ability to drive A's modelling.
+B's and C' ability to drive $A$'s modelling.
 
 In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
-($\Delta_2$ in study 4 and study 5) could also contributes to the idea that A's
+($\Delta_2$ in study 4 and study 5) could also contributes to the idea that $A$'s
 modelling skills is not the key factor. However, between modelling someone and
@@
 modelling: on-task in study 1 versus off-task in study 2, subjective validation
-in study 1 versus objective validation in study 2 (comparing A's model with B's
+in study 1 versus objective validation in study 2 (comparing $A$'s model with $B$'s
 behavior).

### 2015-04-01 mutual-modeling.tex
Proof-read/fixing sections 2/3

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Mutual Modelling}
@@
 instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
-opinion. The mutual model can be implicit ($A$ is not aware of what he knows about
-B), by-default (I believe that $B$ beliefs what I believe unless contrary
-evidence), opportunistic ($A$ does not model $B$ unless the conversation requires
-it), global ($A$ infers $B$'s beliefs based on categories such as age, culture or
-profession) and, of course, it can be incorrect… but it can not remain empty.
+opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
+B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
+evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
+it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
+profession) and, of course, it can be incorrect\ldots but it can not remain empty.
 Dialogues include many instances of utterances such as ``I thought he would do
@@
 which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
-ones for 2 hours.  Here are examples of fragments that constitute $A$'s model of B
-regarding to aspects X, noted $Model(A,B,X)$:
+ones for 2 hours. Examples of fragments that constitute $A$'s model of B
+regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:
 
@@
 
-    \item $Model(A,B, knowledge)$: What does $A$ know about $B$'s knowledge with
+    \item $Model(A,B, knowledge)$: what does $A$ know about $B$'s knowledge with
         respect to the task at hand or, inversely, about $B$'s knowledge gaps?
@@
 
-    \item $Model(A,B, skills)$: What does $A$ know about $B$'s skills with respect to
+    \item $Model(A,B, skills)$: what does $A$ know about $B$'s skills with respect to
         the task at hand? May $A$ expect $B$ to perform well in a specific subtask?
@@
 
-    \item $Model(A,B, goals)$: What does $A$ know about $B$'s intentions with respect
+    \item $Model(A,B, goals)$: what does $A$ know about $B$'s intentions with respect
         to the project, including $B$' motivation and commitment? Can $A$ trust B
@@
 
-    \item $Model(A,B, task)$: What does $A$ know about $B$'s representation of the
+    \item $Model(A,B, task)$: what does $A$ know about $B$'s representation of the
         situation and the task: does $A$ knows whether $B$ has the same
@@
 
-    \item $Model(A,B, plans)$: What does $A$ know about $B$'s strategy. Does A
+    \item $Model(A,B, plans)$: what does $A$ know about $B$'s strategy. Does A
         understand why $B$ did what he did? Is $A$ able to anticipate what $B$ will do
@@
 
-    \item $Model(A,B, ``urgent'')$: What does $A$ about know $B$'s understanding of $A$'s
-        last utterance: does ‘urgent' means now, ASAP and ‘not too lat' ?
-
-    \item We could continue the list of what is $X$ is \model{A}{B}{X}: beliefs,
-        emotions, history, status, …
+    \item $Model(A,B, ``urgent'')$: what does $A$ know about $B$'s understanding of $A$'s
+        last utterance: does ``urgent'' means now, soon or ``not too late''?
 
@@
 
-We have different levels of mutual modelling. If $A$ states ``B thinks I am good in
-maths'', we are the second level of mutual modelling:  $Model(A, $B$, Model(B,A,
-knowledge=``good in math''))$. There is an infinite regress of nested models: If A
-says ``B knows that I don't expect him to solve this statistics problem''
-corresponds to $Model(A, $B$, Model (B, A (Model (A,B, statistic-skills)))$.
-
-A partner model is probably not a ``box'', \ie not a monolithical representation
+The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
+beliefs, emotions, history, status, etc.
+
+Besides, we have different levels of mutual modelling. If $A$ states ``B thinks I am good in
+maths'', $A$ builds a \emph{second level} model:
+\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
+regress of nested models: $A$
+saying ``$B$ knows that I don't expect him to solve this statistics problem''
+corresponds to \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
+
+A partner model is likely not a ``box'', \ie not a monolithic representation
 but rather a mosaic of information fragments about the partner, with various
-grain sizes and various life cycles. This mosaic is elaborated through various
-mechanisms, first for building an initial model of the partner and then for
+granularity and various life cycles. This mosaic is elaborated through a variety
+of mechanisms, first for building an initial model of the partner, then for
 updating this model. As two students meet for the first time, mutual models are
-initialized by the assumptions they make upon each other from cues such as
-his/her membership to large categories (age, culture, profession, ...) include
-stereotypes (sportsmen, junkie, business women, Swiss,...) as well as physical
+initialized by the assumptions they make upon each other based on cues such as
+his/her belonging to broad categories (age, culture, profession, \ldots),
+stereotypes (sportsmen, junkie, business women, Swiss,\ldots) as well as physical
 appearance. Scholars studied how initial modelling impacts communication. In
-their experiments on initial mutual modelling, Slugoski, Lalljee, Lamb \&
-Ginsburg~\citep{slugoski1993attribution} pretended to their subjects that their
-(fake) partner had or had not received the same information. They observed that
-the subjects adapted their dialogue by focusing the explanation on the items
-that (s)he was supposed to ignore. Brennan (1991) \fixme{ref} showed that the subjects used
-different initial strategies in forming queries depending on who they were told
-their partner was.
+their experiments on initial mutual modelling, \citet{slugoski1993attribution}
+pretended to their subjects that their (fake) partner had or had not received
+the same information. They observed that the subjects adapted their dialogue by
+focusing the explanation on the items that he/she was supposed to ignore.
+\citet{brennan1991conversation} showed that the subjects used different initial
+strategies in forming queries depending on who they were told their partner was.
 
 Initial common grounds are also initiated by co-presence: they include events to
-which $A$ and $B$ attended together~\citep{clark2002definite} in the physical space
-or in our cultural space (\eg ‘09-11'). Co-presence means that we can refer to
-a shared objects and event,-s it does of course not imply that we give them the
-same meaning. Namely, a shared screen does not mean a shared
+which $A$ and $B$ attended together~\citep{clark2002definite} in the physical
+space or in their cultural space (\eg ``09-11''). While co-presence means that
+they can refer to shared objects and events, it does not imply that they give
+them the same meaning. Namely, a shared screen does not mean a shared
 understanding~\citep{dillenbourg2006sharing}.
 
-After initialization, mutual models are updated along the collaborative work
+After initialization, mutual models are updated during the collaborative work
 through verbal and non-verbal interactions. A default inference rule is that ``my
-partner agrees with me unless he disagrees'', which reject the critiques that
+partner agrees with me unless he disagrees'', which rejects the critiques that
 mutual modelling generates an unbearable cognitive load. This default rule is
-superseded by the several mechanisms for monitoring and repairing one's partner
+superseded by the several mechanisms for monitoring and repairing the partner
 understanding: acknowledgement, continuous attention, relevance of next turns,
@@
 
-Mutual modelling does not occur in a vacuum but it is highly contextualized. REF
-reviews the features of the collaborative situation, namely the media
-(cotemporality… ), may facilitate or hamper mutual modelling. Hutchins (SILENCE)
-reported a study in which a short silence between two pilots was perfectly
-interpreted because it occurred in a highly constrained communication context.
-Some environments are more productive than others in helping peers to detect
-their misunderstandings. Roschelle~\citep{roschelle1995construction} reformulate
-the design of CSCL interfaces as providing ways for peer to detect and repairs
-their misunderstanding. In CSCW, researchers have explored various so-called
-``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions that A
-could not directly perceived because $B$ was working on a different subset of the
-virtual space. Different awareness tools will be addressed in this contribution
-since they allowed us to manipulate experimentally the mutual modelling activity.
+Mutual modelling does not occur in a vacuum but it is highly contextualized.
+REF\fxfatal{REF missing} reviews how the features of the collaborative
+situation, namely the media (co-temporality,\ldots), may facilitate or hamper
+mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
+short silence between two pilots was perfectly interpreted because it occurred
+in a highly constrained communication context.  Some environments are more
+productive than others in helping peers to detect their misunderstandings.
+\citet{roschelle1995construction} reformulate the design of CSCL
+interfaces to provide ways for peers to detect and repair their
+misunderstanding. In CSCW, researchers have explored various so-called
+``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
+that A could not directly perceive because $B$ was working on a different subset
+of the virtual space. Different awareness tools are presented and relied upon in
+this contribution as well since they allowed us to manipulate experimentally the
+mutual modelling activity.
 
@@
 %
+
 \section{A model for mutual modelling evaluation}
 
-To report experiments on mutual modelling, we suggest
-the notation \model{A}{B}{X} to denote ``A knows that $B$ knows X''. This is a
-simple but useful reduction to reason on mutual modelling. This
-notation does not mean $A$ has an explicit, monolithic representation of B: it
-must be understood as an abstraction referring to complex socio-cognitive
-processes. Besides, we refer to the \emph{degree of accuracy} of the model as
-\Model{A}{B}{X}.
+To report on experiments on mutual modelling, we suggest the notation
+\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This is a simple but
+useful reduction to reason on mutual modelling. This notation does not mean $A$
+has an explicit, monolithic representation of $B$: it must be understood as an
+abstraction referring to complex socio-cognitive processes. Besides, we refer to
+the \emph{degree of accuracy} of the model as \Model{A}{B}{X}.
 
@@
     \item Tasks vary a lot with respect to how much they require mutual
-        understanding.  The grounding
-        criterion -- denoted \groundingcriterion -- refers to how
-        important it is to mutually share a piece of information $X$ to succeed
-        the task $T$. It can be computed as the probability to succeed T despite
-        the fact $X$ is not grounded. $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be
-        estimated from the correlation between \Model{A}{B}{X} and the task
-        performance.
-
-    \item Before any specific grounding action, there is usually a non-null probability
-        that $X$ is mutually understood by $A$ and $B$ (\eg $X$
-        is part of $A$'s and $B$'s cultures, it is manifest to co-present
-        subjects or simply there is not much space for misunderstanding
-        or disagreement about $X$). We simply could not collaborate without a high level of initial
-        grounds. We note the theoretical accuracy of \emph{initial grounds}
-        $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
-
-    \item The cost of grounding $X$ refers to the physical and cognitive effort
-        required to perform a grounding act $\alpha$: a verbal repair (\eg
-        rephrasing), a deictic gesture, a physical move to adopt one partner's
-        viewpoint, etc. This cost varies according to media
-        features~\cite{clark1991grounding}.
+        understanding.  The \emph{grounding criterion} -- denoted
+        \groundingcriterion -- refers to how important it is to mutually share a
+        piece of information $X$ to succeed the task $T$. It can be calculated as
+        the probability to succeed $T$ despite the fact $X$ is not grounded.
+        $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be estimated from the correlation
+        between \Model{A}{B}{X} and the task performance.
+
+    \item Before any specific grounding action, there is generally a non-null
+        probability that $X$ is mutually understood by $A$ and $B$ (\eg $X$ is
+        part of $A$'s and $B$'s cultures, it is manifest to co-present subjects
+        or simply there is not much space for misunderstanding or disagreement
+        about $X$). We simply could not collaborate without a certain level of
+        initial grounds. We note the theoretical accuracy of \emph{initial
+        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
+
+    \item The \emph{cost of grounding} $X$ refers to the physical and cognitive
+        effort required to perform a grounding act $\alpha$: a verbal repair
+        (\eg rephrasing), a deictic gesture, a physical move to adopt one
+        partner's viewpoint, etc. This cost varies according to media
+        features~\citep{clark1991grounding}.
 
@@
 \begin{multline} \label{eq:probrepair}
-p(\alpha_{t_1}(X,T) / M^{\circ}_{t_2}(A,B,X) \nearrow) \approx \\
-\frac{(M^{\circ}_{min}(A,B,X,T) - M^{\circ}_{t_1}(A,B,X))}{cost (\alpha)}
+    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \propto \\
+    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) - \mathcal{M}^{\circ}_{t_1}(A,B,X)}{cost (\alpha)}
 \end{multline}
@@
 This is qualitative summary more that a real equation since several parameters
-are hard to quantify (\eg the cost an communication acts depends upon the user
-as well) but it clarifies the parameters of the experiments we propose hereafter.
+are hard to quantify (\eg the cost of a communication act depends upon the user
+as well) but it clarifies the parameters of the experiments we present hereafter.
 
 \begin{itemize}
-    \item \Model{A}{B}{X}: Our experiments address different contents that can be
+    \item \Model{A}{B}{X}: our experiments address different contents that can be
         represented in mutual models:
@@
         \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
-            (study 3);
+            (study 3),
 
         \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
-            done (study 2) or will do next  (study 1)
+            performed (study 2) or will perform next (study 1),
 
         \item \model{A}{B}{knowledge}: how accurately $A$ estimates $B$'s knowledge
-            with respect to the material they learn together (study 4 and 5)
+            with respect to the material they learn together (study 4 and 5).
 
@@
 
-    \item \Model{A,B}{X}{T} Our studies concern various collaborative tasks:
-        argumentation (study 3), games (study 1 and 2) and concept mapping
-        (study 4 and 5). By varying the tasks, we do actually vary the grounding
-        criterion but they all require a ``reasonably high'' grounding criterion.
-        All tasks are about building a solution or a representation together; we
-        did not address everyday conversations.
-
-    \item $\mathcal{M}^{\circ}_{t1}(A,B,X)$ Along the same reasoning, the initial degree
+    \item $\mathcal{M}^{\circ}_{min}(A,B,X,T)$: our studies build upon various
+        collaborative tasks: argumentation (study 3), games (study 1 and 2) and
+        concept mapping (study 4 and 5). By varying the tasks, we do actually
+        vary the grounding criterion but they all require a ``reasonably high''
+        grounding criterion.  All tasks are about building a solution or a
+        representation together; we did not address everyday conversations.
+
+    \item $\mathcal{M}^{\circ}_{t1}(A,B,X)$: along the same reasoning, the initial degree
         of common grounds should be rather low (and hence the difference between
@@
         with teams of students who did not know each other. They came
-        nonetheless from the same university: they hence had some general common
-        grounds.  For studies 2 and 3, students knew each other before for
+        nonetheless from the same university (and they hence had some general common
+        grounds).  For studies 2 and 3, students knew each other before for
         reasons explained later on. In study 5, we manipulated the initial mutual modelling by
-        using a JIGSAW script.
-
-    \item $cost(\alpha)$ In all studies but study 4, the cost of  grounding is
+        using a {\sc jigsaw} script.
+
+    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
         an independent variable. Study 3 uses media richness as independent
@@
         what $B$ does/knows, this additional information may actually increase
-        cognitive load. Awareness tools constitute as a kind of mutual modelling prosthesis,
-        and, like any prosthesis, they many augment mutual modelling (by facilitating it or
+        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
+        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
         even scaffolding it) or inhibit it (by making it useless).
@@
 
-Each of the studies that we present in the next section have been design in such
-a way that, while tractable on a robot, mutual modelling plays a key role for
-the task performance: the tasks require a ``reasonably high'' grounding
-criterion \groundingcriterion, and all tasks are about building a solution or a
-representation together.
-
-Along the same reasoning, the initial degree of common grounds \inigrounding
-should be rather low (and hence the difference between initial and required
-degrees rather high) in order to make mutual modelling effort more observable.
-While we can exactly control the actual initial knowledge available to the
-robot, estimating \inigrounding still requires special care during the design of
-the experiments: as we have previously shown~\citep{lemaignan2014dynamics,
-lemaignan2014cognitive} and because the interaction takes place between a human and
-a robot, the initial human model of the robot can widely vary depending on
-several factors like the user familiarity with technology, the effect of
-novelty, the user expectations or the anthropomorphic features of the robot.
-
-More generally, while we have explicit access to \model{R}{H}{X} (from the facts
-stored in the robot belief state, as illustrated in
-Table~\ref{table|beliefsfig7}), estimating \Model{R}{H}{X} also requires
-\model{H}{H}{X}, which depends on the meta-cognitive competencies of the human
-partner.
-
-Conversely, measuring \Model{A}{B}{X} is methodologically difficult. We
+Measuring \Model{A}{B}{X} is methodologically difficult. We
 discriminate two steps, first to capture \model{A}{B}{X} and then to
@@
 \begin{itemize}
-    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what (s)he
-        believes about what $B$ knows, feels, intends to do, etc. This raises
-        obvious methodological concerns since such a question triggers a
-        modelling process beyond what would naturally occur. To avoid this
-        bias, one can estimate mutual modelling after task completion. Then, the
-        obvious drawback are memory losses and by post-hoc reconstruction.  The
-        first method was used in study 2 and the second one in the other
-        studies. Another option would be to rely on external behavioural metrics
-        like eye-tracking: we hypothesise for instance that the fixation time
-        reflects the efforts engaged by the human to understand, hence, model,
-        the others. Such an approach has however not yet been investigated in
-        the presented studies.
+    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$
+        what he/she believes about what $B$ knows, feels, intends to do, etc.
+        This raises obvious methodological concerns since such a question
+        triggers a modelling process beyond what would naturally occur. To avoid
+        this bias, one can estimate mutual modelling after task completion.
+        Then, the obvious drawback are memory losses and by post-hoc
+        reconstruction.  The first method was used in study 2 and the second one
+        in the other studies. Another option would be to rely on external
+        behavioural metrics like eye-tracking: we hypothesise for instance that
+        the fixation time reflects the efforts engaged by the human to
+        understand, hence, model, the others. Such an approach has however not
+        yet been investigated in the presented studies.
 
@@
     \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability of
-        effort to help $A$ to model him/herself
+        effort to help $A$ to model him/herself,
 
     \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
-        interactions among $A$ and B
+        interactions among $A$ and $B$.
 
@@
 
-$\mathcal{H}_{2}$ is also called a second level modelling hypothesis since H
-needs to monitor $B$ to see if $A$ understood him/her
-$\mathcal{M}(B,(A,\mathcal{M}(A,B)))$
+$\mathcal{H}_{2}$ is also called a second level modelling hypothesis since $B$
+needs to monitor $A$ to see if $A$ understood him/her
+($\mathcal{M}(B,A,\mathcal{M}(A,B))$).
 

### 2015-04-02 mutual-modeling.tex
Done with proof-reading of sections 1, 2 and 3

@@
 
-Our main issue is the symmetry questions: what is the relationship
-(Fig.~\ref{mm_symmetry}) between \Model{A}{B}{X} and \Model{B}{A}{X}? A low
-symmetry would mean that mutual modelling is mainly an individual
-attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might support
-$\mathcal{H}_{2}$ since there is a low probability that randomly formed pairs
-integrate peers with the same level of mutual modelling skills.
+One interesting question relates to the symmetry of mutual modelling: what is
+the relationship between \Model{A}{B}{X} and
+\Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is mainly an
+individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might
+support $\mathcal{H}_{2}$ since there is a low probability that randomly formed
+pairs integrate peers with the same level of mutual modelling skills.
 
@@
 B have the same accuracy when modelling C? $\Delta_2 = \Delta(\M{A}{C}{X},
-\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gain
+\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and
+$\mathcal{H}_{3}$\fxfatal{why H3?} gain
 over $\mathcal{H}_{1}$
@@
 Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
-question if modelling skills depend upon what aspects are being modeled ($X$ or $Y$),
-which would explain vertical differences ($\Delta_5$ in
-Fig.~\ref{mm_rectangle}). We do not however further develop these questions because we
-don't have the necessary data in our studies.
+question if modelling skills depend upon what aspects are being modeled ($X$ or
+$Y$), which would explain vertical differences ($\Delta_5$ in
+Fig.~\ref{mm_rectangle}). We do not however further explore these questions
+since we do not have the necessary data in our studies.
 
@@
 game}
-
-\subsubsection*{Questions}
 
@@
 \paragraph{Symmetry question} We computed intra-class correlation as described
-by~\citep{kenny1998data} from the answers to the cross-questionnaires.
+by~\citet{kenny1998data} from the answers to the cross-questionnaires.
 Considering all pairs in both conditions, we found a positive and significant
@@
 
-\subsubsection*{Questions}
-
 The aim of this unpublished study was to evaluate the effect of media richness
@@
 
-\subsubsection*{Questions}
-
 This study investigated the effect of a collaboration script on collaborative
-learning~\citep{molinari2008effects}. The script chosen is a JIGSAW: two students
+learning~\citep{molinari2008effects}. The script chosen is a {\sc jigsaw}: two students
 receive different but complementary subsets of the knowledge (texts) which have
@@
 
-\subsubsection*{Questions}
-
 This study investigates if \Model{A}{B}{knowledge} is related to learning
@@
 The peers were located into two different rooms. A complete description of the
-study is provided in~\citep{sangin2008learners}. The experiment lasted 90 minutes.
+study is provided in~\citet{sangin2008learners}. The experiment lasted 90 minutes.
 

### 2015-04-02 mutual-modeling.tex
Work on studies section -- WIP

@@
 \maketitle
+
+\listoffixmes
 
@@
 
-We could go further by comparing self- versus other modelling ($\Delta_4$ in
+We can go further by comparing self- versus other modelling ($\Delta_4$ in
 Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
@@
 $Y$), which would explain vertical differences ($\Delta_5$ in
-Fig.~\ref{mm_rectangle}). We do not however further explore these questions
-since we do not have the necessary data in our studies.
+Fig.~\ref{mm_rectangle}). When possible, we will also discuss these two
+questions in the studies presented hereafter.
 
@@
 other's action plans. During the game, we asked them to anticipate the next
-action of their partner as well as to announce their own action. It is clear
-asking them a question about their partner increases their attention for the
-rest of the game.
+action of their partner as well as to announce their own actions.
 
@@
 
-SpaceMiners is a 3D game that involves two players collecting minerals located
-in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they shoot drones through
+\begin{figure*}
+        \centering
+        \includegraphics[width=0.9\textwidth]{image4.png}
+        \caption{Screenshot of the {\sc SpaceMiners} game}
+        \label{study1:spaceminer}
+\end{figure*}
+
+{\sc SpaceMiners} is a 3D game that involves two players collecting minerals located
+in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch drones through
 the space after choosing their initial direction and speed. Once launched, the
@@
 Thirty-six persons participated in this study, all native French speakers. We
-constituted 18 pairs of participants (N = 18) who were not familiar with each
-other. The pairs were assigned randomly to either the control condition (without
-the awareness tool) or the awareness condition (with the awareness tool).
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=0.9\textwidth]{image4.png}
-        \caption{Snapshot of the SpaceMiner Game}
-        \label{study1:spaceminer}
-\end{figure*}
+constituted 18 dyads who did not know each other beforehand. The pairs were
+randomly assigned to either the control condition (without the awareness tool)
+or the awareness condition (with the awareness tool).
 
@@
 
-Task performance was measured by the score reached by the two subjects after
-three situations. The effort of mutual modelling was measured as the ratio of
-time that players would spend in the scout mode (divided by total time), which
-is the time during which players are not performing their own actions but
-monitoring their partner's actions.
+Task performance was measured by the score reached by the two subjects at the
+end of the game (three levels). The effort of mutual modelling was measured as
+the ratio of time that players would spend in the scout mode (divided by total
+time), which is the time during which players are not performing their own
+actions but monitoring their partner's actions.
 
@@
 while others were reversed (``to guide my partner'' versus ``to guide me''). We
-chose this subjective measure of accuracy ($\Delta(\M{A}{B}{x}, \M{B}{B}{x})$)
+chose a subjective measure of accuracy ($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$)
 rather than an objective measure (\ie the model \gmodel{A}{B} is compared to $B$'s
-next action) because some the activities proposed by the questionnaire were not
-observable by the environment (\eg establishing a strategy). We counted
+next action) because some of the activities proposed by the questionnaire were not
+observable by the environment (\eg establishing a strategy). We calculated
 \Model{A}{B}{activity_{game_{i}}}) as the number of common answers between
-questionnaires \gmodel{A}{B} and \gmodel{B}{B} in each game and computed the
-average value a across the 3 games.
+questionnaires \gmodel{A}{A} and \gmodel{A}{B} in each game and computed the
+average value a across the 3 levels.
 
@@
 constitution of two groups made up of 12 individuals ``short time in scout
-mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
+mode'' and 24 individuals ``long time in scout mode''\fxfatal{Inconsistent! 36
+participants in awareness condition??}. A two-way analysis of
 variance conducted on these contrasted groups revealed that pairs in the
@@
 communicated more.
-
-\paragraph{Triangle questions} We cannot answer the triangle questions since
-this study was conducted with dyads.
 
@@
 space~\citep{nova2006underwhelming}. We studied whether players build an accurate
-model of the path followed by their partners because this path reflected their
+model of the path followed by their partners, assuming this path reflected their
 problem solving strategy. We used an objective measure of \gModel{A}{B}: the
@@
 joint task. The game was played on EPFL campus. Participants had to find a
-virtual object (\emph{Bob}) and ``to catch it'' by forming a triangle around it.
+virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
 Players used a Tablet PC that displayed a map of the campus and an indication of
 their personal distance to Bob. Their annotations on the map were shared with
-the two other players ($A$ could see what $B$ and $C$ wrote). These annotations faded
-out after a few minutes to avoid covering the full display. The awareness tool
-displayed as well the location of the two other players on the map. The
-availability of location awareness tool constituted our independent variable.
+the two other players ($A$ could see what $B$ and $C$ wrote). These annotations
+faded out after a few minutes to avoid covering the full display. The awareness
+tool displayed as well the location of the two other players on the map. This
+resulted in three conditions: the control condition (without tool) and two
+experimental conditions: synchronous awareness (display of the current position
+of each player) and asynchronous awareness (display of current position of each
+player as well as their recent spatial trace).
 
@@
 
-Ninety students participated in this experiment. We chosen only students from
-our university because knowing the campus geography had an impact both on group
-performance but also on mutual modelling: it is hard to represent the path of
-someone across some space without having a mental map of this space. We selected
-groups of students who knew each other. We assigned 10 triads to each of our
-three experimental conditions. Each condition was made up of approximately 25\%
-of women, but we did not control gender repartition within each triad.
+Ninety EPFL students participated in this experiment. We only selected EPFL
+students since knowledge of the campus geography had an impact both on group
+performance and on mutual modelling: to represent the path of someone across
+some space is difficult without an \textit{a priori} mental map of this space.
+We formed groups of students who knew each other. We assigned 10 triads to each
+of our three experimental conditions. Each condition was made up of
+approximately 25\% of women, but we did not control gender repartition within
+each triad.
 
@@
 
-The independent variable was the awareness tool: the control condition (without
-tool) and two experimental conditions: synchronous awareness (display current
-position of each player) and asynchronous awareness (display current position of
-each player and their recent spatial trace).  As a dependent variable, we had
-the task performance which was the distance covered by the team to catch Bob and
-\gModel{A}{B}. To estimate \gModel{A}{B}, we asked players to draw on paper
-their own path and the path of each of their partners after the game. This
-enabled us to calculate the number of errors players made while drawing the path
-of their partners. We compared the path that player $A$ attributed to $B$ with $B$'s
-real path recorded by the system and the same for $A$ \& $C$ and $B$ \& $C$ as depicted
-on Fig.~\ref{study2:paths}.
+The independent variable was the presence and role of the awareness tool. As a
+dependent variable, we had the task performance which was the distance covered
+by the team to catch Bob and \gModel{A}{B}. To estimate \gModel{A}{B}, we asked
+players to draw on paper their own path and the path of each of their partners
+after the game. This enabled us to calculate the number of errors players made
+while drawing the path of their partners. We compared the path that player $A$
+attributed to $B$ with $B$'s real path recorded by the system and the same for
+$A$ \& $C$ and $B$ \& $C$ as depicted on Fig.~\ref{study2:paths}.
 
@@
 \paragraph{Study-specific questions} There was no significant difference
-regarding the task performance. However, our surprise was that the absence of
-the awareness tool was related to a higher \gModel{A}{B}: players better
-remembered their partners' path if they did not see their position permanently!
-We will not enter into the details of these results but, to be short, teams
-without awareness tool made more annotations on the map while permanent MLA has
-an underwhelming effect~\citep{nova2005location}.
+regarding the task performance. However, and surprisingly, the absence of the
+awareness tool was related to a higher \gModel{A}{B}: players tended to better
+remember their partners' paths when they could not constantly monitor their
+positions. As detailed in the original study~\citep{nova2005location}, it
+appeared that teams without awareness tool made more manual annotations on the
+map while permanent monitoring has an underwhelming effect.
 
 \paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
-\gModel{B}{A}  is positive ($r = .41$) and significant ($p < .01$);  the more A
-makes errors about $B$, the more $B$ does the same.
+\gModel{B}{A}  is positive ($r = .41$) and significant ($p < .01$): the more $A$
+makes errors about $B$, the more $B$ does as well.
 
@@
 $\Delta_3$, the correlation between \gModel{A}{C} and \gModel{B}{C} is
-significant as well: $r=0.43, p <.001$.
+significant as well: $r=0.43, p <.001$\fxfatal{Attention! $\Delta_2$ and
+$\Delta_3$ seem to be inverted! Cf fig 2}.
 
@@
 
-This positive result regarding to the symmetry question is even more surprising
-than the previous one: despite the high heterogeneity of spatial skills among
-adults~\citep{liben1981spatial}, a significant correlation could indicate that
-mutual modelling relies more on group processes than on individual processes. The
-significant results on the triangle questions probably go in the same direction:
-either the modeler or the modeled both contribute to accurate modelling, or that
-the quality of teamwork has a global effect. In this experiment, since team
-members did not interact intensively during the task, the factor that
-facilitates accurate mutual modelling is probably less the quality of verbal
-interactions that the group strategy: a clear strategy facilitates memorizing
-one's partner path.
+This positive correlation observed in the symmetry question is surprising as
+well: when acknowledging the high heterogeneity of spatial skills among
+adults~\citep{liben1981spatial}, a significant positive correlation would
+indicate that mutual modelling indeed stems from group processes more than
+individual processes. The significant results on the triangle questions goes in
+the same direction: either both the modeler and the modellee contribute to
+accurate modelling, or the quality of teamwork has a global effect. Since team
+members did not have many opportunities to directly interact during this
+experiment, the factor that facilitates accurate mutual modelling is probably
+more the overall group strategy than the quality of verbal interactions: a clear
+strategy facilitates memorizing one's partner path.
 
@@
         \centering
-        \includegraphics[width=0.9\textwidth]{image8.jpg}
+        \includegraphics[width=0.8\textwidth]{image8.jpg}
         \caption{Example of argumentation graph}
@@
 by homosexual couples. They worked on-line and had to structure their
-argumentation with a shared concept map tool {\sc TeamWave} as illustrated in
+argumentation with the shared concept map tool {\sc TeamWave} as illustrated in
 Fig.~\ref{study3:argumentation_graph}. Ten groups had only an audio connection
-while ten groups had audio plus video. The video communication was provided by a
-web cam and the software {\sc IVisit}. For the audio link, we used microphones,
-headsets and a software called {\sc BattleCom}. In the audio+video condition,
-the screen was divided in three sections. The main part was devoted to the
-concept map window (24x24cm). In the remaining part appeared the image of the
-two peers (8.5x8.5cm each). In the audio condition, this video zone was left
-empty so that the size of the concept map was equal in each condition. The
-subjects were located in the same room, separated by mobile walls. Despite their
-headsets, it is not impossible that they could hear non-verbal audio cues (\eg
-tapping the sole with feet).  The task last in average 61 minutes.
+while ten groups had audio and video. The video communication was provided by a
+webcam and the software {\sc IVisit}. For the audio link, we used microphones,
+headsets and the {\sc BattleCom} software. In the \emph{audio plus video}
+condition, the screen was divided in three sections. The main part was devoted
+to the concept map window, and the images of the two peers appeared next to it.
+In the audio condition, this video zone was left empty so that the size of the
+concept map was equal in each condition. The subjects were located in the same
+room, separated by mobile walls. Despite their headsets, non-verbal audio cues
+(\eg tapping the floor with feet) were possibly heard by the participants. The
+task lasted in average 61 minutes.
 
@@
 
-Twenty triads from the University of Geneva participated to this experiment. We
-formed groups of subjects who knew each other, \ie they had \emph{ab initio}
-mutual models: the task required the discussion of sensitive issues which
-required to feel quite comfortable with peers. There were 36 women and 24 men,
-but, since groups were formed a priori, we could not control the balance of men
-and women in each condition.
+Sixty students (twenty triads) from the University of Geneva participated to
+this experiment (36 women and 24 men). We formed groups of subjects who knew
+each other: the task required the discussion of sensitive issues which required
+to feel quite comfortable with peers. Since groups were formed \textit{a
+priori}, we did not balance gender in each condition.
 
@@
 
-The independent variable was the presence of not of video link. The dependent
-variable \gModel{A}{B} was subjectively by using 3 questionnaires: in the first
-one, $A$ described his or her own emotions \gmodel{A}{A} , while in the two other
-questionnaire, $A$ describes $B$'s and C's emotions. The questionnaire included 18
+The independent variable was the presence or not of video link. The dependent
+variable \gModel{A}{B} was measured subjectively from three questionnaires: in the first
+one, $A$ described his/her own emotions \gmodel{A}{A}, while in the two other
+questionnaire, $A$ described $B$'s and C's emotions. The questionnaire included 18
 items (7-point Likert Scale) describing emotions labeled as adjectives:
@@
 \emph{upset}, \emph{relaxed}, \emph{irritated}, \emph{determined},
-\emph{hostile}, \emph{active}, etc. This list of emotions was
-elaborated from \fixme{REF?}. The proposed emotions were \fixme{?}. \gmodel{A}{B} is a vector
-of 18 numerical values ${x_1 \dots x_i}$ corresponding to their answers on each
-questionnaire. \gModel{A}{B} has been computed as the correlation between the
-two vectors (Eq.~\ref{eq:study3}).
-
-\begin{multline} \label{eq:study3}
-    \Mdeg{A}{B}{emotions}\\
-    = \Delta(\M{A}{B}{\{x_1 \dots x_i\}}, \M{B}{B}{\{y_1 \dots y_i\}}) \\ 
-    = r(\{x_1 \dots x_i\}, \{y_1 \dots y_i\})
-\end{multline}
+\emph{hostile}, \emph{active}, etc. \gmodel{A}{B} was modelled as a vector
+of 18 numerical values corresponding to their answers on each
+questionnaire items, and \gModel{A}{B} was computed as the correlation between the
+two vectors \gmodel{A}{B} and \gmodel{B}{B}.
 
 \subsubsection*{Results}
+
+\paragraph{Grounding criterion} The maps produced by teams were
+ranked\fxfatal{How are they ranked?}. We used
+the average rank as a estimation of the team performance and correlated it with the
+average of the 6 values of \gModel{A}{B} per team (\gModel{A}{B}, \gModel{A}{C},
+\gModel{B}{A},...): The correlation is -0.22 \fxfatal{p value?}: teams with a good \gModel{A}{B}
+team to be better ranked.
+
+\paragraph{Study-specific questions} Our hypothesis about media richness is
+rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
+audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
+0.065$). Our interpretation is that viewing what one's partner sees (shared
+graphical editor) was, in this task, more important than seeing each
+other~\citep{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
+Anderson et al., 2000}.
+
+\paragraph{Symmetry question} Regarding the main items, the mutual modelling of
+emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}
+
+\paragraph{Triangle questions} \fixme{Check ($\Delta_2$)} For $\Delta_3$, if we
+compute for each subject A, the correlation between the vectors of emotions in
+\gmodel{B}{A} and the vector in \gmodel{C}{A}, and take the average across all
+60 subjects, we obtain a high correlation of .66. This average is
+about the same in the two conditions.
 
@@
 
-\paragraph{Grounding criterion} The maps produced by teams were ranked. We use
-the average rank as a estimation of the team performance and correlated with the
-average of the 6 values of \gModel{A}{B} per team (\gModel{A}{B}, \gModel{A}{C},
-\gModel{B}{A},...): The correlation is -0.22: teams with a good \gModel{A}{B}
-team to be better ranked.
-
-\paragraph{Study-specific questions} Our hypothesis about media richness is
-rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
-audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
-0.065$). Our interpretation is that viewing what one's partner sees (shared
-graphical editor) was, in this task, more important than seeing each
-other~\citep{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
-Anderson et al., 2000}.
-
-\paragraph{Symmetry question} Regarding the main items, the mutual modelling of
-emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}
-
-\paragraph{Triangle questions} \fixme{Check ($\Delta_2$)} For $\Delta_3$, if we
-compute for each subject A, the correlation between the vectors of emotions in
-\gmodel{B}{A} and the vector in \gmodel{C}{A}, and take the average across all
-60 subjects, we obtain a very impressive correlation of .66. This average is
-about the same in each condition.
-
 \paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
@@
 according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
-errors: the difference between \gmodel{A}{B} and \gmodel{B}{A}, on a scale of 7,
+errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
 is in average of 0.3 for the emotion \emph{discasted}, and up to 1.9 for the
@@
 0-5$ for \emph{discated} versus $SD=.7$ for \emph{calm}), our point is not to
-interpret this too far, but to show that there are large varations even with one
-area  (perceiving emotions).
+interpret this too far, but to show that there are large variations even with one
+area  (perceiving emotions)\fxfatal{Paragraph to rewrite}.
 
@@
 \fixme{TBD}
-\fixme{Chek si il faut modifier les resultats de la symmetries}
+\fixme{Check si il faut modifier les resultats de la symmetries}
 
@@
 to be integrated to build a shared concept map.  This script increases the
-cognitive effort to build the map, not only to recon ciliate the viewpoints of
+cognitive effort to build the map, not only to conciliate the viewpoints of
 each team members but, before that, to find out what the other knows.
@@
 neurophysiologic phenomenon of \emph{action potential}. The text was divided
-into 3 chapters. The peers were located in two rooms equipped with the same
+into 3 chapters.  In the \emph{same information} (SI) condition, the same text
+was given to both partners. In the \emph{complementary information} (CI)
+condition, it was divided into two sub-texts, one about the electrical processes
+of the neuron while the second one about the chemical processes. These two
+versions were equivalent in terms of number of information pieces.
+
+The peers were located in two rooms equipped with the same
 computer.  The experimental session lasted around 90 minutes and consisted of 6
-phases: Participants used two software components, {\sc
-CmapTools}\footnote{\url{http://cmap.ihmc.us/}} and {\sc
-TeamSpeak}\footnote{\url{http://www.goteamspeak.com/}}.
+phases: Participants used two software components, {\sc CmapTools} and {\sc
+TeamSpeak}.
 
@@
     \item As a pre-test, participants were asked to write down all they knew
-        about the neuron and its functioning (5 minutes).
-
-    \item Participants were instructed to read a text (12 minutes).
+        about the neuron and its functioning (5 minutes),
+
+    \item Participants were instructed to read a text (12 minutes),
 
     \item Participants were asked to build individually a concept map in order to
-        graphically represent what they learnt from the text (10 minutes).
+        graphically represent what they learnt from the text (10 minutes),
 
     \item Dyads had to built a concept map during 20 minutes, communicating by
-        audio.  The screen layout was structured into three areas (see Figure 8)
+        audio.  The screen layout was structured into three areas
+        (Fig.~\ref{study4:concept_map}),
 
     \item Participants were invited to individually complete a knowledge test
-        (15-20 minutes).
+        (15-20 minutes),
 
@@
 
-Fifty-eight first year students from EPFL, 47 men and 11 women, with a mean age
-of 20.46 were remunerated for participation. Dyads were randomly assigned to one
-of the two experimental conditions. The men/women ratio was equivalent in each
-condition. Participants did not know each other before the experiment. Students
-from Life Sciences were not recruited since they had too high background
-knowledge on the learning domain.
+Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
+were remunerated for participation. Dyads were randomly assigned to one of the
+two experimental conditions. Gender was balanced over the conditions.
+Participants did not know each other before the experiment. Students from Life
+Sciences faculty were not recruited to avoid high initial background knowledge on the
+learning domain.
 
@@
 The independent variable, script versus no-script, was implemented by the
-difference of texts that individuals had to read. In the \emph{same information}
-(SI) condition, the same text was given to both partners. In the
-\emph{complementary information} (CI) condition, it was divided into two
-sub-texts, one about the electrical processes of the neuron while the second one
-about the chemical processes. These two versions were equivalent in terms of
-number of information pieces.
-
-The dependent variables were the post-test scores, but this paper is mostly
-concerned by the \Model{A}{B}{knowledge}. In phase 6, participants were ask to
-estimate (7-point Likert scale) their own and their partner's outcome knowledge
-with respect to each chapter of the \fixme{zexz}. The order of questions about
-oneself and about the other was counterbalanced across participants. We measure
-\gModel{A}{B} in an objective (Eq.~\ref{eq:study4.1} and \ref{eq:study4.2}) as
-well as in a subjective way (Eq.~\ref{eq:study4.3}).\fixme{check  subjective
-assess: 2nd equation?}
+difference of texts that individuals had to read.  The dependent variables were
+the post-test scores, used to assess \Model{A}{B}{knowledge}. In phase 6,
+participants were ask to estimate (7-point Likert scale) their own and their
+partner's outcome knowledge with respect to each chapter of the learning
+material. The order of questions about oneself and about the other was
+counterbalanced across participants. We measure \gModel{A}{B} in an objective
+(Eq.~\ref{eq:study4.1} and \ref{eq:study4.2}) as well as in a subjective way
+(Eq.~\ref{eq:study4.3}).
 
@@
 
-\paragraph{Grounding criterion} In this task, the grounding criterion was low,
-but we did not evaluate task performance (\eg the quality of the jointly
-produced concept map) but the learning gains. The correlation between
-\gModel{A}{B} and $A$'s learning gains is not significant ($\beta =.08, ,NS, N =
-60$). It is also not significant within each condition.
+\paragraph{Grounding criterion} In this task, the grounding criterion was low.
+We did not however evaluate task performance (\eg the quality of the jointly
+produced concept map) but learning gains. The correlation between \gModel{A}{B}
+and $A$'s learning gains is not significant ($\beta =.08, ns, N = 60$). It is
+also not significant within each condition.
 
@@
 test on post-test scores for electric questions and a one-way ANOVA on scores
-for ionic questions (Levene tests for homogeneity of variances: $p = .02$ and
+for ionic questions (Levene tests for homogeneity of variances): $p = .02$ and
 $ns$, respectively. Results  did not show any significant difference between the
@@
 
-\paragraph{Symmetry question} The interclass correlation between \gModel{A}{B}
-and \Model{B}{A} is not significant, but almost ($r=.26, F(1,29)=1.71, p=
-0.075$).  Actually, it is significant when students read the same text (SI
+\paragraph{Symmetry question} The inter-class correlation between \gModel{A}{B}
+and \gModel{B}{A} is approaching significance ($r=.26, F(1,29)=1.71, p=
+0.075$). It is indeed significant when students read the same text (SI
 condition: $r=0.43, F(1,15)=2.53, p<.05$) but not when they read different texts
-($r=.13, F(1,12)=1.3, NS$).
-
-\paragraph{Triangle questions} Triangle questions do not apply here since we
-worked with dyads.
-
-\paragraph{Rectangle question}$\Delta_4$: the correlation between \gmodel{A}{A}
-and \gmodel{A}{B} is globally not significant ($r=0.05$) as well as not
+($r=.13, F(1,12)=1.3, ns$).
+
+\paragraph{Rectangle question} The correlation between \gmodel{A}{A}
+and \gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
 significant in each condition: someone good at self-modelling is not necessarily
@@
 of \gModel{A}{B} being a property of interactions in context rather than $A$'s
-cognitive attitude.
+cognitive attitude\fxfatal{to be better justified}.
 
@@
 
-It started with the same two steps as in study 4, followed by:
+It started with the same two first steps as in study 4, followed by:
 
@@
 
-Sixty-four first year EPFL students (18 women, mean age = 21.2 years)
-participated to the study. They were involved in curricula that do not imply
-advanced neurophysiologic notions. They were randomly assigned to conditions and
-did not know each other before.
+Sixty-four first year EPFL students (46 men, 18 women, mean age: 21.2)
+participated to the study. They were randomly assigned to conditions and
+did not know each other before. Like in Study 4, students from Life Science
+faculty were excluded.
 
@@
 \paragraph{Grounding criterion} A linear regression revealed a positive relation
-between \gModel{A}{B} and the collaborative learning gain of the pair $\{A,
+between \gModel{A}{B} and the collaborative learning gain\fxfatal{What is the
+collaborative learning gain} of the pair $\{A,
 B\}$: $\beta= .401, p < .001, r2adj. =.15$, large effect. This relationship was
@@
 group [$M = 3.6\%, t(1, 60) = 2.73, p < .01$, Cohen's $d = 0.7$, medium to large
-effect]. Providing learners with cues about the prior-knowledge of their partner
+effect]: providing learners with cues about the prior-knowledge of their partner
 enhances their collaborative learning. As a treatment check, we found a positive
@@
 analysis revealed that subjects look at KAT to assess their peer's credibility
-when (s)he provided new information.
+when he/she provided new information.
 
@@
 \paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
-0.05, NS$).
-
-\paragraph{Rectangle question} $\Delta_4$: the correlation between \gmodel{A}{A} and \gmodel{A}{B} is
-\fixme{???  MIRWEIS A FAIRE}.
+0.05, ns$).
+
+%\paragraph{Rectangle question} $\Delta_4$: the correlation between \gmodel{A}{A} and \gmodel{A}{B} is
+%\fixme{???  MIRWEIS A FAIRE}.
 
@@
 \textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
-\textit{Shared Editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
+\textit{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
 \textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
@@
 \textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
-\textit{Dependent variable}   & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
-\textit{Independent variable} & Awareness tool vs not.           & Awareness tool vs not.          & Audio only vs Audio+Video & Script vs  not            & Awareness tool vs not.       \\
+\textit{Dependent var.}   & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
+    \textit{Independent var.} & Awareness tool \textit{vs} not.
+& Awareness tool \textit{vs}  not.
+& Audio \textit{vs} Audio+Video & Script \textit{vs} not.
+& Awareness tool \textit{vs} not.
+\\
 \textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
-\textit{$\Delta_1$}                   & +0.38 (Sig)                      & +0.41 (Sig)                     & ???                       & +0.26 (near SIG)          & +0.05 (non SIG)              \\
+\textit{$\Delta_1$}                   & +0.38 (significant)                      & +0.41 (significant)                     & ???
+& +0.26 (near significant)          & +0.05 (ns)              \\
 \textit{$\Delta_2$}                   &                                  & +0.30                           & +0.43                     &                           &                              \\
@@
 Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
-and 3 reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
+and 3\fxfatal{No $\Delta_2$ in study 1! Dyads!} reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
 \emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of

### 2015-04-21 mutual-modeling.tex
Completed citations

@@
 whose occurrence has been related to learning: (self-) explanations in cognitive
-science \fixme{REF:Chi} \fixme{REF: Webb}, conflict resolution in
-socio-cognitive theories \fixme{REF:Doise},
-mutual regulation \fixme{REF: Blaye} in a Vygostkian perspective, etc. The
+science~\cite{chi1989self, webb1991task}, conflict resolution in
+socio-cognitive theories~\cite{doise1975social},
+mutual regulation~\cite{blaye1995collaborative} in a Vygostkian perspective, etc. The
 construction of a shared understanding has been investigated for several years
@@
 Mutual modelling does not occur in a vacuum but it is highly contextualized.
-REF\fxfatal{REF missing} reviews how the features of the collaborative
+\citet{clark1991grounding} review how the features of the collaborative
 situation, namely the media (co-temporality,\ldots), may facilitate or hamper
@@
 graphical editor) was, in this task, more important than seeing each
-other~\citep{gaver1993one}\fixme{check ref with Pierre}\fixme{other cite:
-Anderson et al., 2000}.
+other~\citep{gaver1993one,anderson1997impact}.
 

### 2015-04-22 mutual-modeling.tex
Switched to Springer style (for Journal of CSCL)

@@
-\documentclass[twocolumn]{article}
+\documentclass[natbib]{svjour3}
 
@@
 \usepackage[english]{babel}
-
-\usepackage[round]{natbib}
 
@@
 \title{Do I Need to Model You in order to Learn With You?}
+\subtitle{Mutual Modelling in CSCL}
 
 \author{
-    Pierre Dillenbourg, Séverin Lemaignan \\
-    \small Computer-Human Interaction in Learning and Instruction (CHILI) \\
-    \small Ecole Polytechnique Fédérale de Lausanne (EPFL) \\
-    \small CH-1015 Lausanne, Switzerland \\
-    {\tt firstname.lastname@epfl.ch}
+    Pierre Dillenbourg
     \and
-    Mirweis Sangin \\
+    Séverin Lemaignan
     \and
-    Nicolas Nova \\
-    \small Haute Ecole Arts \& Design de Genève\\
-    {\tt nicolas.nova@hesge.ch}
+    Mirweis Sangin
+    \and
+    Nicolas Nova
     \and
@@
     \and
-    Gaëlle Molinari \\
-    \small Fernuni Schweiz \\
-    {\tt gaelle.molinari@fernuni.ch}
+    Gaëlle Molinari
+}
+\institute{
+    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
+    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
+    \email{firstname.lastname@epfl.ch}
+    \and
+    Mirweis Sangin \at Sony Computer Entertainment Europe
+    \and
+    Nicolas Nova \at Haute École Arts \& Design de Genève \\\email{nicolas.nova@hesge.ch}
+    \and
+    René Glaus
+    \and
+    Gaëlle Molinari \at Fernuni Schweiz \\\email{gaelle.molinari@fernuni.ch}
 }
@@
 \maketitle
-
-\listoffixmes
 
@@
 whose occurrence has been related to learning: (self-) explanations in cognitive
-science~\cite{chi1989self, webb1991task}, conflict resolution in
-socio-cognitive theories~\cite{doise1975social},
-mutual regulation~\cite{blaye1995collaborative} in a Vygostkian perspective, etc. The
+science~\citep{chi1989self, webb1991task}, conflict resolution in
+socio-cognitive theories~\citep{doise1975social},
+mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
 construction of a shared understanding has been investigated for several years
@@
         \node[anchor=south west,inner sep=0] at (0,0)
-            {\includegraphics[width=0.65\textwidth]{image10.png}};
-
-            \node[rectangle,draw, thick] at (11,6.2) {Self map};
-            \node[rectangle,draw, thick] at (11,3) {Partner map};
-            \node[rectangle,draw, thick] at (3,2) {Collaborative map};
+            {\includegraphics[width=0.75\textwidth]{image10.png}};
+
+            \node[rectangle,draw, thick] at (10,5.4) {Self map};
+            \node[rectangle,draw, thick] at (7.5,3.3) {Partner map};
+            \node[rectangle,draw, thick] at (2.5,1.5) {Collaborative map};
     \end{tikzpicture}
@@
 
-\section*{Acknowledgments}
+\begin{acknowledgements}
 
@@
 
+\end{acknowledgements}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{plainnat}
+\bibliographystyle{spbasic}
 \bibliography{biblio}

### 2015-04-22 mutual-modeling.tex
Split Study 5 picture in two
Several minor fix-up

@@
 
-    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability of
+    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability or
         effort to help $A$ to model him/herself,
@@
 This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$ and
-B have the same accuracy when modelling C? $\Delta_2 = \Delta(\M{A}{C}{X},
-\M{B}{C}{X})$. If it is the same, $\mathcal{H}_{2}$ and
-$\mathcal{H}_{3}$\fxfatal{why H3?} gain
-over $\mathcal{H}_{1}$
-
-Conversely, does $C$ model more accurately $A$ than B? $\Delta_3= \Delta(\M{C}{A}{X},
-\M{C}{B}{X})$ A positive answer would support $\mathcal{H}_{3}$.
+$B$ have the same accuracy when modelling $C$ ($\Delta_2 =
+\Delta(\mathcal{M}^{\circ}(A,C,X),
+\mathcal{M}^{\circ}(B,C,X))$)? If it is the same, $\mathcal{H}_{2}$ gains over
+$\mathcal{H}_{1}$
+
+Conversely, does $C$ model more accurately $A$ than $B$ ($\Delta_3=
+\Delta(\mathcal{M}^{\circ}(C,A,X),
+\mathcal{M}^{\circ}(C,B,X))$)? A positive answer would support $\mathcal{H}_{3}$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study 1: Effect of an awareness tool on \gModel{A}{B} in a virtual
+\subsection{{\bf Study 1}: Effect of an awareness tool on \gModel{A}{B} in a virtual
 game}
@@
         \centering
-        \includegraphics[width=0.9\textwidth]{image4.png}
+        \includegraphics[width=\textwidth]{image4.png}
         \caption{Screenshot of the {\sc SpaceMiners} game}
@@
         \centering
-        \includegraphics[width=\columnwidth]{image5.png}
+        \includegraphics[width=0.7\columnwidth]{image5.png}
         \caption{\gmodel{A}{A} and \gmodel{A}{B} questionnaires in SpaceMiners
@@
 constitution of two groups made up of 12 individuals ``short time in scout
-mode'' and 24 individuals ``long time in scout mode''\fxfatal{Inconsistent! 36
-participants in awareness condition??}. A two-way analysis of
+mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
 variance conducted on these contrasted groups revealed that pairs in the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study 2: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
+\subsection{{\bf Study 2}: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
 game}
@@
         \centering
-        \begin{subfigure}{.3\textwidth}
+        \begin{subfigure}{.4\textwidth}
             \includegraphics[width=\linewidth]{image6.png}
@@
         \end{subfigure}
-        \begin{subfigure}{.3\textwidth}
+        \begin{subfigure}{.4\textwidth}
             \includegraphics[width=\linewidth]{image7.png}
@@
 \paragraph{Triangle questions} Regarding $\Delta_2$, the correlation between
-\gModel{A}{B} and \gModel{A}{C} is significant: $r=0.30, p <.01$. Concerning
-$\Delta_3$, the correlation between \gModel{A}{C} and \gModel{B}{C} is
-significant as well: $r=0.43, p <.001$\fxfatal{Attention! $\Delta_2$ and
-$\Delta_3$ seem to be inverted! Cf fig 2}.
+\gModel{A}{C} and \gModel{B}{C} is significant: $r=0.43, p <.001$. Concerning
+$\Delta_3$, the correlation between \gModel{A}{B} and \gModel{A}{C} is
+significant as well: $r=0.30, p <.01$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study 3:  Effect of media richness on \gModel{A}{B} in argumentation}
+\subsection{{\bf Study 3}:  Effect of media richness on \gModel{A}{B} in argumentation}
 
@@
 
-\paragraph{Grounding criterion} The maps produced by teams were
-ranked\fxfatal{How are they ranked?}. We used
-the average rank as a estimation of the team performance and correlated it with the
-average of the 6 values of \gModel{A}{B} per team (\gModel{A}{B}, \gModel{A}{C},
-\gModel{B}{A},...): The correlation is -0.22 \fxfatal{p value?}: teams with a good \gModel{A}{B}
-team to be better ranked.
+\paragraph{Grounding criterion} The maps produced by teams were ranked by three
+independent judges on completeness and structure quality (Kendall's $W=0.474$,
+limited agreement). We used the average rank as a estimation of the team
+performance and correlated it with the average of the 6 values of \gModel{A}{B}
+per team (\gModel{A}{B}, \gModel{A}{C}, \gModel{B}{A},...). The correlation is
+-0.22: teams with a good \gModel{A}{B} tend to be better ranked.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study 4:  Effects of a script on \gModel{A}{B}  in concept mapping}
+\subsection{{\bf Study 4}:  Effects of a script on \gModel{A}{B}  in concept mapping}
 
@@
 \begin{multline} \label{eq:study4.1}
-    \Mdeg{A}{B}{knowledge} = \\
+    \Mdeg{A}{B}{knowledge} = 
         \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
@@
 \begin{multline} \label{eq:study4.2}
-    \Mdeg{A}{A}{knowledge} = \\
+    \Mdeg{A}{A}{knowledge} = 
         \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
@@
 \begin{multline} \label{eq:study4.3}
-    \Mdeg{A}{B}{knowledge} = \\
+    \Mdeg{A}{B}{knowledge} = 
         \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Study 5: \Model{A}{B}{knowledge} in concept mapping}
+\subsection{{\bf Study 5}: \Model{A}{B}{knowledge} in concept mapping}
 
@@
 
-\begin{figure*}
+\begin{figure*}[h!t]
         \centering
-        \includegraphics[width=0.9\textwidth]{image11.png}
+        \begin{subfigure}{.6\textwidth}
+            \includegraphics[width=\linewidth]{study5-conceptmap.png}
+        \end{subfigure} \\
+        \begin{subfigure}{.7\textwidth}
+            \includegraphics[width=\linewidth]{study5-kat.png}
+        \end{subfigure}
         \caption{Screenshot of the KAT condition during the concept-map building
@@
 \paragraph{Grounding criterion} A linear regression revealed a positive relation
-between \gModel{A}{B} and the collaborative learning gain\fxfatal{What is the
-collaborative learning gain} of the pair $\{A,
-B\}$: $\beta= .401, p < .001, r2adj. =.15$, large effect. This relationship was
-not significant in the previous experiment which as been conducted with the same
-task: using the same task should produce the same grounding criterion. This
-probably due to the fact that \gModel{A}{B} was influenced by the KAT.
+between \gModel{A}{B} and the learning gain of the pair $\{A, B\}$ (calculated
+as the average of the individual learning gains): $\beta= .401, p < .001, r2adj.
+=.15$, large effect. This relationship was not significant in the previous
+experiment which as been conducted with the same task (using the same task
+should produce the same grounding criterion). This is probably due to the fact that
+\gModel{A}{B} was influenced by the KAT.
 
@@
 \textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
-\textit{Task}                 & Game in virtual Space            & Game in physical Space          & Building an argument map  & Building a concept map    & Building a concept map       \\
+\textit{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
 \textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
@@
 \textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
-\textit{$\Delta_1$}                   & +0.38 (significant)                      & +0.41 (significant)                     & ???
-& +0.26 (near significant)          & +0.05 (ns)              \\
-\textit{$\Delta_2$}                   &                                  & +0.30                           & +0.43                     &                           &                              \\
-\textit{$\Delta_3$}                   &                                  & ???                             & +0.66                     &                           &                              \\
+\textit{$\Delta_1$}           & +0.38 (significant)              & +0.41 (significant)             & ???                       & +0.26 (near significant)          & +0.05 (ns)              \\
+\textit{$\Delta_2$}           &                                  & +0.43 (significant)             & +0.43                     &                           &                              \\
+    \textit{$\Delta_3$}       &                                  & +0.30 (significant)             & +0.66                     &                           &                              \\
 \textit{$\Delta_4$}           &                                  &                                 &                           & +0.05                     &                             
@@
 
-Coming back to our initial hypotheses, the $\Delta_2$ results in studies 1, 2
-and 3\fxfatal{No $\Delta_2$ in study 1! Dyads!} reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The
-\emph{interaction hypothesis} $\mathcal{H}_{3}$ was that the activity of
-modelling one's partner is not simply the result of an individual cognitive
-effort but overall the side-effect of the quality of interactions. The
-\emph{second-order-modelling hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is
-accurate in modelling $B$, $A$ should notice when $B$ misunderstands $A$ and should be
-able to repair. A good modeler would somehow drive the modelling of his or her
-partner.  This \emph{drive} will however occur through verbal cues and repairs
-and is hence not easy to differentiate from $\mathcal{H}_{3}$.  The
-contradictory $\Delta_1$ results in studies 4 (to some extent) and 5 weaken the
-hypothesis.
+Coming back to our initial hypotheses, the $\Delta_2$ results in studies 2 and 3
+reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
+hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
+is not simply the result of an individual cognitive effort but overall the
+side-effect of the quality of interactions. The \emph{second-order-modelling
+hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
+$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
+good modeler would somehow drive the modelling of his or her partner.  This
+\emph{drive} will however occur through verbal cues and repairs and is hence not
+easy to differentiate from $\mathcal{H}_{3}$.  The contradictory $\Delta_1$
+results in studies 4 (to some extent) and 5 weaken the hypothesis.
 

### 2015-04-22 mutual-modeling.tex
Completed study 3

@@
 rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
-audio+video condition and .71 ($SD = .19$) in the audio condition ($F=3.4, p:
-0.065$). Our interpretation is that viewing what one's partner sees (shared
-graphical editor) was, in this task, more important than seeing each
-other~\citep{gaver1993one,anderson1997impact}.
-
-\paragraph{Symmetry question} Regarding the main items, the mutual modelling of
-emotions, the correlation is .25 or .36  \fixme{a reprendre avec - Patrick}
-
-\paragraph{Triangle questions} \fixme{Check ($\Delta_2$)} For $\Delta_3$, if we
-compute for each subject A, the correlation between the vectors of emotions in
-\gmodel{B}{A} and the vector in \gmodel{C}{A}, and take the average across all
-60 subjects, we obtain a high correlation of .66. This average is
-about the same in the two conditions.
+audio+video condition and 0.71 ($SD = 0.19$) in the audio condition ($F=3.4, p:
+0.065$).
+
+\paragraph{Symmetry question} We computed the absolute differences between
+\gModel{A}{B} and \gModel{B}{A} over all subjects (3 values per triad, 20
+triads) and averaged them. The average difference in the degrees of accuracy of
+symmetrical models is 0.21 ($SD= 0.21$): pairs of subjects tend to assess the
+emotions of each other with similar levels of accuracy.
+
+\paragraph{Triangle questions} $\Delta_2$ is computed in a similar way as the average of
+the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
+subjects. The average difference is 0.17 ($SD= 0.19$), indicating that, inside a
+triad, $A$ and $B$ tend to model $C$ in a similar way. For $\Delta_3$, the
+average absolute difference between \gModel{C}{A} and \gModel{C}{B} over the 60
+subjects is 0.22 ($SD= 0.12$), which means that a given subject tends to model
+the two other member of the triad with a similar degree of accuracy, which
+supports $\mathcal{H}_{1}$.
 
@@
 \paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
-between self-and social accuracy here because we don't have an estimation of
+between self and social accuracy here because we do not have an estimation of
 self-accuracy: subjects describe their emotions but we have no way to check if
-there are correct. However, by measuring \gModel{A}{B} on  18 emotional labels,
+there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
 we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
@@
 errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
-is in average of 0.3 for the emotion \emph{discasted}, and up to 1.9 for the
+is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
 emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
-0-5$ for \emph{discated} versus $SD=.7$ for \emph{calm}), our point is not to
-interpret this too far, but to show that there are large variations even with one
-area  (perceiving emotions)\fxfatal{Paragraph to rewrite}.
+0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
+interpret this too far, but to show that there are large variations even within one
+area  (perceiving emotions).
 
@@
 
-\fixme{TBD}
-\fixme{Check si il faut modifier les resultats de la symmetries}
+The results evidence that richer media does not automatically foster better
+modelling, on the contrary: groups using only the audio modality to exchange
+their views were able to build more accurate models of each others than groups
+using both video and audio, which is especially surprising when modelling
+emotions: one would expect that video feedback indeed help building a better
+model of the emotions of the other. Our interpretation is that viewing what
+one's partner sees (shared graphical editor) was, in this task, more important
+than seeing each other~\citep{gaver1993one,anderson1997impact}.
+
+In this study, the accuracy of mutual modelling between two peers tend to be
+symmetrical, which supports the idea that the quality of modelling is primarily
+building upon the social interaction. This is further supported by a low
+averaged $\Delta_2$: if, for any $A$ and $B$, $A$ and $B$ generally model $C$
+with a similar level of accuracy, it is likely due to the social interaction
+induced by $C$ (hypothesis $\mathcal{H}_{2}$).
+
+On the other hand, the low $\Delta_3$ that we also found ($C$ tends to always
+model $A$ and $B$ with similar levels of accuracy) tend to contradict this, and
+supports the idea that modelling is primarily an individual skill (hypothesis
+$\mathcal{H}_{1}$).
 

### 2015-05-10 mutual-modeling.tex
Add missing '0' in stats values + a bit of house keeping

@@
 \emph{conceptual change} (\ie the acquisition, acceptation and integration of a
-new belief into one's mental model\fxnote{check with Pierre}) over longer
+new belief into one's mental model) over longer
 sessions~\cite{dillenbourg2006sharing}. The cumulative effect of grounding
@@
 confirmed the positive and significant relation between group performance and
-mutual modelling accuracy ($\beta=54, p = .02$).
+mutual modelling accuracy ($\beta=54, p = 0.02$).
 
@@
 Considering all pairs in both conditions, we found a positive and significant
-correlation ($r = .38, p < .05$) between \gModel{A}{B} and \gModel{B}{A}.
+correlation ($r = 0.38, p < 0.05$) between \gModel{A}{B} and \gModel{B}{A}.
 Interestingly, this was higher in the control group ($r = 0.44$) than in the
@@
 \paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
-\gModel{B}{A}  is positive ($r = .41$) and significant ($p < .01$): the more $A$
+\gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more $A$
 makes errors about $B$, the more $B$ does as well.
@@
 \paragraph{Study-specific questions} Our hypothesis about media richness is
-rejected: the average \Model{A}{B}{emotions} was .62 ($SD = .28$) with the
+rejected: the average \Model{A}{B}{emotions} was 0.62 ($SD = 0.28$) with the
 audio+video condition and 0.71 ($SD = 0.19$) in the audio condition ($F=3.4, p:
@@
 produced concept map) but learning gains. The correlation between \gModel{A}{B}
-and $A$'s learning gains is not significant ($\beta =.08, ns, N = 60$). It is
+and $A$'s learning gains is not significant ($\beta = 0.08, ns, N = 60$). It is
 also not significant within each condition.
@@
 test on post-test scores for electric questions and a one-way ANOVA on scores
-for ionic questions (Levene tests for homogeneity of variances): $p = .02$ and
+for ionic questions (Levene tests for homogeneity of variances): $p = 0.02$ and
 $ns$, respectively. Results  did not show any significant difference between the
 \emph{same information} (SI) condition and the \emph{complementary information}
-(CI) condition, neither for electric questions [$U = 388.50, z = -0.88, ns$] nor
-for ionic questions [$F(1, 58) = 0.17, ns$].  The effect of scripts on
+(CI) condition, neither for electric questions ($U = 388.50, z = -0.88, ns$) nor
+for ionic questions ($F(1, 58) = 0.17, ns$).  The effect of scripts on
 \gModel{A}{B} was not significant ($F(1, 58) = 0.78, ns$) when considering the
-absolute difference between \gModel{A}{B} and $B$'s post-test score. However, A
+absolute difference between \gModel{A}{B} and $B$'s post-test score. However, $A$
 tended underestimate $B$'s score in the SI condition ($M = -2.06$) and to
-overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<.01$).
+overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<0.01$).
 Regarding to \gModel{A}{A}, there was no significant difference between
@@
 \paragraph{Symmetry question} The inter-class correlation between \gModel{A}{B}
-and \gModel{B}{A} is approaching significance ($r=.26, F(1,29)=1.71, p=
+and \gModel{B}{A} is approaching significance ($r= 0.26, F(1,29)=1.71, p=
 0.075$). It is indeed significant when students read the same text (SI
-condition: $r=0.43, F(1,15)=2.53, p<.05$) but not when they read different texts
-($r=.13, F(1,12)=1.3, ns$).
+condition: $r=0.43, F(1,15)=2.53, p<0.05$) but not when they read different texts
+($r= 0.13, F(1,12)=1.3, ns$).
 
@@
 of \gModel{A}{B} being a property of interactions in context rather than $A$'s
-cognitive attitude\fxfatal{to be better justified}.
+cognitive attitude.
 
@@
 between \gModel{A}{B} and the learning gain of the pair $\{A, B\}$ (calculated
-as the average of the individual learning gains): $\beta= .401, p < .001, r2adj.
-=.15$, large effect. This relationship was not significant in the previous
+as the average of the individual learning gains): $\beta= 0.401, p < 0.001, r2adj.
+= 0.15$, large effect. This relationship was not significant in the previous
 experiment which as been conducted with the same task (using the same task
@@
 difference between the KAT condition participants ($M = 13.4\%$) and the control
-group [$M = 3.6\%, t(1, 60) = 2.73, p < .01$, Cohen's $d = 0.7$, medium to large
+group [$M = 3.6\%, t(1, 60) = 2.73, p < 0.01$, Cohen's $d = 0.7$, medium to large
 effect]: providing learners with cues about the prior-knowledge of their partner
@@
 and significant correlation between the amount of gazes on KAT (using eye
-tracking devices) and the learning gains ($r(22) = .54, p = .01$). A detailed
+tracking devices) and the learning gains ($r(22) = 0.54, p = 0.01$). A detailed
 analysis revealed that subjects look at KAT to assess their peer's credibility
@@
 estimated their partners knowledge ($M = 1.11$) than those in the control
-condition [$M = .98, t(1, 60) = 3.19, p < .01$, Cohen's $d = 0.83$, large
+condition [$M = 0.98, t(1, 60) = 3.19, p < 0.01$, Cohen's $d = 0.83$, large
 effect]. This is a trivial result since the KAT provided them with an initial
@@
 on the relative learning gain. A linear regression confirmed that \gModel{A}{B}
-was significantly related to the KAT-factor ($\beta= .381, p < .01, r2 = .15$).
+was significantly related to the KAT-factor ($\beta= 0.381, p < 0.01, r2 = 0.15$).
 The KAT-factor was also significantly and positively related to the RLG ($\beta=
-.332, p < .01, r2 = .11$). We then tested the relation between the independent
+.332, p < 0.01, r2 = 0.11$). We then tested the relation between the independent
 variable (KAT) and the dependent variable (gains) when controlling for the
 mediating variable (\gModel{A}{B}). A multiple regression showed that the
-KAT-factor was no longer a significant predictor ($\beta= .210; p = ns$) whereas
-the \gModel{A}{B} was still a significant predictor ($\beta= .32, p < .01$).
+KAT-factor was no longer a significant predictor ($\beta= 0.210; p = ns$) whereas
+the \gModel{A}{B} was still a significant predictor ($\beta= 0.32, p < 0.01$).
 Thus, it can be concluded that \gModel{A}{B} mediated the KAT-factor's effect on
 the learners' RLG. The Sobel significance test for indirect effects was
-significant [$z = 1.99, p < .05$].
+significant [$z = 1.99, p < 0.05$].
 
@@
 \textit{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
-    \textit{Independent var.} & Awareness tool \textit{vs} not.
-& Awareness tool \textit{vs}  not.
-& Audio \textit{vs} Audio+Video & Script \textit{vs} not.
-& Awareness tool \textit{vs} not.
-\\
+\textit{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
 \textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
@@
 Among these complex results, it is important to stress that, for $\Delta_2$, the
-significant correlations are all below .50, not around .90.  So even if all
+significant correlations are all below 0.50, not around 0.90.  So even if all
 $\Delta_2$ correlations were significant in all experiments, we would still not

### 2015-06-05 mutual-modeling.tex
minor
Updated results for study 3, and changed accordingly the discussion

@@
 \newcommand{\gmodel}[2]{{$\mathcal{M}(#1, #2)$}}
+\newcommand{\Gmodel}[2]{{\mathcal{M}(#1, #2)}}
 \newcommand{\refmodel}[2]{{$\mathcal{M}(#1, #2)$}}
@@
 \newcommand{\stmt}[1]{{\footnotesize \tt $\langle$ #1\relax$\rangle$}}
+
+\usepackage{mathtools}
+\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
+\newcommand*\mean[1]{\overline{#1}}
 
@@
 
+\paragraph{The symmetry question}
+
 One interesting question relates to the symmetry of mutual modelling: what is
@@
 
+\paragraph{The triangle questions}
+
 With triads, we may compute the accuracy of 6 models:
@@
 
-This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$ and
-$B$ have the same accuracy when modelling $C$ ($\Delta_2 =
-\Delta(\mathcal{M}^{\circ}(A,C,X),
-\mathcal{M}^{\circ}(B,C,X))$)? If it is the same, $\mathcal{H}_{2}$ gains over
-$\mathcal{H}_{1}$
+This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
+and $B$ have the same accuracy when modelling $C$ ($\Delta_2 =
+\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? If it is the
+same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gains over $\mathcal{H}_{1}$
 
@@
 
-The aim of this unpublished study was to evaluate the effect of media richness
-on \Model{A}{B}{emotions}. The hypothesis was that video communication would
-lead to a better \gModel{A}{B} than audio only since emotions often impact on
-facial expressions.
+The aim of this unpublished\footnote{The data and statistical analyses of this
+study are available online:
+\url{https://github.com/chili-epfl/mutual-modelling-adoption-homosexuality-study}.}
+study was to evaluate the effect of media richness on \Model{A}{B}{emotions}.
+The hypothesis was that video communication would lead to a better \gModel{A}{B}
+than audio only since emotions often impact on facial expressions.
 
@@
 
-The independent variable was the presence or not of video link. The dependent
-variable \gModel{A}{B} was measured subjectively from three questionnaires: in the first
-one, $A$ described his/her own emotions \gmodel{A}{A}, while in the two other
-questionnaire, $A$ described $B$'s and C's emotions. The questionnaire included 18
-items (7-point Likert Scale) describing emotions labeled as adjectives:
-\emph{anxious}, \emph{enthusiastic}, \emph{agitated}, \emph{proud},
+The independent variable was the presence or not of a video link. The dependent
+variable \gModel{A}{B} was measured subjectively from three questionnaires: in
+the first one, $A$ described his/her own emotions \gmodel{A}{A}, while in the
+two other questionnaire, $A$ described $B$'s and $C$'s emotions. The questionnaire
+included 18 items (7-point Likert Scale) describing emotions labeled as
+adjectives: \emph{anxious}, \emph{enthusiastic}, \emph{agitated}, \emph{proud},
 \emph{excited}, \emph{quiet}, \emph{calm}, \emph{stressed}, \emph{bored},
 \emph{upset}, \emph{relaxed}, \emph{irritated}, \emph{determined},
-\emph{hostile}, \emph{active}, etc. \gmodel{A}{B} was modelled as a vector
-of 18 numerical values corresponding to their answers on each
-questionnaire items, and \gModel{A}{B} was computed as the correlation between the
-two vectors \gmodel{A}{B} and \gmodel{B}{B}.
+\emph{hostile}, \emph{active}, etc. \gmodel{A}{B} was modelled as a vector of 18
+numerical values corresponding to their answers on each questionnaire items, and
+\gModel{A}{B} was computed as the distance between the two vectors
+\gmodel{A}{B} and \gmodel{B}{B}:
+
+\[
+    \Mdeg{A}{B}{emotions} = \frac{\displaystyle\sum_{emotions} \abs{\M{A}{B}{e} -
+    \M{B}{B}{e}}}{18}
+\]
+
+The smaller the score, the \emph{more} accurate the model.
 
@@
 \paragraph{Study-specific questions} Our hypothesis about media richness is
-rejected: the average \Model{A}{B}{emotions} was 0.62 ($SD = 0.28$) with the
-audio+video condition and 0.71 ($SD = 0.19$) in the audio condition ($F=3.4, p:
-0.065$).
-
-\paragraph{Symmetry question} We computed the absolute differences between
-\gModel{A}{B} and \gModel{B}{A} over all subjects (3 values per triad, 20
-triads) and averaged them. The average difference in the degrees of accuracy of
-symmetrical models is 0.21 ($SD= 0.21$): pairs of subjects tend to assess the
-emotions of each other with similar levels of accuracy.
+rejected: the average score for \model{A}{B}{emotions} was 1.25 ($SD = 0.53$) in
+the audio+video condition and 1.09 ($SD = 0.41$) in the audio alone condition
+($t=1.89, df=111, p = 0.062$): the average degree of accuracy of
+\model{A}{B}{emotions} was hence \emph{higher} in the audio alone condition.
+
+\paragraph{Symmetry question} We computed the absolute differences $\Delta_1$
+between \gModel{A}{B} and \gModel{B}{A} over all pair of subjects within a triad
+(3 values per triad, 20 triads), and compared them with the same differences
+computed from random gradings (following the same grade distribution as for the
+experimental data).
+
+A t-test on the two sets releaved a significantly lower average difference in
+the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
+$t=-3.3, df=60, p = 0.0016$), supporting the symmetry hypothesis.
 
@@
 the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
-subjects. The average difference is 0.17 ($SD= 0.19$), indicating that, inside a
-triad, $A$ and $B$ tend to model $C$ in a similar way. For $\Delta_3$, the
-average absolute difference between \gModel{C}{A} and \gModel{C}{B} over the 60
-subjects is 0.22 ($SD= 0.12$), which means that a given subject tends to model
-the two other member of the triad with a similar degree of accuracy, which
-supports $\mathcal{H}_{1}$.
+subjects. The average difference is 0.42 ($SD= 0.34$), and if not significantly
+different from the same data obtained from random gradings ($t=-1.61, df=60,
+p=0.11$).
+
+For $\Delta_3$, the average absolute difference between \gModel{C}{A} and
+\gModel{C}{B} over the 60 subjects is 0.40 ($SD= 0.41$) and is significantly
+lower than chance ($t=-2.62, df=60, p=0.01$): a given subject tends to model its
+two partners with similar degrees of accuracy, which supports $\mathcal{H}_{1}$.
 
@@
 In this study, the accuracy of mutual modelling between two peers tend to be
-symmetrical, which supports the idea that the quality of modelling is primarily
-building upon the social interaction. This is further supported by a low
-averaged $\Delta_2$: if, for any $A$ and $B$, $A$ and $B$ generally model $C$
-with a similar level of accuracy, it is likely due to the social interaction
-induced by $C$ (hypothesis $\mathcal{H}_{2}$).
-
-On the other hand, the low $\Delta_3$ that we also found ($C$ tends to always
-model $A$ and $B$ with similar levels of accuracy) tend to contradict this, and
-supports the idea that modelling is primarily an individual skill (hypothesis
-$\mathcal{H}_{1}$).
+symmetrical, which would support the idea that the quality of modelling is primarily
+building upon the social interaction. This is however not fully supported by the
+analysis of $\Delta_2$: $A$ being good at modelling $C$ does not correlate with
+$B$ being good at modelling $C$.
+
+On the other hand, $\Delta_3$ shows that a given $A$ tends to
+model $B$ and $C$ at similar levels of accuracy: this evidences an individual
+skill at modelling others (hypothesis $\mathcal{H}_{1}$).
 
@@
 
-\begin{table*}[h!t]
+\begin{table*}[ht!]
 \centering
 \resizebox{\textwidth}{!}{%
-\begin{tabular}{p{3cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}|p{4cm}}
+\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
 \textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
-\textit{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
-\textit{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
-\textit{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
-\textit{Group Size}           & 2                                & 3                               & 3                         & 3                         & 3                            \\
-\textit{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
-\textit{Awareness tool}       & Partner's position / orientation & Partner's current/past position & -                         & Partner's concept map     & Partner's scores at pre-test \\
-\textit{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge gains & Team/Ind. knowledge gains    \\
-\textit{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
-\textit{MM°-DV correlation}   & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
-\textit{$\Delta_1$}           & +0.38 (significant)              & +0.41 (significant)             & ???                       & +0.26 (near significant)          & +0.05 (ns)              \\
-\textit{$\Delta_2$}           &                                  & +0.43 (significant)             & +0.43                     &                           &                              \\
-    \textit{$\Delta_3$}       &                                  & +0.30 (significant)             & +0.66                     &                           &                              \\
-\textit{$\Delta_4$}           &                                  &                                 &                           & +0.05                     &                             
+\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
+\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
+\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
+\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
+\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
+\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
+\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
+\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
+\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
+$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
+$\mathcal{H}_2$               &                                  & supported                       & supported                 &                           &                              \\
+$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
 \end{tabular}
@@
 
-Coming back to our initial hypotheses, the $\Delta_2$ results in studies 2 and 3
+Coming back to our initial hypotheses, the $\Delta_2$ results in study 2
 reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
@@
 \emph{drive} will however occur through verbal cues and repairs and is hence not
-easy to differentiate from $\mathcal{H}_{3}$.  The contradictory $\Delta_1$
-results in studies 4 (to some extent) and 5 weaken the hypothesis.
+easy to differentiate from $\mathcal{H}_{3}$. The contradictory $\Delta_2$
+analysis in study 3 and $\Delta_1$ results in studies 4 (to some extent) and 5
+weaken the hypothesis.
 
 The triangular questions lead to similar contradictions. The $\Delta_2$
-correlation between the \gmodel{A}{C} and \gmodel{B}{C} (studies 2 and 3?)
+correlation between the \gmodel{A}{C} and \gmodel{B}{C} (study 2)
 support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
@@
 \gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
-the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (studies
-3 and 2?) question $\mathcal{H}_{3}$: if the role of the modellee was so
-important, it would lead be surprising to have by chance a correlation between
-B's and C' ability to drive $A$'s modelling.
+the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (study
+3) question $\mathcal{H}_{3}$: if the role of the modellee was so
+important, it would be surprising to have by chance a correlation between
+$B$'s and $C$' ability to drive $A$'s modelling.
 
 In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
-($\Delta_2$ in study 4 and study 5) could also contributes to the idea that $A$'s
+($\Delta_4$ in study 4 and study 5) could also contributes to the idea that $A$'s
 modelling skills is not the key factor. However, between modelling someone and
 modelling oneself, \ie  metacognition, they are so many differences, that we
-would hardly conclude from that,
+would hardly conclude from that.
 

### 2015-06-17 mutual-modeling.tex
Updated title + list of authors

@@
 
-\title{Do I Need to Model You in order to Learn With You?}
-\subtitle{Mutual Modelling in CSCL}
+\title{Mutual Modelling}
 
@@
     \and
-    Mirweis Sangin
-    \and
     Nicolas Nova
-    \and
-    René Glaus
     \and
@@
     \and
-    Mirweis Sangin \at Sony Computer Entertainment Europe
+    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
     \and
-    Nicolas Nova \at Haute École Arts \& Design de Genève \\\email{nicolas.nova@hesge.ch}
-    \and
-    René Glaus
-    \and
-    Gaëlle Molinari \at Fernuni Schweiz \\\email{gaelle.molinari@fernuni.ch}
+    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
+    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
 }
@@
 
-The experiments were developed with the help of Patrick Jermann,  Fabien
-Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan Bourquin and  Jeremy Goslin.
-The main funding has been provided from a NSF Grant grant \#102511-106940.
+The experiments were developed with the help of Mirweis Sangin, René Glaus,
+Patrick Jermann,  Fabien Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan
+Bourquin and Jeremy Goslin.  We also thank Kshitij Sharma and {\L}ukasz
+Kidzi\'nski for their help with data analysis. The main funding has been
+provided from a NSF Grant grant \#102511-106940.
 

### 2015-06-17 mutual-modeling.tex
Review by Pierre

@@
     of mutual modelling. It integrates several concepts related to the classical
-    idea of knowledge grounding, and we apply it as an operational tool to
+    idea of grounding, and we apply it as an operational tool to
     analyse the interplay of learning and mutual modelling in the five studies.
@@
 new belief into one's mental model) over longer
-sessions~\cite{dillenbourg2006sharing}. The cumulative effect of grounding
+sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
 episodes can probably be better understood from a socio-cultural perspective:
@@
 saying ``$B$ knows that I don't expect him to solve this statistics problem''
-corresponds to \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
+corresponds to \\ \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
 
@@
         understanding.  The \emph{grounding criterion} -- denoted
-        \groundingcriterion -- refers to how important it is to mutually share a
-        piece of information $X$ to succeed the task $T$. It can be calculated as
-        the probability to succeed $T$ despite the fact $X$ is not grounded.
+        \groundingcriterion -- refers to how accurate the models need to be to
+        succeed the task $T$. It can be calculated as the probability to succeed
+        $T$ despite the fact $X$ is not grounded.
         $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be estimated from the correlation
@@
 \begin{multline} \label{eq:probrepair}
-    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \propto \\
+    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \simeq
     \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) - \mathcal{M}^{\circ}_{t_1}(A,B,X)}{cost (\alpha)}
@@
     \begin{enumerate}
+
+        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
+            performed (study 2) or will perform next (study 1),
+
         \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
             (study 3),
-
-        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
-            performed (study 2) or will perform next (study 1),
 
@@
 
-    \item $\mathcal{M}^{\circ}_{t1}(A,B,X)$: along the same reasoning, the initial degree
+    \item $\mathcal{M}^{\circ}_{t0}(A,B,X)$: along the same reasoning, the initial degree
         of common grounds should be rather low (and hence the difference between
@@
         understand, hence, model, the others. Such an approach has however not
-        yet been investigated in the presented studies.
+        been investigated in the presented studies.
 
@@
 
-\subsubsection*{Studies}
+\subsubsection*{Hypotheses}
 
@@
 
-$\mathcal{H}_{2}$ is also called a second level modelling hypothesis since $B$
-needs to monitor $A$ to see if $A$ understood him/her
-($\mathcal{M}(B,A,\mathcal{M}(A,B))$).
+$\mathcal{H}_{2}$ relates to second level modelling since $B$ needs to monitor
+$A$ to see if $A$ understood him/her ($\mathcal{M}(B,A,\mathcal{M}(A,B))$). We
+will see that $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ are actually difficult to
+differentiate.
 
@@
 One interesting question relates to the symmetry of mutual modelling: what is
-the relationship between \Model{A}{B}{X} and
-\Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is mainly an
-individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might
-support $\mathcal{H}_{2}$ since there is a low probability that randomly formed
-pairs integrate peers with the same level of mutual modelling skills.
+the relationship between \Model{A}{B}{X} and \Model{B}{A}{X}
+(Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is
+mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
+might support $\mathcal{H}_{2}$ or $\mathcal{H}_{3}$ since there is a low
+probability that randomly formed pairs integrate peers with the same level of
+mutual modelling skills.
 
@@
 Conversely, does $C$ model more accurately $A$ than $B$ ($\Delta_3=
-\Delta(\mathcal{M}^{\circ}(C,A,X),
-\mathcal{M}^{\circ}(C,B,X))$)? A positive answer would support $\mathcal{H}_{3}$.
+\Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
+answer would support $\mathcal{H}_{3}$. A negative answer would support
+$\mathcal{H}_{2}$ since $\Delta_3$ could mainly be explained by the way $A$ and
+$B$ help $C$ to model them.
 
@@
 Some answers were identical in both questionnaires (like ``adjusting a shot'')
-while others were reversed (``to guide my partner'' versus ``to guide me''). We
-chose a subjective measure of accuracy ($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$)
-rather than an objective measure (\ie the model \gmodel{A}{B} is compared to $B$'s
-next action) because some of the activities proposed by the questionnaire were not
-observable by the environment (\eg establishing a strategy). We calculated
-\Model{A}{B}{activity_{game_{i}}}) as the number of common answers between
-questionnaires \gmodel{A}{A} and \gmodel{A}{B} in each game and computed the
-average value a across the 3 levels.
+while others were reversed (``to guide my partner'' versus ``to guide me'').
+This method provides us with a subjective measure of accuracy
+($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$) rather than an objective measure (\ie the
+model \gmodel{A}{B} is compared to $B$'s next action) because some of the
+activities proposed by the questionnaire were not observable by the environment
+(\eg establishing a strategy). We calculated \Model{A}{B}{activity})
+as the number of common answers between questionnaires \gmodel{A}{A} and
+\gmodel{A}{B} in each game and computed the average value a across the 3 levels.
 
@@
 
+\subsubsection*{The triangle and rectangle questions are not addressed in this study.}
+
 \subsubsection*{Discussion}
@@
 varies according to personality traits or cognitive skills: some persons tend to
-model more accurately their peer while others don't care or can simply not. If
-the correlation was close to 1, we could conclude that the modelling accuracy is
-an emergent property of teams since there would be low probability that, by
-chance, low accurate modellers are paired with accurate modellers and
-conversely. Our correlation is between these two extremes but the very notion of
-significant intra-class correlation tend to indicate that one may not consider
-peers as independent subjects, which is an argument for the \emph{interaction}
-hypothesis $\mathcal{H}_{3}$.
+model more accurately their peer while others don't care or can simply not
+($\mathcal{H}_{1}$). If the correlation was close to 1, we could conclude that
+the modelling accuracy is an emergent property of teams since there would be low
+probability that, by chance, low accurate modellers are paired with accurate
+modellers and conversely ($\mathcal{H}_{2}$). Our correlation is between these
+two extremes but the very notion of significant intra-class correlation tend to
+indicate that one may not consider peers as independent subjects, which is an
+argument for the \emph{interaction} hypothesis $\mathcal{H}_{3}$.
 
@@
         \centering
-        \begin{subfigure}{.4\textwidth}
+        \begin{subfigure}{.45\textwidth}
             \includegraphics[width=\linewidth]{image6.png}
@@
 study are available online:
-\url{https://github.com/chili-epfl/mutual-modelling-adoption-homosexuality-study}.}
+\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.}
 study was to evaluate the effect of media richness on \Model{A}{B}{emotions}.
@@
         \centering
-        \includegraphics[width=0.8\textwidth]{image8.jpg}
+        \includegraphics[width=\textwidth]{image8.jpg}
         \caption{Example of argumentation graph}
@@
 per team (\gModel{A}{B}, \gModel{A}{C}, \gModel{B}{A},...). The correlation is
--0.22: teams with a good \gModel{A}{B} tend to be better ranked.
+0.22: teams with a good \gModel{A}{B} tend to be better ranked.
 
@@
 the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
-$t=-3.3, df=60, p = 0.0016$), supporting the symmetry hypothesis.
+$t=-3.3, df=60, p = 0.0016$), supporting $\mathcal{H}_{2}$ or $\mathcal{H}_{3}$.
 
@@
 different from the same data obtained from random gradings ($t=-1.61, df=60,
-p=0.11$).
+p=0.11$). This rejects $\mathcal{H}_{2}$.
 
@@
 two partners with similar degrees of accuracy, which supports $\mathcal{H}_{1}$.
+It could also support $\mathcal{H}_{3}$, but it does not since $\Delta_2$ is not
+significant.
 
@@
         \centering
-        \includegraphics[width=0.9\textwidth]{image9.pdf}
+        \includegraphics[width=\textwidth]{image9.pdf}
         \caption{Average values of \Model{A}{B}{X} where $X$ is one of the proposed
@@
     \centering
-    \begin{tikzpicture}[scale=0.8]
+    \begin{tikzpicture}[scale=1]
         \node[anchor=south west,inner sep=0] at (0,0)
@@
 
-            \node[rectangle,draw, thick] at (10,5.4) {Self map};
-            \node[rectangle,draw, thick] at (7.5,3.3) {Partner map};
+            \node[rectangle,draw, thick] at (8,4.3) {Self map};
+            \node[rectangle,draw, thick] at (6,2.6) {Partner map};
             \node[rectangle,draw, thick] at (2.5,1.5) {Collaborative map};
@@
 material. The order of questions about oneself and about the other was
-counterbalanced across participants. We measure \gModel{A}{B} in an objective
-(Eq.~\ref{eq:study4.1} and \ref{eq:study4.2}) as well as in a subjective way
-(Eq.~\ref{eq:study4.3}).
-
-\begin{multline} \label{eq:study4.1}
-    \Mdeg{A}{B}{knowledge} = 
-        \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
-\end{multline}
-
-\begin{multline} \label{eq:study4.2}
-    \Mdeg{A}{A}{knowledge} = 
-        \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
-\end{multline}
-
-\begin{multline} \label{eq:study4.3}
-    \Mdeg{A}{B}{knowledge} = 
-        \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
-\end{multline}
+counterbalanced across participants. 
+%We measure \gModel{A}{B} in an objective
+%(Eq.~\ref{eq:study4.1} and \ref{eq:study4.2}) as well as in a subjective way
+%(Eq.~\ref{eq:study4.3}).
+%
+%\begin{multline} \label{eq:study4.1}
+%    \Mdeg{A}{B}{knowledge} = 
+%        \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
+%\end{multline}
+%
+%\begin{multline} \label{eq:study4.2}
+%    \Mdeg{A}{A}{knowledge} = 
+%        \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
+%\end{multline}
+%
+%\begin{multline} \label{eq:study4.3}
+%    \Mdeg{A}{B}{knowledge} = 
+%        \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
+%\end{multline}
 
@@
 are in conditions of symmetry of information. This consolidates the hypothesis
-of \gModel{A}{B} being a property of interactions in context rather than $A$'s
-cognitive attitude.
+of \gModel{A}{B} being a property of interactions in context ($\mathcal{H}_{3}$)
+rather than $A$'s cognitive attitude.
 
@@
 $\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
-$\mathcal{H}_2$               &                                  & supported                       & supported                 &                           &                              \\
+$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
 $\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
@@
 the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (study
-3) question $\mathcal{H}_{3}$: if the role of the modellee was so
+3) question $\mathcal{H}_{2}$: if the role of the modellee was so
 important, it would be surprising to have by chance a correlation between
@@
 In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
-($\Delta_4$ in study 4 and study 5) could also contributes to the idea that $A$'s
+($\Delta_4$ in study 4 and study 5) could also contribute to the idea that $A$'s
 modelling skills is not the key factor. However, between modelling someone and

### 2015-06-18 authors.tex
Prepa for submission: anonymization + page with authors' information

--- 2015-06-17 mutual-modeling.tex
@@
 
-\section{Introduction}
-
-From its very beginning, CSCL research has been following
-\citet{roschelle1995construction} suggestion that collaborative learning has
-something to do with the process of constructing and maintaining a \emph{shared
-understanding} of the task at hand. Building a shared/mutual understanding
-refers to the upper class of collaborative learning situations, those in which
-students should build upon each other's understanding to refine their own
-understanding: what is expected to produce learning is not the mere fact that
-two students build the same understanding but the cognitive effort they have to
-engage to build this shared understanding~\citep{schwartz1995emergence}. This
-effort can be observed by the frequency of rich interactions, \ie interactions
-whose occurrence has been related to learning: (self-) explanations in cognitive
-science~\citep{chi1989self, webb1991task}, conflict resolution in
-socio-cognitive theories~\citep{doise1975social},
-mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
-construction of a shared understanding has been investigated for several years
-in psycholinguistics, under the  notion of
-\emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
-mechanisms for explaining learning outcomes has been questioned in learning
-sciences. The monitoring and repair of misunderstanding explains for instance
-referential failures in short dialogue episodes but does hardly predict
-\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
-new belief into one's mental model) over longer
-sessions~\cite{dillenbourg2006sharing}. The cumulative effect of grounding
-episodes can probably be better understood from a socio-cultural perspective:
-
-\begin{quote}
-Collaborative learning is associated with the increased
-cognitive-interactional effort involved in the transition from \emph{learning to
-understand each other} to \emph{learning to understand the meanings of the semiotic
-tools that constitute the mediators of interpersonal
-interaction}~\citep{baker1999role}
-\end{quote}
-
-Along this line, several scholars suggest that CSCL research should go deeper in
-the understanding of how partners engage into shared meaning
-making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
-making~\citep{suthers2006technology}.
-
-Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
-change viewpoint, it is criticized at the same time as being too cognitivist by
-some psycholinguists, \ie as overestimating the amount of shared knowledge and
-mutual representations actually necessary to conduct a dialogue. The fundamental
-issue, as old as philosophy, is the degree of coupling between the different
-levels of dialogue, mostly between the lexical/syntactical level and the deeper
-semantic levels. \citet{pickering2006alignment} argue that the mutual
-understanding starts mostly with a \emph{superficial alignment} at the level of
-the linguistic representations, due to priming mechanisms, and that this local
-alignment may -- in some cases -- lead to a \emph{global alignment} of the
-semantic level (\emph{deep grounding}).  For these authors, the convergence in
-dialogue, and even the repair of some misunderstandings, is explained by this
-mimetic behavior more than by a monitoring of each other's knowledge:
-\emph{``...interlocutors do not need to monitor and develop full common ground
-as a regular, constant part of routine conversation, as it would be
-unnecessary and far too costly. Establishment of full common ground is, we
-argue, a specialized and non-automatic process that is used primarily in
-times of difficulty (when radical misalignment becomes apparent).''} (p. 179).
-This view is actually not incompatible with Clark's \emph{grounding
-criterion}~\citep{clark1989contributing}: the degree of shared understanding that
-peers need to reach depends upon the task they perform. For instance, a dialogue
-between two surgeons might rely on superficial alignment if they talk about
-their friends but has to guarantee accurate common grounds when talking about
-which intervention will be conducted in which way on which patient. 
-%Some CSCL
-%scholars stressed the ``illusion of shared understanding'' \fixme{REF: Jacco?},
-%\ie the potential decoupling between linguistic alignment and actual shared
-%meanings.
-
-This interesting cognitive science debate occurred mostly outside the field of
-learning. In education, the question is to relate these mechanisms to learning
-outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
-change?} \textbf{Does negotiation of meaning only occurs when partners monitor
-and diagnose each other's knowledge?} If the ratio between shallow alignment and
-deep grounding depends upon the task, and if deep grounding is a condition for
-learning, then the pedagogical challenge is indeed to design tasks that require
-deep grounding. Most empirical studies on grounding and alignment are conducted
-with tasks, which despite being qualified as ``ecologically valid'' by their
-authors, are mere referencing tasks such as asking the way to the train station
-or helping the peer to choose a picture among many. In this contribution, we
-explore several richer tasks such as arguing on a sensitive issue or building a
-concept map.
-
-Deep grounding or shared meaning making requires some cognitive load. For
-\citet{clark1986referring}, what is important is not the individual effort made
-by the receiver of a communicative act, but the overall \emph{least collaborative
-effort}.  The cost of producing a perfect utterance may be higher than the cost
-of repairing the problems that may arise through misunderstandings, and in fact,
-subjects tend to make less efforts adapting their utterances to a specific
-partner when they know that they can later provide feedback on his/her
-understanding~\citep{schober1993spatial}. We introduced the notion of
-\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
-that misunderstanding should not be viewed as something to be avoided (if this
-was possible), but as an opportunity to engage into verbalization, explanation,
-negotiation, and so forth. This issue is related to the global argument
-regarding cognitive load in learning activities, especially in discovery
-learning environments: there is no learning without some cognitive load, but
-overload may hinder learning~\citep{paas2003cognitive}. In the context of
-collaborative learning, we understand the cognitive load induced by mutual
-modelling as part of~\citet{schwartz1995emergence} notion of effort towards a
-shared understanding. For instance, CSCL researchers expanded the use of
-\emph{collaboration scripts}~\citep{kobbe2007specifying}. A script is a pedagogical
-method that frames collaborative learning activities in order to foster the
-emergence of productive interactions such as argumentation, explanation or
-conflict. Conflict-resolution scripts such as the {\sc
-ArgueGraph}~\citep{dillenbourg2008mechanics} form pair of students with
-opposite opinions, which increases the difficulty of consensus building,
-requiring more justifications, more negotiation, more load. Similarly, {\sc
-jigsaw} scripts~\citep{aronson1978jigsaw} provide peers with different but
-complementary knowledge for augmenting (reasonably) the efforts that group members
-have to engage into to reach a shared solution.
-
-%\paragraph{Theory of Mind}
-%
-%Theory of Mind (originally defined in~\citep{premack1978does}) is the cognitive
-%ability that a subject possesses to represent the mental state of another
-%agent, possibly including knowledge that contradicts the subject's own model: for
-%example, a book can be at the same time \emph{visible} for myself, and \emph{not
-%visible} for you.
-%
-%Children develop this skill, which is essential to understand others'
-%perspectives during interactions, around the age of three. It supposes the
-%ability to build, store and retrieve separate models of the knowledge of the
-%interactors.
-%
-%One classical application of this cognitive skill is the so-called
-%\emph{False-Belief} experiment (also known as the \emph{Sally and Ann}
-%experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
-%people, $A$ and $B$, manipulate objects. Then $A$ leaves and $B$ hides away one
-%object. When $A$ comes back, we ask the child ``where do you think $A$ will
-%look for the object?''. Before acquiring a theory of mind, children are not
-%able to separate their own (true) model of the world (where they know that
-%the object was hidden) from the model of A, which contains \emph{false
-%beliefs} on the world ($A$ still thinks the object is at its original
-%position since he did not see $B$ hiding it).
-
-\vspace{2em}
-These preliminary remarks frame the main question that this article address:
-in order to engage into shared understanding, which is a desirable learning
-situation, do partners have to build a representation of each other's knowledge?
-if so, what exactly is the nature of this mutual modelling? We adopt hereafter a
-constructive approach, first introducing a simple yet convenient model of mutual
-modelling, then analyzing five independent studies in terms of the interactions
-between learning and mutual modelling.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Mutual Modelling}
-
-We refer to \emph{mutual modelling} as the process of inferring one's partner
-mental states. Any claim that students carry out a detailed monitoring of their
-peers would be as incorrect as any claim that they do not maintain any
-representation at all. If mutual modelling had to be permanently detailed and
-accurate, subjects would obviously face a huge cognitive load. Conversely, peers
-could not collaborate without some minimal amount of mutual modelling. For
-instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
-opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
-B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
-evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
-it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
-profession) and, of course, it can be incorrect\ldots but it can not remain empty.
-Dialogues include many instances of utterances such as ``I thought he would do
-that'' (first level of mutual modelling) or even ``He thought I would do that
-but I intended something else.'' (second level of mutual modelling).
-
-The content of mutual models ranges from \emph{dispositional} versus
-\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
-representation of $B$'s long term knowledge, skills or traits. It is thus closely
-related to the notion of transactive memory~\citep{wegner1987transactive,
-moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
-B's knowledge, behavior or intentions specifically activated in the situation in
-which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
-ones for 2 hours. Examples of fragments that constitute $A$'s model of B
-regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:
-
-\begin{itemize}
-
-    \item $Model(A,B, knowledge)$: what does $A$ know about $B$'s knowledge with
-        respect to the task at hand or, inversely, about $B$'s knowledge gaps?
-        When can $A$ consider $B$'s statements as reliable?
-
-    \item $Model(A,B, skills)$: what does $A$ know about $B$'s skills with respect to
-        the task at hand? May $A$ expect $B$ to perform well in a specific subtask?
-        The effectiveness of division of labor depends on the quality of this
-        mutual model.
-
-    \item $Model(A,B, goals)$: what does $A$ know about $B$'s intentions with respect
-        to the project, including $B$' motivation and commitment? Can $A$ trust B
-        when $B$ promises to deliver?
-
-    \item $Model(A,B, task)$: what does $A$ know about $B$'s representation of the
-        situation and the task: does $A$ knows whether $B$ has the same
-        understanding of the problem at stake?
-
-    \item $Model(A,B, plans)$: what does $A$ know about $B$'s strategy. Does A
-        understand why $B$ did what he did? Is $A$ able to anticipate what $B$ will do
-        next?
-
-    \item $Model(A,B, ``urgent'')$: what does $A$ know about $B$'s understanding of $A$'s
-        last utterance: does ``urgent'' means now, soon or ``not too late''?
-
-\end{itemize}
-
-The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
-beliefs, emotions, history, status, etc.
-
-Besides, we have different levels of mutual modelling. If $A$ states ``B thinks I am good in
-maths'', $A$ builds a \emph{second level} model:
-\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
-regress of nested models: $A$
-saying ``$B$ knows that I don't expect him to solve this statistics problem''
-corresponds to \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
-
-A partner model is likely not a ``box'', \ie not a monolithic representation
-but rather a mosaic of information fragments about the partner, with various
-granularity and various life cycles. This mosaic is elaborated through a variety
-of mechanisms, first for building an initial model of the partner, then for
-updating this model. As two students meet for the first time, mutual models are
-initialized by the assumptions they make upon each other based on cues such as
-his/her belonging to broad categories (age, culture, profession, \ldots),
-stereotypes (sportsmen, junkie, business women, Swiss,\ldots) as well as physical
-appearance. Scholars studied how initial modelling impacts communication. In
-their experiments on initial mutual modelling, \citet{slugoski1993attribution}
-pretended to their subjects that their (fake) partner had or had not received
-the same information. They observed that the subjects adapted their dialogue by
-focusing the explanation on the items that he/she was supposed to ignore.
-\citet{brennan1991conversation} showed that the subjects used different initial
-strategies in forming queries depending on who they were told their partner was.
-
-Initial common grounds are also initiated by co-presence: they include events to
-which $A$ and $B$ attended together~\citep{clark2002definite} in the physical
-space or in their cultural space (\eg ``09-11''). While co-presence means that
-they can refer to shared objects and events, it does not imply that they give
-them the same meaning. Namely, a shared screen does not mean a shared
-understanding~\citep{dillenbourg2006sharing}.
-
-After initialization, mutual models are updated during the collaborative work
-through verbal and non-verbal interactions. A default inference rule is that ``my
-partner agrees with me unless he disagrees'', which rejects the critiques that
-mutual modelling generates an unbearable cognitive load. This default rule is
-superseded by the several mechanisms for monitoring and repairing the partner
-understanding: acknowledgement, continuous attention, relevance of next turns,
-facial expressions including gaze signals, etc.
-
-Mutual modelling does not occur in a vacuum but it is highly contextualized.
-\citet{clark1991grounding} review how the features of the collaborative
-situation, namely the media (co-temporality,\ldots), may facilitate or hamper
-mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
-short silence between two pilots was perfectly interpreted because it occurred
-in a highly constrained communication context.  Some environments are more
-productive than others in helping peers to detect their misunderstandings.
-\citet{roschelle1995construction} reformulate the design of CSCL
-interfaces to provide ways for peers to detect and repair their
-misunderstanding. In CSCW, researchers have explored various so-called
-``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
-that A could not directly perceive because $B$ was working on a different subset
-of the virtual space. Different awareness tools are presented and relied upon in
-this contribution as well since they allowed us to manipulate experimentally the
-mutual modelling activity.
-
-%\subsection{State of the Art}
-%
-%\paragraph{Building a model of a human during the interaction} Perspective
-%Taking is a human ability which allows one to put him/herself in another
-%person's point of view. Studied in psychology
-%literature~\citep{Flavell1992,Tversky1999}, this ability is crucial when
-%interacting with people by allowing one to reason on others' understanding of
-%the world in terms of visual perception, spatial descriptions, affordances and
-%beliefs, etc.  In the last years these notions have been gradually employed in
-%HRI.~\citep{Breazeal2006} presents a learning algorithm that takes into account
-%information about a teacher's visual perspective in order to learn a task.
-%~\citep{Johnson2005} apply visual perspective taking for action recognition
-%between two robots.~\citep{Trafton2005} use both visual and spatial perspective
-%taking to find out the referent indicated by a human partner.
-%
-%The cognitive model that the robot builds for the agent it interacts with is
-%still simple and mostly focused on geometric features (who sees what? What are
-%our relative positions? etc.). Extending this knowledge with more subtle
-%perceptions (emotional state for instance) remains to be done.
-%
-%\paragraph{Joint action}
-%Several theories dealing with collaboration~\citep{Cohen1991,Grosz1996,Clark1996}
-%emphasize that collaborative tasks have specific requirements compared to
-%individual ones, \eg, since the robot and the person share a common goal, they
-%have to agree on the manner to realize it, they must show their commitment to
-%the goal during execution, etc. Several robotic systems have already been built
-%based on these theories~\citep{Rich1997,Sidner2005,Breazeal2003} and they all
-%have shown benefits of this approach. They have also shown how difficult it is
-%to manage turn-taking between communication partners and to interleave task
-%realization and communication in a generic way. Finally, today only few
-%systems~\citep{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
-%levels.
-%
-
-\section{A model for mutual modelling evaluation}
-
-To report on experiments on mutual modelling, we suggest the notation
-\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This is a simple but
-useful reduction to reason on mutual modelling. This notation does not mean $A$
-has an explicit, monolithic representation of $B$: it must be understood as an
-abstraction referring to complex socio-cognitive processes. Besides, we refer to
-the \emph{degree of accuracy} of the model as \Model{A}{B}{X}.
-
-We parametrize and assess the mutual modelling effort through 3 variables:
-
-\begin{enumerate}
-
-    \item Tasks vary a lot with respect to how much they require mutual
-        understanding.  The \emph{grounding criterion} -- denoted
-        \groundingcriterion -- refers to how important it is to mutually share a
-        piece of information $X$ to succeed the task $T$. It can be calculated as
-        the probability to succeed $T$ despite the fact $X$ is not grounded.
-        $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be estimated from the correlation
-        between \Model{A}{B}{X} and the task performance.
-
-    \item Before any specific grounding action, there is generally a non-null
-        probability that $X$ is mutually understood by $A$ and $B$ (\eg $X$ is
-        part of $A$'s and $B$'s cultures, it is manifest to co-present subjects
-        or simply there is not much space for misunderstanding or disagreement
-        about $X$). We simply could not collaborate without a certain level of
-        initial grounds. We note the theoretical accuracy of \emph{initial
-        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
-
-    \item The \emph{cost of grounding} $X$ refers to the physical and cognitive
-        effort required to perform a grounding act $\alpha$: a verbal repair
-        (\eg rephrasing), a deictic gesture, a physical move to adopt one
-        partner's viewpoint, etc. This cost varies according to media
-        features~\citep{clark1991grounding}.
-
-\end{enumerate}
-
-Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
-content $X$ at time $t_1$ during task T in order to increase \Model{A}{B}{X}
-at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
-costs (3)~\citep{traum1996miscommunication}:
-
-\begin{multline} \label{eq:probrepair}
-    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \propto \\
-    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) - \mathcal{M}^{\circ}_{t_1}(A,B,X)}{cost (\alpha)}
-\end{multline}
-
-This is qualitative summary more that a real equation since several parameters
-are hard to quantify (\eg the cost of a communication act depends upon the user
-as well) but it clarifies the parameters of the experiments we present hereafter.
-
-\begin{itemize}
-    \item \Model{A}{B}{X}: our experiments address different contents that can be
-        represented in mutual models:
-
-    \begin{enumerate}
-        \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
-            (study 3),
-
-        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
-            performed (study 2) or will perform next (study 1),
-
-        \item \model{A}{B}{knowledge}: how accurately $A$ estimates $B$'s knowledge
-            with respect to the material they learn together (study 4 and 5).
-
-    \end{enumerate}
-
-    \item $\mathcal{M}^{\circ}_{min}(A,B,X,T)$: our studies build upon various
-        collaborative tasks: argumentation (study 3), games (study 1 and 2) and
-        concept mapping (study 4 and 5). By varying the tasks, we do actually
-        vary the grounding criterion but they all require a ``reasonably high''
-        grounding criterion.  All tasks are about building a solution or a
-        representation together; we did not address everyday conversations.
-
-    \item $\mathcal{M}^{\circ}_{t1}(A,B,X)$: along the same reasoning, the initial degree
-        of common grounds should be rather low (and hence the difference between
-        initial and required degrees rather high) in order to make mutual
-        modelling effort more observable. Studies 1, 4 and 5 have been conducted
-        with teams of students who did not know each other. They came
-        nonetheless from the same university (and they hence had some general common
-        grounds).  For studies 2 and 3, students knew each other before for
-        reasons explained later on. In study 5, we manipulated the initial mutual modelling by
-        using a {\sc jigsaw} script.
-
-    \item $cost(\alpha)$: in all studies but study 4, the cost of  grounding is
-        an independent variable. Study 3 uses media richness as independent
-        variable, with the hypothesis that modelling emotions is ``cheaper'' with a
-        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
-        use awareness tools which, in principle, reduce the cost of mutual modelling, but do
-        not eliminate all costs: if the tool provide $A$ with information about
-        what $B$ does/knows, this additional information may actually increase
-        cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
-        and, like any prosthesis, they may augment mutual modelling (by facilitating it or
-        even scaffolding it) or inhibit it (by making it useless).
-
-\end{itemize}
-
-Measuring \Model{A}{B}{X} is methodologically difficult. We
-discriminate two steps, first to capture \model{A}{B}{X} and then to
-estimate \Model{A}{B}{X}.
-
-\begin{itemize}
-    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$
-        what he/she believes about what $B$ knows, feels, intends to do, etc.
-        This raises obvious methodological concerns since such a question
-        triggers a modelling process beyond what would naturally occur. To avoid
-        this bias, one can estimate mutual modelling after task completion.
-        Then, the obvious drawback are memory losses and by post-hoc
-        reconstruction.  The first method was used in study 2 and the second one
-        in the other studies. Another option would be to rely on external
-        behavioural metrics like eye-tracking: we hypothesise for instance that
-        the fixation time reflects the efforts engaged by the human to
-        understand, hence, model, the others. Such an approach has however not
-        yet been investigated in the presented studies.
-
-    \item {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
-        need to access the reference model \refmodel{B}{X} to estimate its
-        \emph{accuracy}. Since we can only indirectly access it via what $B$
-        reports (\ie \model{B}{B}{X}), accuracy can be estimated in 2 ways:
-
-        \begin{itemize}
-
-            \item Subjective accuracy: In study 3, we compute
-                \Model{A}{B}{X} by measuring if $A$ describes
-                $B$'s emotions in the same way $B$ reports its emotions 
-                ($\M{A}{B}{X} = \M{B}{B}{X}$).
-
-            \item Objective accuracy: In studies 4 and 5, we compute
-                \Model{A}{B}{X} by comparing \Model{A}{B}{K} to $B$'s
-                actual knowledge as it has measured by a test.
-
-        \end{itemize}
-
-\end{itemize}
-
-\subsubsection*{Studies}
-
-The experiments we report here address mutual modelling across different tasks,
-some with dyads, other with triads. They were conducted along 6 years in two
-different institutions by different researchers. They used different
-independent, intermediate and dependent variables. Nonetheless, we were able to
-address a few questions across these studies. We start with 3 simple hypotheses
-about $\mathcal{M}^{\circ}(A,B)$:
-
-\begin{itemize}
-    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon $A$'s ability or effort
-        to model $B$,
-
-    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability or
-        effort to help $A$ to model him/herself,
-
-    \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
-        interactions among $A$ and $B$.
-
-\end{itemize}
-
-$\mathcal{H}_{2}$ is also called a second level modelling hypothesis since $B$
-needs to monitor $A$ to see if $A$ understood him/her
-($\mathcal{M}(B,A,\mathcal{M}(A,B))$).
-
-\paragraph{The symmetry question}
-
-One interesting question relates to the symmetry of mutual modelling: what is
-the relationship between \Model{A}{B}{X} and
-\Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is mainly an
-individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation might
-support $\mathcal{H}_{2}$ since there is a low probability that randomly formed
-pairs integrate peers with the same level of mutual modelling skills.
-
-\begin{figure}[htb]
-\centering
-
-\begin{tikzpicture}[>=latex, scale=0.5]
-
-\draw(0,0) node[anchor=north] (A) {\scriptsize A};
-\draw(10,0) node[anchor=north] (B) {\scriptsize B};
-\draw(5,2) node[anchor=north] {\scriptsize \Model{A}{ B}{ X}};
-\draw(5,-2) node[anchor=north] {\scriptsize \Model{B}{ A}{ X}};
-\draw[dashed, <->] (5,1) -- (5,-1.9) node[midway, sloped, above] {$\Delta_1$};
-\draw[->] (A) to[bend left] (B);
-\draw[->] (B) to[bend left] (A);
-
-\end{tikzpicture}
-
-\caption{\small Mutual modelling in a dyadic interaction, $\Delta_1 =
-    \Delta(\mathcal{M}^{\circ} (A,B,X),
-\mathcal{M}^{\circ} (B,A,X))$}
-
-\label{mm_symmetry}
-\end{figure}
-
-\paragraph{The triangle questions}
-
-With triads, we may compute the accuracy of 6 models:
-\Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
-\Model{C}{B}{X} and \Model{B}{C}{X}.
-
-This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
-and $B$ have the same accuracy when modelling $C$ ($\Delta_2 =
-\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? If it is the
-same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gains over $\mathcal{H}_{1}$
-
-Conversely, does $C$ model more accurately $A$ than $B$ ($\Delta_3=
-\Delta(\mathcal{M}^{\circ}(C,A,X),
-\mathcal{M}^{\circ}(C,B,X))$)? A positive answer would support $\mathcal{H}_{3}$.
-
-\begin{figure}[htb]
-\centering
-\subcaptionbox{}{ 
-    \begin{tikzpicture}[>=latex, scale=0.5]
-
-    \draw(0,2) node (A) {\scriptsize A};
-    \draw(0,-2) node (B) {\scriptsize B};
-    \draw(10,0) node (C) {\scriptsize C};
-
-    \draw(5,2) node {\scriptsize \Model{A}{ C}{ X}};
-    \draw(5,-2) node {\scriptsize \Model{B}{ C}{ X}};
-    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above] {$\Delta_2$};
-    \draw[->] (A) to (C);
-    \draw[->] (B) to (C);
-
-    \end{tikzpicture}
-}
-\subcaptionbox{}{ 
-    \begin{tikzpicture}[>=latex, scale=0.5]
-
-    \draw(0,2) node (A) {\scriptsize A};
-    \draw(0,-2) node (B) {\scriptsize B};
-    \draw(10,0) node (C) {\scriptsize C};
-
-    \draw(5,2) node {\scriptsize \Model{C}{ A}{ X}};
-    \draw(5,-2) node {\scriptsize \Model{C}{ B}{ X}};
-    \draw[dashed, <->] (3,1.2) -- (3,-1.2) node[midway, sloped, above]
-    {$\Delta_3$};
-    \draw[<-] (A) to (C);
-    \draw[<-] (B) to (C);
-
-    \end{tikzpicture}
-}
-\caption{\small Mutual modelling in a triadic interaction}
-
-\label{mm_triangles}
-\end{figure}
-
-In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
-whether the accuracy of mutual modelling depends more upon the modeller's effort
-or the modellee's behaviour.
-
-\paragraph{The rectangle questions}
-
-We can go further by comparing self- versus other modelling ($\Delta_4$ in
-Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
-question if modelling skills depend upon what aspects are being modeled ($X$ or
-$Y$), which would explain vertical differences ($\Delta_5$ in
-Fig.~\ref{mm_rectangle}). When possible, we will also discuss these two
-questions in the studies presented hereafter.
-
-\begin{figure}[htb]
-\centering
-
-\begin{tikzpicture}[>=latex, scale=0.5]
-
-    \draw(0,0) node (a) {\scriptsize \Model{A}{ B}{ X}};
-    \draw(10,0) node (b) {\scriptsize \Model{A}{ A}{ X}};
-    \draw(10,-4) node (c) {\scriptsize \Model{A}{ A}{ Y}};
-    \draw(0,-4) node (d) {\scriptsize \Model{A}{ B}{ Y}};
-    \draw[<->] (a) -- (b) node[midway, below] {$\Delta_4$};
-    \draw[<->] (b) -- (c);
-    \draw[<->] (c) -- (d);
-    \draw[<->] (d) -- (a) node[midway, sloped, above] {$\Delta_5$};
-
-\end{tikzpicture}
-
-\caption{\small Meta-cognitive skills and domain-dependent modelling}
-
-\label{mm_rectangle}
-\end{figure}
-
-This simple notation does not pretend to provide a mathematical account of
-mutual modelling but to be more systematic in describing the following
-experiments.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Studies}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\bf Study 1}: Effect of an awareness tool on \gModel{A}{B} in a virtual
-game}
-
-We studied the impact of an awareness tool on group performance and mutual
-modelling~\citep{nova2007collaboration}. The availability
-of an awareness tool was our independent variable. In previous studies, we
-replayed a video of the game to subjects who surprised us by their ability to
-remember former states of their mutual model: ``I did that because I thought that
-you would do that''. Hence, this experiment focused the representation of each
-other's action plans. During the game, we asked them to anticipate the next
-action of their partner as well as to announce their own actions.
-
-\subsubsection*{Experimental setting}
-
-\begin{figure*}
-        \centering
-        \includegraphics[width=\textwidth]{image4.png}
-        \caption{Screenshot of the {\sc SpaceMiners} game}
-        \label{study1:spaceminer}
-\end{figure*}
-
-{\sc SpaceMiners} is a 3D game that involves two players collecting minerals located
-in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch drones through
-the space after choosing their initial direction and speed. Once launched, the
-trajectory of drones is only influenced by the gravity of planets and by
-trajectory modification tools.  During the experiment, the teams were confronted
-with three increasingly complex situations. The experiment was 2 hours long: a
-30 minutes tutorial and 3 levels of 30 minutes. Players uses using a regular
-joystick and communicated with each other through an audio channel.
-
-The independent variable was the availability of an awareness tool that shows to
-player $A$  the location and gaze direction of player $B$. In the awareness
-condition, players could switch to the \emph{scout mode} where they could view what
-their partner was looking at. We hypothesize that this would enable subjects to
-more accurately infer their teammate's intentions. Each player sat in front of a
-distinct computer located in different rooms.
-
-\subsubsection*{Subjects}
-
-Thirty-six persons participated in this study, all native French speakers. We
-constituted 18 dyads who did not know each other beforehand. The pairs were
-randomly assigned to either the control condition (without the awareness tool)
-or the awareness condition (with the awareness tool).
-
-\subsubsection*{Variables}
-
-Task performance was measured by the score reached by the two subjects at the
-end of the game (three levels). The effort of mutual modelling was measured as
-the ratio of time that players would spend in the scout mode (divided by total
-time), which is the time during which players are not performing their own
-actions but monitoring their partner's actions.
-
-In order to evaluate \gModel{A}{B} during the task, we used two questionnaires
-(Fig.~\ref{study1:questionnaires}) that were displayed during each of the three
-games, as a transparent layer appearing over the game display. The first
-questionnaire concerned the player's intended actions. The second questionnaire
-asked each player about what he thought her or his partner was intending to do.
-Some answers were identical in both questionnaires (like ``adjusting a shot'')
-while others were reversed (``to guide my partner'' versus ``to guide me''). We
-chose a subjective measure of accuracy ($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$)
-rather than an objective measure (\ie the model \gmodel{A}{B} is compared to $B$'s
-next action) because some of the activities proposed by the questionnaire were not
-observable by the environment (\eg establishing a strategy). We calculated
-\Model{A}{B}{activity_{game_{i}}}) as the number of common answers between
-questionnaires \gmodel{A}{A} and \gmodel{A}{B} in each game and computed the
-average value a across the 3 levels.
-
-\begin{figure}[ht!]
-        \centering
-        \includegraphics[width=0.7\columnwidth]{image5.png}
-        \caption{\gmodel{A}{A} and \gmodel{A}{B} questionnaires in SpaceMiners
-        (translated from French).}
-
-        \label{study1:questionnaires}
-\end{figure}
-
-\subsubsection*{Results}
-
-\paragraph{Grounding criterion} The grounding criterion was high: the
-correlation between \gModel{A}{B} and task performance was 0.42, $p = 0.05$.
-Pairs with an accurate mutual model reached higher scores. A regression analysis
-confirmed the positive and significant relation between group performance and
-mutual modelling accuracy ($\beta=54, p = 0.02$).
-
-\paragraph{Study-specific questions} The awareness tool permitted higher group
-performance, but it did not improve the accuracy of the mutual model. Since
-teams were free to use the awareness tool or not (the \emph{scout} mode), we
-performed a post-hoc split of players depending on how much time they used it.
-The split point was the mean of time spent in the scout mode and it led to the
-constitution of two groups made up of 12 individuals ``short time in scout
-mode'' and 24 individuals ``long time in scout mode''. A two-way analysis of
-variance conducted on these contrasted groups revealed that pairs in the
-awareness condition who spent more time in the scout mode reached higher levels
-of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
-does not support a causal direction. An alternative explanation could be that
-good modellers are more social and hence appreciate the awareness tool.
-
-\paragraph{Symmetry question} We computed intra-class correlation as described
-by~\citet{kenny1998data} from the answers to the cross-questionnaires.
-Considering all pairs in both conditions, we found a positive and significant
-correlation ($r = 0.38, p < 0.05$) between \gModel{A}{B} and \gModel{B}{A}.
-Interestingly, this was higher in the control group ($r = 0.44$) than in the
-experimental group ($r = 0.24$). Actually, $\Delta(\gMdeg{A}{B},\gMdeg{B}{A})$,
-\ie the absolute differences between the models accuracy, was not significantly
-different with or without the awareness tool ($F [1,13]= 0.144, p > 0.5$). This
-result could be explained by the fact that the players without awareness tools
-communicated more.
-
-\subsubsection*{Discussion}
-
-How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
-If it was null, modelling would be interpreted as an individual process that
-varies according to personality traits or cognitive skills: some persons tend to
-model more accurately their peer while others don't care or can simply not. If
-the correlation was close to 1, we could conclude that the modelling accuracy is
-an emergent property of teams since there would be low probability that, by
-chance, low accurate modellers are paired with accurate modellers and
-conversely. Our correlation is between these two extremes but the very notion of
-significant intra-class correlation tend to indicate that one may not consider
-peers as independent subjects, which is an argument for the \emph{interaction}
-hypothesis $\mathcal{H}_{3}$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\bf Study 2}: Effect of an awareness tool on \gModel{A}{B}  in a pervasive
-game}
-
-This study concerns a collaborative game that occurred in the physical
-space~\citep{nova2006underwhelming}. We studied whether players build an accurate
-model of the path followed by their partners, assuming this path reflected their
-problem solving strategy. We used an objective measure of \gModel{A}{B}: the
-distance between where $A$ believes $B$ has been walking and where $B$ actually went.
-The main hypothesis concerned the effect of awareness tools on group performance
-and on \gModel{A}{B}.
-
-\subsubsection*{Experimental setting}
-
-{\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
-joint task. The game was played on EPFL campus. Participants had to find a
-virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
-Players used a Tablet PC that displayed a map of the campus and an indication of
-their personal distance to Bob. Their annotations on the map were shared with
-the two other players ($A$ could see what $B$ and $C$ wrote). These annotations
-faded out after a few minutes to avoid covering the full display. The awareness
-tool displayed as well the location of the two other players on the map. This
-resulted in three conditions: the control condition (without tool) and two
-experimental conditions: synchronous awareness (display of the current position
-of each player) and asynchronous awareness (display of current position of each
-player as well as their recent spatial trace).
-
-\begin{figure*}[h!t]
-        \centering
-        \begin{subfigure}{.4\textwidth}
-            \includegraphics[width=\linewidth]{image6.png}
-            \caption{A's drawing of $B$'s path.}
-        \end{subfigure}
-        \begin{subfigure}{.4\textwidth}
-            \includegraphics[width=\linewidth]{image7.png}
-            \caption{Actual path followed by $B$.}
-        \end{subfigure}
-        \caption{Reported and actual path of one of the player, during the {\sc
-        CatchBob} game.}
-        \label{study2:paths}
-\end{figure*}
-
-\subsubsection*{Subjects}
-
-Ninety EPFL students participated in this experiment. We only selected EPFL
-students since knowledge of the campus geography had an impact both on group
-performance and on mutual modelling: to represent the path of someone across
-some space is difficult without an \textit{a priori} mental map of this space.
-We formed groups of students who knew each other. We assigned 10 triads to each
-of our three experimental conditions. Each condition was made up of
-approximately 25\% of women, but we did not control gender repartition within
-each triad.
-
-\subsubsection*{Variables}
-
-The independent variable was the presence and role of the awareness tool. As a
-dependent variable, we had the task performance which was the distance covered
-by the team to catch Bob and \gModel{A}{B}. To estimate \gModel{A}{B}, we asked
-players to draw on paper their own path and the path of each of their partners
-after the game. This enabled us to calculate the number of errors players made
-while drawing the path of their partners. We compared the path that player $A$
-attributed to $B$ with $B$'s real path recorded by the system and the same for
-$A$ \& $C$ and $B$ \& $C$ as depicted on Fig.~\ref{study2:paths}.
-
-\gModel{A}{B} is the sum of errors made by $A$ about $B$'s paths. An error was
-either drawing a place where the partner had not been or not drawing a place
-where he/she had gone. One could argue that \gModel{A}{B} is biased by the
-subjects' ability to translate their trajectories souvenir into a map drawing.
-However, 85\% of subjects made no mistake at all when drawing their own path. We
-therefore consider mistakes in their partners' path as being due to a lack of
-mutual modelling accuracy instead of being due to spatial reasoning skills.
-
-\subsubsection*{Results}
-
-\paragraph{Grounding criterion} The correlation between \gModel{A}{B} and the
-task performance (path lengths) was low: 0.15. Using a \emph{post-hoc} split on
-\gModel{A}{B}, there is no significant difference between the performance of the
-groups with high and low \gModel{A}{B}  ($F = 1.45, p = 0.24$). Conversely, a
-\emph{post-hoc} split of the groups according to their performance did not show
-any significant differences on \gModel{A}{B} ($F = 1.16, p = 0.29$).
-
-\paragraph{Study-specific questions} There was no significant difference
-regarding the task performance. However, and surprisingly, the absence of the
-awareness tool was related to a higher \gModel{A}{B}: players tended to better
-remember their partners' paths when they could not constantly monitor their
-positions. As detailed in the original study~\citep{nova2005location}, it
-appeared that teams without awareness tool made more manual annotations on the
-map while permanent monitoring has an underwhelming effect.
-
-\paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
-\gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more $A$
-makes errors about $B$, the more $B$ does as well.
-
-\paragraph{Triangle questions} Regarding $\Delta_2$, the correlation between
-\gModel{A}{C} and \gModel{B}{C} is significant: $r=0.43, p <.001$. Concerning
-$\Delta_3$, the correlation between \gModel{A}{B} and \gModel{A}{C} is
-significant as well: $r=0.30, p <.01$.
-
-\subsubsection*{Discussion}
-
-This positive correlation observed in the symmetry question is surprising as
-well: when acknowledging the high heterogeneity of spatial skills among
-adults~\citep{liben1981spatial}, a significant positive correlation would
-indicate that mutual modelling indeed stems from group processes more than
-individual processes. The significant results on the triangle questions goes in
-the same direction: either both the modeler and the modellee contribute to
-accurate modelling, or the quality of teamwork has a global effect. Since team
-members did not have many opportunities to directly interact during this
-experiment, the factor that facilitates accurate mutual modelling is probably
-more the overall group strategy than the quality of verbal interactions: a clear
-strategy facilitates memorizing one's partner path.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\bf Study 3}:  Effect of media richness on \gModel{A}{B} in argumentation}
-
-The aim of this unpublished\footnote{The data and statistical analyses of this
-study are available online:
-\url{https://github.com/chili-epfl/mutual-modelling-adoption-homosexuality-study}.}
-study was to evaluate the effect of media richness on \Model{A}{B}{emotions}.
-The hypothesis was that video communication would lead to a better \gModel{A}{B}
-than audio only since emotions often impact on facial expressions.
-
-\begin{figure*}[ht!]
-        \centering
-        \includegraphics[width=0.8\textwidth]{image8.jpg}
-        \caption{Example of argumentation graph}
-        \label{study3:argumentation_graph}
-\end{figure*}
-
-\subsubsection*{Experimental settings}
-
-Triads had to address an emotional societal debate: authorizing or not adoption
-by homosexual couples. They worked on-line and had to structure their
-argumentation with the shared concept map tool {\sc TeamWave} as illustrated in
-Fig.~\ref{study3:argumentation_graph}. Ten groups had only an audio connection
-while ten groups had audio and video. The video communication was provided by a
-webcam and the software {\sc IVisit}. For the audio link, we used microphones,
-headsets and the {\sc BattleCom} software. In the \emph{audio plus video}
-condition, the screen was divided in three sections. The main part was devoted
-to the concept map window, and the images of the two peers appeared next to it.
-In the audio condition, this video zone was left empty so that the size of the
-concept map was equal in each condition. The subjects were located in the same
-room, separated by mobile walls. Despite their headsets, non-verbal audio cues
-(\eg tapping the floor with feet) were possibly heard by the participants. The
-task lasted in average 61 minutes.
-
-\subsubsection*{Subjects}
-
-Sixty students (twenty triads) from the University of Geneva participated to
-this experiment (36 women and 24 men). We formed groups of subjects who knew
-each other: the task required the discussion of sensitive issues which required
-to feel quite comfortable with peers. Since groups were formed \textit{a
-priori}, we did not balance gender in each condition.
-
-\subsubsection*{Variables}
-
-The independent variable was the presence or not of a video link. The dependent
-variable \gModel{A}{B} was measured subjectively from three questionnaires: in
-the first one, $A$ described his/her own emotions \gmodel{A}{A}, while in the
-two other questionnaire, $A$ described $B$'s and $C$'s emotions. The questionnaire
-included 18 items (7-point Likert Scale) describing emotions labeled as
-adjectives: \emph{anxious}, \emph{enthusiastic}, \emph{agitated}, \emph{proud},
-\emph{excited}, \emph{quiet}, \emph{calm}, \emph{stressed}, \emph{bored},
-\emph{upset}, \emph{relaxed}, \emph{irritated}, \emph{determined},
-\emph{hostile}, \emph{active}, etc. \gmodel{A}{B} was modelled as a vector of 18
-numerical values corresponding to their answers on each questionnaire items, and
-\gModel{A}{B} was computed as the distance between the two vectors
-\gmodel{A}{B} and \gmodel{B}{B}:
-
-\[
-    \Mdeg{A}{B}{emotions} = \frac{\displaystyle\sum_{emotions} \abs{\M{A}{B}{e} -
-    \M{B}{B}{e}}}{18}
-\]
-
-The smaller the score, the \emph{more} accurate the model.
-
-\subsubsection*{Results}
-
-\paragraph{Grounding criterion} The maps produced by teams were ranked by three
-independent judges on completeness and structure quality (Kendall's $W=0.474$,
-limited agreement). We used the average rank as a estimation of the team
-performance and correlated it with the average of the 6 values of \gModel{A}{B}
-per team (\gModel{A}{B}, \gModel{A}{C}, \gModel{B}{A},...). The correlation is
--0.22: teams with a good \gModel{A}{B} tend to be better ranked.
-
-\paragraph{Study-specific questions} Our hypothesis about media richness is
-rejected: the average score for \model{A}{B}{emotions} was 1.25 ($SD = 0.53$) in
-the audio+video condition and 1.09 ($SD = 0.41$) in the audio alone condition
-($t=1.89, df=111, p = 0.062$): the average degree of accuracy of
-\model{A}{B}{emotions} was hence \emph{higher} in the audio alone condition.
-
-\paragraph{Symmetry question} We computed the absolute differences $\Delta_1$
-between \gModel{A}{B} and \gModel{B}{A} over all pair of subjects within a triad
-(3 values per triad, 20 triads), and compared them with the same differences
-computed from random gradings (following the same grade distribution as for the
-experimental data).
-
-A t-test on the two sets releaved a significantly lower average difference in
-the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
-$t=-3.3, df=60, p = 0.0016$), supporting the symmetry hypothesis.
-
-\paragraph{Triangle questions} $\Delta_2$ is computed in a similar way as the average of
-the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
-subjects. The average difference is 0.42 ($SD= 0.34$), and if not significantly
-different from the same data obtained from random gradings ($t=-1.61, df=60,
-p=0.11$).
-
-For $\Delta_3$, the average absolute difference between \gModel{C}{A} and
-\gModel{C}{B} over the 60 subjects is 0.40 ($SD= 0.41$) and is significantly
-lower than chance ($t=-2.62, df=60, p=0.01$): a given subject tends to model its
-two partners with similar degrees of accuracy, which supports $\mathcal{H}_{1}$.
-
-\begin{figure*}[ht!]
-        \centering
-        \includegraphics[width=0.9\textwidth]{image9.pdf}
-        \caption{Average values of \Model{A}{B}{X} where $X$ is one of the proposed
-        emotions (max = 7).}
-        \label{study3:deg_m_values}
-\end{figure*}
-
-\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
-between self and social accuracy here because we do not have an estimation of
-self-accuracy: subjects describe their emotions but we have no way to check if
-there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
-we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
-according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
-errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
-is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
-emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
-0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
-interpret this too far, but to show that there are large variations even within one
-area  (perceiving emotions).
-
-\subsubsection*{Discussion}
-
-The results evidence that richer media does not automatically foster better
-modelling, on the contrary: groups using only the audio modality to exchange
-their views were able to build more accurate models of each others than groups
-using both video and audio, which is especially surprising when modelling
-emotions: one would expect that video feedback indeed help building a better
-model of the emotions of the other. Our interpretation is that viewing what
-one's partner sees (shared graphical editor) was, in this task, more important
-than seeing each other~\citep{gaver1993one,anderson1997impact}.
-
-In this study, the accuracy of mutual modelling between two peers tend to be
-symmetrical, which would support the idea that the quality of modelling is primarily
-building upon the social interaction. This is however not fully supported by the
-analysis of $\Delta_2$: $A$ being good at modelling $C$ does not correlate with
-$B$ being good at modelling $C$.
-
-On the other hand, $\Delta_3$ shows that a given $A$ tends to
-model $B$ and $C$ at similar levels of accuracy: this evidences an individual
-skill at modelling others (hypothesis $\mathcal{H}_{1}$).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\bf Study 4}:  Effects of a script on \gModel{A}{B}  in concept mapping}
-
-This study investigated the effect of a collaboration script on collaborative
-learning~\citep{molinari2008effects}. The script chosen is a {\sc jigsaw}: two students
-receive different but complementary subsets of the knowledge (texts) which have
-to be integrated to build a shared concept map.  This script increases the
-cognitive effort to build the map, not only to conciliate the viewpoints of
-each team members but, before that, to find out what the other knows.
-
-\subsubsection*{Experimental settings}
-
-The instructional material consisted of an explanatory text about the
-neurophysiologic phenomenon of \emph{action potential}. The text was divided
-into 3 chapters.  In the \emph{same information} (SI) condition, the same text
-was given to both partners. In the \emph{complementary information} (CI)
-condition, it was divided into two sub-texts, one about the electrical processes
-of the neuron while the second one about the chemical processes. These two
-versions were equivalent in terms of number of information pieces.
-
-The peers were located in two rooms equipped with the same
-computer.  The experimental session lasted around 90 minutes and consisted of 6
-phases: Participants used two software components, {\sc CmapTools} and {\sc
-TeamSpeak}.
-
-\begin{enumerate}
-
-    \item As a pre-test, participants were asked to write down all they knew
-        about the neuron and its functioning (5 minutes),
-
-    \item Participants were instructed to read a text (12 minutes),
-
-    \item Participants were asked to build individually a concept map in order to
-        graphically represent what they learnt from the text (10 minutes),
-
-    \item Dyads had to built a concept map during 20 minutes, communicating by
-        audio.  The screen layout was structured into three areas
-        (Fig.~\ref{study4:concept_map}),
-
-    \item Participants were invited to individually complete a knowledge test
-        (15-20 minutes),
-
-    \item Participants where asked to estimate their own- and their partner's
-        final knowledge in a questionnaire.
-
-\end{enumerate}
-
-\begin{figure*}
-    \centering
-    \begin{tikzpicture}[scale=0.8]
-        \node[anchor=south west,inner sep=0] at (0,0)
-            {\includegraphics[width=0.75\textwidth]{image10.png}};
-
-            \node[rectangle,draw, thick] at (10,5.4) {Self map};
-            \node[rectangle,draw, thick] at (7.5,3.3) {Partner map};
-            \node[rectangle,draw, thick] at (2.5,1.5) {Collaborative map};
-    \end{tikzpicture}
-    \caption{The group concept map and the individual concept maps in study
-    4}
-    \label{study4:concept_map}
-\end{figure*}
-
-\subsubsection*{Subjects}
-
-Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
-were remunerated for participation. Dyads were randomly assigned to one of the
-two experimental conditions. Gender was balanced over the conditions.
-Participants did not know each other before the experiment. Students from Life
-Sciences faculty were not recruited to avoid high initial background knowledge on the
-learning domain.
-
-\subsubsection*{Variables}
-
-The independent variable, script versus no-script, was implemented by the
-difference of texts that individuals had to read.  The dependent variables were
-the post-test scores, used to assess \Model{A}{B}{knowledge}. In phase 6,
-participants were ask to estimate (7-point Likert scale) their own and their
-partner's outcome knowledge with respect to each chapter of the learning
-material. The order of questions about oneself and about the other was
-counterbalanced across participants. We measure \gModel{A}{B} in an objective
-(Eq.~\ref{eq:study4.1} and \ref{eq:study4.2}) as well as in a subjective way
-(Eq.~\ref{eq:study4.3}).
-
-\begin{multline} \label{eq:study4.1}
-    \Mdeg{A}{B}{knowledge} = 
-        \sum_{i=1}^{3} (\M{A}{B}{chapter_i} - score(B, chapter_i))
-\end{multline}
-
-\begin{multline} \label{eq:study4.2}
-    \Mdeg{A}{A}{knowledge} = 
-        \sum_{i=1}^{3}  (\M{A}{A}{chapter_i} - score(A, chapter_i))
-\end{multline}
-
-\begin{multline} \label{eq:study4.3}
-    \Mdeg{A}{B}{knowledge} = 
-        \sum_{i=1}^{3}  (\M{A}{B}{chapter_i} - \M{B}{A}{chapter_i})
-\end{multline}
-
-\subsubsection*{Results}
-
-\paragraph{Grounding criterion} In this task, the grounding criterion was low.
-We did not however evaluate task performance (\eg the quality of the jointly
-produced concept map) but learning gains. The correlation between \gModel{A}{B}
-and $A$'s learning gains is not significant ($\beta = 0.08, ns, N = 60$). It is
-also not significant within each condition.
-
-\paragraph{Study-specific questions} We performed a non-parametric Mann-Whitney
-test on post-test scores for electric questions and a one-way ANOVA on scores
-for ionic questions (Levene tests for homogeneity of variances): $p = 0.02$ and
-$ns$, respectively. Results  did not show any significant difference between the
-\emph{same information} (SI) condition and the \emph{complementary information}
-(CI) condition, neither for electric questions ($U = 388.50, z = -0.88, ns$) nor
-for ionic questions ($F(1, 58) = 0.17, ns$).  The effect of scripts on
-\gModel{A}{B} was not significant ($F(1, 58) = 0.78, ns$) when considering the
-absolute difference between \gModel{A}{B} and $B$'s post-test score. However, $A$
-tended underestimate $B$'s score in the SI condition ($M = -2.06$) and to
-overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<0.01$).
-Regarding to \gModel{A}{A}, there was no significant difference between
-conditions.
-
-\paragraph{Symmetry question} The inter-class correlation between \gModel{A}{B}
-and \gModel{B}{A} is approaching significance ($r= 0.26, F(1,29)=1.71, p=
-0.075$). It is indeed significant when students read the same text (SI
-condition: $r=0.43, F(1,15)=2.53, p<0.05$) but not when they read different texts
-($r= 0.13, F(1,12)=1.3, ns$).
-
-\paragraph{Rectangle question} The correlation between \gmodel{A}{A}
-and \gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
-significant in each condition: someone good at self-modelling is not necessarily
-good at modelling someone else and vice-versa.
-
-\subsubsection*{Discussion}
-
-In this experiment, the symmetry of mutual modelling is only found when subjects
-are in conditions of symmetry of information. This consolidates the hypothesis
-of \gModel{A}{B} being a property of interactions in context rather than $A$'s
-cognitive attitude.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{{\bf Study 5}: \Model{A}{B}{knowledge} in concept mapping}
-
-This study investigates if \Model{A}{B}{knowledge} is related to learning
-outcomes by comparing teams with or without a knowledge awareness tool (KAT),
-\ie a tool that informs $A$ about $B$'s knowledge as measured through a pre-test.
-
-\subsubsection*{Experimental setting}
-
-The peers were located into two different rooms. A complete description of the
-study is provided in~\citet{sangin2008learners}. The experiment lasted 90 minutes.
-
-It started with the same two first steps as in study 4, followed by:
-
-\begin{enumerate}
-    \setcounter{enumi}{2}
-
-    \item Subjects passed a pre-test, with ten questions per chapter.
-
-    \item Participants had 20 minutes to draw a collaborative concept map
-        reporting the content of the texts. They were able to communicate
-        orally through headsets.  We used Tobbii eye tracking devices to
-        record their gazes.
-
-    \item The post-test included the same items than the pre-test but in a
-        different order.
-
-    \item  Finally, participants were asked to estimate their partner's
-        knowledge at the post-test for each of the three chapters on a 7-point
-        Likert-like survey.
-
-\end{enumerate}
-
-\begin{figure*}[h!t]
-        \centering
-        \begin{subfigure}{.6\textwidth}
-            \includegraphics[width=\linewidth]{study5-conceptmap.png}
-        \end{subfigure} \\
-        \begin{subfigure}{.7\textwidth}
-            \includegraphics[width=\linewidth]{study5-kat.png}
-        \end{subfigure}
-        \caption{Screenshot of the KAT condition during the concept-map building
-        phase.}
-        \label{study5:kat}
-\end{figure*}
-
-\subsubsection*{Subjects}
-
-Sixty-four first year EPFL students (46 men, 18 women, mean age: 21.2)
-participated to the study. They were randomly assigned to conditions and
-did not know each other before. Like in Study 4, students from Life Science
-faculty were excluded.
-
-\subsubsection*{Variables}
-
-The participants of the experimental condition group were provided with the KAT
-on the bottom part of the screen (Fig.~\ref{study5:kat}): each line represents
-the score obtained by the partner at the pre-test for a chapter. Participants
-did not see they own score, but they usually started their discussion by
-exchanging this information.
-
-\subsubsection*{Results}
-
-\paragraph{Grounding criterion} A linear regression revealed a positive relation
-between \gModel{A}{B} and the learning gain of the pair $\{A, B\}$ (calculated
-as the average of the individual learning gains): $\beta= 0.401, p < 0.001, r2adj.
-= 0.15$, large effect. This relationship was not significant in the previous
-experiment which as been conducted with the same task (using the same task
-should produce the same grounding criterion). This is probably due to the fact that
-\gModel{A}{B} was influenced by the KAT.
-
-\paragraph{Study-specific questions} The $t$-test reported a significant
-difference between the KAT condition participants ($M = 13.4\%$) and the control
-group [$M = 3.6\%, t(1, 60) = 2.73, p < 0.01$, Cohen's $d = 0.7$, medium to large
-effect]: providing learners with cues about the prior-knowledge of their partner
-enhances their collaborative learning. As a treatment check, we found a positive
-and significant correlation between the amount of gazes on KAT (using eye
-tracking devices) and the learning gains ($r(22) = 0.54, p = 0.01$). A detailed
-analysis revealed that subjects look at KAT to assess their peer's credibility
-when he/she provided new information.
-
-The KAT has a significant effect on \gModel{A}{B}: peers more accurately
-estimated their partners knowledge ($M = 1.11$) than those in the control
-condition [$M = 0.98, t(1, 60) = 3.19, p < 0.01$, Cohen's $d = 0.83$, large
-effect]. This is a trivial result since the KAT provided them with an initial
-\gmodel{A}{B}. However, subjects have to predict the post-test score while the
-KAT informed them about the pre-test score. Actually, pairs in the KAT condition
-produced significantly more instances of 3 specific categories of interactions:
-(1) elaborated utterances with rich contents,  (2) utterances asking about the
-other's knowledge such as ``Did you understand how transmission works?'' (3)
-utterances describing one's own knowledge (\gmodel{A}{A}) such as ``I don't
-remember the Ranvier's thing...''. These 3 categories provide different account
-of \gModel{A}{B}: (1) as an effect of the quality of interaction, (2) as an
-effect of $A$'s effort to model $B$ or (3) as $B$'s effort to give cues to $A$ about his
-own knowledge.
-
-We examined \gModel{A}{B} as potentially mediating the effect of the KAT factor
-on the relative learning gain. A linear regression confirmed that \gModel{A}{B}
-was significantly related to the KAT-factor ($\beta= 0.381, p < 0.01, r2 = 0.15$).
-The KAT-factor was also significantly and positively related to the RLG ($\beta=
-.332, p < 0.01, r2 = 0.11$). We then tested the relation between the independent
-variable (KAT) and the dependent variable (gains) when controlling for the
-mediating variable (\gModel{A}{B}). A multiple regression showed that the
-KAT-factor was no longer a significant predictor ($\beta= 0.210; p = ns$) whereas
-the \gModel{A}{B} was still a significant predictor ($\beta= 0.32, p < 0.01$).
-Thus, it can be concluded that \gModel{A}{B} mediated the KAT-factor's effect on
-the learners' RLG. The Sobel significance test for indirect effects was
-significant [$z = 1.99, p < 0.05$].
-
-\paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
-0.05, ns$).
-
-%\paragraph{Rectangle question} $\Delta_4$: the correlation between \gmodel{A}{A} and \gmodel{A}{B} is
-%\fixme{???  MIRWEIS A FAIRE}.
-
-\subsubsection*{Discussion}
-
-Despite the fact that the learning task was the same as in study 4, the
-conditions of collaboration (viewing multiple maps or not) and the conditions
-(scripted or not, awareness tool or not) probably explain differences in terms
-of mutual modelling.
-
-\section{Synthesis}
-
-Out of 5 studies, 3 revealed a correlation between the model peers built about
-each other: when $A$ builds an accurate model of $B$, $B$ also tends to build an
-accurate model of A.  In one study, this was the case only in one condition.
-The conclusion is hence that modelling is mutual but only in certain conditions,
-summarized in Table~\ref{synthesis_table}.
-
-\begin{table*}[ht!]
-\centering
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
-\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
-\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
-\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
-\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
-\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
-\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
-\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
-\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
-\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
-\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
-$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
-$\mathcal{H}_2$               &                                  & supported                       & supported                 &                           &                              \\
-$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
-\end{tabular}
-}
-\caption{Comparison of the studies}
-\label{synthesis_table}
-\end{table*}
-
-Coming back to our initial hypotheses, the $\Delta_2$ results in study 2
-reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
-hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
-is not simply the result of an individual cognitive effort but overall the
-side-effect of the quality of interactions. The \emph{second-order-modelling
-hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
-$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
-good modeler would somehow drive the modelling of his or her partner.  This
-\emph{drive} will however occur through verbal cues and repairs and is hence not
-easy to differentiate from $\mathcal{H}_{3}$. The contradictory $\Delta_2$
-analysis in study 3 and $\Delta_1$ results in studies 4 (to some extent) and 5
-weaken the hypothesis.
-
-The triangular questions lead to similar contradictions. The $\Delta_2$
-correlation between the \gmodel{A}{C} and \gmodel{B}{C} (study 2)
-support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
-be no reason for modelling skills of $A$ and $B$ to be correlated.  The role of $C$ in
-\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
-the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (study
-3) question $\mathcal{H}_{3}$: if the role of the modellee was so
-important, it would be surprising to have by chance a correlation between
-$B$'s and $C$' ability to drive $A$'s modelling.
-
-In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
-($\Delta_4$ in study 4 and study 5) could also contributes to the idea that $A$'s
-modelling skills is not the key factor. However, between modelling someone and
-modelling oneself, \ie  metacognition, they are so many differences, that we
-would hardly conclude from that.
-
-Among these complex results, it is important to stress that, for $\Delta_2$, the
-significant correlations are all below 0.50, not around 0.90.  So even if all
-$\Delta_2$ correlations were significant in all experiments, we would still not
-conclude that modelling purely reflects the quality of social interactions.
-There is obviously a large part of individual variance; some of us are much
-better in modelling peers. This is not new. The contribution of these studies is
-rather to establish the importance of team processes on something that a priori
-seemed to be more of a personal cognitive skill.  In other words, we stress that
-the bottle is half-full.
-
-These conclusions must be presented with multiple disclaimers. First, they
-heavily rely on correlations; hence we cannot identify causal links. Second, we
-faced hard methodological issues. Providing learners with on-task questionnaires
-introduce a bias: they will pay more attention to their partners in the
-remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
-and rationalization biases. The nature of mutual modelling implies methodological
-challenges that call for new  measurement methods. We have promising results for
-using eye tracking methods to address this challenge.
-
-Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
-imply we have a mechanical view of modelling. It's just a language for referring
-to different components of modelling each other in teams.  This simple notation
-emerges from the need to compare five mutual modelling studies; it did not
-generate a consistent program of consistent studies.  Our results are not very
-robust because they emerge from a \emph{post-hoc} reinterpretation of studies
-that addressed different research questions (media richness, awareness tools,
-scripts,...). This diversity makes our results difficult to integrate, partly
-contradictory. Nonetheless, this diversity gives some generalizability: mutual
-modelling has been investigated in different contexts (virtual space versus real
-space), with different groups sizes (pairs and triads) and different tasks.
-Moreover, these results have been found using different methods for measuring
-modelling: on-task in study 1 versus off-task in study 2, subjective validation
-in study 1 versus objective validation in study 2 (comparing $A$'s model with $B$'s
-behavior).
-
-\begin{acknowledgements}
-
-The experiments were developed with the help of Mirweis Sangin, René Glaus,
-Patrick Jermann,  Fabien Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan
-Bourquin and Jeremy Goslin.  We also thank Kshitij Sharma and {\L}ukasz
-Kidzi\'nski for their help with data analysis. The main funding has been
-provided from a NSF Grant grant \#102511-106940.
-
-\end{acknowledgements}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{spbasic}
-\bibliography{biblio}
-
+\subsection*{Corresponding author}
+Séverin Lemaignan \texttt{severin.lemaignan@epfl.ch}
 \end{document}

### 2015-06-18 mutual-modeling.tex
Added keywords
Prepa for submission: anonymization + page with authors' information

@@
 
-\author{
-    Pierre Dillenbourg
-    \and
-    Séverin Lemaignan
-    \and
-    Nicolas Nova
-    \and
-    Gaëlle Molinari
-}
-\institute{
-    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
-    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
-    \email{firstname.lastname@epfl.ch}
-    \and
-    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
-    \and
-    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
-    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
-}
+% ANON
+%\author{
+%    Pierre Dillenbourg
+%    \and
+%    Séverin Lemaignan
+%    \and
+%    Nicolas Nova
+%    \and
+%    Gaëlle Molinari
+%}
+%\institute{
+%    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
+%    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
+%    \email{firstname.lastname@epfl.ch}
+%    \and
+%    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
+%    \and
+%    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
+%    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
+%}
+\author{}
 \begin{document}
@@
 \end{abstract}
+
+\keywords{Cognitive Modelling; Grounding; Theory of Mind}
 
@@
 partner when they know that they can later provide feedback on his/her
-understanding~\citep{schober1993spatial}. We introduced the notion of
-\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
+%ANON
+%understanding~\citep{schober1993spatial}. We introduced the notion of
+%\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
+understanding~\citep{schober1993spatial}. \citet{dillenbourg1995evolution} introduced the notion of
+\emph{optimal collaborative effort} to stress
 that misunderstanding should not be viewed as something to be avoided (if this
@@
 We studied the impact of an awareness tool on group performance and mutual
-modelling~\citep{nova2007collaboration}. The availability
+%ANON
+%modelling~\citep{nova2007collaboration}. The availability
+modelling \textit{[citation removed for blind review]}. The availability
 of an awareness tool was our independent variable. In previous studies, we
@@
 This study concerns a collaborative game that occurred in the physical
-space~\citep{nova2006underwhelming}. We studied whether players build an accurate
+%ANON
+%space~\citep{nova2006underwhelming}. We studied whether players build an accurate
+space \textit{[citation removed for blind review]}. We studied whether players build an accurate
 model of the path followed by their partners, assuming this path reflected their
@@
 {\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
-joint task. The game was played on EPFL campus. Participants had to find a
+% ANON
+%joint task. The game was played on EPFL campus. Participants had to find a
+joint task. The game was played on our campus. Participants had to find a
 virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
@@
 \subsubsection*{Subjects}
-
-Ninety EPFL students participated in this experiment. We only selected EPFL
-students since knowledge of the campus geography had an impact both on group
+%ANON
+%Ninety EPFL students participated in this experiment. We only selected EPFL
+%students since knowledge of the campus geography had an impact both on group
+Ninety students participated in this experiment. We only selected students from
+our campus since knowledge of the campus geography had an impact both on group
 performance and on mutual modelling: to represent the path of someone across
@@
 remember their partners' paths when they could not constantly monitor their
-positions. As detailed in the original study~\citep{nova2005location}, it
+%ANON
+%positions. As detailed in the original study~\citep{nova2005location}, it
+positions. As detailed in the original study, it
 appeared that teams without awareness tool made more manual annotations on the
@@
 
-The aim of this unpublished\footnote{The data and statistical analyses of this
-study are available online:
-\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.}
+%ANON
+%The aim of this unpublished\footnote{The data and statistical analyses of this
+%study are available online:
+%
+%\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.}
+The aim of this unpublished
 study was to evaluate the effect of media richness on \Model{A}{B}{emotions}.
@@
 This study investigated the effect of a collaboration script on collaborative
-learning~\citep{molinari2008effects}. The script chosen is a {\sc jigsaw}: two students
+%ANON
+%learning~\citep{molinari2008effects}. The script chosen is a {\sc jigsaw}: two students
+learning \textit{[citation removed for blind review]}. The script chosen is a {\sc jigsaw}: two students
 receive different but complementary subsets of the knowledge (texts) which have
@@
 
-            \node[rectangle,draw, thick] at (8,4.3) {Self map};
-            \node[rectangle,draw, thick] at (6,2.6) {Partner map};
-            \node[rectangle,draw, thick] at (2.5,1.5) {Collaborative map};
+            \node[rectangle,draw, thick] at (8,4) {Self map};
+            \node[rectangle,draw, thick] at (6,2.2) {Partner map};
+            \node[rectangle,draw, thick] at (2.3,1.3) {Collaborative map};
     \end{tikzpicture}
@@
 
-Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
+%ANON
+%Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
+Fifty-eight first year students (47 men and 11 women, mean age: 20.46)
 were remunerated for participation. Dyads were randomly assigned to one of the
@@
 The peers were located into two different rooms. A complete description of the
-study is provided in~\citet{sangin2008learners}. The experiment lasted 90 minutes.
+%ANON
+%study is provided in~\citet{sangin2008learners}. The experiment lasted 90 minutes.
+study is provided in \textit{[citation removed for blind review]}. The experiment lasted 90 minutes.
 
@@
 \subsubsection*{Subjects}
-
-Sixty-four first year EPFL students (46 men, 18 women, mean age: 21.2)
+%ANON
+%Sixty-four first year EPFL students (46 men, 18 women, mean age: 21.2)
+Sixty-four first year students (46 men, 18 women, mean age: 21.2)
 participated to the study. They were randomly assigned to conditions and
@@
 
-The experiments were developed with the help of Mirweis Sangin, René Glaus,
-Patrick Jermann,  Fabien Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan
-Bourquin and Jeremy Goslin.  We also thank Kshitij Sharma and {\L}ukasz
-Kidzi\'nski for their help with data analysis. The main funding has been
-provided from a NSF Grant grant \#102511-106940.
+%ANON
+%The experiments were developed with the help of Mirweis Sangin, René Glaus,
+%Patrick Jermann,  Fabien Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan
+%Bourquin and Jeremy Goslin.  We also thank Kshitij Sharma and {\L}ukasz
+%Kidzi\'nski for their help with data analysis. The main funding has been
+%provided from a NSF Grant grant \#102511-106940.
+
+\textit{Removed for blind review.}
 

### 2016-01-25 mutual-modeling.tex
Added back Mirweis
Typos reported by reviewers

@@
 %    \and
+%    Mirweis Sangin
+%    \and
 %    Nicolas Nova
@@
 %    \and
+%    Mirweis Sangin \at Use-able consulting \\ CH-1004 Lausanne, Switzerland\\\email{mirweis@gmail.com}
+%    \and
 %    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
@@
 From its very beginning, CSCL research has been following
-\citet{roschelle1995construction} suggestion that collaborative learning has
+the suggestion by \citet{roschelle1995construction} that collaborative learning has
 something to do with the process of constructing and maintaining a \emph{shared
@@
 mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
-construction of a shared understanding has been investigated for several years
+construction of a shared understanding has been investigated for several decades
 in psycholinguistics, under the  notion of
 \emph{grounding}~\citep{clark1986referring}. However, the relevance of grounding
-mechanisms for explaining learning outcomes has been questioned in learning
+mechanisms for explaining learning outcomes has been questioned in the learning
 sciences. The monitoring and repair of misunderstanding explains for instance
@@
 outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
-change?} \textbf{Does negotiation of meaning only occurs when partners monitor
+change?} \textbf{Does negotiation of meaning only occur when partners monitor
 and diagnose each other's knowledge?} If the ratio between shallow alignment and
@@
 deep grounding. Most empirical studies on grounding and alignment are conducted
-with tasks, which despite being qualified as ``ecologically valid'' by their
+with tasks that, despite being qualified as ``ecologically valid'' by their
 authors, are mere referencing tasks such as asking the way to the train station
@@
 
-The content of mutual models ranges from \emph{dispositional} versus
+The content of mutual models ranges from \emph{dispositional} to
 \emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
@@
         use awareness tools which, in principle, reduce the cost of mutual modelling, but do
-        not eliminate all costs: if the tool provide $A$ with information about
+        not eliminate all costs: if the tool provides $A$ with information about
         what $B$ does/knows, this additional information may actually increase
@@
 with three increasingly complex situations. The experiment was 2 hours long: a
-30 minutes tutorial and 3 levels of 30 minutes. Players uses using a regular
+30 minute tutorial and 3 levels of 30 minutes. Players uses using a regular
 joystick and communicated with each other through an audio channel.
@@
 modelling, on the contrary: groups using only the audio modality to exchange
-their views were able to build more accurate models of each others than groups
+their views were able to build more accurate models of each other than groups
 using both video and audio, which is especially surprising when modelling

### 2016-02-04 mutual-modeling.tex
Changed title
Updated abstract

@@
 
-\title{Mutual Modelling}
+\title{The symmetry of partner modelling}
 
@@
 \begin{abstract}
-    Collaborative learning has something to do with building a shared
-    understanding of the situation at hand. Yet, the psycholinguistics mechanisms
-    at work while establishing common grounds are complex and certainly involve
-    modelling what the partner knows/wants/intends at a given time.
-
-    Building on a set of five independent studies that where conducted over
-    several years, we propose in this article a simple model of the mechanisms
-    of mutual modelling. It integrates several concepts related to the classical
-    idea of grounding, and we apply it as an operational tool to
-    analyse the interplay of learning and mutual modelling in the five studies.
-
-    We come up with mixed experimental results that evidence the dual nature of
-    mutual modelling: a individual meta-cognitive aptitude that co-exists with
-    modelling elicited and fostered by social interactions.
+    Collaborative learning has often been associated to the construction of  a shared   understanding of the situation at hand. The psycholinguistics mechanisms
+    at work while establishing common grounds are the object of scientific controversy. We postulate that collaborative tasks require some level of mutual 
+    modeling, i.e.
+that each partner needs some model of what the other partners know/want/intend at a given time.
+We use the term 'some model' to stress the fact that this model is not necessarily details or complete, but that we acquire some representations of the persons we interact with.
+The question we address is:
+Does the quality of the partner model depend upon the modeler's ability to represent his or her partner, upon the modelee's ability to make his state clear to the modeler or upon the quality of their interactions?
+We address this question by analyzing differences between the  accuracy of the models built by different team members.
+We therefore report 5 experiments on collaborative problem solving or collaborative learning that vary in terms of tasks (how important it is to build an accurate model) and settings (how difficult it is to build an accurate model).
+In 4 studies, the accuracy of the model that A built about B was correlated with the accuracy of the model that B built about A,  which seems to imply that the quality of interactions matters more than individual abilities when building mutual models.
+However, these finding do not discard the fact individual abilities also  participate in the quality of modellng process .
 

### 2016-02-04 mutual-modeling.tex
Updated section 1

@@
 mechanisms for explaining learning outcomes has been questioned in the learning
-sciences. The monitoring and repair of misunderstanding explains for instance
-referential failures in short dialogue episodes but does hardly predict
+sciences. Grounding mechanisms  are appropriate for explains conversational events, such
+referential failures in short dialogue episodes, but they do  hardly predict deeper phenomena such as
 \emph{conceptual change} (\ie the acquisition, acceptation and integration of a
@@
 
-Paradoxically, while Clark's theory is somewhat too linguistic from a conceptual
-change viewpoint, it is criticized at the same time as being too cognitivist by
+Paradoxically, while Clark's theory is somewhat too linguistic from a learning  viewpoint, it is criticized at the same time as being too cognitivist by
 some psycholinguists, \ie as overestimating the amount of shared knowledge and
@@
 which intervention will be conducted in which way on which patient.
-%Some CSCL
-%scholars stressed the ``illusion of shared understanding'' \fixme{REF: Jacco?},
-%\ie the potential decoupling between linguistic alignment and actual shared
-%meanings.
+In this paper, we operationalized the grounding criterion, i.e.
+the necessity for accurate modelling, as the correlation between the accuracy of partner models and various measures of team performance.
 
@@
 deep grounding. Most empirical studies on grounding and alignment are conducted
-with tasks that, despite being qualified as ``ecologically valid'' by their
-authors, are mere referencing tasks such as asking the way to the train station
-or helping the peer to choose a picture among many. In this contribution, we
+with simple referencing tasks such as asking the way to the train station
+or helping the peer to choose a picture among many. In the studies we report here, we
 explore several richer tasks such as arguing on a sensitive issue or building a
@@
 \vspace{2em}
-These preliminary remarks frame the main question that this article address:
-in order to engage into shared understanding, which is a desirable learning
-situation, do partners have to build a representation of each other's knowledge?
-if so, what exactly is the nature of this mutual modelling? We adopt hereafter a
-constructive approach, first introducing a simple yet convenient model of mutual
-modelling, then analyzing five independent studies in terms of the interactions
-between learning and mutual modelling.
+In summary, the controversy around the cognitive depth of shared understanding pertains to psycholinguistic, which investigates 'natural conversations, but  in collaborative learning, the situation is different; tasks are precisely designed for requirint a deeper negotiation fo meaning.
+The craft of CSCL is to design collaborative situations that create interdependence, avoid groupthink, and allows students to detect any 'illusion of shared understanding' (Cherubini et al, 2005).
+Our question is hence not  anymore"do peers build a shared understanding ?", but becomes "when peers have to build a shared understanding, how do they do it ?".
+This question addresses the mechanisms of grounding, in which the basic sequence is: make a proposition, detect misunderstanding and, if any, repair them.
+This paper focuses on the middle part, the detection of misunderstandings.
+Detecting a misunderstanding means that the emitter of a message identifies a mismatch between his communicative intention and the way his message is understood by his or her partner.
+Detecting one peer's understanding is investigate hereafter under the general umbrella of partner modelling.
+
+Cherubini, M., Van Der Pol, J., & Dillenbourg, P.
+(2005).
+Grounding is not shared understanding: Distinguishing grounding at an utterance and knowledge level.
+In CONTEXT'05, the Fifth International and Interdisciplinary Conference on Modeling and Using Context (pp.
+11-23).
+
+%These preliminary remarks frame the main question that this article address:
+%in order to engage into shared understanding, which is a desirable learning
+%situation, do partners have to build a representation of each other's knowledge?
+%if so, what exactly is the nature of this mutual modelling? We adopt hereafter a
+%constructive approach, first introducing a simple yet convenient model of mutual
+%modelling, then analyzing five independent studies in terms of the interactions
+%between learning and mutual modelling.
 

### 2016-02-04 mutual-modeling.tex
Updated section 2

@@
 
-\section{Mutual Modelling}
-
-We refer to \emph{mutual modelling} as the process of inferring one's partner
+\section{Partner Modelling}
+
+We refer to \emph{Partner modelling} as the process of inferring one's partner
 mental states. Any claim that students carry out a detailed monitoring of their
@@
 accurate, subjects would obviously face a huge cognitive load. Conversely, peers
-could not collaborate without some minimal amount of mutual modelling. For
-instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
-opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
-B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
-evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
-it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
-profession) and, of course, it can be incorrect\ldots but it can not remain empty.
-Dialogues include many instances of utterances such as ``I thought he would do
+could not collaborate without some minimal amount of mutual modelling. 
+Collaborative learning dialogue include many instances of utterances such as ``I thought he would do
 that'' (first level of mutual modelling) or even ``He thought I would do that
@@
 
-The content of mutual models ranges from \emph{dispositional} to
+%For instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
+%opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
+%B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
+%evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
+%it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
+%profession) and, of course, it can be incorrect\ldots but it can not remain empty.
+
+The content of the partner model ranges from \emph{dispositional} to
 \emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
@@
 The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
-beliefs, emotions, history, status, etc.
-
-Besides, we have different levels of mutual modelling. If $A$ states ``B thinks I am good in
-maths'', $A$ builds a \emph{second level} model:
-\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
-regress of nested models: $A$
-saying ``$B$ knows that I don't expect him to solve this statistics problem''
-corresponds to \\ \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
-
-A partner model is likely not a ``box'', \ie not a monolithic representation
+beliefs, emotions, history, status, etc. A partner model is likely not a ``box'', \ie not a monolithic representation
 but rather a mosaic of information fragments about the partner, with various
@@
 of mechanisms, first for building an initial model of the partner, then for
-updating this model. As two students meet for the first time, mutual models are
+updating this model. As two students meet for the first time, partners models are
 initialized by the assumptions they make upon each other based on cues such as
@@
 appearance. Scholars studied how initial modelling impacts communication. In
-their experiments on initial mutual modelling, \citet{slugoski1993attribution}
+their experiment, \citet{slugoski1993attribution}
 pretended to their subjects that their (fake) partner had or had not received
@@
 
-After initialization, mutual models are updated during the collaborative work
+After initialization, partners models are updated during the collaborative work
 through verbal and non-verbal interactions. A default inference rule is that ``my
 partner agrees with me unless he disagrees'', which rejects the critiques that
-mutual modelling generates an unbearable cognitive load. This default rule is
+partner modelling generates an unbearable cognitive load. This default rule is
 superseded by the several mechanisms for monitoring and repairing the partner
@@
 
-Mutual modelling does not occur in a vacuum but it is highly contextualized.
+Finally, partner modelling does not occur in a vacuum but it is highly contextualized.
 \citet{clark1991grounding} review how the features of the collaborative
@@
 interfaces to provide ways for peers to detect and repair their
-misunderstanding. In CSCW, researchers have explored various so-called
-``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
-that A could not directly perceive because $B$ was working on a different subset
-of the virtual space. Different awareness tools are presented and relied upon in
-this contribution as well since they allowed us to manipulate experimentally the
-mutual modelling activity.
-
-%\subsection{State of the Art}
-%
-%\paragraph{Building a model of a human during the interaction} Perspective
-%Taking is a human ability which allows one to put him/herself in another
-%person's point of view. Studied in psychology
-%literature~\citep{Flavell1992,Tversky1999}, this ability is crucial when
-%interacting with people by allowing one to reason on others' understanding of
-%the world in terms of visual perception, spatial descriptions, affordances and
-%beliefs, etc.  In the last years these notions have been gradually employed in
-%HRI.~\citep{Breazeal2006} presents a learning algorithm that takes into account
-%information about a teacher's visual perspective in order to learn a task.
-%~\citep{Johnson2005} apply visual perspective taking for action recognition
-%between two robots.~\citep{Trafton2005} use both visual and spatial perspective
-%taking to find out the referent indicated by a human partner.
-%
-%The cognitive model that the robot builds for the agent it interacts with is
-%still simple and mostly focused on geometric features (who sees what? What are
-%our relative positions? etc.). Extending this knowledge with more subtle
-%perceptions (emotional state for instance) remains to be done.
-%
-%\paragraph{Joint action}
-%Several theories dealing with collaboration~\citep{Cohen1991,Grosz1996,Clark1996}
-%emphasize that collaborative tasks have specific requirements compared to
-%individual ones, \eg, since the robot and the person share a common goal, they
-%have to agree on the manner to realize it, they must show their commitment to
-%the goal during execution, etc. Several robotic systems have already been built
-%based on these theories~\citep{Rich1997,Sidner2005,Breazeal2003} and they all
-%have shown benefits of this approach. They have also shown how difficult it is
-%to manage turn-taking between communication partners and to interleave task
-%realization and communication in a generic way. Finally, today only few
-%systems~\citep{Fong2006,Breazeal2003,Sisbot2008} take humans into account at all
-%levels.
-%
+misunderstanding.
 

### 2016-02-04 mutual-modeling.tex
Updated section 3

@@
 
-\section{A model for mutual modelling evaluation}
-
-To report on experiments on mutual modelling, we suggest the notation
-\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This is a simple but
-useful reduction to reason on mutual modelling. This notation does not mean $A$
+\section{Mutual Modelling}
+
+Mutual modelling is bi-directional.
+During dyadic problem solving,   partner A builds some model of B and B build some model of A.
+Moreover, these two processes are not independent: A's model of what B knows, include what B knows about A.
+This leads to nested levels of  modelling.
+If $A$ states ``B thinks I am good in
+maths'', $A$ builds a \emph{second level} model:
+\model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
+regress of nested models: $A$
+saying ``$B$ knows that I don't expect him to solve this statistics problem''
+corresponds to \\ \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
+As we will see in one study we report, mutuality also applies to triads in which A will elaborate a model of B and of C, and reciprocally.
+This is probably also true for larger groups although one may hypothesize that there exists (yet undefined) group size from which it is not possible to model all partners individually and, therefore, members  model  the group as a whole instead as a collection of individuals.
+We do not explore this hypothesis here and limit ourselves to dyads and triads.
+
+This mutuality allows us to address a fundamental question: \emph{does the quality of partner modelling depend upon the cognitive skills of each partner (some people being better in perceiving other's states) or does it result from the quality of interactions among them ?} Behind his question, the reader may perceive the long lasting debate between tenets of, respectively, the individual and social views of human cognition.
+The simple hypothesis is that when individual are randomly paired for an experiment, there is no reason for which their individual cognitive skills would correlate.
+Therefore, if it occurs that the quality of the Model (A,B) is correlated with Model (B,A), one may infers that this quality depends upon what A and B have in common, their verbal interactions.
+
+To answer this question, we went back to five previous studies that addressed various other research questions, but in which the correlation between the quality of these two models could be computed.
+The studies we report do hence not constitue a clean sequence of experiment to investigate mutual modelling but the ad-hoc revisiting of previous experiments to explore a question that we had then neglected.
+Some of these studies are about collaborative learning while others are only collaborative problem solving, but the latter rely on rich tasks that are similar to those we use in CSCL.
+
+\section{A notation for discussing mutual modelling }
+
+Natural language becomes cumbersome when describing things such as "the model that A builds about the model that B builds about A".
+Therefore, to define hypotheses and report on experiments on mutual modelling, we use  the notation
+\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''.
+This  notation is not proposed as a formal theory of mutual modelling but as useful simplification for communicating about mutual modelling.
+This notation does not mean $A$
 has an explicit, monolithic representation of $B$: it must be understood as an
-abstraction referring to complex socio-cognitive processes. Besides, we refer to
-the \emph{degree of accuracy} of the model as \Model{A}{B}{X}.
+abstraction referring to complex socio-cognitive processes.
+As explained in the previous section, $A$ can be fragmented, multi-dimensional, etc.
+This notation is neither presented as a computational model of mutual modelling, nor as some universal formalism; it's usefulness is internal to this paper.
+Besides, we refer to
+the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We discuss in the next section the methodological difficulty in measuring this accuracy.
 
@@
 
-This is qualitative summary more that a real equation since several parameters
+This formula is presented as a qualitative summary, not as a real equation, since several parameters
 are hard to quantify (\eg the cost of a communication act depends upon the user
-as well) but it clarifies the parameters of the experiments we present hereafter.
+as well). We use it for clarifying the parameters of the experiments we present hereafter.
 
@@
         grounding criterion.  All tasks are about building a solution or a
-        representation together; we did not address everyday conversations.
+        representation together. As stated earlier, we did not address everyday conversations.
 

### 2016-02-04 mutual-modeling.tex
Split section 3 and created a new section 'Methodological Issues'

@@
 
-Measuring \Model{A}{B}{X} is methodologically difficult. We
-discriminate two steps, first to capture \model{A}{B}{X} and then to
+\section{Methodological Issues}
+
+Since a mental model is not directly observable,  the study of mutual modelling is methodologically difficult.
+How can we for instance measure \Model{A}{B}{X}.
+We proceed in two steps, first to capture \model{A}{B}{X} and then to
 estimate \Model{A}{B}{X}.
 
-\begin{itemize}
-    \item {\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$
+{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$
         what he/she believes about what $B$ knows, feels, intends to do, etc.
@@
 
-    \item {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
+    {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
         need to access the reference model \refmodel{B}{X} to estimate its
@@
 
-\end{itemize}
+Our method for investigating mutual modelling is observe the variations of  accuracy that result from variations of external parameters that appear as variable in the formula presented in the previous section: the accuracy should for go down if we increase the cost of grounding acts and goes up if we increase the grounding criterion (the necessity to build a shared understanding).
+One way to produce these variations is what CSCW,researchers have called
+``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
+that A could not directly perceive because $B$ was working on a different subset
+of the virtual space. Different awareness tools are used in the following studies as methodological tricks  to manipulate experimentally the  mutual modelling activity.
 

### 2016-02-04 mutual-modeling.tex
Split section 4 and created a new section 'Hypotheses and questions'

@@
 
-\subsubsection*{Hypotheses}
+\section*{Hypotheses and questions}
 
@@
 different institutions by different researchers. They used different
-independent, intermediate and dependent variables. Nonetheless, we were able to
-address a few questions across these studies. We start with 3 simple hypotheses
-about $\mathcal{M}^{\circ}(A,B)$:
+independent, intermediate and dependent variables. Nonetheless, we were able to retroactively
+address 3 research questions across these studies.
+
+\paragraph{The symmetry question}
+
+As stated earlier, the fundamental challenge is to determine if mutual modelling is an individual skill (hypothesis 1) or  the emergent property of social interactions (hypothesis 3).
+This question is empirically translated into the symmetry of mutual modelling: what is
+the relationship between \Model{A}{B}{X} and \Model{B}{A}{X}
+(Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is
+mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
+might support $\mathcal{H}_{3}$ since there is a low
+probability that randomly formed pairs integrate peers with the same level of
+mutual modelling skills.
+However, a high correlation could also have another explanation, stated in  $\mathcal{H}_{2}$: it could be that A is good at modelling B and good at helping B to model herself or himself.
 
@@
 differentiate.
-
-\paragraph{The symmetry question}
-
-One interesting question relates to the symmetry of mutual modelling: what is
-the relationship between \Model{A}{B}{X} and \Model{B}{A}{X}
-(Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is
-mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
-might support $\mathcal{H}_{2}$ or $\mathcal{H}_{3}$ since there is a low
-probability that randomly formed pairs integrate peers with the same level of
-mutual modelling skills.
 
@@
 \Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
-\Model{C}{B}{X} and \Model{B}{C}{X}.
-
-This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
+\Model{C}{B}{X} and \Model{B}{C}{X}.This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
 and $B$ have the same accuracy when modelling $C$ ($\Delta_2 =
-\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? If it is the
-same, $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ gains over $\mathcal{H}_{1}$
+\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? A significant correlation would support
+, $\mathcal{H}_{2}$  (C has helped both A and B to model C) or support $\mathcal{H}_{3}$ (the quality of triad interactions enables all partners to model each other accurately)
 
@@
 \Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
-answer would support $\mathcal{H}_{3}$. A negative answer would support
-$\mathcal{H}_{2}$ since $\Delta_3$ could mainly be explained by the way $A$ and
-$B$ help $C$ to model them.
+answer would support $\mathcal{H}_{1}$, C being simply good at modelling any partner.
+It could also support  $\mathcal{H}_{3}$,  since the quality of interactions influences the accuracy of two models that C is building.
+A negative answer would support
+$\mathcal{H}_{2}$ since $\Delta_3$ could mainly be explained by the fact that  $A$ helped more $C$ to model him than 
+$B$ did.
+
+In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
+whether the accuracy of mutual modelling depends more upon the modeller's effort  $\mathcal{H}_{1}$
+or the modellee's behaviour  $\mathcal{H}_{2}$.
+
+Please note that we consider the quality of interactions at triad level (ABC), neglecting the cases where AB interact better for instance than BC, since there was no 'private' communication channel in the following studies.
+We do nonetheless acknowledge that this point could be debated.
 
@@
 
-In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
-whether the accuracy of mutual modelling depends more upon the modeller's effort
-or the modellee's behaviour.
-
 \paragraph{The rectangle questions}
@@
 We can go further by comparing self- versus other modelling ($\Delta_4$ in
-Fig.~\ref{mm_rectangle}) as an indication of meta-cognitive skills. We can also
-question if modelling skills depend upon what aspects are being modeled ($X$ or
+Fig.~\ref{mm_rectangle}).
+A large difference would indicate that meta-cognitive skills (self-modelling) and partner modelling skills are rather different skills, while a small $\Delta_4$ could be interpreted as the indications that these are two specific instance of a more general cognitive process.
+This questions is however not central to this paper, and the value of $\Delta_4$ us only available in one of the reported studies.
+
+ We can also question if modelling skills depend upon what aspects are being modeled ($X$ or
 $Y$), which would explain vertical differences ($\Delta_5$ in
-Fig.~\ref{mm_rectangle}). When possible, we will also discuss these two
-questions in the studies presented hereafter.
+Fig.~\ref{mm_rectangle}).
+These differences would allow refining the notion of modelling skills, namely whether there exist some general ability to model partners or whether this is only the abstraction of a beam of more specific skills such as detecting emotions versus identifying references from deictic gestures.
 
@@
 
-\caption{\small Meta-cognitive skills and domain-dependent modelling}
+\caption{\small Meta-cognitive skills (horizontally) and domain-dependent modelling (vertically}
 
@@
 \end{figure}
-
-This simple notation does not pretend to provide a mathematical account of
-mutual modelling but to be more systematic in describing the following
-experiments.
 

### 2016-02-04 mutual-modeling.tex
Updated section 'Studies'

@@
 
+We report five studies conducted  by different researchers in different contexts between 2000 and 2015.
+They do not form a consistent research strategy but actually the fact that some trends emerge despite their diversity constitutes indeed the richness of this line of work.
+
+\begin{table*}[ht!]
+\centering
+\resizebox{\textwidth}{!}{%
+\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
+\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
+\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
+\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
+\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
+\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
+\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
+\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
+\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
+\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
+%\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
+%$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
+%$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
+%$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
+\end{tabular}
+}
+\caption{Comparison of the studies}
+\label{synthesis_table}
+\end{table*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
-If it was null, modelling would be interpreted as an individual process that
-varies according to personality traits or cognitive skills: some persons tend to
-model more accurately their peer while others don't care or can simply not
-($\mathcal{H}_{1}$). If the correlation was close to 1, we could conclude that
-the modelling accuracy is an emergent property of teams since there would be low
-probability that, by chance, low accurate modellers are paired with accurate
-modellers and conversely ($\mathcal{H}_{2}$). Our correlation is between these
-two extremes but the very notion of significant intra-class correlation tend to
-indicate that one may not consider peers as independent subjects, which is an
-argument for the \emph{interaction} hypothesis $\mathcal{H}_{3}$.
+It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$.
+It is nonetheless far from 1 which implies we cannot discard the individual modelling skills.
+We collect more evidence along the next studies.
+
+$\mathcal{H}_{3}$
+
+%If it was null, modelling would be interpreted as an individual process that
+%varies according to personality traits or cognitive skills: some persons tend to
+%model more accurately their peer while others don't care or can simply not
+%($\mathcal{H}_{1}$). If the correlation was close to 1, we could conclude that
+%the modelling accuracy is an emergent property of teams since there would be low
+%probability that, by chance, low accurate modellers are paired with accurate
+%modellers and conversely ($\mathcal{H}_{2}$). Our correlation is between these
+%two extremes but the very notion of significant intra-class correlation tend to
+%indicate that one may not consider peers as independent subjects, which is an
+%argument for the \emph{interaction} hypothesis $\mathcal{H}_{3}$.
 
@@
 \subsubsection*{Subjects}
-%ANON
-%Ninety EPFL students participated in this experiment. We only selected EPFL
-%students since knowledge of the campus geography had an impact both on group
+
 Ninety students participated in this experiment. We only selected students from
@@
 \gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more $A$
-makes errors about $B$, the more $B$ does as well.
+makes errors about $B$, the more $B$ does as well (and vice-versa).
 
@@
 
-This positive correlation observed in the symmetry question is surprising as
-well: when acknowledging the high heterogeneity of spatial skills among
-adults~\citep{liben1981spatial}, a significant positive correlation would
-indicate that mutual modelling indeed stems from group processes more than
-individual processes. The significant results on the triangle questions goes in
-the same direction: either both the modeler and the modellee contribute to
-accurate modelling, or the quality of teamwork has a global effect. Since team
-members did not have many opportunities to directly interact during this
-experiment, the factor that facilitates accurate mutual modelling is probably
-more the overall group strategy than the quality of verbal interactions: a clear
-strategy facilitates memorizing one's partner path.
+This positive correlation observed in the symmetry question confirms the first study.
+In this case, this is actually more surprising given  the high heterogeneity of spatial skills among adults~\citep{liben1981spatial}.
+This result supports  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$.
+The results regarding $\Delta_2$ may support both  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$  but discards   $\mathcal{H}_{1}$  (if the modeller's skill would matter a lot,  $\Delta_2$ would be lower).
+Conversely, $\Delta_3$  supports $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$ but discards   $\mathcal{H}_{2}$  (if the modeller's skill would matter a lot,  $\Delta_3$ would be lower).
+In summary, various indices support the 3 hypotheses, which implies there is some truth in each of them, but $\mathcal{H}_{3}$ is the only hypothesis that is not rejected by any index.
+We may hence, with great caution, concluded that the social perspective ($\mathcal{H}_{3}$) is moderately reinforced by this second study.
+Since the values of $\Delta_2$ and $\Delta_3$ are very similar, we do not interpret their minor difference as evidence for a stronger role of the modeller ($\mathcal{H}_{1}$) or the modellee ($\mathcal{H}_{1}$).
+
+We have however to bring some nuances to the social viewpoint ($\mathcal{H}_{3}).
+The main feature that can be associated to the team level in this experiment is probably not the quality of their verbal interactions per se (they interact mostly by drawing on a shared map), but rather the consistency of  the spatial exploration strategy: a clear strategy facilitates memorizing one's partner path.
+One could argue whether the team strategy can be dissociated from the team interactions quality or constitues one of its components.
 
@@
 computed from random gradings (following the same grade distribution as for the
-experimental data).
-
-A t-test on the two sets releaved a significantly lower average difference in
+experimental data). A t-test on the two sets revealed a significantly lower average difference in
 the experimental data (mean difference: $0.40$ vs $0.54$ with random gradings,
-$t=-3.3, df=60, p = 0.0016$), supporting $\mathcal{H}_{2}$ or $\mathcal{H}_{3}$.
+$t=-3.3, df=60, p = 0.0016$), which confirms the symmetry of mutual modelling.
 
@@
 the absolute differences between \gModel{A}{C} and \gModel{B}{C} over the 60
-subjects. The average difference is 0.42 ($SD= 0.34$), and if not significantly
-different from the same data obtained from random gradings ($t=-1.61, df=60,
-p=0.11$). This rejects $\mathcal{H}_{2}$.
+subjects. The average difference is 0.42 ($SD= 0.34$), and is not significantly
+different from the same index computed from random gradings ($t=-1.61, df=60,
+p=0.11$).
 
@@
 lower than chance ($t=-2.62, df=60, p=0.01$): a given subject tends to model its
-two partners with similar degrees of accuracy, which supports $\mathcal{H}_{1}$.
-It could also support $\mathcal{H}_{3}$, but it does not since $\Delta_2$ is not
-significant.
+two partners with similar degrees of accuracy.
 
@@
 interpret this too far, but to show that there are large variations even within one
-area  (perceiving emotions).
+area  (perceiving emotions). These variations shomehow question the existence of a general aptitude to model others ( $\mathcal{H}_{1}$).
 
@@
 
-The results evidence that richer media does not automatically foster better
-modelling, on the contrary: groups using only the audio modality to exchange
-their views were able to build more accurate models of each other than groups
-using both video and audio, which is especially surprising when modelling
-emotions: one would expect that video feedback indeed help building a better
-model of the emotions of the other. Our interpretation is that viewing what
-one's partner sees (shared graphical editor) was, in this task, more important
+In this third study, the accuracy of mutual modelling  between two peers tend to be symmetrical, which  supports $\mathcal{H}_{2}$ and ($\mathcal{H}_{3}$.)  This result is contradicted by the fact that ($\Delta_2$) is not significant, which rejects both
+ $\mathcal{H}_{2}$ and, to a lower extent,,  $\mathcal{H}_{3}$.
+Finally,  ($\Delta_3$) supports both  $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$.
+In summary, this study brings some supports to  $\mathcal{H}_{3}$, but reveal again that $\mathcal{H}_{3}$ is only part of the explanation.
+
+Like the previous one, this study leads us to refine what we mean by quality of interactions in $\mathcal{H}_{3}$.
+We  expected that the video channel would help peers building a more accurate model of the each other's emotions.
+The results show the opposite: peers in the audio-only condition built more accurate models, probably because they concentrated more on the shared concept map.
+This confirms other studies that revealed that viewing what one's partner sees (shared graphical editor) is more important
 than seeing each other~\citep{gaver1993one,anderson1997impact}.
-
-In this study, the accuracy of mutual modelling between two peers tend to be
-symmetrical, which would support the idea that the quality of modelling is primarily
-building upon the social interaction. This is however not fully supported by the
-analysis of $\Delta_2$: $A$ being good at modelling $C$ does not correlate with
-$B$ being good at modelling $C$.
-
-On the other hand, $\Delta_3$ shows that a given $A$ tends to
-model $B$ and $C$ at similar levels of accuracy: this evidences an individual
-skill at modelling others (hypothesis $\mathcal{H}_{1}$).
+Therefore, what is meant by the quality of interactions is $\mathcal{H}_{3} is more than the linguistic features of dialogue but includes the way these interactions are articulated to the task.
 
@@
 significant in each condition: someone good at self-modelling is not necessarily
-good at modelling someone else and vice-versa.
+good at modelling someone else and vice-versa.  This seems to indicate that partner modelling
+requires different skills than meta-cognition, despite their similarity at some level of abstraction.
 
@@
 
-In this experiment, the symmetry of mutual modelling is only found when subjects
-are in conditions of symmetry of information. This consolidates the hypothesis
-of \gModel{A}{B} being a property of interactions in context ($\mathcal{H}_{3}$)
-rather than $A$'s cognitive attitude.
+In this experiment, the symmetry of mutual modelling is  found but  in one condition, namely when subjects
+ receive the same information before the task.  This condition corresponds to the situation tested in the three
+previous studies and supports  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$.
+
+How do we interpret the fact that the symmetry vanishes when peers receive different texts to read before the task?
+One explanation would be the difficult of mutual modelling when peers don't know what others read, but we found no significance of \gModel{A}{A} between conditions.
+Since texts were partly overlapping, another explanation is that, in absence of these initial common grounds, mutual modelling requires A to make evident to B what A thinks B does not know about A, which is the second level of modelling described in $\mathcal{H}_{2}$ .
+A low symmetry means that some peers are better than others at this second level of modelling, which supports the existence of such an individual skill, as stated in  $\mathcal{H}_{2}$
 
@@
 KAT informed them about the pre-test score. Actually, pairs in the KAT condition
-produced significantly more instances of 3 specific categories of interactions:
-(1) elaborated utterances with rich contents,  (2) utterances asking about the
-other's knowledge such as ``Did you understand how transmission works?'' (3)
+produced significantly more instances of 3 interesting categories of interactions:
+ (1) utterances asking about the
+other's knowledge such as ``Did you understand how transmission works?'' (2)
 utterances describing one's own knowledge (\gmodel{A}{A}) such as ``I don't
-remember the Ranvier's thing...''. These 3 categories provide different account
-of \gModel{A}{B}: (1) as an effect of the quality of interaction, (2) as an
-effect of $A$'s effort to model $B$ or (3) as $B$'s effort to give cues to $A$ about his
-own knowledge.
+remember the Ranvier's thing...'' and (3) elaborated utterances with rich contents,  These 3 categories provide different account
+of \gModel{A}{B}: (1) as an
+effect of $A$'s effort to model $B$  ($\mathcal{H}_{1}$), (2) as $B$'s effort to give cues to $A$ about his
+own knowledge  ($\mathcal{H}_{2}$) and (3) as an effect of the quality of interaction  ($\mathcal{H}_{3}$) .
 
@@
 \paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
-0.05, ns$).
+0.05, ns$), which discards  ($\mathcal{H}_{2}$) and  ($\mathcal{H}_{3}$).
+In addition, the KAT supports the processing of modelling at the first level ($\mathcal{H}_{1}$) and not at the second level ($\mathcal{H}_{2}$).
+Hence the fact that the KAT enhances  \gModel{A}{B}  brings additional supports to $\mathcal{H}_{1}$.
+However, the analysis of verbal interactions find elements that actually supports the three hypotheses.
+
+ ($\mathcal{H}_{3}$)
 

### 2016-02-04 mutual-modeling.tex
Updated section 'Synthesis'

@@
 
-Out of 5 studies, 3 revealed a correlation between the model peers built about
-each other: when $A$ builds an accurate model of $B$, $B$ also tends to build an
-accurate model of A.  In one study, this was the case only in one condition.
-The conclusion is hence that modelling is mutual but only in certain conditions,
-summarized in Table~\ref{synthesis_table}.
-
-\begin{table*}[ht!]
-\centering
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
-\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
-\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
-\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
-\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
-\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
-\textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
-\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
-\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
-\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
-\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
-$\mathcal{H}_1$               &                                  & supported                       & supported                 &                           &                              \\
-$\mathcal{H}_2$               &                                  & supported&                         &                           &                              \\
-$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
-\end{tabular}
-}
-\caption{Comparison of the studies}
-\label{synthesis_table}
-\end{table*}
-
-Coming back to our initial hypotheses, the $\Delta_2$ results in study 2
-reinforce both $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The \emph{interaction
-hypothesis} $\mathcal{H}_{3}$ was that the activity of modelling one's partner
-is not simply the result of an individual cognitive effort but overall the
-side-effect of the quality of interactions. The \emph{second-order-modelling
-hypothesis} $\mathcal{H}_{2}$ stated that, if $A$ is accurate in modelling $B$,
-$A$ should notice when $B$ misunderstands $A$ and should be able to repair. A
-good modeler would somehow drive the modelling of his or her partner.  This
-\emph{drive} will however occur through verbal cues and repairs and is hence not
-easy to differentiate from $\mathcal{H}_{3}$. The contradictory $\Delta_2$
-analysis in study 3 and $\Delta_1$ results in studies 4 (to some extent) and 5
-weaken the hypothesis.
-
-The triangular questions lead to similar contradictions. The $\Delta_2$
-correlation between the \gmodel{A}{C} and \gmodel{B}{C} (study 2)
-support $\mathcal{H}_{2}$ if modelling skills were purely individual, there would
-be no reason for modelling skills of $A$ and $B$ to be correlated.  The role of $C$ in
-\gmodel{A}{C} and \gmodel{B}{C} probably explains this correlation.  Conversely,
-the $\Delta_3$ correlation between the \gmodel{A}{B} and \gmodel{A}{C} (study
-3) question $\mathcal{H}_{2}$: if the role of the modellee was so
-important, it would be surprising to have by chance a correlation between
-$B$'s and $C$' ability to drive $A$'s modelling.
-
-In addition, the lack of correlation between \gmodel{A}{A} and \gmodel{A}{B}
-($\Delta_4$ in study 4 and study 5) could also contribute to the idea that $A$'s
-modelling skills is not the key factor. However, between modelling someone and
-modelling oneself, \ie  metacognition, they are so many differences, that we
-would hardly conclude from that.
-
-Among these complex results, it is important to stress that, for $\Delta_2$, the
-significant correlations are all below 0.50, not around 0.90.  So even if all
-$\Delta_2$ correlations were significant in all experiments, we would still not
-conclude that modelling purely reflects the quality of social interactions.
-There is obviously a large part of individual variance; some of us are much
-better in modelling peers. This is not new. The contribution of these studies is
-rather to establish the importance of team processes on something that a priori
-seemed to be more of a personal cognitive skill.  In other words, we stress that
-the bottle is half-full.
+We found symmetry of the mutual model on 4 studies out of 5.
+In study 4, this only applies to the control group (having read the same text before), which corresponds to the situation of the 3 first studies.
+Finding this symmetry per se constitutes an interesting result as we don't know earlier studies who have established this relationship.
+The problem is that this symmetry does not allow use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$.
+The second hypothesis, the skill of helping one's peer to model oneself, is questioned by study 3 where C could be modelled  accurately by A and not accurately by B.
+The same hypothesis is however supported by the results of study 2, and indirectly by study 4.
+This does not discard \mathcal{H}_{1}$: even when we found significant correlations, they were all below 0.50, not around 0.90.
+Hence, even of the quality of social interaction matters, there is obviously a large part of individual variance within teams.
+
+In other words, these studies do conclude that one hypothesis is right and the others are wrong, which is per se the main contribution of this paper.
+We started from the idea that partner modelling is essentially an individual skill  and that we therefore improve the quality of collaboration by providing the so-called awareness tools, which act as some kind of prostheses for partner modelling.
+This individual skill cannot be discarded.
+Everyone has experienced the pleasure of interacting with colleagues who answer precisely to the question we asked them  and even guess the reason why we asked this question.
+Conversely, everyone also experienced the frustration of someone referring about a third person by his name, say Mike Smith,  while knowing perfectly that there is no chance that we know this Mike Smith.
+The existence of these individual skills is not denied by our studies; the novelty of this paper is to show, thanks to symmetry values, that individual skills only account  for a part of the accuracy in mutual modelling.
+
+We acknowledge that the difference between these hypotheses is  rather theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$) and the process of helping one's partner to model onself ($\mathcal{H}_{3}$) are mediated by verbal interactions in the team.
+One could hardly imagine someone managing a very accurate modelling despite the very bad quality of interactions.
+There is a bidirectional causal link between the accuracy of mutual modelling and the quality of interactions.
+This rather artificial distinction does however participate to the more fundamental discussion of the role of individual and social mechanisms in human cognition.
+In the field of social cognition, it is a common place to state that 'the whole is greater than the sum of the parts'.
+This refers to the emergence of team properties than cannot be reduced to the set of individual contributions.
+Our paper illustrates this emergence by showing the symmetry of mutual modelling
 
@@
 using eye tracking methods to address this challenge. 
-
-Finally, let us stress the expressions we used, such as \gmodel{A}{B}, do not
-imply we have a mechanical view of modelling. It's just a language for referring
-to different components of modelling each other in teams.  This simple notation
-emerges from the need to compare five mutual modelling studies; it did not
-generate a consistent program of consistent studies.  Our results are not very
-robust because they emerge from a \emph{post-hoc} reinterpretation of studies
+Third, our results emerge from a \emph{post-hoc} reinterpretation of studies
 that addressed different research questions (media richness, awareness tools,
@@
 space), with different groups sizes (pairs and triads) and different tasks.
-Moreover, these results have been found using different methods for measuring
-modelling: on-task in study 1 versus off-task in study 2, subjective validation
-in study 1 versus objective validation in study 2 (comparing $A$'s model with $B$'s
-behavior).
+
+Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do not
+imply we have a mechanical view of modelling. It's just a language for referring
+to different components of modelling each other in teams.  This simple notation
+emerged from the need to compare five mutual modelling studies. It could however pave the road for elaborating a systematic agenda for research on mutual modelling.
 

### 2016-02-04 mutual-modeling.tex
Text reflow at 80 characters

@@
 
-\title{The symmetry of partner modelling}
+\title{The Symmetry of Partner Modelling}
 
@@
 \begin{abstract}
-    Collaborative learning has often been associated to the construction of  a shared   understanding of the situation at hand. The psycholinguistics mechanisms
-    at work while establishing common grounds are the object of scientific controversy. We postulate that collaborative tasks require some level of mutual 
-    modeling, i.e.
-that each partner needs some model of what the other partners know/want/intend at a given time.
-We use the term 'some model' to stress the fact that this model is not necessarily details or complete, but that we acquire some representations of the persons we interact with.
-The question we address is:
-Does the quality of the partner model depend upon the modeler's ability to represent his or her partner, upon the modelee's ability to make his state clear to the modeler or upon the quality of their interactions?
-We address this question by analyzing differences between the  accuracy of the models built by different team members.
-We therefore report 5 experiments on collaborative problem solving or collaborative learning that vary in terms of tasks (how important it is to build an accurate model) and settings (how difficult it is to build an accurate model).
-In 4 studies, the accuracy of the model that A built about B was correlated with the accuracy of the model that B built about A,  which seems to imply that the quality of interactions matters more than individual abilities when building mutual models.
-However, these finding do not discard the fact individual abilities also  participate in the quality of modellng process .
+
+Collaborative learning has often been associated to the construction of  a
+shared   understanding of the situation at hand. The psycholinguistics
+mechanisms at work while establishing common grounds are the object of
+scientific controversy. We postulate that collaborative tasks require some
+level of mutual modeling, i.e. that each partner needs some model of what
+the other partners know/want/intend at a given time.  We use the term 'some
+model' to stress the fact that this model is not necessarily details or
+complete, but that we acquire some representations of the persons we
+interact with. The question we address is: Does the quality of the partner
+model depend upon the modeler's ability to represent his or her partner,
+upon the modelee's ability to make his state clear to the modeler or upon
+the quality of their interactions? We address this question by analyzing
+differences between the  accuracy of the models built by different team
+members. We therefore report 5 experiments on collaborative problem solving
+or collaborative learning that vary in terms of tasks (how important it is
+to build an accurate model) and settings (how difficult it is to build an
+accurate model). In 4 studies, the accuracy of the model that A built about
+B was correlated with the accuracy of the model that B built about A,  which
+seems to imply that the quality of interactions matters more than individual
+abilities when building mutual models.  However, these finding do not
+discard the fact individual abilities also  participate in the quality of
+modellng process .
 
@@
 
-From its very beginning, CSCL research has been following
-the suggestion by \citet{roschelle1995construction} that collaborative learning has
-something to do with the process of constructing and maintaining a \emph{shared
+From its very beginning, CSCL research has been following the suggestion by
+\citet{roschelle1995construction} that collaborative learning has something to
+do with the process of constructing and maintaining a \emph{shared
 understanding} of the task at hand. Building a shared/mutual understanding
@@
 science~\citep{chi1989self, webb1991task}, conflict resolution in
-socio-cognitive theories~\citep{doise1975social},
-mutual regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
+socio-cognitive theories~\citep{doise1975social}, mutual
+regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
 construction of a shared understanding has been investigated for several decades
@@
 mechanisms for explaining learning outcomes has been questioned in the learning
-sciences. Grounding mechanisms  are appropriate for explains conversational events, such
-referential failures in short dialogue episodes, but they do  hardly predict deeper phenomena such as
-\emph{conceptual change} (\ie the acquisition, acceptation and integration of a
-new belief into one's mental model) over longer
-sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
-episodes can probably be better understood from a socio-cultural perspective:
+sciences. Grounding mechanisms  are appropriate for explains conversational
+events, such referential failures in short dialogue episodes, but they do
+hardly predict deeper phenomena such as \emph{conceptual change} (\ie the
+acquisition, acceptation and integration of a new belief into one's mental
+model) over longer sessions~\citep{dillenbourg2006sharing}. The cumulative
+effect of grounding episodes can probably be better understood from a
+socio-cultural perspective:
 
@@
 
-Paradoxically, while Clark's theory is somewhat too linguistic from a learning  viewpoint, it is criticized at the same time as being too cognitivist by
-some psycholinguists, \ie as overestimating the amount of shared knowledge and
-mutual representations actually necessary to conduct a dialogue. The fundamental
-issue, as old as philosophy, is the degree of coupling between the different
-levels of dialogue, mostly between the lexical/syntactical level and the deeper
-semantic levels. \citet{pickering2006alignment} argue that the mutual
-understanding starts mostly with a \emph{superficial alignment} at the level of
-the linguistic representations, due to priming mechanisms, and that this local
-alignment may -- in some cases -- lead to a \emph{global alignment} of the
-semantic level (\emph{deep grounding}).  For these authors, the convergence in
-dialogue, and even the repair of some misunderstandings, is explained by this
-mimetic behavior more than by a monitoring of each other's knowledge:
-\emph{``...interlocutors do not need to monitor and develop full common ground
-as a regular, constant part of routine conversation, as it would be
-unnecessary and far too costly. Establishment of full common ground is, we
-argue, a specialized and non-automatic process that is used primarily in
-times of difficulty (when radical misalignment becomes apparent).''} (p. 179).
-This view is actually not incompatible with Clark's \emph{grounding
-criterion}~\citep{clark1989contributing}: the degree of shared understanding that
-peers need to reach depends upon the task they perform. For instance, a dialogue
-between two surgeons might rely on superficial alignment if they talk about
-their friends but has to guarantee accurate common grounds when talking about
-which intervention will be conducted in which way on which patient.
-In this paper, we operationalized the grounding criterion, i.e.
-the necessity for accurate modelling, as the correlation between the accuracy of partner models and various measures of team performance.
+Paradoxically, while Clark's theory is somewhat too linguistic from a learning
+viewpoint, it is criticized at the same time as being too cognitivist by some
+psycholinguists, \ie as overestimating the amount of shared knowledge and mutual
+representations actually necessary to conduct a dialogue. The fundamental issue,
+as old as philosophy, is the degree of coupling between the different levels of
+dialogue, mostly between the lexical/syntactical level and the deeper semantic
+levels. \citet{pickering2006alignment} argue that the mutual understanding
+starts mostly with a \emph{superficial alignment} at the level of the linguistic
+representations, due to priming mechanisms, and that this local alignment may --
+in some cases -- lead to a \emph{global alignment} of the semantic level
+(\emph{deep grounding}).  For these authors, the convergence in dialogue, and
+even the repair of some misunderstandings, is explained by this mimetic behavior
+more than by a monitoring of each other's knowledge: \emph{``...interlocutors do
+not need to monitor and develop full common ground as a regular, constant
+part of routine conversation, as it would be unnecessary and far too costly.
+Establishment of full common ground is, we argue, a specialized and
+non-automatic process that is used primarily in times of difficulty (when
+radical misalignment becomes apparent).''} (p. 179).  This view is actually not
+incompatible with Clark's \emph{grounding
+criterion}~\citep{clark1989contributing}: the degree of shared understanding
+that peers need to reach depends upon the task they perform. For instance, a
+dialogue between two surgeons might rely on superficial alignment if they talk
+about their friends but has to guarantee accurate common grounds when talking
+about which intervention will be conducted in which way on which patient.  In
+this paper, we operationalized the grounding criterion, i.e. the necessity for
+accurate modelling, as the correlation between the accuracy of partner models
+and various measures of team performance.
 
@@
 \vspace{2em}
-In summary, the controversy around the cognitive depth of shared understanding pertains to psycholinguistic, which investigates 'natural conversations, but  in collaborative learning, the situation is different; tasks are precisely designed for requirint a deeper negotiation fo meaning.
-The craft of CSCL is to design collaborative situations that create interdependence, avoid groupthink, and allows students to detect any 'illusion of shared understanding' (Cherubini et al, 2005).
-Our question is hence not  anymore"do peers build a shared understanding ?", but becomes "when peers have to build a shared understanding, how do they do it ?".
-This question addresses the mechanisms of grounding, in which the basic sequence is: make a proposition, detect misunderstanding and, if any, repair them.
-This paper focuses on the middle part, the detection of misunderstandings.
-Detecting a misunderstanding means that the emitter of a message identifies a mismatch between his communicative intention and the way his message is understood by his or her partner.
-Detecting one peer's understanding is investigate hereafter under the general umbrella of partner modelling.
-
-Cherubini, M., Van Der Pol, J., & Dillenbourg, P.
-(2005).
-Grounding is not shared understanding: Distinguishing grounding at an utterance and knowledge level.
-In CONTEXT'05, the Fifth International and Interdisciplinary Conference on Modeling and Using Context (pp.
-11-23).
+
+In summary, the controversy around the cognitive depth of shared understanding
+pertains to psycholinguistic, which investigates 'natural conversations, but  in
+collaborative learning, the situation is different; tasks are precisely designed
+for requirint a deeper negotiation fo meaning.  The craft of CSCL is to design
+collaborative situations that create interdependence, avoid groupthink, and
+allows students to detect any 'illusion of shared understanding' (Cherubini et
+al, 2005). Our question is hence not  anymore"do peers build a shared
+understanding ?", but becomes "when peers have to build a shared understanding,
+how do they do it ?".  This question addresses the mechanisms of grounding, in
+which the basic sequence is: make a proposition, detect misunderstanding and, if
+any, repair them.  This paper focuses on the middle part, the detection of
+misunderstandings.  Detecting a misunderstanding means that the emitter of a
+message identifies a mismatch between his communicative intention and the way
+his message is understood by his or her partner. Detecting one peer's
+understanding is investigate hereafter under the general umbrella of partner
+modelling.
+
+%TODO
+Cherubini, M., Van Der Pol, J., \& Dillenbourg, P. (2005). Grounding is not
+shared understanding: Distinguishing grounding at an utterance and knowledge
+level. In CONTEXT'05, the Fifth International and Interdisciplinary Conference
+on Modeling and Using Context (pp. 11-23).
 
@@
 could not collaborate without some minimal amount of mutual modelling.
-Collaborative learning dialogue include many instances of utterances such as ``I thought he would do
-that'' (first level of mutual modelling) or even ``He thought I would do that
-but I intended something else.'' (second level of mutual modelling).
+Collaborative learning dialogue include many instances of utterances such as ``I
+thought he would do that'' (first level of mutual modelling) or even ``He
+thought I would do that but I intended something else.'' (second level of mutual
+modelling).
 
@@
 The list of what $X$ stands for in \model{A}{B}{X} is possibly infinite:
-beliefs, emotions, history, status, etc. A partner model is likely not a ``box'', \ie not a monolithic representation
-but rather a mosaic of information fragments about the partner, with various
-granularity and various life cycles. This mosaic is elaborated through a variety
-of mechanisms, first for building an initial model of the partner, then for
-updating this model. As two students meet for the first time, partners models are
-initialized by the assumptions they make upon each other based on cues such as
-his/her belonging to broad categories (age, culture, profession, \ldots),
-stereotypes (sportsmen, junkie, business women, Swiss,\ldots) as well as physical
-appearance. Scholars studied how initial modelling impacts communication. In
-their experiment, \citet{slugoski1993attribution}
-pretended to their subjects that their (fake) partner had or had not received
-the same information. They observed that the subjects adapted their dialogue by
-focusing the explanation on the items that he/she was supposed to ignore.
-\citet{brennan1991conversation} showed that the subjects used different initial
-strategies in forming queries depending on who they were told their partner was.
+beliefs, emotions, history, status, etc. A partner model is likely not a
+``box'', \ie not a monolithic representation but rather a mosaic of information
+fragments about the partner, with various granularity and various life cycles.
+This mosaic is elaborated through a variety of mechanisms, first for building an
+initial model of the partner, then for updating this model. As two students meet
+for the first time, partners models are initialized by the assumptions they make
+upon each other based on cues such as his/her belonging to broad categories
+(age, culture, profession, \ldots), stereotypes (sportsmen, junkie, business
+women, Swiss,\ldots) as well as physical appearance. Scholars studied how
+initial modelling impacts communication. In their experiment,
+\citet{slugoski1993attribution} pretended to their subjects that their (fake)
+partner had or had not received the same information. They observed that the
+subjects adapted their dialogue by focusing the explanation on the items that
+he/she was supposed to ignore.  \citet{brennan1991conversation} showed that the
+subjects used different initial strategies in forming queries depending on who
+they were told their partner was.
 
@@
 
-Finally, partner modelling does not occur in a vacuum but it is highly contextualized.
-\citet{clark1991grounding} review how the features of the collaborative
-situation, namely the media (co-temporality,\ldots), may facilitate or hamper
-mutual modelling.  \citet{hutchins1997constructing} reported a study in which a
-short silence between two pilots was perfectly interpreted because it occurred
-in a highly constrained communication context.  Some environments are more
-productive than others in helping peers to detect their misunderstandings.
-\citet{roschelle1995construction} reformulate the design of CSCL
-interfaces to provide ways for peers to detect and repair their
+Finally, partner modelling does not occur in a vacuum but it is highly
+contextualized.  \citet{clark1991grounding} review how the features of the
+collaborative situation, namely the media (co-temporality,\ldots), may
+facilitate or hamper mutual modelling.  \citet{hutchins1997constructing}
+reported a study in which a short silence between two pilots was perfectly
+interpreted because it occurred in a highly constrained communication context.
+Some environments are more productive than others in helping peers to detect
+their misunderstandings.  \citet{roschelle1995construction} reformulate the
+design of CSCL interfaces to provide ways for peers to detect and repair their
 misunderstanding.
@@
 
-Mutual modelling is bi-directional.
-During dyadic problem solving,   partner A builds some model of B and B build some model of A.
-Moreover, these two processes are not independent: A's model of what B knows, include what B knows about A.
-This leads to nested levels of  modelling.
-If $A$ states ``B thinks I am good in
-maths'', $A$ builds a \emph{second level} model:
+Mutual modelling is bi-directional. During dyadic problem solving,   partner A
+builds some model of B and B build some model of A.  Moreover, these two
+processes are not independent: A's model of what B knows, include what B knows
+about A.  This leads to nested levels of  modelling. If $A$ states ``B thinks I
+am good in maths'', $A$ builds a \emph{second level} model:
 \model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
-regress of nested models: $A$
-saying ``$B$ knows that I don't expect him to solve this statistics problem''
-corresponds to \\ \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.
-As we will see in one study we report, mutuality also applies to triads in which A will elaborate a model of B and of C, and reciprocally.
-This is probably also true for larger groups although one may hypothesize that there exists (yet undefined) group size from which it is not possible to model all partners individually and, therefore, members  model  the group as a whole instead as a collection of individuals.
-We do not explore this hypothesis here and limit ourselves to dyads and triads.
-
-This mutuality allows us to address a fundamental question: \emph{does the quality of partner modelling depend upon the cognitive skills of each partner (some people being better in perceiving other's states) or does it result from the quality of interactions among them ?} Behind his question, the reader may perceive the long lasting debate between tenets of, respectively, the individual and social views of human cognition.
-The simple hypothesis is that when individual are randomly paired for an experiment, there is no reason for which their individual cognitive skills would correlate.
-Therefore, if it occurs that the quality of the Model (A,B) is correlated with Model (B,A), one may infers that this quality depends upon what A and B have in common, their verbal interactions.
-
-To answer this question, we went back to five previous studies that addressed various other research questions, but in which the correlation between the quality of these two models could be computed.
-The studies we report do hence not constitue a clean sequence of experiment to investigate mutual modelling but the ad-hoc revisiting of previous experiments to explore a question that we had then neglected.
-Some of these studies are about collaborative learning while others are only collaborative problem solving, but the latter rely on rich tasks that are similar to those we use in CSCL.
+regress of nested models: $A$ saying ``$B$ knows that I don't expect him to
+solve this statistics problem'' corresponds to \\
+\model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.  As we will see in
+one study we report, mutuality also applies to triads in which A will elaborate
+a model of B and of C, and reciprocally. This is probably also true for larger
+groups although one may hypothesize that there exists (yet undefined) group size
+from which it is not possible to model all partners individually and, therefore,
+members  model  the group as a whole instead as a collection of individuals.  We
+do not explore this hypothesis here and limit ourselves to dyads and triads.
+
+This mutuality allows us to address a fundamental question: \emph{does the
+quality of partner modelling depend upon the cognitive skills of each partner
+(some people being better in perceiving other's states) or does it result from
+the quality of interactions among them ?} Behind his question, the reader may
+perceive the long lasting debate between tenets of, respectively, the individual
+and social views of human cognition.  The simple hypothesis is that when
+individual are randomly paired for an experiment, there is no reason for which
+their individual cognitive skills would correlate. Therefore, if it occurs that
+the quality of the Model (A,B) is correlated with Model (B,A), one may infers
+that this quality depends upon what A and B have in common, their verbal
+interactions.
+
+To answer this question, we went back to five previous studies that addressed
+various other research questions, but in which the correlation between the
+quality of these two models could be computed.  The studies we report do hence
+not constitue a clean sequence of experiment to investigate mutual modelling but
+the ad-hoc revisiting of previous experiments to explore a question that we had
+then neglected. Some of these studies are about collaborative learning while
+others are only collaborative problem solving, but the latter rely on rich tasks
+that are similar to those we use in CSCL.
 
@@
 
-Natural language becomes cumbersome when describing things such as "the model that A builds about the model that B builds about A".
-Therefore, to define hypotheses and report on experiments on mutual modelling, we use  the notation
-\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''.
-This  notation is not proposed as a formal theory of mutual modelling but as useful simplification for communicating about mutual modelling.
-This notation does not mean $A$
-has an explicit, monolithic representation of $B$: it must be understood as an
-abstraction referring to complex socio-cognitive processes.
-As explained in the previous section, $A$ can be fragmented, multi-dimensional, etc.
-This notation is neither presented as a computational model of mutual modelling, nor as some universal formalism; it's usefulness is internal to this paper.
-Besides, we refer to
-the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We discuss in the next section the methodological difficulty in measuring this accuracy.
+Natural language becomes cumbersome when describing things such as "the model
+that A builds about the model that B builds about A". Therefore, to define
+hypotheses and report on experiments on mutual modelling, we use  the notation
+\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This  notation is not
+proposed as a formal theory of mutual modelling but as useful simplification for
+communicating about mutual modelling. This notation does not mean $A$ has an
+explicit, monolithic representation of $B$: it must be understood as an
+abstraction referring to complex socio-cognitive processes.  As explained in the
+previous section, $A$ can be fragmented, multi-dimensional, etc. This notation
+is neither presented as a computational model of mutual modelling, nor as some
+universal formalism; it's usefulness is internal to this paper.  Besides, we
+refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
+discuss in the next section the methodological difficulty in measuring this
+accuracy.
 
@@
 
-Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$ about
-content $X$ at time $t_1$ during task T in order to increase \Model{A}{B}{X}
-at time $t_2$ is the ratio between how much it is needed  ((2)-(1)) and how much it
-costs (3)~\citep{traum1996miscommunication}:
+Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$
+about content $X$ at time $t_1$ during task T in order to increase
+\Model{A}{B}{X} at time $t_2$ is the ratio between how much it is needed
+((2)-(1)) and how much it costs (3)~\citep{traum1996miscommunication}:
 
@@
 
-This formula is presented as a qualitative summary, not as a real equation, since several parameters
-are hard to quantify (\eg the cost of a communication act depends upon the user
-as well). We use it for clarifying the parameters of the experiments we present hereafter.
+This formula is presented as a qualitative summary, not as a real equation,
+since several parameters are hard to quantify (\eg the cost of a communication
+act depends upon the user as well). We use it for clarifying the parameters of
+the experiments we present hereafter.
 
@@
 
-Since a mental model is not directly observable,  the study of mutual modelling is methodologically difficult.
-How can we for instance measure \Model{A}{B}{X}.
-We proceed in two steps, first to capture \model{A}{B}{X} and then to
-estimate \Model{A}{B}{X}.
-
-{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$
-        what he/she believes about what $B$ knows, feels, intends to do, etc.
-        This raises obvious methodological concerns since such a question
-        triggers a modelling process beyond what would naturally occur. To avoid
-        this bias, one can estimate mutual modelling after task completion.
-        Then, the obvious drawback are memory losses and by post-hoc
-        reconstruction.  The first method was used in study 2 and the second one
-        in the other studies. Another option would be to rely on external
-        behavioural metrics like eye-tracking: we hypothesise for instance that
-        the fixation time reflects the efforts engaged by the human to
-        understand, hence, model, the others. Such an approach has however not
-        been investigated in the presented studies.
-
-    {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we
-        need to access the reference model \refmodel{B}{X} to estimate its
-        \emph{accuracy}. Since we can only indirectly access it via what $B$
-        reports (\ie \model{B}{B}{X}), accuracy can be estimated in 2 ways:
+Since a mental model is not directly observable,  the study of mutual modelling
+is methodologically difficult. How can we for instance measure \Model{A}{B}{X}.
+We proceed in two steps, first to capture \model{A}{B}{X} and then to estimate
+\Model{A}{B}{X}.
+
+{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what he/she
+believes about what $B$ knows, feels, intends to do, etc.  This raises obvious
+methodological concerns since such a question triggers a modelling process
+beyond what would naturally occur. To avoid this bias, one can estimate mutual
+modelling after task completion.  Then, the obvious drawback are memory losses
+and by post-hoc reconstruction.  The first method was used in study 2 and the
+second one in the other studies. Another option would be to rely on external
+behavioural metrics like eye-tracking: we hypothesise for instance that the
+fixation time reflects the efforts engaged by the human to understand, hence,
+model, the others. Such an approach has however not been investigated in the
+presented studies.
+
+{\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we need to
+access the reference model \refmodel{B}{X} to estimate its \emph{accuracy}.
+Since we can only indirectly access it via what $B$ reports (\ie
+\model{B}{B}{X}), accuracy can be estimated in 2 ways:
 
@@
 
-            \item Subjective accuracy: In study 3, we compute
-                \Model{A}{B}{X} by measuring if $A$ describes
-                $B$'s emotions in the same way $B$ reports its emotions 
-                ($\M{A}{B}{X} = \M{B}{B}{X}$).
-
-            \item Objective accuracy: In studies 4 and 5, we compute
-                \Model{A}{B}{X} by comparing \Model{A}{B}{K} to $B$'s
-                actual knowledge as it has measured by a test.
+    \item Subjective accuracy: In study 3, we compute \Model{A}{B}{X} by
+        measuring if $A$ describes $B$'s emotions in the same way $B$ reports
+        its emotions ($\M{A}{B}{X} = \M{B}{B}{X}$).
+
+    \item Objective accuracy: In studies 4 and 5, we compute \Model{A}{B}{X} by
+        comparing \Model{A}{B}{K} to $B$'s actual knowledge as it has measured
+        by a test.
 
@@
 
-Our method for investigating mutual modelling is observe the variations of  accuracy that result from variations of external parameters that appear as variable in the formula presented in the previous section: the accuracy should for go down if we increase the cost of grounding acts and goes up if we increase the grounding criterion (the necessity to build a shared understanding).
-One way to produce these variations is what CSCW,researchers have called
-``awareness tools'', \ie functionalities that inform $A$ is about $B$'s actions
-that A could not directly perceive because $B$ was working on a different subset
-of the virtual space. Different awareness tools are used in the following studies as methodological tricks  to manipulate experimentally the  mutual modelling activity.
+Our method for investigating mutual modelling is observe the variations of
+accuracy that result from variations of external parameters that appear as
+variable in the formula presented in the previous section: the accuracy should
+for go down if we increase the cost of grounding acts and goes up if we increase
+the grounding criterion (the necessity to build a shared understanding). One way
+to produce these variations is what CSCW,researchers have called ``awareness
+tools'', \ie functionalities that inform $A$ is about $B$'s actions that A could
+not directly perceive because $B$ was working on a different subset of the
+virtual space. Different awareness tools are used in the following studies as
+methodological tricks  to manipulate experimentally the  mutual modelling
+activity.
 
@@
 different institutions by different researchers. They used different
-independent, intermediate and dependent variables. Nonetheless, we were able to retroactively
-address 3 research questions across these studies.
+independent, intermediate and dependent variables. Nonetheless, we were able to
+retroactively address 3 research questions across these studies.
 
@@
 
-As stated earlier, the fundamental challenge is to determine if mutual modelling is an individual skill (hypothesis 1) or  the emergent property of social interactions (hypothesis 3).
-This question is empirically translated into the symmetry of mutual modelling: what is
-the relationship between \Model{A}{B}{X} and \Model{B}{A}{X}
-(Fig.~\ref{mm_symmetry})? A low symmetry would mean that mutual modelling is
-mainly an individual attitude/aptitude ($\mathcal{H}_{1}$). A high correlation
-might support $\mathcal{H}_{3}$ since there is a low
+As stated earlier, the fundamental challenge is to determine if mutual modelling
+is an individual skill (hypothesis 1) or  the emergent property of social
+interactions (hypothesis 3).  This question is empirically translated into the
+symmetry of mutual modelling: what is the relationship between \Model{A}{B}{X}
+and \Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that
+mutual modelling is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$).
+A high correlation might support $\mathcal{H}_{3}$ since there is a low
 probability that randomly formed pairs integrate peers with the same level of
-mutual modelling skills.
-However, a high correlation could also have another explanation, stated in  $\mathcal{H}_{2}$: it could be that A is good at modelling B and good at helping B to model herself or himself.
+mutual modelling skills. However, a high correlation could also have another
+explanation, stated in  $\mathcal{H}_{2}$: it could be that A is good at
+modelling B and good at helping B to model herself or himself.
 
@@
 
-With triads, we may compute the accuracy of 6 models:
-\Model{A}{B}{X}, \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X},
-\Model{C}{B}{X} and \Model{B}{C}{X}.This leads to two \emph{triangle questions} (Fig.~\ref{mm_triangles}): Do $A$
-and $B$ have the same accuracy when modelling $C$ ($\Delta_2 =
-\Delta(\mathcal{M}^{\circ}(A,C,X), \mathcal{M}^{\circ}(B,C,X))$)? A significant correlation would support
-, $\mathcal{H}_{2}$  (C has helped both A and B to model C) or support $\mathcal{H}_{3}$ (the quality of triad interactions enables all partners to model each other accurately)
+With triads, we may compute the accuracy of 6 models: \Model{A}{B}{X},
+\Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X}, \Model{C}{B}{X} and
+\Model{B}{C}{X}.This leads to two \emph{triangle questions}
+(Fig.~\ref{mm_triangles}): Do $A$ and $B$ have the same accuracy when modelling
+$C$ ($\Delta_2 = \Delta(\mathcal{M}^{\circ}(A,C,X),
+\mathcal{M}^{\circ}(B,C,X))$)? A significant correlation would support ,
+$\mathcal{H}_{2}$  (C has helped both A and B to model C) or support
+$\mathcal{H}_{3}$ (the quality of triad interactions enables all partners to
+model each other accurately)
 
@@
 \Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
-answer would support $\mathcal{H}_{1}$, C being simply good at modelling any partner.
-It could also support  $\mathcal{H}_{3}$,  since the quality of interactions influences the accuracy of two models that C is building.
-A negative answer would support
-$\mathcal{H}_{2}$ since $\Delta_3$ could mainly be explained by the fact that  $A$ helped more $C$ to model him than 
-$B$ did.
+answer would support $\mathcal{H}_{1}$, C being simply good at modelling any
+partner.  It could also support  $\mathcal{H}_{3}$,  since the quality of
+interactions influences the accuracy of two models that C is building. A
+negative answer would support $\mathcal{H}_{2}$ since $\Delta_3$ could mainly be
+explained by the fact that  $A$ helped more $C$ to model him than $B$ did.
 
 In addition, the comparison between $\Delta_2$ and $\Delta_3$ could tell us
-whether the accuracy of mutual modelling depends more upon the modeller's effort  $\mathcal{H}_{1}$
-or the modellee's behaviour  $\mathcal{H}_{2}$.
-
-Please note that we consider the quality of interactions at triad level (ABC), neglecting the cases where AB interact better for instance than BC, since there was no 'private' communication channel in the following studies.
-We do nonetheless acknowledge that this point could be debated.
+whether the accuracy of mutual modelling depends more upon the modeller's effort
+$\mathcal{H}_{1}$ or the modellee's behaviour  $\mathcal{H}_{2}$.
+
+Please note that we consider the quality of interactions at triad level (ABC),
+neglecting the cases where AB interact better for instance than BC, since there
+was no 'private' communication channel in the following studies. We do
+nonetheless acknowledge that this point could be debated.
 
@@
 We can go further by comparing self- versus other modelling ($\Delta_4$ in
-Fig.~\ref{mm_rectangle}).
-A large difference would indicate that meta-cognitive skills (self-modelling) and partner modelling skills are rather different skills, while a small $\Delta_4$ could be interpreted as the indications that these are two specific instance of a more general cognitive process.
-This questions is however not central to this paper, and the value of $\Delta_4$ us only available in one of the reported studies.
-
- We can also question if modelling skills depend upon what aspects are being modeled ($X$ or
-$Y$), which would explain vertical differences ($\Delta_5$ in
-Fig.~\ref{mm_rectangle}).
-These differences would allow refining the notion of modelling skills, namely whether there exist some general ability to model partners or whether this is only the abstraction of a beam of more specific skills such as detecting emotions versus identifying references from deictic gestures.
+Fig.~\ref{mm_rectangle}). A large difference would indicate that meta-cognitive
+skills (self-modelling) and partner modelling skills are rather different
+skills, while a small $\Delta_4$ could be interpreted as the indications that
+these are two specific instance of a more general cognitive process. This
+questions is however not central to this paper, and the value of $\Delta_4$ us
+only available in one of the reported studies.
+
+We can also question if modelling skills depend upon what aspects are being
+modeled ($X$ or $Y$), which would explain vertical differences ($\Delta_5$ in
+Fig.~\ref{mm_rectangle}).  These differences would allow refining the notion of
+modelling skills, namely whether there exist some general ability to model
+partners or whether this is only the abstraction of a beam of more specific
+skills such as detecting emotions versus identifying references from deictic
+gestures.
 
@@
 
-We report five studies conducted  by different researchers in different contexts between 2000 and 2015.
-They do not form a consistent research strategy but actually the fact that some trends emerge despite their diversity constitutes indeed the richness of this line of work.
+We report five studies conducted  by different researchers in different contexts
+between 2000 and 2015. They do not form a consistent research strategy but
+actually the fact that some trends emerge despite their diversity constitutes
+indeed the richness of this line of work.
 
@@
 %modelling~\citep{nova2007collaboration}. The availability
-modelling \textit{[citation removed for blind review]}. The availability
-of an awareness tool was our independent variable. In previous studies, we
-replayed a video of the game to subjects who surprised us by their ability to
-remember former states of their mutual model: ``I did that because I thought that
-you would do that''. Hence, this experiment focused the representation of each
+modelling \textit{[citation removed for blind review]}. The availability of an
+awareness tool was our independent variable. In previous studies, we replayed a
+video of the game to subjects who surprised us by their ability to remember
+former states of their mutual model: ``I did that because I thought that you
+would do that''. Hence, this experiment focused the representation of each
 other's action plans. During the game, we asked them to anticipate the next
@@
 
-{\sc SpaceMiners} is a 3D game that involves two players collecting minerals located
-in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch drones through
-the space after choosing their initial direction and speed. Once launched, the
-trajectory of drones is only influenced by the gravity of planets and by
-trajectory modification tools.  During the experiment, the teams were confronted
-with three increasingly complex situations. The experiment was 2 hours long: a
-30 minute tutorial and 3 levels of 30 minutes. Players uses using a regular
-joystick and communicated with each other through an audio channel.
+{\sc SpaceMiners} is a 3D game that involves two players collecting minerals
+located in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch
+drones through the space after choosing their initial direction and speed. Once
+launched, the trajectory of drones is only influenced by the gravity of planets
+and by trajectory modification tools.  During the experiment, the teams were
+confronted with three increasingly complex situations. The experiment was 2
+hours long: a 30 minute tutorial and 3 levels of 30 minutes. Players uses using
+a regular joystick and communicated with each other through an audio channel.
 
@@
 player $A$  the location and gaze direction of player $B$. In the awareness
-condition, players could switch to the \emph{scout mode} where they could view what
-their partner was looking at. We hypothesize that this would enable subjects to
-more accurately infer their teammate's intentions. Each player sat in front of a
-distinct computer located in different rooms.
+condition, players could switch to the \emph{scout mode} where they could view
+what their partner was looking at. We hypothesize that this would enable
+subjects to more accurately infer their teammate's intentions. Each player sat
+in front of a distinct computer located in different rooms.
 
@@
 How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
-It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$.
-It is nonetheless far from 1 which implies we cannot discard the individual modelling skills.
-We collect more evidence along the next studies.
+It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. It is
+nonetheless far from 1 which implies we cannot discard the individual modelling
+skills. We collect more evidence along the next studies.
 
@@
 %space~\citep{nova2006underwhelming}. We studied whether players build an accurate
-space \textit{[citation removed for blind review]}. We studied whether players build an accurate
-model of the path followed by their partners, assuming this path reflected their
-problem solving strategy. We used an objective measure of \gModel{A}{B}: the
-distance between where $A$ believes $B$ has been walking and where $B$ actually went.
-The main hypothesis concerned the effect of awareness tools on group performance
-and on \gModel{A}{B}.
+space \textit{[citation removed for blind review]}. We studied whether players
+build an accurate model of the path followed by their partners, assuming this
+path reflected their problem solving strategy. We used an objective measure of
+\gModel{A}{B}: the distance between where $A$ believes $B$ has been walking and
+where $B$ actually went.  The main hypothesis concerned the effect of awareness
+tools on group performance and on \gModel{A}{B}.
 
@@
 
-This positive correlation observed in the symmetry question confirms the first study.
-In this case, this is actually more surprising given  the high heterogeneity of spatial skills among adults~\citep{liben1981spatial}.
-This result supports  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$.
-The results regarding $\Delta_2$ may support both  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$  but discards   $\mathcal{H}_{1}$  (if the modeller's skill would matter a lot,  $\Delta_2$ would be lower).
-Conversely, $\Delta_3$  supports $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$ but discards   $\mathcal{H}_{2}$  (if the modeller's skill would matter a lot,  $\Delta_3$ would be lower).
-In summary, various indices support the 3 hypotheses, which implies there is some truth in each of them, but $\mathcal{H}_{3}$ is the only hypothesis that is not rejected by any index.
-We may hence, with great caution, concluded that the social perspective ($\mathcal{H}_{3}$) is moderately reinforced by this second study.
-Since the values of $\Delta_2$ and $\Delta_3$ are very similar, we do not interpret their minor difference as evidence for a stronger role of the modeller ($\mathcal{H}_{1}$) or the modellee ($\mathcal{H}_{1}$).
-
-We have however to bring some nuances to the social viewpoint ($\mathcal{H}_{3}).
-The main feature that can be associated to the team level in this experiment is probably not the quality of their verbal interactions per se (they interact mostly by drawing on a shared map), but rather the consistency of  the spatial exploration strategy: a clear strategy facilitates memorizing one's partner path.
-One could argue whether the team strategy can be dissociated from the team interactions quality or constitues one of its components.
+This positive correlation observed in the symmetry question confirms the first
+study.  In this case, this is actually more surprising given  the high
+heterogeneity of spatial skills among adults~\citep{liben1981spatial}. This
+result supports  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The results regarding
+$\Delta_2$ may support both  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$  but
+discards   $\mathcal{H}_{1}$  (if the modeller's skill would matter a lot,
+$\Delta_2$ would be lower). Conversely, $\Delta_3$  supports $\mathcal{H}_{1}$
+and  $\mathcal{H}_{3}$ but discards   $\mathcal{H}_{2}$  (if the modeller's
+skill would matter a lot,  $\Delta_3$ would be lower). In summary, various
+indices support the 3 hypotheses, which implies there is some truth in each of
+them, but $\mathcal{H}_{3}$ is the only hypothesis that is not rejected by any
+index. We may hence, with great caution, concluded that the social perspective
+($\mathcal{H}_{3}$) is moderately reinforced by this second study. Since the
+values of $\Delta_2$ and $\Delta_3$ are very similar, we do not interpret their
+minor difference as evidence for a stronger role of the modeller
+($\mathcal{H}_{1}$) or the modellee ($\mathcal{H}_{1}$).
+
+We have however to bring some nuances to the social viewpoint
+($\mathcal{H}_{3}$).  The main feature that can be associated to the team level
+in this experiment is probably not the quality of their verbal interactions per
+se (they interact mostly by drawing on a shared map), but rather the consistency
+of  the spatial exploration strategy: a clear strategy facilitates memorizing
+one's partner path.  One could argue whether the team strategy can be
+dissociated from the team interactions quality or constitues one of its
+components.
 
@@
 %\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.}
-The aim of this unpublished
-study was to evaluate the effect of media richness on \Model{A}{B}{emotions}.
-The hypothesis was that video communication would lead to a better \gModel{A}{B}
-than audio only since emotions often impact on facial expressions.
+The aim of this unpublished study was to evaluate the effect of media richness
+on \Model{A}{B}{emotions}.  The hypothesis was that video communication would
+lead to a better \gModel{A}{B} than audio only since emotions often impact on
+facial expressions.
 
@@
 interpret this too far, but to show that there are large variations even within one
-area  (perceiving emotions). These variations shomehow question the existence of a general aptitude to model others ( $\mathcal{H}_{1}$).
+area  (perceiving emotions). These variations shomehow question the existence of
+a general aptitude to model others ( $\mathcal{H}_{1}$).
 
@@
 
-In this third study, the accuracy of mutual modelling  between two peers tend to be symmetrical, which  supports $\mathcal{H}_{2}$ and ($\mathcal{H}_{3}$.)  This result is contradicted by the fact that ($\Delta_2$) is not significant, which rejects both
- $\mathcal{H}_{2}$ and, to a lower extent,,  $\mathcal{H}_{3}$.
+In this third study, the accuracy of mutual modelling  between two peers tend to
+be symmetrical, which  supports $\mathcal{H}_{2}$ and ($\mathcal{H}_{3}$.)  This
+result is contradicted by the fact that ($\Delta_2$) is not significant, which
+rejects both $\mathcal{H}_{2}$ and, to a lower extent,,  $\mathcal{H}_{3}$.
 Finally,  ($\Delta_3$) supports both  $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$.
-In summary, this study brings some supports to  $\mathcal{H}_{3}$, but reveal again that $\mathcal{H}_{3}$ is only part of the explanation.
-
-Like the previous one, this study leads us to refine what we mean by quality of interactions in $\mathcal{H}_{3}$.
-We  expected that the video channel would help peers building a more accurate model of the each other's emotions.
-The results show the opposite: peers in the audio-only condition built more accurate models, probably because they concentrated more on the shared concept map.
-This confirms other studies that revealed that viewing what one's partner sees (shared graphical editor) is more important
-than seeing each other~\citep{gaver1993one,anderson1997impact}.
-Therefore, what is meant by the quality of interactions is $\mathcal{H}_{3} is more than the linguistic features of dialogue but includes the way these interactions are articulated to the task.
+In summary, this study brings some supports to  $\mathcal{H}_{3}$, but reveal
+again that $\mathcal{H}_{3}$ is only part of the explanation.
+
+Like the previous one, this study leads us to refine what we mean by quality of
+interactions in $\mathcal{H}_{3}$. We  expected that the video channel would
+help peers building a more accurate model of the each other's emotions.  The
+results show the opposite: peers in the audio-only condition built more accurate
+models, probably because they concentrated more on the shared concept map. This
+confirms other studies that revealed that viewing what one's partner sees
+(shared graphical editor) is more important than seeing each
+other~\citep{gaver1993one,anderson1997impact}. Therefore, what is meant by the
+quality of interactions is $\mathcal{H}_{3}$ is more than the linguistic features
+of dialogue but includes the way these interactions are articulated to the task.
 
@@
 %learning~\citep{molinari2008effects}. The script chosen is a {\sc jigsaw}: two students
-learning \textit{[citation removed for blind review]}. The script chosen is a {\sc jigsaw}: two students
-receive different but complementary subsets of the knowledge (texts) which have
-to be integrated to build a shared concept map.  This script increases the
-cognitive effort to build the map, not only to conciliate the viewpoints of
-each team members but, before that, to find out what the other knows.
+learning \textit{[citation removed for blind review]}. The script chosen is a
+{\sc jigsaw}: two students receive different but complementary subsets of the
+knowledge (texts) which have to be integrated to build a shared concept map.
+This script increases the cognitive effort to build the map, not only to
+conciliate the viewpoints of each team members but, before that, to find out
+what the other knows.
 
@@
 
-\paragraph{Rectangle question} The correlation between \gmodel{A}{A}
-and \gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
+\paragraph{Rectangle question} The correlation between \gmodel{A}{A} and
+\gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
 significant in each condition: someone good at self-modelling is not necessarily
-good at modelling someone else and vice-versa.  This seems to indicate that partner modelling
-requires different skills than meta-cognition, despite their similarity at some level of abstraction.
+good at modelling someone else and vice-versa.  This seems to indicate that
+partner modelling requires different skills than meta-cognition, despite their
+similarity at some level of abstraction.
 
@@
 
-In this experiment, the symmetry of mutual modelling is  found but  in one condition, namely when subjects
- receive the same information before the task.  This condition corresponds to the situation tested in the three
-previous studies and supports  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$.
-
-How do we interpret the fact that the symmetry vanishes when peers receive different texts to read before the task?
-One explanation would be the difficult of mutual modelling when peers don't know what others read, but we found no significance of \gModel{A}{A} between conditions.
-Since texts were partly overlapping, another explanation is that, in absence of these initial common grounds, mutual modelling requires A to make evident to B what A thinks B does not know about A, which is the second level of modelling described in $\mathcal{H}_{2}$ .
-A low symmetry means that some peers are better than others at this second level of modelling, which supports the existence of such an individual skill, as stated in  $\mathcal{H}_{2}$
+In this experiment, the symmetry of mutual modelling is  found but  in one
+condition, namely when subjects receive the same information before the task.
+This condition corresponds to the situation tested in the three previous studies
+and supports  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$.
+
+How do we interpret the fact that the symmetry vanishes when peers receive
+different texts to read before the task? One explanation would be the difficult
+of mutual modelling when peers don't know what others read, but we found no
+significance of \gModel{A}{A} between conditions. Since texts were partly
+overlapping, another explanation is that, in absence of these initial common
+grounds, mutual modelling requires A to make evident to B what A thinks B does
+not know about A, which is the second level of modelling described in
+$\mathcal{H}_{2}$ . A low symmetry means that some peers are better than others
+at this second level of modelling, which supports the existence of such an
+individual skill, as stated in  $\mathcal{H}_{2}$
 
@@
 KAT informed them about the pre-test score. Actually, pairs in the KAT condition
-produced significantly more instances of 3 interesting categories of interactions:
- (1) utterances asking about the
-other's knowledge such as ``Did you understand how transmission works?'' (2)
-utterances describing one's own knowledge (\gmodel{A}{A}) such as ``I don't
-remember the Ranvier's thing...'' and (3) elaborated utterances with rich contents,  These 3 categories provide different account
-of \gModel{A}{B}: (1) as an
-effect of $A$'s effort to model $B$  ($\mathcal{H}_{1}$), (2) as $B$'s effort to give cues to $A$ about his
-own knowledge  ($\mathcal{H}_{2}$) and (3) as an effect of the quality of interaction  ($\mathcal{H}_{3}$) .
+produced significantly more instances of 3 interesting categories of
+interactions: (1) utterances asking about the other's knowledge such as ``Did
+you understand how transmission works?'' (2) utterances describing one's own
+knowledge (\gmodel{A}{A}) such as ``I don't remember the Ranvier's thing...''
+and (3) elaborated utterances with rich contents,  These 3 categories provide
+different account of \gModel{A}{B}: (1) as an effect of $A$'s effort to model
+$B$  ($\mathcal{H}_{1}$), (2) as $B$'s effort to give cues to $A$ about his own
+knowledge  ($\mathcal{H}_{2}$) and (3) as an effect of the quality of
+interaction  ($\mathcal{H}_{3}$) .
 
@@
 \paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
-0.05, ns$), which discards  ($\mathcal{H}_{2}$) and  ($\mathcal{H}_{3}$).
-In addition, the KAT supports the processing of modelling at the first level ($\mathcal{H}_{1}$) and not at the second level ($\mathcal{H}_{2}$).
-Hence the fact that the KAT enhances  \gModel{A}{B}  brings additional supports to $\mathcal{H}_{1}$.
-However, the analysis of verbal interactions find elements that actually supports the three hypotheses.
+0.05, ns$), which discards  ($\mathcal{H}_{2}$) and  ($\mathcal{H}_{3}$). In
+addition, the KAT supports the processing of modelling at the first level
+($\mathcal{H}_{1}$) and not at the second level ($\mathcal{H}_{2}$). Hence the
+fact that the KAT enhances  \gModel{A}{B}  brings additional supports to
+$\mathcal{H}_{1}$. However, the analysis of verbal interactions find elements
+that actually supports the three hypotheses.
 
@@
 
-We found symmetry of the mutual model on 4 studies out of 5.
-In study 4, this only applies to the control group (having read the same text before), which corresponds to the situation of the 3 first studies.
-Finding this symmetry per se constitutes an interesting result as we don't know earlier studies who have established this relationship.
-The problem is that this symmetry does not allow use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$.
-The second hypothesis, the skill of helping one's peer to model oneself, is questioned by study 3 where C could be modelled  accurately by A and not accurately by B.
-The same hypothesis is however supported by the results of study 2, and indirectly by study 4.
-This does not discard \mathcal{H}_{1}$: even when we found significant correlations, they were all below 0.50, not around 0.90.
-Hence, even of the quality of social interaction matters, there is obviously a large part of individual variance within teams.
-
-In other words, these studies do conclude that one hypothesis is right and the others are wrong, which is per se the main contribution of this paper.
-We started from the idea that partner modelling is essentially an individual skill  and that we therefore improve the quality of collaboration by providing the so-called awareness tools, which act as some kind of prostheses for partner modelling.
-This individual skill cannot be discarded.
-Everyone has experienced the pleasure of interacting with colleagues who answer precisely to the question we asked them  and even guess the reason why we asked this question.
-Conversely, everyone also experienced the frustration of someone referring about a third person by his name, say Mike Smith,  while knowing perfectly that there is no chance that we know this Mike Smith.
-The existence of these individual skills is not denied by our studies; the novelty of this paper is to show, thanks to symmetry values, that individual skills only account  for a part of the accuracy in mutual modelling.
-
-We acknowledge that the difference between these hypotheses is  rather theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$) and the process of helping one's partner to model onself ($\mathcal{H}_{3}$) are mediated by verbal interactions in the team.
-One could hardly imagine someone managing a very accurate modelling despite the very bad quality of interactions.
-There is a bidirectional causal link between the accuracy of mutual modelling and the quality of interactions.
-This rather artificial distinction does however participate to the more fundamental discussion of the role of individual and social mechanisms in human cognition.
-In the field of social cognition, it is a common place to state that 'the whole is greater than the sum of the parts'.
-This refers to the emergence of team properties than cannot be reduced to the set of individual contributions.
-Our paper illustrates this emergence by showing the symmetry of mutual modelling
+We found symmetry of the mutual model on 4 studies out of 5. In study 4, this
+only applies to the control group (having read the same text before), which
+corresponds to the situation of the 3 first studies.  Finding this symmetry per
+se constitutes an interesting result as we don't know earlier studies who have
+established this relationship. The problem is that this symmetry does not allow
+use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The second
+hypothesis, the skill of helping one's peer to model oneself, is questioned by
+study 3 where C could be modelled  accurately by A and not accurately by B. The
+same hypothesis is however supported by the results of study 2, and indirectly
+by study 4.  This does not discard $\mathcal{H}_{1}$: even when we found
+significant correlations, they were all below 0.50, not around 0.90.  Hence,
+even of the quality of social interaction matters, there is obviously a large
+part of individual variance within teams.
+
+In other words, these studies do conclude that one hypothesis is right and the
+others are wrong, which is per se the main contribution of this paper. We
+started from the idea that partner modelling is essentially an individual skill
+and that we therefore improve the quality of collaboration by providing the
+so-called awareness tools, which act as some kind of prostheses for partner
+modelling. This individual skill cannot be discarded.  Everyone has experienced
+the pleasure of interacting with colleagues who answer precisely to the question
+we asked them  and even guess the reason why we asked this question. Conversely,
+everyone also experienced the frustration of someone referring about a third
+person by his name, say Mike Smith,  while knowing perfectly that there is no
+chance that we know this Mike Smith. The existence of these individual skills is
+not denied by our studies; the novelty of this paper is to show, thanks to
+symmetry values, that individual skills only account  for a part of the accuracy
+in mutual modelling.
+
+We acknowledge that the difference between these hypotheses is  rather
+theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$)
+and the process of helping one's partner to model onself ($\mathcal{H}_{3}$) are
+mediated by verbal interactions in the team. One could hardly imagine someone
+managing a very accurate modelling despite the very bad quality of interactions.
+There is a bidirectional causal link between the accuracy of mutual modelling
+and the quality of interactions. This rather artificial distinction does however
+participate to the more fundamental discussion of the role of individual and
+social mechanisms in human cognition. In the field of social cognition, it is a
+common place to state that 'the whole is greater than the sum of the parts'.
+This refers to the emergence of team properties than cannot be reduced to the
+set of individual contributions.  Our paper illustrates this emergence by
+showing the symmetry of mutual modelling
 
@@
 
-Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do not
-imply we have a mechanical view of modelling. It's just a language for referring
-to different components of modelling each other in teams.  This simple notation
-emerged from the need to compare five mutual modelling studies. It could however pave the road for elaborating a systematic agenda for research on mutual modelling.
+Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do
+not imply we have a mechanical view of modelling. It's just a language for
+referring to different components of modelling each other in teams.  This simple
+notation emerged from the need to compare five mutual modelling studies. It
+could however pave the road for elaborating a systematic agenda for research on
+mutual modelling.
 

### 2016-02-04 mutual-modeling.tex
Typos

@@
 scientific controversy. We postulate that collaborative tasks require some
-level of mutual modeling, i.e. that each partner needs some model of what
+level of mutual modelling, i.e. that each partner needs some model of what
 the other partners know/want/intend at a given time.  We use the term 'some
@@
 discard the fact individual abilities also  participate in the quality of
-modellng process .
+modelling process .
 
@@
 collaborative learning, the situation is different; tasks are precisely designed
-for requirint a deeper negotiation fo meaning.  The craft of CSCL is to design
-collaborative situations that create interdependence, avoid groupthink, and
+for requiring a deeper negotiation of meaning.  The craft of CSCL is to design
+collaborative situations that create interdependence, avoid group think, and
 allows students to detect any 'illusion of shared understanding' (Cherubini et
@@
 quality of these two models could be computed.  The studies we report do hence
-not constitue a clean sequence of experiment to investigate mutual modelling but
+not constitute a clean sequence of experiment to investigate mutual modelling but
 the ad-hoc revisiting of previous experiments to explore a question that we had
@@
 the grounding criterion (the necessity to build a shared understanding). One way
-to produce these variations is what CSCW,researchers have called ``awareness
+to produce these variations is what CSCW researchers have called ``awareness
 tools'', \ie functionalities that inform $A$ is about $B$'s actions that A could
@@
 one's partner path.  One could argue whether the team strategy can be
-dissociated from the team interactions quality or constitues one of its
+dissociated from the team interactions quality or constitutes one of its
 components.
@@
 interpret this too far, but to show that there are large variations even within one
-area  (perceiving emotions). These variations shomehow question the existence of
+area  (perceiving emotions). These variations somehow question the existence of
 a general aptitude to model others ( $\mathcal{H}_{1}$).
@@
 theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$)
-and the process of helping one's partner to model onself ($\mathcal{H}_{3}$) are
+and the process of helping one's partner to model oneself ($\mathcal{H}_{3}$) are
 mediated by verbal interactions in the team. One could hardly imagine someone

### 2016-02-04 mutual-modeling.tex
Proof-reading sections 1 2 3

@@
 
-\newcommand{\ie}{{\textit{i.e.\ }}}
-\newcommand{\cf}{{\textit{cf\ }}}
-\newcommand{\eg}{{\textit{e.g.\ }}}
-\newcommand{\al}{{\textit{et al.\ }}}
+\usepackage{xspace}
+\newcommand{\ie}{i.e.\xspace}
+\newcommand{\cf}{{\textit{cf\xspace}}}
+\newcommand{\eg}{{\textit{e.g.\xspace}}}
+\newcommand{\al}{{\textit{et al.\xspace}}}
+
+\newcommand{\A}{A\xspace}
+\newcommand{\B}{B\xspace}
+\newcommand{\C}{C\xspace}
 
@@
 scientific controversy. We postulate that collaborative tasks require some
-level of mutual modelling, i.e. that each partner needs some model of what
-the other partners know/want/intend at a given time.  We use the term 'some
-model' to stress the fact that this model is not necessarily details or
+level of mutual modelling, \ie that each partner needs some model of what
+the other partners know/want/intend at a given time.  We use the term ``some
+model'' to stress the fact that this model is not necessarily detailed or
 complete, but that we acquire some representations of the persons we
 interact with. The question we address is: Does the quality of the partner
-model depend upon the modeler's ability to represent his or her partner,
-upon the modelee's ability to make his state clear to the modeler or upon
-the quality of their interactions? We address this question by analyzing
-differences between the  accuracy of the models built by different team
-members. We therefore report 5 experiments on collaborative problem solving
+model depend upon the modeler's ability to represent his or her partner?
+Upon the modelee's ability to make his state clear to the modeler? Or rather upon
+the quality of their interactions? We address this question by comparing
+the respective accuracies of the models built by different team
+members. We therefore report on 5 experiments on collaborative problem solving
 or collaborative learning that vary in terms of tasks (how important it is
 to build an accurate model) and settings (how difficult it is to build an
-accurate model). In 4 studies, the accuracy of the model that A built about
-B was correlated with the accuracy of the model that B built about A,  which
+accurate model). In 4 studies, the accuracy of the model that \A built about
+\B was correlated with the accuracy of the model that \B built about \A,  which
 seems to imply that the quality of interactions matters more than individual
 abilities when building mutual models.  However, these finding do not
-discard the fact individual abilities also  participate in the quality of
-modelling process .
+rule out the fact that individual abilities also participate in the quality of
+modelling process.
 
@@
 
-From its very beginning, CSCL research has been following the suggestion by
-\citet{roschelle1995construction} that collaborative learning has something to
-do with the process of constructing and maintaining a \emph{shared
-understanding} of the task at hand. Building a shared/mutual understanding
-refers to the upper class of collaborative learning situations, those in which
-students should build upon each other's understanding to refine their own
-understanding: what is expected to produce learning is not the mere fact that
-two students build the same understanding but the cognitive effort they have to
-engage to build this shared understanding~\citep{schwartz1995emergence}. This
-effort can be observed by the frequency of rich interactions, \ie interactions
-whose occurrence has been related to learning: (self-) explanations in cognitive
+From its inception, Computer Supported Collaborative Learning (CSCL)
+research has been following the suggestion by \citet{roschelle1995construction}
+that collaborative learning has something to do with the process of constructing
+and maintaining a \emph{shared understanding} of the task at hand. Building a
+shared/mutual understanding refers to the upper class of collaborative learning
+situations, those in which students should build upon each other's understanding
+to refine their own understanding: what is expected to produce learning is not
+the mere fact that two students build the same understanding but the cognitive
+effort they have to engage to build this shared
+understanding~\citep{schwartz1995emergence}. This effort can be observed by the
+frequency of rich interactions, \ie interactions whose occurrence has been
+related to learning: (self-) explanations in cognitive
 science~\citep{chi1989self, webb1991task}, conflict resolution in
@@
 mechanisms for explaining learning outcomes has been questioned in the learning
-sciences. Grounding mechanisms  are appropriate for explains conversational
-events, such referential failures in short dialogue episodes, but they do
-hardly predict deeper phenomena such as \emph{conceptual change} (\ie the
-acquisition, acceptation and integration of a new belief into one's mental
-model) over longer sessions~\citep{dillenbourg2006sharing}. The cumulative
-effect of grounding episodes can probably be better understood from a
-socio-cultural perspective:
+sciences. Grounding mechanisms are appropriate to explain conversational
+events, such as referential failures in short dialogue episodes, but they do hardly
+predict deeper phenomena such as \emph{conceptual change} (\ie the acquisition,
+acceptation and integration of a new belief into one's mental model) over longer
+sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
+episodes can probably be better understood from a socio-cultural perspective:
 
@@
 dialogue, mostly between the lexical/syntactical level and the deeper semantic
-levels. \citet{pickering2006alignment} argue that the mutual understanding
+levels. \citet{pickering2006alignment} argue that mutual understanding
 starts mostly with a \emph{superficial alignment} at the level of the linguistic
@@
 about which intervention will be conducted in which way on which patient.  In
-this paper, we operationalized the grounding criterion, i.e. the necessity for
-accurate modelling, as the correlation between the accuracy of partner models
-and various measures of team performance.
+this paper, we operationalized the grounding criterion, \ie the necessity for
+accurate modelling, as \emph{the correlation between the accuracy of partner models
+and measures of team performance}.
 
@@
 learning. In education, the question is to relate these mechanisms to learning
-outcomes: \textbf{Is linguistic alignment sufficient to trigger conceptual
-change?} \textbf{Does negotiation of meaning only occur when partners monitor
-and diagnose each other's knowledge?} If the ratio between shallow alignment and
+outcomes: Is linguistic alignment sufficient to trigger conceptual
+change? Does negotiation of meaning only occur when partners monitor
+and diagnose each other's knowledge? If the ratio between shallow alignment and
 deep grounding depends upon the task, and if deep grounding is a condition for
@@
 %experiment)~\citep{Leslie2000}: a child is asked to watch a scene where two
-%people, $A$ and $B$, manipulate objects. Then $A$ leaves and $B$ hides away one
-%object. When $A$ comes back, we ask the child ``where do you think $A$ will
+%people, \A and \B, manipulate objects. Then \A leaves and \B hides away one
+%object. When \A comes back, we ask the child ``where do you think \A will
 %look for the object?''. Before acquiring a theory of mind, children are not
@@
 %the object was hidden) from the model of A, which contains \emph{false
-%beliefs} on the world ($A$ still thinks the object is at its original
-%position since he did not see $B$ hiding it).
+%beliefs} on the world (\A still thinks the object is at its original
+%position since he did not see \B hiding it).
 
@@
 In summary, the controversy around the cognitive depth of shared understanding
-pertains to psycholinguistic, which investigates 'natural conversations, but  in
-collaborative learning, the situation is different; tasks are precisely designed
-for requiring a deeper negotiation of meaning.  The craft of CSCL is to design
+pertains to psycholinguistic, which investigates natural conversations. The
+situation is different in collaborative learning; tasks are actually designed
+for requiring a deeper negotiation of meaning.  The expertise of CSCL is to design
 collaborative situations that create interdependence, avoid group think, and
-allows students to detect any 'illusion of shared understanding' (Cherubini et
-al, 2005). Our question is hence not  anymore"do peers build a shared
-understanding ?", but becomes "when peers have to build a shared understanding,
-how do they do it ?".  This question addresses the mechanisms of grounding, in
+allows students to detect any ``illusion of shared
+understanding''~\citet{cherubini2005grounding}. Our question is hence not
+anymore ``do peers build a shared
+understanding ?'', but rather ``whenever peers have to build a shared understanding,
+how do they achieve it ?''.  This question addresses the mechanisms of grounding, in
 which the basic sequence is: make a proposition, detect misunderstanding and, if
@@
 his message is understood by his or her partner. Detecting one peer's
-understanding is investigate hereafter under the general umbrella of partner
+misunderstanding is investigated hereafter under the general umbrella of partner
 modelling.
-
-%TODO
-Cherubini, M., Van Der Pol, J., \& Dillenbourg, P. (2005). Grounding is not
-shared understanding: Distinguishing grounding at an utterance and knowledge
-level. In CONTEXT'05, the Fifth International and Interdisciplinary Conference
-on Modeling and Using Context (pp. 11-23).
 
@@
 peers would be as incorrect as any claim that they do not maintain any
-representation at all. If mutual modelling had to be permanently detailed and
+representation at all. If mental modelling had to be permanently detailed and
 accurate, subjects would obviously face a huge cognitive load. Conversely, peers
@@
 Collaborative learning dialogue include many instances of utterances such as ``I
-thought he would do that'' (first level of mutual modelling) or even ``He
-thought I would do that but I intended something else.'' (second level of mutual
+thought he would do that'' (first order modelling) or even ``He
+thought I would do that but I intended something else.'' (second order
 modelling).
 
-%For instance, $A$ cannot disagree with $B$ without knowing that $B$ has a different
-%opinion. The mutual model can be \emph{implicit} ($A$ is not aware of what he knows about
-%B), \emph{by-default} (I believe that $B$ believes what I believe unless contrary
-%evidence), \emph{opportunistic} ($A$ does not model $B$ unless the conversation requires
-%it), \emph{global} ($A$ infers $B$'s beliefs based on categories such as age, culture or
+%For instance, \A cannot disagree with \B without knowing that \B has a different
+%opinion. The mutual model can be \emph{implicit} (\A is not aware of what he knows about
+%B), \emph{by-default} (I believe that \B believes what I believe unless contrary
+%evidence), \emph{opportunistic} (\A does not model \B unless the conversation requires
+%it), \emph{global} (\A infers \B's beliefs based on categories such as age, culture or
 %profession) and, of course, it can be incorrect\ldots but it can not remain empty.
@@
 The content of the partner model ranges from \emph{dispositional} to
-\emph{situational} aspects. The \emph{dispositional} aspects refer to $A$'s
-representation of $B$'s long term knowledge, skills or traits. It is thus closely
+\emph{situational} aspects. The \emph{dispositional} aspects refer to \A's
+representation of \B's long term knowledge, skills or traits. It is thus closely
 related to the notion of transactive memory~\citep{wegner1987transactive,
-moreland1999transactive}.  \emph{Situational} aspects refer to $A$'s representation of
+moreland1999transactive}.  \emph{Situational} aspects refer to \A's representation of
 B's knowledge, behavior or intentions specifically activated in the situation in
-which $A$ and $B$ are collaborating, some of them being valid for 2 seconds, other
-ones for 2 hours. Examples of fragments that constitute $A$'s model of B
+which \A and \B are collaborating, some of them being valid for 2 seconds, other
+ones for 2 hours. Examples of fragments that constitute \A's model of B
 regarding to aspects X, \ie $Model(A,B,X)$, abbreviated \model{A}{B}{X}, could be:
@@
 
-    \item $Model(A,B, knowledge)$: what does $A$ know about $B$'s knowledge with
-        respect to the task at hand or, inversely, about $B$'s knowledge gaps?
-        When can $A$ consider $B$'s statements as reliable?
-
-    \item $Model(A,B, skills)$: what does $A$ know about $B$'s skills with respect to
-        the task at hand? May $A$ expect $B$ to perform well in a specific subtask?
+    \item $Model(A,B, knowledge)$: what does \A know about \B's knowledge with
+        respect to the task at hand or, inversely, about \B's knowledge gaps?
+        When can \A consider \B's statements as reliable?
+
+    \item $Model(A,B, skills)$: what does \A know about \B's skills with respect to
+        the task at hand? May \A expect \B to perform well in a specific subtask?
         The effectiveness of division of labor depends on the quality of this
@@
 
-    \item $Model(A,B, goals)$: what does $A$ know about $B$'s intentions with respect
-        to the project, including $B$' motivation and commitment? Can $A$ trust B
-        when $B$ promises to deliver?
-
-    \item $Model(A,B, task)$: what does $A$ know about $B$'s representation of the
-        situation and the task: does $A$ knows whether $B$ has the same
+    \item $Model(A,B, goals)$: what does \A know about \B's intentions with respect
+        to the project, including \B' motivation and commitment? Can \A trust
+        \B
+        when \B promises to deliver?
+
+    \item $Model(A,B, task)$: what does \A know about \B's representation of the
+        situation and the task: does \A knows whether \B has the same
         understanding of the problem at stake?
 
-    \item $Model(A,B, plans)$: what does $A$ know about $B$'s strategy. Does A
-        understand why $B$ did what he did? Is $A$ able to anticipate what $B$ will do
+    \item $Model(A,B, plans)$: what does \A know about \B's strategy. Does A
+        understand why \B did what he did? Is \A able to anticipate what \B will do
         next?
 
-    \item $Model(A,B, ``urgent'')$: what does $A$ know about $B$'s understanding of $A$'s
+    \item $Model(A,B, ``urgent")$: what does \A know about \B's understanding of \A's
         last utterance: does ``urgent'' means now, soon or ``not too late''?
@@
 initial modelling impacts communication. In their experiment,
-\citet{slugoski1993attribution} pretended to their subjects that their (fake)
-partner had or had not received the same information. They observed that the
-subjects adapted their dialogue by focusing the explanation on the items that
-he/she was supposed to ignore.  \citet{brennan1991conversation} showed that the
-subjects used different initial strategies in forming queries depending on who
-they were told their partner was.
+\citet{slugoski1993attribution} pretended to their subjects that their
+(confederate) partner had or had not received the same information. They
+observed that the subjects adapted their dialogue by focusing the explanation on
+the items that he/she was supposed to ignore.  \citet{brennan1991conversation}
+showed that the subjects used different initial strategies in forming queries
+depending on who they were told their partner was.
 
 Initial common grounds are also initiated by co-presence: they include events to
-which $A$ and $B$ attended together~\citep{clark2002definite} in the physical
+which \A and \B attended together~\citep{clark2002definite} in the physical
 space or in their cultural space (\eg ``09-11''). While co-presence means that
@@
 Mutual modelling is bi-directional. During dyadic problem solving,   partner A
-builds some model of B and B build some model of A.  Moreover, these two
-processes are not independent: A's model of what B knows, include what B knows
-about A.  This leads to nested levels of  modelling. If $A$ states ``B thinks I
-am good in maths'', $A$ builds a \emph{second level} model:
+builds some model of \B and \B build some model of \A.  Moreover, these two
+processes are not independent: \A's model of what \B knows, include what \B knows
+about \A.  This leads to nested levels of  modelling. If \A states ``\B thinks I
+am good in maths'', \A builds a \emph{second level} model:
 \model{A}{B}{\M{B}{A}{\text{maths-skill}}}. This leads to possibly infinite
-regress of nested models: $A$ saying ``$B$ knows that I don't expect him to
+regress of nested models: \A saying ``\B knows that I don't expect him to
 solve this statistics problem'' corresponds to \\
 \model{A}{B}{\M{B}{A}{\M{A}{B}{\text{statistic-skills}}}}.  As we will see in
-one study we report, mutuality also applies to triads in which A will elaborate
-a model of B and of C, and reciprocally. This is probably also true for larger
+one study we report on, mutuality also applies to triads in which \A will elaborate
+a model of \B and of C, and reciprocally. This is probably also true for larger
 groups although one may hypothesize that there exists (yet undefined) group size
@@
 their individual cognitive skills would correlate. Therefore, if it occurs that
-the quality of the Model (A,B) is correlated with Model (B,A), one may infers
-that this quality depends upon what A and B have in common, their verbal
-interactions.
+the quality of the $Model(A,B)$ is correlated with $Model(B,A)$, one may infers
+that this quality depends upon what \A and \B have built together while
+interacting.
 
@@
 
-Natural language becomes cumbersome when describing things such as "the model
-that A builds about the model that B builds about A". Therefore, to define
+Natural language becomes cumbersome when describing things such as ``the model
+that \A builds about the model that \B builds about A''. Therefore, to define
 hypotheses and report on experiments on mutual modelling, we use  the notation
-\model{A}{B}{X} to denote ``$A$ knows that $B$ knows X''. This  notation is not
+\model{A}{B}{X} to denote ``\A knows that \B knows X''. This  notation is not
 proposed as a formal theory of mutual modelling but as useful simplification for
-communicating about mutual modelling. This notation does not mean $A$ has an
-explicit, monolithic representation of $B$: it must be understood as an
+communicating about mutual modelling. This notation does not mean \A has an
+explicit, monolithic representation of \B: it must be understood as an
 abstraction referring to complex socio-cognitive processes.  As explained in the
-previous section, $A$ can be fragmented, multi-dimensional, etc. This notation
+previous section, \A can be fragmented, multi-dimensional, etc. This notation
 is neither presented as a computational model of mutual modelling, nor as some
@@
     \item Before any specific grounding action, there is generally a non-null
-        probability that $X$ is mutually understood by $A$ and $B$ (\eg $X$ is
-        part of $A$'s and $B$'s cultures, it is manifest to co-present subjects
+        probability that $X$ is mutually understood by \A and \B (\eg $X$ is
+        part of \A's and \B's cultures, it is manifest to co-present subjects
         or simply there is not much space for misunderstanding or disagreement
@@
 
-        \item \model{A}{B}{actions} is about how well $A$ guesses what action $B$ has
+        \item \model{A}{B}{actions} is about how well \A guesses what action \B has
             performed (study 2) or will perform next (study 1),
 
-        \item \model{A}{B}{emotion}: how accurately $A$ perceives $B$'s emotional state
+        \item \model{A}{B}{emotion}: how accurately \A perceives \B's emotional state
             (study 3),
 
-        \item \model{A}{B}{knowledge}: how accurately $A$ estimates $B$'s knowledge
+        \item \model{A}{B}{knowledge}: how accurately \A estimates \B's knowledge
             with respect to the material they learn together (study 4 and 5).
@@
         use awareness tools which, in principle, reduce the cost of mutual modelling, but do
-        not eliminate all costs: if the tool provides $A$ with information about
-        what $B$ does/knows, this additional information may actually increase
+        not eliminate all costs: if the tool provides \A with information about
+        what \B does/knows, this additional information may actually increase
         cognitive load. Awareness tools constitute a kind of mutual modelling prosthesis,
@@
 
-{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask $A$ what he/she
-believes about what $B$ knows, feels, intends to do, etc.  This raises obvious
+{\bf Capturing \model{A}{B}{X}}: The simplest method is to ask \A what he/she
+believes about what \B knows, feels, intends to do, etc.  This raises obvious
 methodological concerns since such a question triggers a modelling process
@@
 access the reference model \refmodel{B}{X} to estimate its \emph{accuracy}.
-Since we can only indirectly access it via what $B$ reports (\ie
+Since we can only indirectly access it via what \B reports (\ie
 \model{B}{B}{X}), accuracy can be estimated in 2 ways:
@@
     \item Subjective accuracy: In study 3, we compute \Model{A}{B}{X} by
-        measuring if $A$ describes $B$'s emotions in the same way $B$ reports
+        measuring if \A describes \B's emotions in the same way \B reports
         its emotions ($\M{A}{B}{X} = \M{B}{B}{X}$).
@@
     \item Objective accuracy: In studies 4 and 5, we compute \Model{A}{B}{X} by
-        comparing \Model{A}{B}{K} to $B$'s actual knowledge as it has measured
+        comparing \Model{A}{B}{K} to \B's actual knowledge as it has measured
         by a test.
@@
 to produce these variations is what CSCW researchers have called ``awareness
-tools'', \ie functionalities that inform $A$ is about $B$'s actions that A could
-not directly perceive because $B$ was working on a different subset of the
+tools'', \ie functionalities that inform \A is about \B's actions that \A could
+not directly perceive because \B was working on a different subset of the
 virtual space. Different awareness tools are used in the following studies as
@@
 mutual modelling skills. However, a high correlation could also have another
-explanation, stated in  $\mathcal{H}_{2}$: it could be that A is good at
-modelling B and good at helping B to model herself or himself.
+explanation, stated in  $\mathcal{H}_{2}$: it could be that \A is good at
+modelling \B and good at helping \B to model herself or himself.
 
 \begin{itemize}
-    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon $A$'s ability or effort
-        to model $B$,
-
-    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  $B$'s ability or
-        effort to help $A$ to model him/herself,
+    \item $\mathcal{H}_{1}$: $\mathcal{M}^{\circ}(A,B)$ depends upon \A's ability or effort
+        to model \B,
+
+    \item $\mathcal{H}_{2}$: $\mathcal{M}^{\circ}(A,B)$ depends upon  \B's ability or
+        effort to help \A to model him/herself,
 
     \item $\mathcal{H}_{3}$: $\mathcal{M}^{\circ}(A,B)$ depends upon the quality of
-        interactions among $A$ and $B$.
+        interactions among \A and \B.
 
@@
 
-$\mathcal{H}_{2}$ relates to second level modelling since $B$ needs to monitor
-$A$ to see if $A$ understood him/her ($\mathcal{M}(B,A,\mathcal{M}(A,B))$). We
+$\mathcal{H}_{2}$ relates to second level modelling since \B needs to monitor
+\A to see if \A understood him/her ($\mathcal{M}(B,A,\mathcal{M}(A,B))$). We
 will see that $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$ are actually difficult to
@@
 \Model{B}{C}{X}.This leads to two \emph{triangle questions}
-(Fig.~\ref{mm_triangles}): Do $A$ and $B$ have the same accuracy when modelling
-$C$ ($\Delta_2 = \Delta(\mathcal{M}^{\circ}(A,C,X),
+(Fig.~\ref{mm_triangles}): Do \A and \B have the same accuracy when modelling
+\C ($\Delta_2 = \Delta(\mathcal{M}^{\circ}(A,C,X),
 \mathcal{M}^{\circ}(B,C,X))$)? A significant correlation would support ,
-$\mathcal{H}_{2}$  (C has helped both A and B to model C) or support
+$\mathcal{H}_{2}$  (C has helped both \A and \B to model C) or support
 $\mathcal{H}_{3}$ (the quality of triad interactions enables all partners to
@@
 
-Conversely, does $C$ model more accurately $A$ than $B$ ($\Delta_3=
+Conversely, does \C model more accurately \A than \B ($\Delta_3=
 \Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
-answer would support $\mathcal{H}_{1}$, C being simply good at modelling any
+answer would support $\mathcal{H}_{1}$, \C being simply good at modelling any
 partner.  It could also support  $\mathcal{H}_{3}$,  since the quality of
-interactions influences the accuracy of two models that C is building. A
+interactions influences the accuracy of two models that \C is building. A
 negative answer would support $\mathcal{H}_{2}$ since $\Delta_3$ could mainly be
-explained by the fact that  $A$ helped more $C$ to model him than $B$ did.
+explained by the fact that  \A helped more \C to model him than \B did.
 
@@
 neglecting the cases where AB interact better for instance than BC, since there
-was no 'private' communication channel in the following studies. We do
+was no ``private'' communication channel in the following studies. We do
 nonetheless acknowledge that this point could be debated.
@@
 The independent variable was the availability of an awareness tool that shows to
-player $A$  the location and gaze direction of player $B$. In the awareness
+player \A  the location and gaze direction of player \B. In the awareness
 condition, players could switch to the \emph{scout mode} where they could view
@@
 ($\Delta(\M{A}{A}{X}, \M{A}{B}{X})$) rather than an objective measure (\ie the
-model \gmodel{A}{B} is compared to $B$'s next action) because some of the
+model \gmodel{A}{B} is compared to \B's next action) because some of the
 activities proposed by the questionnaire were not observable by the environment
@@
 path reflected their problem solving strategy. We used an objective measure of
-\gModel{A}{B}: the distance between where $A$ believes $B$ has been walking and
-where $B$ actually went.  The main hypothesis concerned the effect of awareness
+\gModel{A}{B}: the distance between where \A believes \B has been walking and
+where \B actually went.  The main hypothesis concerned the effect of awareness
 tools on group performance and on \gModel{A}{B}.
@@
 their personal distance to Bob. Their annotations on the map were shared with
-the two other players ($A$ could see what $B$ and $C$ wrote). These annotations
+the two other players (\A could see what \B and \C wrote). These annotations
 faded out after a few minutes to avoid covering the full display. The awareness
@@
             \includegraphics[width=\linewidth]{image6.png}
-            \caption{A's drawing of $B$'s path.}
+            \caption{A's drawing of \B's path.}
         \end{subfigure}
@@
             \includegraphics[width=\linewidth]{image7.png}
-            \caption{Actual path followed by $B$.}
+            \caption{Actual path followed by \B.}
         \end{subfigure}
@@
 after the game. This enabled us to calculate the number of errors players made
-while drawing the path of their partners. We compared the path that player $A$
-attributed to $B$ with $B$'s real path recorded by the system and the same for
-$A$ \& $C$ and $B$ \& $C$ as depicted on Fig.~\ref{study2:paths}.
-
-\gModel{A}{B} is the sum of errors made by $A$ about $B$'s paths. An error was
+while drawing the path of their partners. We compared the path that player \A
+attributed to \B with \B's real path recorded by the system and the same for
+\A \& \C and \B \& \C as depicted on Fig.~\ref{study2:paths}.
+
+\gModel{A}{B} is the sum of errors made by \A about \B's paths. An error was
 either drawing a place where the partner had not been or not drawing a place
@@
 \paragraph{Symmetry question} The correlation between \gModel{A}{B}  and
-\gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more $A$
-makes errors about $B$, the more $B$ does as well (and vice-versa).
+\gModel{B}{A}  is positive ($r = 0.41$) and significant ($p < 0.01$): the more \A
+makes errors about \B, the more \B does as well (and vice-versa).
 
@@
 variable \gModel{A}{B} was measured subjectively from three questionnaires: in
-the first one, $A$ described his/her own emotions \gmodel{A}{A}, while in the
-two other questionnaire, $A$ described $B$'s and $C$'s emotions. The questionnaire
+the first one, \A described his/her own emotions \gmodel{A}{A}, while in the
+two other questionnaire, \A described \B's and \C's emotions. The questionnaire
 included 18 items (7-point Likert Scale) describing emotions labeled as
@@
 produced concept map) but learning gains. The correlation between \gModel{A}{B}
-and $A$'s learning gains is not significant ($\beta = 0.08, ns, N = 60$). It is
+and \A's learning gains is not significant ($\beta = 0.08, ns, N = 60$). It is
 also not significant within each condition.
@@
 \gModel{A}{B} was not significant ($F(1, 58) = 0.78, ns$) when considering the
-absolute difference between \gModel{A}{B} and $B$'s post-test score. However, $A$
-tended underestimate $B$'s score in the SI condition ($M = -2.06$) and to
+absolute difference between \gModel{A}{B} and \B's post-test score. However, \A
+tended underestimate \B's score in the SI condition ($M = -2.06$) and to
 overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<0.01$).
@@
 overlapping, another explanation is that, in absence of these initial common
-grounds, mutual modelling requires A to make evident to B what A thinks B does
+grounds, mutual modelling requires \A to make evident to \B what \A thinks \B does
 not know about A, which is the second level of modelling described in
@@
 outcomes by comparing teams with or without a knowledge awareness tool (KAT),
-\ie a tool that informs $A$ about $B$'s knowledge as measured through a pre-test.
+\ie a tool that informs \A about \B's knowledge as measured through a pre-test.
 
@@
 and (3) elaborated utterances with rich contents,  These 3 categories provide
-different account of \gModel{A}{B}: (1) as an effect of $A$'s effort to model
-$B$  ($\mathcal{H}_{1}$), (2) as $B$'s effort to give cues to $A$ about his own
+different account of \gModel{A}{B}: (1) as an effect of \A's effort to model
+\B  ($\mathcal{H}_{1}$), (2) as \B's effort to give cues to \A about his own
 knowledge  ($\mathcal{H}_{2}$) and (3) as an effect of the quality of
@@
 hypothesis, the skill of helping one's peer to model oneself, is questioned by
-study 3 where C could be modelled  accurately by A and not accurately by B. The
+study 3 where \C could be modelled  accurately by \A and not accurately by \B. The
 same hypothesis is however supported by the results of study 2, and indirectly
@@
 social mechanisms in human cognition. In the field of social cognition, it is a
-common place to state that 'the whole is greater than the sum of the parts'.
+common place to state that ``the whole is greater than the sum of the parts''.
 This refers to the emergence of team properties than cannot be reduced to the

### 2016-02-04 rebuttal.tex
Added draft rebuttal

\documentclass{article}
\usepackage[utf8]{inputenc}

\begin{document}

Dear Editor, Dear Reviewers,

\vspace{2em}

We would like to thank both the reviewers and the editor for going through our
manuscript. As a follow-up of your comments, the manuscript underwent
several significant changes, that we summarize hereafter.

\begin{itemize}

    \item The title and the abstract have been completely changed to clarify the
        claims of the paper, which, as the editor pointed out, were not clear
        for some reviewers. In a nutshell, we don’t know what’s going on in he
        brain when building a shared solution, we leave this scientific debate
        to psycholinguists, but we observe how they vary (or not) across
        situations (across the 5 experiments)  and this enables some inferences.

    \item   Section 1, Introduction. We did minor edits except for the last
        paragraph has been completely rewritten it  that was somehow misleading
        regarding to the goals of the paper.  I think the point is now made
        clearly that the issue is not whether it is alway necessary to build a
        shared understanding, but that CSCL folds design tasks specifically for
        requiring shared understanding.

    \item We clarified the notion of grounding criterion.

    \item Section 2 has been split into partner modeling and mutual modeling (now
        section 3), to emphasized the fact that the question is the mutuality. The
        main research question is then clearly stated to address some of the
        reviewers cirtciisms

    \item Section 4. We clarified the status of the proposed formalism, which some
        reviewers find convenient and others criticize. Clearly, we are not
        proposing here a computational theory of mutual modeling. This notation is
        only useful because the text become quickly unreadable as one talks about
        the model that A build about the model that B build about A, while
        M(A,B,M(B,A,X)) is clear. So, it’s just a language for discussing
        experiments and hypotheses in a simple way. It’s function is internal to
        this paper

    \item For the 5 studies, we have re-written all the ‘discussion’ sections in a
        way that they cumulate naturally in alignment with the research questions

    \item As requested by the reviewers, we have completely rewritten  the
        conclusions to make them more explicit.  We cannot of course not pretend
        that our results are clearcut. We have to be careful not to publish only
        papers with significant results (statisticians have proved that some
        meta-analyses are biased by the fact that papers with significant results
        get more easily published). However, despite the fact that none of the 3
        hypothesis is declared as winner, the key results is that we prove that
        individual skills only explain a part of the variance, that the quality of
        social interactions matter significantly. This is why Gerry asked me (a
        few years ago I must say) to present these results that I mentionned in a
        ICLS panel.

\end{itemize}

\end{document}

### 2016-02-04 rebuttal.tex
WIP on the rebuttal

@@
 We would like to thank both the reviewers and the editor for going through our
-manuscript. As a follow-up of your comments, the manuscript underwent
-several significant changes, that we summarize hereafter.
+manuscript and providing in-depth review of our submission.
+
+The main points that were raised relate to:
+\begin{itemize}
+    \item the need for a better presentation of the concepts that we build upon,
+    \item the need for a ``deeper synthetic discussion'', also ``accounting for the
+        conflicting results'', that would present the ``theoretical
+        picture'' resulting from our work,
+    \item discuss a possible research agenda
+\end{itemize}
+
+In line with these comments and suggestions, the manuscript underwent several
+significant changes that we summarize hereafter.
 
@@
 
-    \item The title and the abstract have been completely changed to clarify the
+    \item The title and the abstract have been changed to clarify the
         claims of the paper, which, as the editor pointed out, were not clear
-        for some reviewers. In a nutshell, we don’t know what’s going on in he
-        brain when building a shared solution, we leave this scientific debate
-        to psycholinguists, but we observe how they vary (or not) across
-        situations (across the 5 experiments)  and this enables some inferences.
+        for some reviewers. In a nutshell, we do not know what is going on in the
+        brain when building a shared solution (we leave this scientific debate
+        to psycholinguists), but we observe how they vary (or not) across
+        situations (across the 5 experiments) and this enables some reflexions
+        that we put in the context of collaborative learning.
 
-    \item   Section 1, Introduction. We did minor edits except for the last
+    \item \emph{Section 1 -- Introduction} We did minor edits except for the last
         paragraph has been completely rewritten it  that was somehow misleading
-        regarding to the goals of the paper.  I think the point is now made
+        regarding to the goals of the paper.  We think the point is now made
         clearly that the issue is not whether it is alway necessary to build a
@@
 
-    \item We clarified the notion of grounding criterion.
+    \item We clarified the notion of grounding criterion, as requested by
+        reviewer 2 (page [TODO!!])
 
-    \item Section 2 has been split into partner modeling and mutual modeling (now
-        section 3), to emphasized the fact that the question is the mutuality. The
-        main research question is then clearly stated to address some of the
-        reviewers cirtciisms
+    \item \emph{Section 2} has been split into partner modeling and mutual modeling (now
+        section 3) to emphasize the fact that the question is the mutuality. The
+        main research question is then clearly stated to address some [WHICH?] of the
+        reviewers' criticisms.
 
-    \item Section 4. We clarified the status of the proposed formalism, which some
-        reviewers find convenient and others criticize. Clearly, we are not
+    \item \emph{Section 'Notations'} We clarified the status of the proposed formalism, which some
+        reviewers found useful and others criticized. We are not
         proposing here a computational theory of mutual modeling. This notation is
@@
         the model that A build about the model that B build about A, while
-        M(A,B,M(B,A,X)) is clear. So, it’s just a language for discussing
-        experiments and hypotheses in a simple way. It’s function is internal to
-        this paper
+        M(A,B,M(B,A,X)) is less ambiguous. We mainly see it as a language for discussing
+        experiments and hypotheses in a simple way. Its function is internal to
+        this paper.
 
-    \item For the 5 studies, we have re-written all the ‘discussion’ sections in a
-        way that they cumulate naturally in alignment with the research questions
+    \item \emph{Section 'Studies'} For all the 5 studies, we have re-written all
+        the \emph{discussion} sub-sections in a
+        way that they naturally align with the research questions, leading to a
+        clearer overall message.
 
-    \item As requested by the reviewers, we have completely rewritten  the
+    \item As requested by the reviewers, we have entirely rewritten  the
         conclusions to make them more explicit.  We cannot of course not pretend
@@
         individual skills only explain a part of the variance, that the quality of
-        social interactions matter significantly. This is why Gerry asked me (a
-        few years ago I must say) to present these results that I mentionned in a
-        ICLS panel.
+        social interactions matter significantly.
 

### 2016-02-05 rebuttal.tex
WIP on the rebuttal

@@
 
-    \item We clarified the notion of grounding criterion, as requested by
-        reviewer 2 (page [TODO!!])
+    \item As requested by the reviewer 2, we have clarified the notion of
+        grounding criterion (p.6), and clarified as well that we have not attempted to
+        quantify the cost of grounding in the presented studies (p.8).
 

### 2016-02-05 mutual-modeling.tex
Polish section 4, 5, 6, 7.1

@@
 
+\usepackage{booktabs}
+\usepackage{array}
+
 \usepackage{amsmath}
@@
 \newcommand{\ie}{i.e.\xspace}
-\newcommand{\cf}{{\textit{cf\xspace}}}
-\newcommand{\eg}{{\textit{e.g.\xspace}}}
-\newcommand{\al}{{\textit{et al.\xspace}}}
+\newcommand{\cf}{{\textit{cf}\xspace}}
+\newcommand{\eg}{e.g.\xspace}
+\newcommand{\al}{{\textit{et al.}\xspace}}
 
@@
 is neither presented as a computational model of mutual modelling, nor as some
-universal formalism; it's usefulness is internal to this paper.  Besides, we
+universal formalism; its usefulness is internal to this paper.  Besides, we
 refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
@@
         understanding.  The \emph{grounding criterion} -- denoted
-        \groundingcriterion -- refers to how accurate the models need to be to
-        succeed the task $T$. It can be calculated as the probability to succeed
-        $T$ despite the fact $X$ is not grounded.
-        $\mathcal{M}^{\circ}_{min}(A,B,X)$ can be estimated from the correlation
-        between \Model{A}{B}{X} and the task performance.
+        \groundingcriterion -- represents the minimum level of modelling
+        accuracy required for a task $T$ to succeed. Qualitatively, if the
+        performance on a given task $T$ is correlated to \Model{A}{B}{X}, then
+        \model{A}{B}{X} is significant to $T$ success, and the
+        grounding criterion of $X$ for $T$
+        ($\mathcal{M}^{\circ}_{min}(A,B,X,T)$) is non-zero.
+
+        Under the assumption that the higher the correlation, the more
+        critical \model{A}{B}{X} is to $T$, we hereafter use the correlation
+        coefficient as an estimate of \groundingcriterion.
+        \vspace{1em}
 
@@
         grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
+        \vspace{1em}
 
@@
 
-Based on these 3 parameters, the probability of making an action $\alpha_{t_1}$
-about content $X$ at time $t_1$ during task T in order to increase
-\Model{A}{B}{X} at time $t_2$ is the ratio between how much it is needed
-((2)-(1)) and how much it costs (3)~\citep{traum1996miscommunication}:
-
-\begin{multline} \label{eq:probrepair}
-    p(\alpha_{t_1}(X,T) / \mathcal{M}^{\circ}_{t_2}(A,B,X) \nearrow) \simeq
-    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) - \mathcal{M}^{\circ}_{t_1}(A,B,X)}{cost (\alpha)}
-\end{multline}
+Based on these 3 parameters, the probability of making an action $\alpha_{t}$
+about content $X$ at time $t$ during task T in order to increase \Model{A}{B}{X}
+is the ratio between how much it is needed and how much it
+costs~\citep{traum1996miscommunication}:
+
+\begin{equation}\label{eq:probrepair}
+    p(\alpha_{t}(X,T) /_{\mathcal{M}^{\circ}_{t+1}(A,B,X)\nearrow}) \simeq
+    \frac{\mathcal{M}^{\circ}_{min}(A,B,X,T) -
+    \mathcal{M}^{\circ}_{t}(A,B,X)}{cost (\alpha_{t})}
+\end{equation}
 
@@
 since several parameters are hard to quantify (\eg the cost of a communication
-act depends upon the user as well). We use it for clarifying the parameters of
-the experiments we present hereafter.
+act depends upon the user as well).
+
+We can further clarify the parameters in the context of the experiments we
+present hereafter:
 
@@
 
-        \item \model{A}{B}{actions} is about how well \A guesses what action \B has
+        \item \Model{A}{B}{actions} is about how well \A guesses what action \B has
             performed (study 2) or will perform next (study 1),
 
-        \item \model{A}{B}{emotion}: how accurately \A perceives \B's emotional state
+        \item \Model{A}{B}{emotion}: how accurately \A perceives \B's emotional state
             (study 3),
 
-        \item \model{A}{B}{knowledge}: how accurately \A estimates \B's knowledge
+        \item \Model{A}{B}{knowledge}: how accurately \A estimates \B's knowledge
             with respect to the material they learn together (study 4 and 5).
@@
     \end{enumerate}
+        \vspace{1em}
 
@@
         concept mapping (study 4 and 5). By varying the tasks, we do actually
-        vary the grounding criterion but they all require a ``reasonably high''
-        grounding criterion.  All tasks are about building a solution or a
-        representation together. As stated earlier, we did not address everyday conversations.
+        vary the grounding criterion. The tasks were all designed to require a reasonably high
+        grounding criterion, as they are meant for the participants to have to
+        actually build a solution or a representation together.
+        \vspace{1em}
 
@@
         using a {\sc jigsaw} script.
+        \vspace{1em}
 
@@
 
+        While we introduce here formally the cost of grounding $cost(\alpha)$ as
+        one relevant variable for the discussion of mutual modelling situations,
+        we will not attempt to characterize it beyond these qualitative
+        observations in the studies we present hereafter.
+
 \end{itemize}
@@
 Since a mental model is not directly observable,  the study of mutual modelling
-is methodologically difficult. How can we for instance measure \Model{A}{B}{X}.
+is methodologically difficult. How can we for instance measure \Model{A}{B}{X}?
 We proceed in two steps, first to capture \model{A}{B}{X} and then to estimate
@@
 modelling after task completion. Then, the obvious drawback are memory losses
-and by post-hoc reconstruction.  The first method was used in study 2 and the
+and post-hoc reconstruction. The first method was used in study 2 and the
 second one in the other studies. Another option would be to rely on external
@@
 {\bf Estimating \Model{A}{B}{X}}: Once \model{A}{B}{X} is captured, we need to
-access the reference model \refmodel{B}{X} to estimate its \emph{accuracy}.
+access the reference model \refmodel{B}{X} to estimate its accuracy.
 Since we can only indirectly access it via what \B reports (\ie
@@
 
-    \item Subjective accuracy: In study 3, we compute \Model{A}{B}{X} by
+    \item Subjective accuracy: In study 3, for instance, we compute \Model{A}{B}{X} by
         measuring if \A describes \B's emotions in the same way \B reports
@@
 
-Our method for investigating mutual modelling is observe the variations of
-accuracy that result from variations of external parameters that appear as
-variable in the formula presented in the previous section: the accuracy should
-for go down if we increase the cost of grounding acts and goes up if we increase
-the grounding criterion (the necessity to build a shared understanding). One way
-to produce these variations is what CSCW researchers have called ``awareness
-tools'', \ie functionalities that inform \A is about \B's actions that \A could
-not directly perceive because \B was working on a different subset of the
+Our method for investigating mutual modelling relies on the observation of the variations of
+accuracy that result from variations of external parameters (the variables of
+the formula~\ref{eq:probrepair} above): for instance, the accuracy should
+go down if we increase the cost of grounding acts, and conversly go up if we increase
+the grounding criterion (\ie the necessity to build a shared understanding). One way
+to produce these variations relies what CSCW researchers call ``awareness
+tools'', \ie functionalities that inform \A about \B's actions that \A can
+not directly perceive due to \B working in a different subset of the
 virtual space. Different awareness tools are used in the following studies as
-methodological tricks  to manipulate experimentally the  mutual modelling
+methodological levers to experimentally manipulate the mutual modelling
 activity.
 
-\section*{Hypotheses and questions}
+\section{Hypotheses and questions}
 
@@
 As stated earlier, the fundamental challenge is to determine if mutual modelling
-is an individual skill (hypothesis 1) or  the emergent property of social
-interactions (hypothesis 3).  This question is empirically translated into the
-symmetry of mutual modelling: what is the relationship between \Model{A}{B}{X}
-and \Model{B}{A}{X} (Fig.~\ref{mm_symmetry})? A low symmetry would mean that
-mutual modelling is mainly an individual attitude/aptitude ($\mathcal{H}_{1}$).
-A high correlation might support $\mathcal{H}_{3}$ since there is a low
-probability that randomly formed pairs integrate peers with the same level of
-mutual modelling skills. However, a high correlation could also have another
-explanation, stated in  $\mathcal{H}_{2}$: it could be that \A is good at
-modelling \B and good at helping \B to model herself or himself.
+is an individual skill (hypothesis $\mathcal{H}_{1}$ below) or  the emergent
+property of social interactions (hypothesis $\mathcal{H}_{3}$).  This question
+is empirically translated into the symmetry of mutual modelling
+(Fig.~\ref{mm_symmetry}): what is the relationship between \Model{A}{B}{X} and
+\Model{B}{A}{X}? A low symmetry would mean that mutual modelling is mainly an
+individual attitude/aptitude ($\mathcal{H}_{1}$).  A high correlation might
+support $\mathcal{H}_{3}$ since there is a low probability that randomly formed
+pairs integrate peers with the same level of mutual modelling skills. However, a
+high correlation could also have another explanation, stated in
+$\mathcal{H}_{2}$: it could be that \A is good at modelling \B and good at
+helping \B to model herself or himself.
 
@@
 \Model{B}{A}{X}, \Model{A}{C}{X}, \Model{C}{A}{X}, \Model{C}{B}{X} and
-\Model{B}{C}{X}.This leads to two \emph{triangle questions}
+\Model{B}{C}{X}. This leads to two \emph{triangle questions}
 (Fig.~\ref{mm_triangles}): Do \A and \B have the same accuracy when modelling
@@
 
-Conversely, does \C model more accurately \A than \B ($\Delta_3=
+Conversely, does \C model with similar accuracy \A and \B? (low $\Delta_3=
 \Delta(\mathcal{M}^{\circ}(C,A,X), \mathcal{M}^{\circ}(C,B,X))$)? A positive
@@
 whether the accuracy of mutual modelling depends more upon the modeller's effort
-$\mathcal{H}_{1}$ or the modellee's behaviour  $\mathcal{H}_{2}$.
-
-Please note that we consider the quality of interactions at triad level (ABC),
-neglecting the cases where AB interact better for instance than BC, since there
+($\mathcal{H}_{1}$) or the modellee's behaviour ($\mathcal{H}_{2}$).
+
+Note that we consider the quality of interactions at triad level (\A, \B, \C),
+neglecting the cases where \A and \B interact better for instance than \B and \C, since there
 was no ``private'' communication channel in the following studies. We do
@@
 these are two specific instance of a more general cognitive process. This
-questions is however not central to this paper, and the value of $\Delta_4$ us
+questions is however not central to this paper, and the value of $\Delta_4$ is
 only available in one of the reported studies.
@@
 
-We report five studies conducted  by different researchers in different contexts
+We report on five studies conducted  by different researchers in different contexts
 between 2000 and 2015. They do not form a consistent research strategy but
@@
 \begin{table*}[ht!]
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 \centering
-\resizebox{\textwidth}{!}{%
-\begin{tabular}{p{2.6cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}}
-\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ \hline
-\textbf{Task}                 & Game in virtual space            & Game in physical space          & Building an argument map  & Building a concept map    & Building a concept map       \\
-\textbf{Interactions}         & Audio                            & Written                         & Audio / Video             & Audio                     & Audio                        \\
-\textbf{Shared editor}        & 3D Space                         & 2D Map                          & Concept Map               & Concept Map               & Concept Map                  \\
-\textbf{Group Size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
+%\resizebox{\textwidth}{!}{%
+\begin{tabular}{p{2.1cm}L{1.65cm}L{1.7cm}L{1.8cm}L{1.6cm}L{1.6cm}}
+\toprule
+\textit{}                     & \textbf{Study 1}                 & \textbf{Study 2}                & \textbf{Study 3}          & \textbf{Study 4}          & \textbf{Study 5}             \\ 
+\midrule
+\textbf{Task}                 & game in virtual space            & game in physical space          & building an argument map  & building a concept map    & building a concept map       \\
+\midrule
+\textbf{Interactions}         & audio                            & written                         & audio / video             & audio                     & audio                        \\
+\midrule
+\textbf{Shared editor}        & 3D space                         & 2D map                          & concept map               & concept map               & concept map                  \\
+\midrule
+\textbf{Group size}           & dyads                            & \textbf{triads}                 & \textbf{triads}           & dyads                     & dyads                            \\
+\midrule
 \textbf{Duration (mean)}      & 90min                            & 16min                           & 61min                     & 90min                     & 90min                        \\
-\textbf{Awareness tool}       & Partner's position               & Partner's current/past pos.     & -                         & Partner's concept map     & Partner's scores at pre-test \\
-\textbf{Dependent var.}       & Team/Ind. performance            & Team/Ind. performance           & Team performance          & Team/Ind. knowledge       & Team/Ind. knowledge          \\
-\textbf{Independent var.}     & Awareness tool \textit{vs} none  & Awareness tool \textit{vs} none & Audio \textit{vs} Audio+Video & Script \textit{vs} none & Awareness tool \textit{vs} none       \\
+\midrule
+\textbf{Awareness tool}       & partner's position               & partner's current/past pos.     & -                         & partner's concept map     & partner's scores at pre-test \\
+\midrule
+\textbf{Dependent variable}       & team/indiv. performance            & team/indiv. performance           & team performance          & team/indiv. knowledge       & team/indiv. knowledge          \\
+\midrule
+\textbf{Independent variable}     & awareness tool \textit{vs} none  & awareness tool \textit{vs} none & audio \textit{vs} audio+video & script \textit{vs} none & awareness tool \textit{vs} none       \\
 %\textbf{Grounding crit.}      & +0.42                            & +0.15                           & +0.22                     & +0.08                     & +0.40                        \\
@@
 %$\mathcal{H}_3$               & supported                        & supported                       &                           & supported                 &                              \\
+\bottomrule
 \end{tabular}
-}
-\caption{Comparison of the studies}
+%}
+\caption{Overview of the studies}
 \label{synthesis_table}
@@
 
-{\sc SpaceMiners} is a 3D game that involves two players collecting minerals
-located in asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch
+{\sc SpaceMiners} is a 3D game that involves two players harvesting ore
+found on asteroids (Fig.~\ref{study1:spaceminer}). To do so, they must launch
 drones through the space after choosing their initial direction and speed. Once
-launched, the trajectory of drones is only influenced by the gravity of planets
-and by trajectory modification tools.  During the experiment, the teams were
+launched, the trajectory of drones is influenced by the gravity of planets
+and by ``trajectory modification'' tools. During the experiment, the teams were
 confronted with three increasingly complex situations. The experiment was 2
-hours long: a 30 minute tutorial and 3 levels of 30 minutes. Players uses using
+hours long: a 30 minute tutorial and 3 levels of 30 minutes. The players were using
 a regular joystick and communicated with each other through an audio channel.
@@
 The independent variable was the availability of an awareness tool that shows to
-player \A  the location and gaze direction of player \B. In the awareness
+player \A  the location and gaze direction of player \B: in this ``awareness''
 condition, players could switch to the \emph{scout mode} where they could view
@@
 as the number of common answers between questionnaires \gmodel{A}{A} and
-\gmodel{A}{B} in each game and computed the average value a across the 3 levels.
+\gmodel{A}{B} in each game and computed the average value across the 3 levels.
 
@@
 awareness condition who spent more time in the scout mode reached higher levels
-of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a \textit{post-hoc} split
+of \gModel{A}{B}($F = 8.02, p = 0.015$). Of course, a post-hoc split
 does not support a causal direction. An alternative explanation could be that
@@
 skills. We collect more evidence along the next studies.
-
-$\mathcal{H}_{3}$
 
@@
 \paragraph{Grounding criterion} The correlation between \gModel{A}{B} and the
-task performance (path lengths) was low: 0.15. Using a \emph{post-hoc} split on
+task performance (path lengths) was low: 0.15. Using a post-hoc split on
 \gModel{A}{B}, there is no significant difference between the performance of the
 groups with high and low \gModel{A}{B}  ($F = 1.45, p = 0.24$). Conversely, a
-\emph{post-hoc} split of the groups according to their performance did not show
+post-hoc split of the groups according to their performance did not show
 any significant differences on \gModel{A}{B} ($F = 1.16, p = 0.29$).
@@
 using eye tracking methods to address this challenge. 
-Third, our results emerge from a \emph{post-hoc} reinterpretation of studies
+Third, our results emerge from a post-hoc reinterpretation of studies
 that addressed different research questions (media richness, awareness tools,

### 2016-02-05 mutual-modeling.tex
Completed proof-reading

@@
 space \textit{[citation removed for blind review]}. We studied whether players
-build an accurate model of the path followed by their partners, assuming this
-path reflected their problem solving strategy. We used an objective measure of
+build an accurate model of the path followed by their partners, assuming that this
+path would reflect their problem solving strategy. We used an objective measure of
 \gModel{A}{B}: the distance between where \A believes \B has been walking and
@@
 virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
-Players used a Tablet PC that displayed a map of the campus and an indication of
+The players used a Tablet PC that displayed a map of the campus and an indication of
 their personal distance to Bob. Their annotations on the map were shared with
@@
 faded out after a few minutes to avoid covering the full display. The awareness
-tool displayed as well the location of the two other players on the map. This
-resulted in three conditions: the control condition (without tool) and two
+tool displayed as well the location of the two other players on the map. Three
+conditions were considered: the control condition (without tool) and two
 experimental conditions: synchronous awareness (display of the current position
@@
 
-\begin{figure*}[h!t]
-        \centering
-        \begin{subfigure}{.45\textwidth}
-            \includegraphics[width=\linewidth]{image6.png}
-            \caption{A's drawing of \B's path.}
-        \end{subfigure}
-        \begin{subfigure}{.4\textwidth}
-            \includegraphics[width=\linewidth]{image7.png}
-            \caption{Actual path followed by \B.}
-        \end{subfigure}
-        \caption{Reported and actual path of one of the player, during the {\sc
-        CatchBob} game.}
-        \label{study2:paths}
-\end{figure*}
+\begin{figure*}[h!t] \centering \begin{subfigure}{.45\textwidth}
+\includegraphics[width=\linewidth]{image6.png} \caption{A's drawing of \B's
+path.} \end{subfigure} \begin{subfigure}{.4\textwidth}
+\includegraphics[width=\linewidth]{image7.png} \caption{Actual path followed by
+\B.} \end{subfigure} \caption{Reported and actual path of one of the player,
+during the {\sc CatchBob} game.} \label{study2:paths} \end{figure*}
 
@@
 while drawing the path of their partners. We compared the path that player \A
-attributed to \B with \B's real path recorded by the system and the same for
-\A \& \C and \B \& \C as depicted on Fig.~\ref{study2:paths}.
+attributed to \B with \B's real path recorded by the system and the same for \A
+\& \C and \B \& \C as depicted on Fig.~\ref{study2:paths}.
 
@@
 where he/she had gone. One could argue that \gModel{A}{B} is biased by the
-subjects' ability to translate their trajectories souvenir into a map drawing.
-However, 85\% of subjects made no mistake at all when drawing their own path. We
-therefore consider mistakes in their partners' path as being due to a lack of
-mutual modelling accuracy instead of being due to spatial reasoning skills.
+subjects' ability to translate the memory of their trajectories into a map drawing.
+However, 85\% of subjects made no mistake at all when drawing their own path.
+We therefore consider mistakes in their partners' path as being due to a lack
+of mutual modelling accuracy instead of being due to spatial reasoning skills.
 
@@
 task performance (path lengths) was low: 0.15. Using a post-hoc split on
-\gModel{A}{B}, there is no significant difference between the performance of the
+\gModel{A}{B}, we found no significant difference between the performance of the
 groups with high and low \gModel{A}{B}  ($F = 1.45, p = 0.24$). Conversely, a
@@
 
-This positive correlation observed in the symmetry question confirms the first
-study.  In this case, this is actually more surprising given  the high
-heterogeneity of spatial skills among adults~\citep{liben1981spatial}. This
-result supports  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The results regarding
-$\Delta_2$ may support both  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$  but
-discards   $\mathcal{H}_{1}$  (if the modeller's skill would matter a lot,
-$\Delta_2$ would be lower). Conversely, $\Delta_3$  supports $\mathcal{H}_{1}$
-and  $\mathcal{H}_{3}$ but discards   $\mathcal{H}_{2}$  (if the modeller's
-skill would matter a lot,  $\Delta_3$ would be lower). In summary, various
-indices support the 3 hypotheses, which implies there is some truth in each of
-them, but $\mathcal{H}_{3}$ is the only hypothesis that is not rejected by any
-index. We may hence, with great caution, concluded that the social perspective
-($\mathcal{H}_{3}$) is moderately reinforced by this second study. Since the
-values of $\Delta_2$ and $\Delta_3$ are very similar, we do not interpret their
-minor difference as evidence for a stronger role of the modeller
-($\mathcal{H}_{1}$) or the modellee ($\mathcal{H}_{1}$).
-
-We have however to bring some nuances to the social viewpoint
+The positive correlation observed in the symmetry question confirms the first
+study.  In this case, this was not expected given the high heterogeneity of
+spatial skills among adults~\citep{liben1981spatial}. This result therefore
+supports  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The results regarding
+$\Delta_2$ support both  $\mathcal{H}_{2}$ and  $\mathcal{H}_{3}$  but discards
+$\mathcal{H}_{1}$: if the skill of the modeller would dominate -- as
+hypothesized by $\mathcal{H}_{1}$, \gModel{A}{C}  and \gModel{B}{C} would tend
+\emph{not} to be generally correlated.  Conversely, $\Delta_3$  supports
+$\mathcal{H}_{1}$ and $\mathcal{H}_{3}$ but discards $\mathcal{H}_{2}$  (if the
+modellee's skill were to dominate, \gModel{A}{B}  and \gModel{A}{C} would tend
+\emph{not} to correlate). In summary, various indices support the 3 hypotheses,
+which implies there is some truth in each of them, but $\mathcal{H}_{3}$ is the
+only hypothesis that is not rejected by any index. We may hence, with great
+caution, conclude that the social perspective ($\mathcal{H}_{3}$) is moderately
+reinforced by this second study. Since the correlation values for $\Delta_2$ and
+$\Delta_3$ are similar, we do not interpret their minor difference as evidence
+for a stronger role of the modeller ($\mathcal{H}_{1}$) or the modellee
+($\mathcal{H}_{2}$).
+
+We have also to bring some nuances to the social viewpoint
 ($\mathcal{H}_{3}$).  The main feature that can be associated to the team level
@@
 on \Model{A}{B}{emotions}.  The hypothesis was that video communication would
-lead to a better \gModel{A}{B} than audio only since emotions often impact on
+lead to a better \gModel{A}{B} than audio only since emotions often impact 
 facial expressions.
@@
 webcam and the software {\sc IVisit}. For the audio link, we used microphones,
-headsets and the {\sc BattleCom} software. In the \emph{audio plus video}
+headsets and the {\sc BattleCom} software. In the \emph{audio + video}
 condition, the screen was divided in three sections. The main part was devoted
@@
 
-Sixty students (twenty triads) from the University of Geneva participated to
+%ANON
+%Sixty students (twenty triads) from the University of Geneva participated to
+Sixty students (twenty triads) from \textit{[hidden for blind review]} participated to
 this experiment (36 women and 24 men). We formed groups of subjects who knew
@@
 \gModel{A}{B} was computed as the distance between the two vectors
-\gmodel{A}{B} and \gmodel{B}{B}:
+\gmodel{A}{B} and \gmodel{B}{B}, hence \emph{the smaller the score, the more
+accurate the model}:
 
@@
 \]
-
-The smaller the score, the \emph{more} accurate the model.
 
@@
 \paragraph{Study-specific questions} Our hypothesis about media richness is
-rejected: the average score for \model{A}{B}{emotions} was 1.25 ($SD = 0.53$) in
-the audio+video condition and 1.09 ($SD = 0.41$) in the audio alone condition
-($t=1.89, df=111, p = 0.062$): the average degree of accuracy of
-\model{A}{B}{emotions} was hence \emph{higher} in the audio alone condition.
+rejected: the average degree of accuracy for \model{A}{B}{emotions} was 1.25
+($SD = 0.53$) in the audio+video condition and 1.09 ($SD = 0.41$) in the audio
+alone condition ($t=1.89, df=111, p = 0.062$). The smaller distance between
+\model{A}{B}{emotions} and \model{B}{B}{emotions} in the audio condition shows
+that, in average, the degree of accuracy of \model{A}{B}{emotions} is
+\emph{higher} in the audio alone condition.
 
@@
 
+\paragraph{Rectangle question} We cannot address the relationship $\Delta_4$
+between self and social accuracy here because we do not have an estimation of
+self-accuracy: subjects describe their own emotions but we have no way to check if
+they are correct. By measuring \gModel{A}{B} on  18 emotional labels,
+we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
+according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
+errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
+is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
+emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
+0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
+interpret this too far, but to show that there are large variations even within one
+area  (perceiving emotions). These variations still question the existence of
+a general aptitude to model others ( $\mathcal{H}_{1}$).
+
 \begin{figure*}[ht!]
@@
 
-\paragraph{Rectangle question} We cannot address the relationship ($\Delta_4$)
-between self and social accuracy here because we do not have an estimation of
-self-accuracy: subjects describe their emotions but we have no way to check if
-there are correct. By measuring \gModel{A}{B} on  18 emotional labels,
-we can however have a glimpse about $\Delta_5$: how \Model{A}{B}{X} varies
-according to X.  Fig.~\ref{study3:deg_m_values} shows the range of modelling
-errors: the difference between \gmodel{A}{B} and \gmodel{B}{B}, on a scale of 7,
-is 0.3 in average for the emotion \emph{discasted}, and up to 1.9 for the
-emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
-0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
-interpret this too far, but to show that there are large variations even within one
-area  (perceiving emotions). These variations somehow question the existence of
-a general aptitude to model others ( $\mathcal{H}_{1}$).
-
 \subsubsection*{Discussion}
@@
 In this third study, the accuracy of mutual modelling  between two peers tend to
-be symmetrical, which  supports $\mathcal{H}_{2}$ and ($\mathcal{H}_{3}$.)  This
-result is contradicted by the fact that ($\Delta_2$) is not significant, which
-rejects both $\mathcal{H}_{2}$ and, to a lower extent,,  $\mathcal{H}_{3}$.
-Finally,  ($\Delta_3$) supports both  $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$.
-In summary, this study brings some supports to  $\mathcal{H}_{3}$, but reveal
-again that $\mathcal{H}_{3}$ is only part of the explanation.
-
-Like the previous one, this study leads us to refine what we mean by quality of
-interactions in $\mathcal{H}_{3}$. We  expected that the video channel would
-help peers building a more accurate model of the each other's emotions.  The
+be symmetrical, which  supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. This
+result is contradicted by the fact that $\Delta_2$ is not significant, which
+contradicts both $\mathcal{H}_{2}$ and, to a lower extent, $\mathcal{H}_{3}$.
+Finally, $\Delta_3$ supports both $\mathcal{H}_{1}$ and  $\mathcal{H}_{3}$: this
+study brings some supports to  $\mathcal{H}_{3}$, but reveals again that
+$\mathcal{H}_{3}$ is only part of the explanation.
+
+Like the previous one, this study leads us to refine what we mean by ``quality
+of interactions'' in $\mathcal{H}_{3}$. We  expected that the video channel
+would help peers building a more accurate model of each other's emotions. The
 results show the opposite: peers in the audio-only condition built more accurate
@@
 other~\citep{gaver1993one,anderson1997impact}. Therefore, what is meant by the
-quality of interactions is $\mathcal{H}_{3}$ is more than the linguistic features
-of dialogue but includes the way these interactions are articulated to the task.
+quality of interactions in $\mathcal{H}_{3}$ is more than the linguistic
+features of dialogue but includes the way these interactions are articulated to
+the task.
 
@@
 the post-test scores, used to assess \Model{A}{B}{knowledge}. In phase 6,
-participants were ask to estimate (7-point Likert scale) their own and their
+participants were asked to estimate (7-point Likert scale) their own and their
 partner's outcome knowledge with respect to each chapter of the learning
@@
 \paragraph{Study-specific questions} We performed a non-parametric Mann-Whitney
-test on post-test scores for electric questions and a one-way ANOVA on scores
-for ionic questions (Levene tests for homogeneity of variances): $p = 0.02$ and
-$ns$, respectively. Results  did not show any significant difference between the
-\emph{same information} (SI) condition and the \emph{complementary information}
-(CI) condition, neither for electric questions ($U = 388.50, z = -0.88, ns$) nor
-for ionic questions ($F(1, 58) = 0.17, ns$).  The effect of scripts on
-\gModel{A}{B} was not significant ($F(1, 58) = 0.78, ns$) when considering the
-absolute difference between \gModel{A}{B} and \B's post-test score. However, \A
-tended underestimate \B's score in the SI condition ($M = -2.06$) and to
-overestimate it in the CI condition ($M = 1.21$) ($F(1, 58) = 6.44, p<0.01$).
-Regarding to \gModel{A}{A}, there was no significant difference between
-conditions.
+test on post-test scores for the questions touching the electrical inner working
+of neurons, and a one-way ANOVA on scores for chemistry-related questions
+(Levene tests for homogeneity of variances): $p = 0.02$ and $ns$, respectively.
+Results did not show any significant difference between the \emph{same
+information} (SI) condition and the \emph{complementary information} (CI)
+condition, neither for electrical-related questions ($U = 388.50, z = -0.88,
+ns$), nor for chemistry-related questions ($F(1, 58) = 0.17, ns$).
+
+The effect of scripts on \gModel{A}{B} was not significant ($F(1, 58) = 0.78,
+ns$) when considering the absolute difference between \gModel{A}{B} and \B's
+post-test score. However, \A tended to underestimate \B's score in the SI
+condition ($M = -2.06$) and to overestimate it in the CI condition ($M = 1.21$)
+($F(1, 58) = 6.44, p<0.01$).  Regarding \gModel{A}{A}, there was no
+significant difference between conditions.
 
@@
 
-\paragraph{Rectangle question} The correlation between \gmodel{A}{A} and
-\gmodel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$) as well as not
-significant in each condition: someone good at self-modelling is not necessarily
+\paragraph{Rectangle question} The correlation between \gModel{A}{A} and
+\gModel{A}{B} ($\Delta_4$) is globally not significant ($r=0.05$), and neither
+it is in each of the conditions: someone good at self-modelling is not necessarily
 good at modelling someone else and vice-versa.  This seems to indicate that
@@
 How do we interpret the fact that the symmetry vanishes when peers receive
-different texts to read before the task? One explanation would be the difficult
-of mutual modelling when peers don't know what others read, but we found no
+different texts to read before the task? One explanation would be the difficulty
+of mutual modelling when peers do not know what the others read, but we found no
 significance of \gModel{A}{A} between conditions. Since texts were partly
@@
 at this second level of modelling, which supports the existence of such an
-individual skill, as stated in  $\mathcal{H}_{2}$
+individual skill, as stated in  $\mathcal{H}_{2}$.
 
@@
 This study investigates if \Model{A}{B}{knowledge} is related to learning
-outcomes by comparing teams with or without a knowledge awareness tool (KAT),
+outcomes by comparing teams with or without a Knowledge Awareness Tool (KAT),
 \ie a tool that informs \A about \B's knowledge as measured through a pre-test.
@@
 
-The participants of the experimental condition group were provided with the KAT
-on the bottom part of the screen (Fig.~\ref{study5:kat}): each line represents
-the score obtained by the partner at the pre-test for a chapter. Participants
-did not see they own score, but they usually started their discussion by
-exchanging this information.
+The participants of the experimental condition group were provided with the
+Knowledge Awareness Tool on the bottom part of the screen
+(Fig.~\ref{study5:kat}): each line represents the score obtained by the partner
+at the pre-test for a chapter. Participants did not see their own score, but they
+usually started their discussion by exchanging this information.
 
@@
 difference between the KAT condition participants ($M = 13.4\%$) and the control
-group [$M = 3.6\%, t(1, 60) = 2.73, p < 0.01$, Cohen's $d = 0.7$, medium to large
+group $M = 3.6\%$ [$t(1, 60) = 2.73, p < 0.01$, Cohen's $d = 0.7$, medium to large
 effect]: providing learners with cues about the prior-knowledge of their partner
@@
 tracking devices) and the learning gains ($r(22) = 0.54, p = 0.01$). A detailed
-analysis revealed that subjects look at KAT to assess their peer's credibility
+analysis revealed that the participants look at KAT to assess their peer's credibility
 when he/she provided new information.
@@
 estimated their partners knowledge ($M = 1.11$) than those in the control
-condition [$M = 0.98, t(1, 60) = 3.19, p < 0.01$, Cohen's $d = 0.83$, large
+condition $M = 0.98$ [$t(1, 60) = 3.19, p < 0.01$, Cohen's $d = 0.83$, large
 effect]. This is a trivial result since the KAT provided them with an initial
-\gmodel{A}{B}. However, subjects have to predict the post-test score while the
+\gmodel{A}{B}. However, the participants have to predict the post-test score while the
 KAT informed them about the pre-test score. Actually, pairs in the KAT condition
@@
 knowledge (\gmodel{A}{A}) such as ``I don't remember the Ranvier's thing...''
-and (3) elaborated utterances with rich contents,  These 3 categories provide
+and (3) elaborated utterances with rich contents. These three categories provide
 different account of \gModel{A}{B}: (1) as an effect of \A's effort to model
@@
 \paragraph{Symmetry question} We did not find an intra-pair correlation ($ICC =
-0.05, ns$), which discards  ($\mathcal{H}_{2}$) and  ($\mathcal{H}_{3}$). In
+0.05, ns$), which does not support $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. In
 addition, the KAT supports the processing of modelling at the first level
@@
 
- ($\mathcal{H}_{3}$)
-
 %\paragraph{Rectangle question} $\Delta_4$: the correlation between \gmodel{A}{A} and \gmodel{A}{B} is
@@
 
-We found symmetry of the mutual model on 4 studies out of 5. In study 4, this
+We found a symmetry of the mutual models on 4 studies out of 5. In study 4, this
 only applies to the control group (having read the same text before), which
-corresponds to the situation of the 3 first studies.  Finding this symmetry per
-se constitutes an interesting result as we don't know earlier studies who have
-established this relationship. The problem is that this symmetry does not allow
-use to discriminate  $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. The second
-hypothesis, the skill of helping one's peer to model oneself, is questioned by
-study 3 where \C could be modelled  accurately by \A and not accurately by \B. The
-same hypothesis is however supported by the results of study 2, and indirectly
-by study 4.  This does not discard $\mathcal{H}_{1}$: even when we found
-significant correlations, they were all below 0.50, not around 0.90.  Hence,
-even of the quality of social interaction matters, there is obviously a large
-part of individual variance within teams.
-
-In other words, these studies do conclude that one hypothesis is right and the
-others are wrong, which is per se the main contribution of this paper. We
+corresponds to the situation of the 3 first studies. Evidencing this symmetry
+constitutes \textit{per se} an interesting result as we are not aware of earlier
+studies who have established this relationship.
+Still, the symmetry alone does not allow to discriminate $\mathcal{H}_{2}$ from
+$\mathcal{H}_{3}$. The second hypothesis (\ie mutual modelling is primarily
+supported by the capacity of the modellee to help the modellers to form a
+correct image of herself), is questioned by study 3 where \C could be modelled
+accurately by \A and not accurately by \B. The same hypothesis is however
+supported by the results of study 2, and indirectly by study 4.
+
+This does not rule out entirely $\mathcal{H}_{1}$ (\ie partner modelling is
+primarily a individual skill) either: even where we found significant
+correlations, they were all below 0.50, not around 0.90.  Hence, even if the
+quality of the social interaction matters, there is obviously a large part of
+individual variance within teams.
+
+In other words, these studies do not conclude that one hypothesis is right and the
+others are wrong, and this is indeed the main contribution of this article. We
 started from the idea that partner modelling is essentially an individual skill
-and that we therefore improve the quality of collaboration by providing the
-so-called awareness tools, which act as some kind of prostheses for partner
-modelling. This individual skill cannot be discarded.  Everyone has experienced
-the pleasure of interacting with colleagues who answer precisely to the question
+and that we would therefore improve the quality of collaboration by providing 
+awareness tools, as those tools would act as a kind of prosthesis for partner
+modelling. The role and importance of this individual skill cannot be discarded:
+everyone has experienced the pleasure of interacting with colleagues who answer precisely to the question
 we asked them and even guess the reason why we asked this question. Conversely,
@@
 person by his name, say Mike Smith, while knowing perfectly that there is no
-chance that we know this Mike Smith. The existence of these individual skills is
+chance that we know this Mike Smith. In fact, the role of this individual skill is
 not denied by our studies; the novelty of this paper is to show, thanks to
@@
 We acknowledge that the difference between these hypotheses is rather
-theoretical since the  process of modelling one's partner( $\mathcal{H}_{1}$)
-and the process of helping one's partner to model oneself ($\mathcal{H}_{3}$) are
+theoretical since the  process of modelling one's partner $\mathcal{H}_{1}$ and
+the process of helping one's partner to model oneself $\mathcal{H}_{2}$ are
 mediated by verbal interactions in the team. One could hardly imagine someone
@@
 There is a bidirectional causal link between the accuracy of mutual modelling
-and the quality of interactions. This rather artificial distinction does however
-participate to the more fundamental discussion of the role of individual and
-social mechanisms in human cognition. In the field of social cognition, it is a
-common place to state that ``the whole is greater than the sum of the parts''.
-This refers to the emergence of team properties than cannot be reduced to the
-set of individual contributions.  Our paper illustrates this emergence by
-showing the symmetry of mutual modelling
+and the quality of the interactions. This rather artificial distinction does
+however participate to the more fundamental discussion of the role of individual
+and social mechanisms in human cognition. In the field of social cognition, it
+is a common place to state that ``the whole is greater than the sum of the
+parts''.  This refers to the emergence of team properties than cannot be reduced
+to the set of individual contributions. Our paper illustrates this emergence by
+showing the symmetry of mutual modelling.
 
@@
 faced hard methodological issues. Providing learners with on-task questionnaires
-introduce a bias: they will pay more attention to their partners in the
-remaining time. Providing them with ``after-task'' questionnaires implies mnemonic
-and rationalization biases. The nature of mutual modelling implies methodological
-challenges that call for new  measurement methods. We have promising results for
-using eye tracking methods to address this challenge. 
-Third, our results emerge from a post-hoc reinterpretation of studies
-that addressed different research questions (media richness, awareness tools,
+introduces a bias: they will pay more attention to their partners in the
+remaining time. Providing them with ``after-task'' questionnaires implies
+mnemonic and rationalization biases. The nature of mutual modelling implies
+methodological challenges that call for new measurement methods. We have
+promising results for using eye tracking methods to address this challenge.
+Third, our results emerge from a post-hoc reinterpretation of studies that
+addressed different research questions (media richness, awareness tools,
 scripts,...). This diversity makes our results difficult to integrate, partly
-contradictory. Nonetheless, this diversity gives some generalizability: mutual
-modelling has been investigated in different contexts (virtual space versus real
-space), with different groups sizes (pairs and triads) and different tasks.
+contradictory. Nonetheless, this diversity also incidentally provides some
+generalizability: mutual modelling has been investigated in different contexts
+(virtual space versus real space), with different groups sizes (pairs and
+triads) and different tasks.
 

### 2016-02-15 rebuttal.tex
Re-submitted version

--- 2016-02-04 rebuttal.tex
@@
 We would like to thank both the reviewers and the editor for going through our
-manuscript and providing in-depth review of our submission.
+manuscript and providing in-depth reviews of our submission.
 
@@
     \item the need for a better presentation of the concepts that we build upon,
-    \item the need for a ``deeper synthetic discussion'', also ``accounting for the
-        conflicting results'', that would present the ``theoretical
-        picture'' resulting from our work,
-    \item discuss a possible research agenda
+    \item the need for a deeper synthetic discussion (also accounting for the
+        conflicting results) that would clarify the theoretical
+        picture resulting from our work.
 \end{itemize}
 
-In line with these comments and suggestions, the manuscript underwent several
-significant changes that we summarize hereafter.
+In line with these comments and the other suggestions made by the reviewers, the
+manuscript underwent several significant changes that we summarize hereafter.
 
@@
     \item \emph{Section 1 -- Introduction} We did minor edits except for the last
-        paragraph has been completely rewritten it  that was somehow misleading
+        paragraph that has been completely rewritten, as it was somewhat misleading
         regarding to the goals of the paper.  We think the point is now made
@@
 
-    \item We clarified the notion of grounding criterion, as requested by
-        reviewer 2 (page [TODO!!])
-
-    \item \emph{Section 2} has been split into partner modeling and mutual modeling (now
+    \item \emph{Section 2} has been split into partner modelling and mutual modelling (now
         section 3) to emphasize the fact that the question is the mutuality. The
-        main research question is then clearly stated to address some [WHICH?] of the
-        reviewers' criticisms.
+        main research question is then clearly stated so as to address the
+        perceived lack of clarity raised by some reviewers.
 
@@
         reviewers found useful and others criticized. We are not
-        proposing here a computational theory of mutual modeling. This notation is
+        proposing here a computational theory of mutual modelling. This notation is
         only useful because the text become quickly unreadable as one talks about
@@
         this paper.
+
+    \item As requested by the reviewer 2, we have also clarified how we estimate the
+        grounding criterion (p.6), and clarified as well that we have not attempted to
+        quantify the cost of grounding in the presented studies (p.8).
 
@@
     \item As requested by the reviewers, we have entirely rewritten  the
-        conclusions to make them more explicit.  We cannot of course not pretend
-        that our results are clearcut. We have to be careful not to publish only
-        papers with significant results (statisticians have proved that some
+        discussion: our results are not clearcut (and, in that regard, we argue
+        that as a community, we have to be careful not to publish only
+        papers with significant results -- statisticians have proved that some
         meta-analyses are biased by the fact that papers with significant results
-        get more easily published). However, despite the fact that none of the 3
-        hypothesis is declared as winner, the key results is that we prove that
-        individual skills only explain a part of the variance, that the quality of
-        social interactions matter significantly.
+        get more easily published). However, the fact that none of our three
+        main hypotheses (the quality of mutual modelling depends on the modeller's
+        skills \emph{vs} it depends on the modellee's capability of making herself easy
+        to model \emph{vs} it depends on the overall quality of the interaction)
+        emerges as a clearly dominant factor is the key result of our work: we indeed show that
+        individual skills only explain a part of the observed variance in
+        collaborative tasks performance, and that the quality of
+        social interactions does have a significant impact as well.
 
@@
 
+We do hope that the sum of the changes brought to the article clarifies and
+strengthens our contribution, and we remain of course dedicated to account for further
+suggestions if necessary.
+
 \end{document}

### 2016-02-15 authors.tex
Added back Mirweis
Re-submitted version

@@
 
+\usepackage{xspace}
+\newcommand{\A}{A\xspace}
+\newcommand{\B}{B\xspace}
+\newcommand{\C}{C\xspace}
+
 \newcommand{\M}[3]{{\mathcal{M}(#1, #2, #3)}}
@@
 
-\title{Mutual Modelling}
+\title{The Symmetry of Partner Modelling}
 
@@
     Séverin Lemaignan
+    \and
+    Mirweis Sangin
     \and
@@
     \and
+    Mirweis Sangin \at Use-able consulting \\ CH-1004 Lausanne, Switzerland\\\email{mirweis@gmail.com}
+    \and
     Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
@@
 \begin{abstract}
-    Collaborative learning has something to do with building a shared
-    understanding of the situation at hand. Yet, the psycholinguistics mechanisms
-    at work while establishing common grounds are complex and certainly involve
-    modelling what the partner knows/wants/intends at a given time.
-
-    Building on a set of five independent studies that where conducted over
-    several years, we propose in this article a simple model of the mechanisms
-    of mutual modelling. It integrates several concepts related to the classical
-    idea of grounding, and we apply it as an operational tool to
-    analyse the interplay of learning and mutual modelling in the five studies.
-
-    We come up with mixed experimental results that evidence the dual nature of
-    mutual modelling: a individual meta-cognitive aptitude that co-exists with
-    modelling elicited and fostered by social interactions.
-
+Collaborative learning has often been associated to the construction of  a
+shared   understanding of the situation at hand. The psycholinguistics
+mechanisms at work while establishing common grounds are the object of
+scientific controversy. We postulate that collaborative tasks require some
+level of mutual modelling, \ie that each partner needs some model of what
+the other partners know/want/intend at a given time.  We use the term ``some
+model'' to stress the fact that this model is not necessarily detailed or
+complete, but that we acquire some representations of the persons we
+interact with. The question we address is: Does the quality of the partner
+model depend upon the modeler's ability to represent his or her partner?
+Upon the modelee's ability to make his state clear to the modeler? Or rather upon
+the quality of their interactions? We address this question by comparing
+the respective accuracies of the models built by different team
+members. We therefore report on 5 experiments on collaborative problem solving
+or collaborative learning that vary in terms of tasks (how important it is
+to build an accurate model) and settings (how difficult it is to build an
+accurate model). In 4 studies, the accuracy of the model that \A built about
+\B was correlated with the accuracy of the model that \B built about \A,  which
+seems to imply that the quality of interactions matters more than individual
+abilities when building mutual models.  However, these finding do not
+rule out the fact that individual abilities also participate in the quality of
+modelling process.
 \end{abstract}
@@
 \subsection*{Corresponding author}
-Séverin Lemaignan \texttt{severin.lemaignan@epfl.ch}
+Séverin Lemaignan \texttt{severin.lemaignan@plymouth.ac.uk}
 \end{document}

### 2016-02-23 mutual-modeling.tex
Title case on sections
De-anonymized

@@
 
-% ANON
-%\author{
-%    Pierre Dillenbourg
-%    \and
-%    Séverin Lemaignan
-%    \and
-%    Mirweis Sangin
-%    \and
-%    Nicolas Nova
-%    \and
-%    Gaëlle Molinari
-%}
-%\institute{
-%    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
-%    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
-%    \email{firstname.lastname@epfl.ch}
-%    \and
-%    Mirweis Sangin \at Use-able consulting \\ CH-1004 Lausanne, Switzerland\\\email{mirweis@gmail.com}
-%    \and
-%    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
-%    \and
-%    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
-%    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
-%}
+\author{
+    Pierre Dillenbourg
+    \and
+    Séverin Lemaignan
+    \and
+    Mirweis Sangin
+    \and
+    Nicolas Nova
+    \and
+    Gaëlle Molinari
+}
+\institute{
+    Pierre Dillenbourg, Séverin Lemaignan \at Computer-Human Interaction in Learning and Instruction \\
+    École Polytechnique Fédérale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
+    \email{firstname.lastname@epfl.ch}
+    \and
+    Mirweis Sangin \at Use-able consulting \\ CH-1004 Lausanne, Switzerland\\\email{mirweis@gmail.com}
+    \and
+    Nicolas Nova \at Haute École Arts \& Design de Genève \\ CH-1201 Genève, Switzerland\\\email{nicolas.nova@hesge.ch}
+    \and
+    Gaëlle Molinari \at Distance Learning University Switzerland (Unidistance),
+    CH-3960 Sierre, Switzerland\\\email{gaelle.molinari@unidistance.ch}
+}
 \author{}
@@
 partner when they know that they can later provide feedback on his/her
-%ANON
-%understanding~\citep{schober1993spatial}. We introduced the notion of
-%\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
-understanding~\citep{schober1993spatial}. \citet{dillenbourg1995evolution} introduced the notion of
-\emph{optimal collaborative effort} to stress
+
+understanding~\citep{schober1993spatial}. We introduced the notion of
+\emph{optimal collaborative effort}~\citep{dillenbourg1995evolution} to stress
 that misunderstanding should not be viewed as something to be avoided (if this
@@
 
-\section{A notation for discussing mutual modelling }
+\section{A Notation for Discussing Mutual Modelling }
 
@@
 
-\section{Hypotheses and questions}
+\section{Hypotheses and Questions}
 
@@
 We studied the impact of an awareness tool on group performance and mutual
-%ANON
-%modelling~\citep{nova2007collaboration}. The availability
-modelling \textit{[citation removed for blind review]}. The availability of an
+modelling~\citep{nova2007collaboration}. The availability of an
 awareness tool was our independent variable. In previous studies, we replayed a
@@
 This study concerns a collaborative game that occurred in the physical
-%ANON
-%space~\citep{nova2006underwhelming}. We studied whether players build an accurate
-space \textit{[citation removed for blind review]}. We studied whether players
+space~\citep{nova2006underwhelming}. We studied whether players
 build an accurate model of the path followed by their partners, assuming that this
@@
 {\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
-% ANON
-%joint task. The game was played on EPFL campus. Participants had to find a
 joint task. The game was played on our campus. Participants had to find a
@@
 remember their partners' paths when they could not constantly monitor their
-%ANON
-%positions. As detailed in the original study~\citep{nova2005location}, it
-positions. As detailed in the original study, it
+positions. As detailed in the original study~\citep{nova2005location}, it
 appeared that teams without awareness tool made more manual annotations on the
@@
 
-%ANON
-%The aim of this unpublished\footnote{The data and statistical analyses of this
-%study are available online:
-%
-%\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.}
-The aim of this unpublished study was to evaluate the effect of media richness
-on \Model{A}{B}{emotions}.  The hypothesis was that video communication would
-lead to a better \gModel{A}{B} than audio only since emotions often impact 
-facial expressions.
+The aim of this unpublished\footnote{The data and statistical analyses of this
+study are available online:
+\url{https://github.com/chili-epfl/mutual-modelling-emotions-study}.} study was
+to evaluate the effect of media richness on \Model{A}{B}{emotions}.  The
+hypothesis was that video communication would lead to a better \gModel{A}{B}
+than audio only since emotions often impact facial expressions.
 
@@
 
-%ANON
-%Sixty students (twenty triads) from the University of Geneva participated to
-Sixty students (twenty triads) from \textit{[hidden for blind review]} participated to
+Sixty students (twenty triads) from the University of Geneva participated to
 this experiment (36 women and 24 men). We formed groups of subjects who knew
@@
 This study investigated the effect of a collaboration script on collaborative
-%ANON
-%learning~\citep{molinari2008effects}. The script chosen is a {\sc jigsaw}: two students
-learning \textit{[citation removed for blind review]}. The script chosen is a
+learning~\citep{molinari2008effects}. The script chosen is a
 {\sc jigsaw}: two students receive different but complementary subsets of the
@@
 
-%ANON
-%Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
-Fifty-eight first year students (47 men and 11 women, mean age: 20.46)
+Fifty-eight first year students from EPFL (47 men and 11 women, mean age: 20.46)
 were remunerated for participation. Dyads were randomly assigned to one of the
@@
 The peers were located into two different rooms. A complete description of the
-%ANON
-%study is provided in~\citet{sangin2008learners}. The experiment lasted 90 minutes.
-study is provided in \textit{[citation removed for blind review]}. The experiment lasted 90 minutes.
+study is provided in~\citet{sangin2008learners}. The experiment lasted 90 minutes.
 
@@
 \subsubsection*{Subjects}
-%ANON
-%Sixty-four first year EPFL students (46 men, 18 women, mean age: 21.2)
-Sixty-four first year students (46 men, 18 women, mean age: 21.2)
+
+Sixty-four first year EPFL students (46 men, 18 women, mean age: 21.2)
 participated to the study. They were randomly assigned to conditions and
@@
 
-%ANON
-%The experiments were developed with the help of Mirweis Sangin, René Glaus,
-%Patrick Jermann,  Fabien Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan
-%Bourquin and Jeremy Goslin.  We also thank Kshitij Sharma and {\L}ukasz
-%Kidzi\'nski for their help with data analysis. The main funding has been
-%provided from a NSF Grant grant \#102511-106940.
-
-\textit{Removed for blind review.}
+The experiments were developed with the help of Mirweis Sangin, René Glaus,
+Patrick Jermann,  Fabien Girardin, Marc-Antoine Nüssli, Thomas Werhle, Yvan
+Bourquin and Jeremy Goslin.  We also thank Kshitij Sharma and {\L}ukasz
+Kidzi\'nski for their help with data analysis. The main funding has been
+provided from a NSF Grant grant \#102511-106940.
 

### 2016-02-23 mutual-modeling.tex
Fixed grammar errors pointed by editors

@@
 
-Collaborative learning has often been associated to the construction of  a
+Collaborative learning has often been associated with the construction of  a
 shared   understanding of the situation at hand. The psycholinguistics
@@
 the respective accuracies of the models built by different team
-members. We therefore report on 5 experiments on collaborative problem solving
+members. We report on 5 experiments on collaborative problem solving
 or collaborative learning that vary in terms of tasks (how important it is
@@
 seems to imply that the quality of interactions matters more than individual
-abilities when building mutual models.  However, these finding do not
+abilities when building mutual models.  However, these findings do not
 rule out the fact that individual abilities also participate in the quality of
@@
 science~\citep{chi1989self, webb1991task}, conflict resolution in
-socio-cognitive theories~\citep{doise1975social}, mutual
-regulation~\citep{blaye1995collaborative} in a Vygostkian perspective, etc. The
+socio-cognitive theories~\citep{doise1975social} and mutual
+regulation~\citep{blaye1995collaborative} in a Vygostkian perspective. The
 construction of a shared understanding has been investigated for several decades
@@
 sciences. Grounding mechanisms are appropriate to explain conversational
-events, such as referential failures in short dialogue episodes, but they do hardly
+events, such as referential failures in short dialogue episodes, but they hardly
 predict deeper phenomena such as \emph{conceptual change} (\ie the acquisition,
@@
 In summary, the controversy around the cognitive depth of shared understanding
-pertains to psycholinguistic, which investigates natural conversations. The
+pertains to psycholinguistics, which investigates natural conversations. The
 situation is different in collaborative learning; tasks are actually designed
@@
 collaborative situations that create interdependence, avoid group think, and
-allows students to detect any ``illusion of shared
+allow students to detect any ``illusion of shared
 understanding''~\citet{cherubini2005grounding}. Our question is hence not
@@
     \item $Model(A,B, ``urgent")$: what does \A know about \B's understanding of \A's
-        last utterance: does ``urgent'' means now, soon or ``not too late''?
+        last utterance: does ``urgent'' mean now, soon or ``not too late''?
 
@@
 and social views of human cognition.  The simple hypothesis is that when
-individual are randomly paired for an experiment, there is no reason for which
+individuals are randomly paired for an experiment, there is no reason for which
 their individual cognitive skills would correlate. Therefore, if it occurs that
-the quality of the $Model(A,B)$ is correlated with $Model(B,A)$, one may infers
+the quality of the $Model(A,B)$ is correlated with $Model(B,A)$, one may infer
 that this quality depends upon what \A and \B have built together while
@@
 various other research questions, but in which the correlation between the
-quality of these two models could be computed.  The studies we report do hence
-not constitute a clean sequence of experiment to investigate mutual modelling but
+quality of these two models could be computed.  The studies we report do not hence
+constitute a clean sequence of experiments to investigate mutual modelling but
 the ad-hoc revisiting of previous experiments to explore a question that we had
@@
 abstraction referring to complex socio-cognitive processes.  As explained in the
-previous section, \A can be fragmented, multi-dimensional, etc. This notation
+previous section, the model built by \A can be fragmented, multi-dimensional, etc. This notation
 is neither presented as a computational model of mutual modelling, nor as some
@@
         about $X$). We simply could not collaborate without a certain level of
-        initial grounds. We note the theoretical accuracy of \emph{initial
-        grounds} $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
+        initial grounds. We denote the theoretical accuracy of \emph{initial
+        grounds} with $\mathcal{M}^{\circ}_{t_0}(A,B,X)$.
         \vspace{1em}
@@
 between 2000 and 2015. They do not form a consistent research strategy but
-actually the fact that some trends emerge despite their diversity constitutes
-indeed the richness of this line of work.
+the fact that some trends emerge despite their diversity constitutes the
+richness of this line of work.
 
@@
 It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. It is
-nonetheless far from 1 which implies we cannot discard the individual modelling
-skills. We collect more evidence along the next studies.
+nonetheless far from 1, which implies we cannot discard the individual modelling
+skills. We collect more evidence in the next studies.
 
@@
 emotion \emph{calm}. This is probably specific to the variety of scales  ($SD=
-0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}), our point is not to
+0.5$ for \emph{discated} versus $SD=0.7$ for \emph{calm}). Our point is not to
 interpret this too far, but to show that there are large variations even within one
@@
 = 0.15$, large effect. This relationship was not significant in the previous
-experiment which as been conducted with the same task (using the same task
+experiment which was conducted with the same task (using the same task
 should produce the same grounding criterion). This is probably due to the fact that
@@
 fact that the KAT enhances  \gModel{A}{B}  brings additional supports to
-$\mathcal{H}_{1}$. However, the analysis of verbal interactions find elements
+$\mathcal{H}_{1}$. However, the analysis of verbal interactions finds elements
 that actually supports the three hypotheses.
@@
 constitutes \textit{per se} an interesting result as we are not aware of earlier
-studies who have established this relationship.
+studies that have established this relationship.
 Still, the symmetry alone does not allow to discriminate $\mathcal{H}_{2}$ from
@@
 we asked them and even guess the reason why we asked this question. Conversely,
-everyone also experienced the frustration of someone referring about a third
+everyone also experienced the frustration of someone referring to a third
 person by his name, say Mike Smith, while knowing perfectly that there is no
@@
 and the quality of the interactions. This rather artificial distinction does
-however participate to the more fundamental discussion of the role of individual
+however participate in the more fundamental discussion of the role of individual
 and social mechanisms in human cognition. In the field of social cognition, it
-is a common place to state that ``the whole is greater than the sum of the
+is a commonplace to state that ``the whole is greater than the sum of the
 parts''.  This refers to the emergence of team properties than cannot be reduced
@@
 scripts,...). This diversity makes our results difficult to integrate, partly
-contradictory. Nonetheless, this diversity also incidentally provides some
+and contradictory. Nonetheless, this diversity also incidentally provides some
 generalizability: mutual modelling has been investigated in different contexts
@@
 Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do
-not imply we have a mechanical view of modelling. It's just a language for
+not imply we have a mechanical view of modelling. The notation is just a language for
 referring to different components of modelling each other in teams.  This simple

### 2016-02-29 mutual-modeling.tex
Improved a bit the quality of 2 figures
Last changes to the synthesis

@@
         \centering
-        \includegraphics[width=\textwidth]{image4.png}
+        \includegraphics[width=\textwidth]{study1}
         \caption{Screenshot of the {\sc SpaceMiners} game}
@@
         \begin{subfigure}{.6\textwidth}
-            \includegraphics[width=\linewidth]{study5-conceptmap.png}
+            \includegraphics[width=\linewidth]{study5-conceptmap.pdf}
         \end{subfigure} \\
@@
 
+In the introduction, we mentioned the controversy around the cognitive depth of
+dialogues: do they build upon shallow syntactic alignment or do they entail
+deeper semantic processes, which would require modelling of the partner. Our
+studies rather support the second viewpoint since the quality of modelling
+emerged as an intermediate variable, sensitive to some independent variables
+(like the presence of awareness tools or media richness), and predictive of
+dependent variables, such as task performance. We do not claim that this deeper
+semantic modelling is always present in everyday conversations, but rather that
+it \emph{was} indeed present in the collaborative tasks designed for our
+experiments (which are representative of the tasks assigned to learners in CSCL
+environments).
+
+Secondarily, we questioned whether the accuracy of the partner's model depends on
+the cognitive skills of each partner or instead results from the quality of
+the interactions among them. We respectively refer to these hypothesis as
+$\mathcal{H}_1$ and $\mathcal{H}_3$. Our rationale was that a symmetry of mutual
+modelling (correlation between the accuracy of each other's model) would favor
+$\mathcal{H}_3$ over $\mathcal{H}_1$. We also formulated $\mathcal{H}_2$, by
+which a learner \A who is good at modelling \B would also help \B to repair his
+inaccurate representations of \A.
+
 We found a symmetry of the mutual models on 4 studies out of 5. In study 4, this
@@
 constitutes \textit{per se} an interesting result as we are not aware of earlier
-studies that have established this relationship.
-Still, the symmetry alone does not allow to discriminate $\mathcal{H}_{2}$ from
-$\mathcal{H}_{3}$. The second hypothesis (\ie mutual modelling is primarily
-supported by the capacity of the modellee to help the modellers to form a
-correct image of herself), is questioned by study 3 where \C could be modelled
-accurately by \A and not accurately by \B. The same hypothesis is however
-supported by the results of study 2, and indirectly by study 4.
+studies that have established this relationship.  Still, the symmetry alone does
+not allow to discriminate $\mathcal{H}_{2}$ from $\mathcal{H}_{3}$. The second
+hypothesis is questioned by study 3 where \C could be modelled accurately by \A
+and not accurately by \B. The same hypothesis is however supported by the
+results of study 2, and indirectly by study 4.
 
@@
 
-In other words, these studies do not conclude that one hypothesis is right and the
-others are wrong, and this is indeed the main contribution of this article. We
-started from the idea that partner modelling is essentially an individual skill
-and that we would therefore improve the quality of collaboration by providing 
-awareness tools, as those tools would act as a kind of prosthesis for partner
-modelling. The role and importance of this individual skill cannot be discarded:
-everyone has experienced the pleasure of interacting with colleagues who answer precisely to the question
-we asked them and even guess the reason why we asked this question. Conversely,
-everyone also experienced the frustration of someone referring to a third
-person by his name, say Mike Smith, while knowing perfectly that there is no
-chance that we know this Mike Smith. In fact, the role of this individual skill is
-not denied by our studies; the novelty of this paper is to show, thanks to
-symmetry values, that individual skills only account for a part of the accuracy
-in mutual modelling.
+In other words, these studies do not conclude that one hypothesis is right and
+the others are wrong, and this is indeed the main contribution of this article.
+We started from the idea that partner modelling is essentially an individual
+skill and that we would therefore improve the quality of collaboration by
+providing awareness tools, as those tools would act as a kind of prosthesis for
+partner modelling. The role and importance of this individual skill cannot be
+discarded: everyone has experienced the pleasure of interacting with colleagues
+who answer precisely to the question we asked them and even guess the reason why
+we asked this question. Conversely, everyone also experienced the frustration of
+someone referring to a third person by his name, say Mike Smith, while knowing
+perfectly that there is no chance that we know this Mike Smith. In fact, the
+role of this individual skill is not denied by our studies; the novelty of this
+paper is to show, thanks to symmetry values, that individual skills only account
+for a part of the accuracy in mutual modelling.
 
@@
 to the set of individual contributions. Our paper illustrates this emergence by
-showing the symmetry of mutual modelling.
+showing the symmetry of mutual modelling. In a nutshell, yes there is a
+non-negligible component of mutuality in modelling one's partner.
 
@@
 
-Finally, let us repeat that the expressions we used, such as \gmodel{A}{B}, do
-not imply we have a mechanical view of modelling. The notation is just a language for
-referring to different components of modelling each other in teams.  This simple
-notation emerged from the need to compare five mutual modelling studies. It
-could however pave the road for elaborating a systematic agenda for research on
-mutual modelling.
+These limitations and our difficulty to provide clear-cut conclusions comes from
+the fact that this series of experiments was not planned a priori. This paper
+relies on the post-hoc comparison of experiments conducted across various
+contexts. The overall conclusion is that this research question would deserve a
+specific research agenda for 3 reasons: (1) it raises fundamental theoretical
+dilemma on the social nature of cognition, (2) it raises methodological
+challenges and (3) it could provide empirical grounds to design decisions for
+CSCL environments. The eye tracking tools we developed and the notation we used
+for referring to different components of mutual modelling could pave the road
+for elaborating a systematic agenda for research on mutual modelling. Let us
+repeat that the expressions we used, such as \gmodel{A}{B}, do not imply we have a
+mechanical view of modelling, but were simply useful ways to talk about mutual
+modelling.
 

### 2016-03-13 mutual-modeling.tex
Rephrase synthesis to refer to lines of thoughts from intro

@@
 In the introduction, we mentioned the controversy around the cognitive depth of
-dialogues: do they build upon shallow syntactic alignment or do they entail
-deeper semantic processes, which would require modelling of the partner. Our
-studies rather support the second viewpoint since the quality of modelling
-emerged as an intermediate variable, sensitive to some independent variables
+dialogues: does efficient dialogue require some modeling of what the partner
+understands or intends to convey, as initially postulated by
+\citet{clark1986referring}, or can dialogue simply rely on some shallow
+syntactic alignment, as objected by \citet{pickering2006alignment}. While the
+shallow hypothesis may be relevant to simple chat situations, we investigated
+this issue in richer problem solving tasks that are more representative of the
+tasks assigned to learners in CSCL environments. In four different tasks, we
+found evidence that partner model each other since the quality of modelling
+emerged as an intermediate variable, sensitive to several independent variables
 (like the presence of awareness tools or media richness), and predictive of
-dependent variables, such as task performance. We do not claim that this deeper
-semantic modelling is always present in everyday conversations, but rather that
-it \emph{was} indeed present in the collaborative tasks designed for our
-experiments (which are representative of the tasks assigned to learners in CSCL
-environments).
+several dependent variables, such as task performance or learning gains.
+Thereby, even if we do not bring any definite conclusion to this debate, our
+results support the CSCL school of thought in which shared understanding or
+intersubjective meaning making have been a foundational concept
+\citep{roschelle1995construction,schwartz1995emergence,dillenbourg2006sharing,suthers2006technology,stahl2007meaning}.
 

### 2016-04-18 mutual-modeling.tex
Integrated proof-reading from CSCL editors

@@
 \usepackage{subcaption}
+\usepackage{endnotes}
 
@@
 model depend upon the modeler's ability to represent his or her partner?
-Upon the modelee's ability to make his state clear to the modeler? Or rather upon
+Upon the modelee's ability to make his state clear to the modeler? Or rather, upon
 the quality of their interactions? We address this question by comparing
@@
 abilities when building mutual models.  However, these findings do not
-rule out the fact that individual abilities also participate in the quality of
+rule out the fact that individual abilities also contribute to the quality of
 modelling process.
@@
 regulation~\citep{blaye1995collaborative} in a Vygostkian perspective. The
-construction of a shared understanding has been investigated for several decades
+construction of shared understanding has been investigated for several decades
 in psycholinguistics, under the  notion of
@@
 predict deeper phenomena such as \emph{conceptual change} (\ie the acquisition,
-acceptation and integration of a new belief into one's mental model) over longer
+acceptance and integration of a new belief into one's mental model) over longer
 sessions~\citep{dillenbourg2006sharing}. The cumulative effect of grounding
@@
 tools that constitute the mediators of interpersonal
-interaction}~\citep{baker1999role}
+interaction}~\citep[p.31]{baker1999role}
 \end{quote}
 
-Along this line, several scholars suggest that CSCL research should go deeper in
-the understanding of how partners engage into shared meaning
+\noindent Along this line, several scholars suggest that CSCL research should go deeper 
+towards understanding how partners engage in shared meaning
 making~\citep{stahl2007meaning} or \emph{intersubjective} meaning
@@
 even the repair of some misunderstandings, is explained by this mimetic behavior
-more than by a monitoring of each other's knowledge: \emph{``...interlocutors do
+more than by a monitoring of each other's knowledge:
+
+\begin{quote}
+[\ldots] interlocutors do
 not need to monitor and develop full common ground as a regular, constant
@@
 non-automatic process that is used primarily in times of difficulty (when
-radical misalignment becomes apparent).''} (p. 179).  This view is actually not
+radical misalignment becomes apparent).~\citep[p.179]{pickering2006alignment}
+\end{quote}
+
+\noindent This view is actually not
 incompatible with Clark's \emph{grounding
@@
 or helping the peer to choose a picture among many. In the studies we report here, we
-explore several richer tasks such as arguing on a sensitive issue or building a
+explore several richer tasks such as arguing about a sensitive issue or building a
 concept map.
@@
 partner when they know that they can later provide feedback on his/her
-
 understanding~\citep{schober1993spatial}. We introduced the notion of
@@
 conflict. Conflict-resolution scripts such as the {\sc
-ArgueGraph}~\citep{dillenbourg2008mechanics} form pair of students with
+ArgueGraph}~\citep{dillenbourg2008mechanics} form pairs of students with
 opposite opinions, which increases the difficulty of consensus building,
-requiring more justifications, more negotiation, more load. Similarly, {\sc
+requiring more justifications, more negotiation, and more load. Similarly, {\sc
 jigsaw} scripts~\citep{aronson1978jigsaw} provide peers with different but
@@
 
-We refer to \emph{Partner modelling} as the process of inferring one's partner
+We refer to \emph{Partner modelling} as the process of inferring one's partner's
 mental states. Any claim that students carry out a detailed monitoring of their
@@
 builds some model of \B and \B build some model of \A.  Moreover, these two
-processes are not independent: \A's model of what \B knows, include what \B knows
+processes are not independent: \A's model of what \B knows, includes what \B knows
 about \A.  This leads to nested levels of  modelling. If \A states ``\B thinks I
@@
 is neither presented as a computational model of mutual modelling, nor as some
-universal formalism; its usefulness is internal to this paper.  Besides, we
+universal formalism; its usefulness is internal to this paper.  Additionally, we
 refer to the \emph{degree of accuracy} of the model as \Model{A}{B}{X}. We
@@
         initial and required degrees rather high) in order to make mutual
-        modelling effort more observable. Studies 1, 4 and 5 have been conducted
+        modelling effort more observable. Studies 1, 4, and 5 have been conducted
         with teams of students who did not know each other. They came
@@
         variable, with the hypothesis that modelling emotions is ``cheaper'' with a
-        richer medium, \ie  when peers can see each other.  Studies 1,2 and 4
+        richer medium, \ie  when peers can see each other.  Studies 1, 2, and 4
         use awareness tools which, in principle, reduce the cost of mutual modelling, but do
@@
 The experiments we report here address mutual modelling across different tasks,
-some with dyads, other with triads. They were conducted along 6 years in two
+some with dyads, others with triads. They were conducted over 6 years in two
 different institutions by different researchers. They used different
-independent, intermediate and dependent variables. Nonetheless, we were able to
+independent, intermediate, and dependent variables. Nonetheless, we were able to
 retroactively address 3 research questions across these studies.
@@
 %}
-\caption{Overview of the studies}
+\caption{Overview of the studies and main characteristics}
 \label{synthesis_table}
@@
 Thirty-six persons participated in this study, all native French speakers. We
-constituted 18 dyads who did not know each other beforehand. The pairs were
+formed 18 dyads who did not know each other beforehand. The pairs were
 randomly assigned to either the control condition (without the awareness tool)
@@
 
-How to interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
+How do we interpret a correlation of 0.38 between \gModel{A}{B} and \gModel{B}{A}?
 It is significant, which supports $\mathcal{H}_{2}$ and $\mathcal{H}_{3}$. It is
@@
 
-This study concerns a collaborative game that occurred in the physical
+This study concerns a collaborative game that occurred in physical
 space~\citep{nova2006underwhelming}. We studied whether players
@@
 {\sc CatchBob} is a mobile game in which groups of 3 players have to solve a
-joint task. The game was played on our campus. Participants had to find a
+joint task. The game was played on a university campus. Participants had to find a
 virtual object (\emph{Bob}) and to ``catch'' it by forming a triangle around it.
@@
 faded out after a few minutes to avoid covering the full display. The awareness
-tool displayed as well the location of the two other players on the map. Three
+tool also displayed the location of the two other players on the map. Three
 conditions were considered: the control condition (without tool) and two
@@
 Ninety students participated in this experiment. We only selected students from
-our campus since knowledge of the campus geography had an impact both on group
+university campus since knowledge of the campus geography had an impact both on group
 performance and on mutual modelling: to represent the path of someone across
@@
 
-The aim of this unpublished\footnote{The data and statistical analyses of this
+The aim of this unpublished\endnote{The data and statistical analyses of this
 study are available online:
@@
 This script increases the cognitive effort to build the map, not only to
-conciliate the viewpoints of each team members but, before that, to find out
+conciliate the viewpoints of each team member but, before that, to find out
 what the other knows.
@@
 two experimental conditions. Gender was balanced over the conditions.
-Participants did not know each other before the experiment. Students from Life
+Participants did not know each other before the experiment. Students from the Life
 Sciences faculty were not recruited to avoid high initial background knowledge on the
@@
 \paragraph{Grounding criterion} In this task, the grounding criterion was low.
-We did not however evaluate task performance (\eg the quality of the jointly
+We did not, however, evaluate task performance (\eg the quality of the jointly
 produced concept map) but learning gains. The correlation between \gModel{A}{B}
@@
 $\mathcal{H}_{1}$. However, the analysis of verbal interactions finds elements
-that actually supports the three hypotheses.
+that actually support the three hypotheses.
 
@@
 tasks assigned to learners in CSCL environments. In four different tasks, we
-found evidence that partner model each other since the quality of modelling
+found evidence that partners model each other since the quality of modelling
 emerged as an intermediate variable, sensitive to several independent variables
@@
 several dependent variables, such as task performance or learning gains.
-Thereby, even if we do not bring any definite conclusion to this debate, our
+Therefore, even if we do not bring any definite conclusion to this debate, our
 results support the CSCL school of thought in which shared understanding or
@@
 
-Secondarily, we questioned whether the accuracy of the partner's model depends on
+Second, we questioned whether the accuracy of the partner's model depends on
 the cognitive skills of each partner or instead results from the quality of
@@
 the process of helping one's partner to model oneself $\mathcal{H}_{2}$ are
-mediated by verbal interactions in the team. One could hardly imagine someone
-managing a very accurate modelling despite the very bad quality of interactions.
+mediated by verbal interactions in the team. It is difficult to imagine someone
+managing a very accurate modeling despite low quality interaction.
 There is a bidirectional causal link between the accuracy of mutual modelling
 and the quality of the interactions. This rather artificial distinction does
-however participate in the more fundamental discussion of the role of individual
+however contribute to the more fundamental discussion of the role of individual
 and social mechanisms in human cognition. In the field of social cognition, it
-is a commonplace to state that ``the whole is greater than the sum of the
+is commonplace to state that ``the whole is greater than the sum of the
 parts''.  This refers to the emergence of team properties than cannot be reduced
@@
 heavily rely on correlations; hence we cannot identify causal links. Second, we
-faced hard methodological issues. Providing learners with on-task questionnaires
+faced difficult methodological issues. Providing learners with on-task questionnaires
 introduces a bias: they will pay more attention to their partners in the
@@
 addressed different research questions (media richness, awareness tools,
-scripts,...). This diversity makes our results difficult to integrate, partly
-and contradictory. Nonetheless, this diversity also incidentally provides some
+scripts,...). This diversity makes our results difficult to integrate as they
+appear partly contradictory. Nonetheless, this diversity also incidentally provides some
 generalizability: mutual modelling has been investigated in different contexts
@@
 
+\theendnotes
+
 \end{document}
